### 2020-06-04 analysis/001_jn_list.Rmd
get journal info from apc price list and explore coverage with crossref index

---
title: "Elsevier APC Journal List information"
output: github_document
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

## Portfolio overview

### Load mined data

```{r}
library(tidyverse)
library(here)
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
```

### Journals per business model

```{r}
els_jns_df %>%
  count(oa_type) %>%
  mutate(prop = n / sum(n))
```

### List prices per business model

Only journals with APC in USD considered

```{r}
els_jns_df %>%
 # filter(apc_currency == "USD") %>%
  group_by(oa_type, apc_currency) %>%
  summarise(n_journals = n(),
            mean_apc = mean(apc, na.rm = TRUE),
            sd = sd(apc, na.rm = TRUE)) %>%
  arrange(desc(apc_currency))
```
```{r}
els_jns_df %>%
  filter(apc_currency == "USD") %>%
  ggplot(aes(oa_type, apc, fill = oa_type)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
  colorblindr::scale_fill_OkabeIto() +
  cowplot::theme_minimal_hgrid(12) +
  theme(legend.position = "none", 
        plot.title.position = "plot") +
  labs(x = "OA Business Model", y = "APC (in USD)",
       title = "Elsevier APC List Prices by Open Access Business Model",
       caption = "Elsevier: 30 May 2020")
```

### 2020-06-04 analysis/002_cr_els_matching.Rmd
get journal info from apc price list and explore coverage with crossref index

---
title: "Obtain metadata from Crossref Dump"
output: html_notebook
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

### Connect to database

```{r}
# connect to google bg where we imported the json lines Unpaywall dump
library(DBI)
library(bigrquery)
con <- dbConnect(
  bigrquery::bigquery(),
  project = "api-project-764811344545",
  dataset = "cr_dump_march_20"
)
```

### Journals associated with Elsevier

Determined by the Crossref member id (78).

```{sql connection=con, output.var="els_2020_jns"}
SELECT DISTINCT(ISSN)
FROM `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` 
WHERE member = '78' AND issued_year BETWEEN 2015 AND 2019
```

### Comparision with Crossref Dump

Journals from hybrid Elsevier list not indexed in Crossref

```{r}
library(tidyverse)
library(here)
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
els_jns_df %>% 
  filter(!issn %in% els_2020_jns$ISSN, oa_type == "Hybrid")
```

Possible reason:

a) Journal transfer including metadata ownership. Example: Canadian Association of Radiologists Journal (transferred from Elsevier to SAGE Publications as of 2020)

```{sql connection=con}
SELECT member, publisher, COUNT(DISTINCT(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` 
WHERE container_title = 'Canadian Association of Radiologists Journal'
GROUP BY member, publisher
```

However, there are cases, where metadata ownership was not fully transferred. Example: Acta Mathematica Scientia (transferred to Springer Nature as of 2019)

```{sql connection=con}
SELECT member, publisher, COUNT(DISTINCT(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` 
WHERE container_title = 'Acta Mathematica Scientia'
GROUP BY member, publisher
```

b) Journal started in 2020

### 2020-06-13 analysis/002_cr_els_matching.Rmd
match apc pricing list and elsevier data

@@
 title: "Obtain metadata from Crossref Dump"
-output: html_notebook
+output: github_document
 ---
@@
 
-Possible reason:
+Possible reasons:
 
@@
 b) Journal started in 2020
+
+### Create Crossref Elsevier subset in BQ
+
+Upload list of Elsevier hybrid journals to BQ.
+
+```{r}
+els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
+els_jn_hybrid <- els_jns_df %>% 
+  filter(oa_type == "Hybrid")
+bq_els <- bq_table_create(
+  "api-project-764811344545.cr_dump_march_20.elshybrid",
+  fields =  els_jn_hybrid,
+  friendly_name = "Elsevier hybrid journals May 2020",
+  description = "The data was extracted from Elsevier APC pricing list, 
+    downloaded 30 May 2020"
+)
+bq_table_upload(bq_els, els_jn_hybrid)
+```
+
+Create table with Crossref metadata for these journals since 2008.
+
+```{sql, connection=con}
+CREATE OR REPLACE TABLE `api-project-764811344545.cr_dump_march_20.els_hybrid_cr` AS
+SELECT * FROM `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` as cr
+WHERE EXISTS ( 
+  SELECT 1 FROM `api-project-764811344545.cr_dump_march_20.elshybrid` as els_hybrid 
+  WHERE (cr.issn = els_hybrid.issn)
+  )
+```
+
+Descriptive:
+
+By members
+
+```{sql, connection=con}
+SELECT  member, publisher, COUNT(DISTINCT(doi)) as articles
+FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`
+WHERE issued_year > 2014 and issued_year < 2020
+GROUP BY member, publisher
+ORDER BY articles DESC
+```
+
+License prevalence
+
+```{sql, connection=con}
+SELECT REGEXP_REPLACE(license.URL, "http://|https://", "") as license_url, COUNT(DISTINCT(doi)) as articles
+FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`, unnest(license) as license
+WHERE issued_year > 2014 and issued_year < 2020 AND license.content_version = 'vor'
+GROUP BY license_url
+ORDER BY articles DESC
+```
+
+License prevalence by days with delay
+
+```{sql, connection=con, output.var="els_license_delay"}
+SELECT 
+  REGEXP_REPLACE(license.URL, "http://|https://", "") as license_url,
+  license.delay_in_days,
+  COUNT(DISTINCT(doi)) as articles
+FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`, unnest(license) as license
+WHERE issued_year > 2014 and issued_year < 2020 AND license.content_version = 'vor'
+GROUP BY license_url, license.delay_in_days
+ORDER BY articles DESC
+```
+
+Immediate Licenses
+
+```{r}
+els_license_delay %>%
+  filter(delay_in_days == 0)
+```

### 2020-06-14 analysis/002_cr_els_matching.Rmd
prepare parsing for invoice data

@@
   filter(oa_type == "Hybrid")
+if(!bq_table_exists("api-project-764811344545.cr_dump_march_20.elshybrid")) {
 bq_els <- bq_table_create(
@@
 bq_table_upload(bq_els, els_jn_hybrid)
+} else {
+  NULL
+}
 ```
@@
 ```
+
+A few records have a date formats without day, which is used for the delayed calculation. We should therefore allow for a delay of up to  30 days.
+
+```{r}
+els_license_delay %>%
+  filter(delay_in_days < 31) 
+```
+
+### Download open access articles in hybrid journals
+
+```{sql, connection=con, output.var="els_hybrid_articles"}
+SELECT *
+FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`, unnest(license) as license
+WHERE (REGEXP_CONTAINS(license.URL, 'creativecommons') AND license.content_version = 'vor') AND 
+  (issued_year BETWEEN 2015 and 2019) AND
+  license.delay_in_days < 31
+```
+
+Backup
+
+```{r}
+require(jsonlite)
+jsonlite::stream_out(els_hybrid_articles, file(here::here("data", "els_hybrid_articles.json")))
+```

### 2020-06-14 analysis/003_cr_xml_fetch.Rmd
prepare parsing for invoice data

---
title: "003_ftx_mining.Rmd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Obtain Elsevier ftxt links for OA articles

Libraries

```{r}
require(jsonlite)
require(tidyverse)
require(crminer)
require(xml2)
require(here)
```

Load Elsevier articles:

```{r}
els_hybrid <- jsonlite::stream_in(file(here::here("data", "els_hybrid_articles.json")))
```

get tdm links

```{r}
elsevier_tdm <- els_hybrid %>%
  select(doi, link) %>%
  unnest(link) %>%
  filter(content_type == "text/xml",
         intended_application == "text-mining",
         content_version == "vor")
```

Apply function on every XML full-text link

```{r}
tt <- purrr::map(sample(elsevier_tdm$URL, 10), .f = purrr::safely(elsevier_parse))
```

### 2020-06-14 R/parse_els_xml.R

@@ -1,0 +1,30 @@
+#' Get OA info from Elsevier full-text
+#'
+#' @param tdm_url link to Elsevier XML full-text
+#'
+#' @return
+#' @importFrom crminer crm_xml
+#' @importFrom xml2 xml_find_first xml_text
+#' @importFrom tibble tibble
+#' @export
+#'
+#' @examples
+elsevier_parse <- function(tdm_url) {
+  req <- crminer::crm_xml(tdm_url)
+  oa_sponsor_name <-  xml2::xml_text(
+    xml2::xml_find_first(req, "//d1:coredata//d1:openaccessSponsorName")
+    )
+  oa_sponsor_type <- xml2::xml_text(
+    xml2::xml_find_first(req, "//d1:coredata//d1:openaccessSponsorType")
+    )
+  oa_article <-  xml2::xml_text(
+    xml2::xml_find_first(req, "//d1:coredata//d1:openaccessArticle")
+  )
+  oa_type <- xml2::xml_text(
+    xml2::xml_find_first(req, "//d1:coredata//d1:openaccessType")
+  )
+  oa_archive <-  xml2::xml_text(
+    xml2::xml_find_first(req, "//d1:coredata//d1:openArchiveArticle")
+    )
+  tibble(oa_sponsor_name, oa_sponsor_type, oa_article, oa_type, oa_archive, tdm_url)
+}

### 2020-06-14 analysis/002_cr_els_matching.Rmd
no delay info from crossref. instead, the open archive info from elsevier is used

@@
 A few records have a date formats without day, which is used for the delayed calculation. We should therefore allow for a delay of up to  30 days. 
-
 ```{r}
@@
 ```
+
+But it seems that also some articles misses months. I will therefore obtain all cc tagged articles and validate immediate oa provision via the open archive tag
 
@@
 WHERE (REGEXP_CONTAINS(license.URL, 'creativecommons') AND license.content_version = 'vor') AND 
-  (issued_year BETWEEN 2015 and 2019) AND
-  license.delay_in_days < 31
+  (issued_year BETWEEN 2015 and 2019) 
 ```
 
-Backup
+Dataset with TDM Links
 
 ```{r}
-require(jsonlite)
-jsonlite::stream_out(els_hybrid_articles, file(here::here("data", "els_hybrid_articles.json")))
+els_tdm <- els_hybrid_articles %>%
+  select(doi, link) %>%
+  unnest(link) %>%
+  filter(content_type == "text/xml",
+         intended_application == "text-mining",
+         content_version == "vor")
+write_csv(els_tdm, here::here("data", "els_tmd_links.csv"))
 ```

### 2020-06-14 analysis/003_cr_xml_fetch.Rmd
no delay info from crossref. instead, the open archive info from elsevier is used

@@
 ```{r}
-require(jsonlite)
 require(tidyverse)
@@
 require(here)
+source(here::here("R", "parse_els_xml.R"))
 ```
@@
 ```{r}
-els_hybrid <- jsonlite::stream_in(file(here::here("data", "els_hybrid_articles.json")))
+els_hybrid <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
 ```
 
-get tdm links
-
 ```{r}
-elsevier_tdm <- els_hybrid %>%
-  select(doi, link) %>%
-  unnest(link) %>%
-  filter(content_type == "text/xml",
-         intended_application == "text-mining",
-         content_version == "vor")
+tt <- plyr::llply(sample(elsevier_tdm$URL, 10), .f = purrr::safely(elsevier_parse), .progress = "text")
 ```
-
-Apply function on every XML full-text link
-
-```{r}
-tt <- purrr::map(sample(elsevier_tdm$URL, 10), .f = purrr::safely(elsevier_parse))
-```

### 2020-06-14 analysis/003_cr_xml_fetch.Rmd
invoice data

@@
 ```{r}
-tt <- plyr::llply(sample(elsevier_tdm$URL, 10), .f = purrr::safely(elsevier_parse), .progress = "text")
+tt <- plyr::llply(els_hybrid$URL, .f = purrr::safely(elsevier_parse), .progress = "text")
+tt_df <- purrr::map_df(tt, "result")
+write_csv(tt_df, here::here("data", "invoice_df.csv"))
 ```

### 2020-06-15 analysis/004_EDA.Rmd
starting eda

---
title: "EDA Hybrid OA"
author: "Najko Jahn"
date: "6/15/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)
require(here)
els_invoicing <- readr::read_csv(here::here("data", "invoice_df.csv"))
els_tdm <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
els_df <- els_tdm %>%
  select(doi, URL) %>%
  inner_join(els_invoicing, by = c("URL" = "tdm_url")) %>% 
  distinct()
```

### By sponsor type and archive

```{r}
els_df %>%
  count(oa_sponsor_type, oa_archive)
```

### By invoice recipient

```{r}
els_df %>%
  count(oa_sponsor_name) %>%
  arrange(desc(n))
```

### compare with existing table

```{r}
library(googlesheets4)
library(writexl)
googlesheets4::gs4_deauth()
gs_df <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/")
miss_funder <- els_df %>%
  filter(!oa_sponsor_name %in% gs_df$oa_sponsor_name) %>%
  group_by(oa_sponsor_name) %>%
  summarise(n = n_distinct(doi))

writexl::write_xlsx(miss_funder, here::here("data", "new_funders.xlsx"))
```

### 2020-07-07 analysis/005_article_volume.Rmd
obtain article volume

---
title: "Obtain Article Volume"
author: "Najko Jahn"
date: "7/7/2020"
output: github_document
---

```{r, setup, echo=FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
library(tidyverse)
library(here)
```

### Connect to database

```{r}
# connect to google bg where we imported the json lines Unpaywall dump
library(DBI)
library(bigrquery)
con <- dbConnect(
  bigrquery::bigquery(),
  project = "api-project-764811344545",
  dataset = "cr_dump_march_20"
)
```

### Overall article volume per journal and year

```{sql, connection = "con", output.var = "article_volume"}
SELECT issued_year, issn, count(distinct(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`
WHERE issued_year > 2014 and issued_year < 2020
group by issued_year, issn
```

How many articles by year?

```{r}
article_volume %>%
  group_by(issued_year) %>%
  summarise(n = sum(articles))
```

backup

```{r}
write_csv(article_volume, here::here("data", "article_volume.csv"))
```

### Article volume with at least one reference per journal and year

```{sql, connection = "con", output.var = "ref_article_volume"}
SELECT issued_year, issn, count(distinct(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`
WHERE reference_count > 0 AND (issued_year > 2014 and issued_year < 2020)
group by issued_year, issn
```

```{r}
ref_article_volume %>%
  group_by(issued_year) %>%
  summarise(n = sum(articles))
```

backup

```{r}
write_csv(ref_article_volume, here::here("data", "ref_article_volume.csv"))
```

### Validation: Do these journals actually deposit references with Crossref

```{r}
zero_ref_jns <- article_volume %>% 
  filter(!issn %in% ref_article_volume$issn) %>% 
  distinct(issn)
```

Do these journals deposit references?

```{r}
library(rcrossref)
cr_jns <- rcrossref::cr_journals(zero_ref_jns$issn)
table(current = cr_jns$data$deposits_references_current, back_file =  cr_jns$data$deposits_references_backfile)
```

All journals deposited references with Crossref!

### 2020-07-07 analysis/results.Rmd
start answering research questions

---
title: "results"
author: "Najko Jahn"
date: "7/7/2020"
output: github_document
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

## Results

### Uptake

#### How many Elsevier journals supported the hybrid model? How many of them flipped to fully open access?

to do

#### What is the uptake of open access in Elsevier's hybrid journal portfolio?

*Number and percentage of journals with at least one article*

```{r}
library(tidyverse)
library(here)
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
```

To determine the number of Elsevier hybrid journals with at least one open access article, we downloaded and parsed the publisher's APC pricing list from 31 May 2020. The list comprised `r length(unique(els_jns_df$issn))` journals, of which `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model, representing a share of `r round(els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() / length(unique(els_jns_df$issn)) * 100, 0)`%. 
```{r}
hybrid_md <- readr::read_csv(here::here("data", "hybrid_md.csv"))
els_tdm <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
invoice_df <-  readr::read_csv(here::here("data", "invoice_df.csv"))

hybrid_df <- els_tdm %>% 
  select(doi, URL) %>% 
  inner_join(invoice_df, by = c("URL" = "tdm_url")) %>% 
  select(-URL) %>% 
  inner_join(hybrid_md, by = c("doi")) %>% 
  distinct()
```

Of those, `r hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article immediately under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio. At the same time, `r 100 - round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of hybrid journals did not published an open access article in this period and were, thus, excluded from further analysis.

```{r}
eligible_jns <- hybrid_df %>% 
  filter(oa_archive == FALSE) %>% 
  distinct(issn)
```

*Number and percentage of open access articles per hybrid journal*

We used the Crossref index to determine the overall article volume of Elsevier's hybrid journal portfolio, and to related it to the number of open access articles.
Because DOIs can be also issued to abstracts presented at scientific meetings and non-scholarly journal content including table of contents and list of reviewers, we furthermore distinguished between citing and non-citing articles defined by the number of deposited references in Crossref per article, regardless of whether they are public or not.

Table presents the high-level findings by year of publication. illustrating a growth

### 2020-07-18 analysis/results.Rmd
development of elseviers oa journal portfolio

@@
 date: "7/7/2020"
+bibliography: references.bib
+csl: chicago.csl
 output: github_document
@@
   message = FALSE,
-  echo = TRUE,
+  echo = FALSE,
   fig.width = 6,
@@
 
+```{r}
+# libraries
+library(tidyverse)
+library(here)
+library(cowplot)
+```
+
 ## Results
@@
 
-to do
+Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio including hybrid journals.
+Figure shows that the majority of journals supported the hybrid model.
+The year 2016 marked the highest number of journals.
+After that, the trend reversed.
+Since 2018, the number of journals stagnated.
 
-#### What is the uptake of open access in Elsevier's hybrid journal portfolio?
+```{r}
+els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv")) 
+els_jns_count <- els_jns %>%
+  group_by(year, oa_model) %>%
+  count(year) %>%
+  group_by(year) %>%
+  mutate(prop = n / sum(n)) %>%
+  mutate(oa_model_fct = fct_rev(as_factor(oa_model))) 
+ggplot(els_jns_count,
+       aes(as.character(year), n,  fill = oa_model_fct, group = oa_model_fct)) +
+  geom_area(alpha = 0.8,
+            colour = "white") +
+  scale_fill_manual(
+    values = c("#cccccca0", "#56b4e9"),
+    name = NULL,
+    labels = c("Fully", "Hybrid")
+  ) +
+  labs(x = "Year", y = "Elsevier Journals") +
+  scale_y_continuous(
+    labels = scales::number_format(big.mark = ","),
+    expand = expansion(mult = c(0, 0.05)),
+    breaks =  scales::extended_breaks()
+  ) +
+  theme_minimal_hgrid() +
+  theme(legend.position = "top",
+        legend.justification = "right") +
+  guides(fill = guide_legend("OA Business Model"))
+```
+
+Compared to 2015, the number of hybrid journals grew by  `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$n) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$n)` journals, while its proportion relative to Elsevier's overall journal portfolio slightly decreased from `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$prop, 2) * 100`% to `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$prop, 2) * 100`%.
+
+```{r}
+flipped_jns <- els_jns %>%
+  distinct(issn, oa_model) %>%
+  group_by(issn) %>%
+  filter(n() > 1)
+flipped_jns_ind <- els_jns %>%
+  filter(issn %in% flipped_jns$issn) %>%
+  group_by(issn) %>%
+  slice(which.max(year)) %>%
+  ungroup() %>%
+  count(oa_model)
+```
+
+Following the approach of an earlier study [@Matthias_2019], we furthermore used APC pricing list information to determine the number of open access business model flips between 2015-2020.
+Although the transition from hybrid to a fully open access business model has drawn most attention, also the reverse can be observed.
+A total number of `r sum(flipped_jns_ind$n)` journals changed the business model under the ownership of Elsevier.
+Of these, `r filter(flipped_jns_ind, oa_model == "Open Access")$n` journals flipped to fully open access, while `r filter(flipped_jns_ind, oa_model != "Open Access")$n`changed to hybrid open access.
+Examples include the physical journal *Materials Today* and the life science journal *Meta Gene*.
+
+#### What is the uptake of open access in Elsevier's current hybrid journal portfolio?
 
@@
 
+<!-- here, we change to the most current pricing list, which seems to be identical to 31 May-->
+
 ```{r}
-library(tidyverse)
-library(here)
 els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
@@
 
-To determine the number of Elsevier hybrid journals with at least one open access article, we downloaded and parsed the publisher's APC pricing list from 31 May 2020.
-The list comprised `r length(unique(els_jns_df$issn))` journals, of which `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model, representing a share of `r round(els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() / length(unique(els_jns_df$issn)) * 100, 0)`%.
 ```{r}
@@
 
+To determine the number of Elsevier hybrid journals with at least one open access article, we downloaded and parsed the publisher's APC pricing list from 31 May 2020.
+The list comprised `r length(unique(els_jns_df$issn))` journals, of which `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model, representing a share of `r round(els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() / length(unique(els_jns_df$issn)) * 100, 0)`%.
 Of those, `r hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article immediately under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio.
-At the same time, `r 100 - round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of hybrid journals did not published an open access article in this period and were, thus, excluded from further analysis.
+At the same time, `r 100 - round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of hybrid journals did not published an open access article.
 

### 2020-07-18 R/fetch_els_historic.R

@@ -1,0 +1,7 @@
+library(tidyverse)
+library(here)
+u <- "https://raw.githubusercontent.com/lmatthia/publisher-oa-portfolios/master/elsevier_oa_and_hybrid.csv"
+els_jns <- readr::read_delim(u, delim = ";")
+tt <- els_jns %>%
+  filter(year > 2014) 
+write_csv(tt, here::here("data", "els_historic_jns.csv"))

### 2020-07-24 analysis/001_jn_list.Rmd
check with @lmatthia apc list

@@
 ```
+
+## Comparision with Lisa Matthias list
+
+Are both lists equal in terms of hybrid journals?
+
+```{r}
+u <- "https://raw.githubusercontent.com/lmatthia/publisher-oa-portfolios/master/elsevier_oa_and_hybrid.csv"
+l_apc <- readr::read_delim(u, delim = ";")
+l_apc %>%
+  filter(year == 2020) %>%
+  count(oa_model)
+els_hybrid <- els_jns_df %>%
+  filter(oa_type == "Hybrid")
+l_hybrid <- l_apc %>%
+  filter(oa_model == "Hybrid", year == 2020)
+base::setequal(els_hybrid$issn, l_hybrid$issn)
+```

### 2020-07-24 analysis/results.Rmd
check with @lmatthia apc list

@@
 
-#### How many Elsevier journals supported the hybrid model? How many of them flipped to fully open access?
-
-Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio including hybrid journals.
-Figure shows that the majority of journals supported the hybrid model.
-The year 2016 marked the highest number of journals.
-After that, the trend reversed.
-Since 2018, the number of journals stagnated.
-
 ```{r}
@@
   mutate(oa_model_fct = fct_rev(as_factor(oa_model))) 
+```
+
+#### How many Elsevier journals supported the hybrid model?
+
+Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio including hybrid journals.
+Figure shows that the majority of journals supported the hybrid model.
+The year 2016 marked the highest number of journals.
+After that, the trend reversed.
+Since 2018, the number of journals stagnated.
+Compared to 2015, the number of hybrid journals grew by  `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$n) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$n)` journals, while its proportion relative to Elsevier's overall journal portfolio slightly decreased from `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$prop, 2) * 100`% to `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$prop, 2) * 100`%.
+
+```{r}
 ggplot(els_jns_count,
@@
 
-Compared to 2015, the number of hybrid journals grew by  `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$n) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$n)` journals, while its proportion relative to Elsevier's overall journal portfolio slightly decreased from `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$prop, 2) * 100`% to `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$prop, 2) * 100`%.
+#### How many of hybrid journals flipped to fully open access?
 
@@
 Of these, `r filter(flipped_jns_ind, oa_model == "Open Access")$n` journals flipped to fully open access, while `r filter(flipped_jns_ind, oa_model != "Open Access")$n`changed to hybrid open access.
-Examples include the physical journal *Materials Today* and the life science journal *Meta Gene*.
+Examples for such reverse flipped open access journals include the physical journal *Materials Today* and the life science journal *Meta Gene*.
 
@@
 *Number and percentage of journals with at least one article*
-
-<!-- here, we change to the most current pricing list, which seems to be identical to 31 May-->
 
@@
 
-To determine the number of Elsevier hybrid journals with at least one open access article, we downloaded and parsed the publisher's APC pricing list from 31 May 2020.
-The list comprised `r length(unique(els_jns_df$issn))` journals, of which `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model, representing a share of `r round(els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() / length(unique(els_jns_df$issn)) * 100, 0)`%.
-Of those, `r hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article immediately under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio.
+To determine the number of Elsevier hybrid journals with at least one open access article, we, again, started with APC pricing list information form Elsevier [@Matthias_2020], lastly updated on 23 May 2020.
+We used the Crossref index to determine the overall article volume of Elsevier's hybrid journal portfolio, and to related it to the number of immediate open access articles indicated by Creative Commons licenses.
+Because DOIs can be also issued to  non-scholarly journal content including table of contents and list of reviewers, we excluded these records using Unpaywall's paratext recognition approach [^1].
+They account for about 2.5% of DOIs registered by Crossref.
+
+[^1]: <https://support.unpaywall.org/support/solutions/articles/44001894783-what-does-is-paratext-mean-in-the-api->
+
+Elsevier APC pricing list comprised `r length(unique(els_jns_df$issn))` journals, of which `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model, representing a share of `r round(els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() / length(unique(els_jns_df$issn)) * 100, 0)`%.
+Of those, `r hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio.
 At the same time, `r 100 - round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of hybrid journals did not published an open access article.
@@
 
-We used the Crossref index to determine the overall article volume of Elsevier's hybrid journal portfolio, and to related it to the number of open access articles.
-Because DOIs can be also issued to abstracts presented at scientific meetings and non-scholarly journal content including table of contents and list of reviewers, we furthermore distinguished between citing and non-citing articles defined by the number of deposited references in Crossref per article, regardless of whether they are public or not.
-
 Table presents the high-level findings by year of publication. illustrating a growth

### 2020-07-24 analysis/005_article_volume.Rmd
include paratext recognition

@@
 ```{r}
-article_volume %>%
+els_vol <- article_volume %>%
   group_by(issued_year) %>%
   summarise(n = sum(articles))
+els_vol
 ```
@@
 write_csv(article_volume, here::here("data", "article_volume.csv"))
+```
+
+### Article volume excluding paratext
+
+```{sql, connection = "con", output.var = "article_volume_para"}
+SELECT
+        issued_year,
+        issn,
+        count(distinct(doi)) as articles     
+    FROM
+        `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`      
+    WHERE
+        issued_year > 2014 
+        and issued_year < 2020     
+        AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
+    GROUP BY
+        issued_year,
+        issn
+```
+
+How many articles by year?
+
+```{r}
+article_volume_para %>%
+  group_by(issued_year) %>%
+  summarise(articles = sum(articles)) %>%
+  inner_join(els_vol, by = "issued_year") %>%
+  mutate(para = n - articles) %>%
+  mutate(para / n * 100)
+```
+backup
+
+```{r}
+write_csv(article_volume_para, here::here("data", "article_volume_para.csv"))
 ```

### 2020-07-25 analysis/004_EDA.Rmd
starting with summary table, formating issue.

@@
 ```
+
+### Master dataset
+
+```{r}
+els_invoicing <- readr::read_csv(here::here("data", "invoice_df.csv"))
+els_tdm <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
+hybrid_md <- readr::read_csv(here::here("data", "hybrid_md.csv"))
+els_df <- els_tdm %>%
+  select(doi, URL) %>%
+  inner_join(els_invoicing, by = c("URL" = "tdm_url")) %>% 
+  distinct()
+els_md <- els_df %>% 
+  rename(tdm_url = URL) %>% 
+  inner_join(hybrid_md, by = "doi") %>% 
+  distinct(doi, .keep_all = TRUE)
+readr::write_csv(els_md, here::here("data", "els_analyses_main.csv"))
+````

### 2020-07-25 analysis/results.Rmd
starting with summary table, formating issue.

@@
     if (is.numeric(x)) {
-      return(prettyNum(x, big.mark = ","))
+      return(formatC(x, big.mark = ",", format = "g"))
     } else{
@@
 ```{r}
+els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
 els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv")) 
@@
 
+```{r}
+els_overall <- els_jns %>% 
+  group_by(oa_model) %>% 
+  summarise(n = n_distinct(issn))
+```
+
 Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio including hybrid journals.
 Figure shows that the majority of journals supported the hybrid model.
+Overall, Elsevier published `r length(unique(els_jns$issn))`, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.
+
+Elseviers journal portfolio changed over the period of observation.
+The number of hybrid journals grew by  `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$n) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$n)` journals between 2015 and May 2020, while its proportion relative to Elsevier's overall journal portfolio slightly decreased from `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$prop, 2) * 100`% to `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$prop, 2) * 100`%.
 The year 2016 marked the highest number of journals.
@@
 Since 2018, the number of journals stagnated.
-Compared to 2015, the number of hybrid journals grew by  `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$n) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$n)` journals, while its proportion relative to Elsevier's overall journal portfolio slightly decreased from `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$prop, 2) * 100`% to `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$prop, 2) * 100`%.
+On 23 May 2020, Elsevier APC pricing list comprised `r length(unique(els_jns_df$issn))` journals, of which `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model, representing a share of `r round(els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() / length(unique(els_jns_df$issn)) * 100, 0)`%.
 
@@
 
-Following the approach of an earlier study [@Matthias_2019], we furthermore used APC pricing list information to determine the number of open access business model flips between 2015-2020.
+Following the approach of our earlier study [@Matthias_2019], we furthermore used APC pricing list information to determine the number of open access business model flips between 2015-2020.
 Although the transition from hybrid to a fully open access business model has drawn most attention, also the reverse can be observed.
 A total number of `r sum(flipped_jns_ind$n)` journals changed the business model under the ownership of Elsevier.
-Of these, `r filter(flipped_jns_ind, oa_model == "Open Access")$n` journals flipped to fully open access, while `r filter(flipped_jns_ind, oa_model != "Open Access")$n`changed to hybrid open access.
+Of these, `r filter(flipped_jns_ind, oa_model == "Open Access")$n` journals flipped to fully open access, while `r filter(flipped_jns_ind, oa_model != "Open Access")$n` changed to hybrid open access.
 Examples for such reverse flipped open access journals include the physical journal *Materials Today* and the life science journal *Meta Gene*.
@@
 
-To determine the number of Elsevier hybrid journals with at least one open access article, we, again, started with APC pricing list information form Elsevier [@Matthias_2020], lastly updated on 23 May 2020.
-We used the Crossref index to determine the overall article volume of Elsevier's hybrid journal portfolio, and to related it to the number of immediate open access articles indicated by Creative Commons licenses.
-Because DOIs can be also issued to  non-scholarly journal content including table of contents and list of reviewers, we excluded these records using Unpaywall's paratext recognition approach [^1].
+To determine the number of Elsevier hybrid journals with at least one open access article, we focused on the most current version of Elsevier's APC pricing list, lastly updated on 23 May 2020 [@Matthias_2020].
+We used the Crossref index to determine the overall article volume of Elsevier's hybrid journal portfolio, and related it to the number of open access articles indicated by Creative Commons licenses.
+We only considered open access articles that were provided without delay, which we determined from mined full-text metadata.
+Because Elsevier also issued DOIs to non-scholarly journal content including table of contents and list of reviewers, we excluded these records using Unpaywall's paratext recognition approach [^1].
 They account for about 2.5% of DOIs registered by Crossref.
@@
 
-Elsevier APC pricing list comprised `r length(unique(els_jns_df$issn))` journals, of which `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model, representing a share of `r round(els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() / length(unique(els_jns_df$issn)) * 100, 0)`%.
+On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model.
 Of those, `r hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio.
@@
 
-*Number and percentage of open access articles per hybrid journal*
+*Number and percentage of immediate open access articles per hybrid journal*
+
+```{r}
+# required data spreadsheets
+# pub volume
+els_pub_volume <- readr::read_csv(
+  here::here("data", "article_volume_para.csv")
+  ) 
+# eligble journals 
+eligible_jns <- hybrid_df %>% 
+  filter(oa_archive == FALSE) %>% 
+  distinct(issn)
+# pub volume eligble data
+els_yr <- els_pub_volume %>%
+  filter(issn %in% eligible_jns$issn) %>%
+  group_by(issued_year) %>%
+  summarise(articles = sum(articles),
+            journals = n_distinct(issn))
+
+# oa
+ind_per_year <- hybrid_df %>%
+  filter(oa_archive == FALSE) %>%
+  group_by(issued_year) %>%
+  summarise(oa_articles = n()) %>%
+  inner_join(els_yr, by = "issued_year") %>%
+  ungroup() %>%
+  mutate(prop = oa_articles / articles * 100)
+ind_summary_table <- ind_per_year %>% 
+  pivot_longer(oa_articles:prop, names_to = "indicator", values_to = "count") %>%
+  pivot_wider(names_from = issued_year, values_from = count) %>%
+  mutate(indicator = factor(indicator, levels = c("journals", "articles", "oa_articles", "prop"))) %>%
+  arrange(indicator) %>%
+  mutate(indicator = recode(indicator,
+                            oa_articles = "Hybrid OA Articles", 
+                            articles = "All Articles", 
+                            journals = "Hybrid journals with at least one OA article", 
+                            prop = "Percentage"))
+`Total` <- c(
+  hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow(),
+  sum(ind_summary_table[2,2:6]),
+  sum(ind_summary_table[3,2:6]),
+  sum(ind_per_year$oa_articles) / sum(ind_per_year$articles) * 100)
+bind_cols(ind_summary_table, Total = `Total`) %>%
+  rename(` ` = indicator) %>%
+  knitr::kable(digits = 1)
+```
 

### 2020-07-26 analysis/results.Rmd
tmp save

@@
     if (is.numeric(x)) {
-      return(formatC(x, big.mark = ",", format = "g"))
+      return(formatC(x, big.mark = ",", format = "fg", digits = 2))
     } else{
@@
 
-Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio including hybrid journals.
+Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio.
 Figure shows that the majority of journals supported the hybrid model.
-Overall, Elsevier published `r length(unique(els_jns$issn))`, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.
+Overall, Elsevier published `r length(unique(els_jns$issn))` open access journals, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.
 
-Elseviers journal portfolio changed over the period of observation.
+Elsevier's journal portfolio changed over the period of observation.
 The number of hybrid journals grew by  `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$n) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$n)` journals between 2015 and May 2020, while its proportion relative to Elsevier's overall journal portfolio slightly decreased from `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$prop, 2) * 100`% to `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$prop, 2) * 100`%.
@@
 
-Following the approach of our earlier study [@Matthias_2019], we furthermore used APC pricing list information to determine the number of open access business model flips between 2015-2020.
+Following the approach of our earlier study [@Matthias_2019], we started with APC pricing list information to determine the number of open access business model flips between 2015-2020.
 Although the transition from hybrid to a fully open access business model has drawn most attention, also the reverse can be observed.
@@
   sum(ind_per_year$oa_articles) / sum(ind_per_year$articles) * 100)
-bind_cols(ind_summary_table, Total = `Total`) %>%
+uptake_table <- bind_cols(ind_summary_table, Total = `Total`) %>%
   rename(` ` = indicator) %>%
@@
 
-Table presents the high-level findings by year of publication. illustrating a growth
+Overall, `r Total[1]` hybrid journals provided immediate open access to `r Total[3]` articles between 2015 and 2019.
+The total share of hybrid open access in the publication volume of Elsevier journals was `r Total[4]`%.
+Table present the high-level findings by year, illustrating a moderate growth in hybrid open access.
+Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_articles)` (2015) to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_articles)` (2019).
+However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(prop)`% (2015) to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(prop)`% (2019), because the number of articles published also increased during this five-years period.
+
+```{r}
+uptake_table
+```
+
+```{r}
+els_jn_pub_volume <- els_pub_volume %>%
+  filter(issn %in% eligible_jns$issn) %>%
+  group_by(issued_year, issn) %>%
+  summarise(articles = sum(articles),
+            journals = n_distinct(issn))
+hybrid_oa_per_jn <- hybrid_df %>%
+  filter(oa_archive == FALSE) %>%
+  group_by(issued_year, issn) %>%
+  summarise(oa_articles = n())%>%
+  inner_join(els_jn_pub_volume, by = c("issued_year", "issn")) %>%
+  ungroup() %>%
+  mutate(prop = oa_articles / articles * 100) %>% View()
+```
+
+The open access share varied across Elsevier hybrid journals.
+Since 2015, most hybrid journals have had a slow uptake rate of open access articles.
+In general, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume.
+
+# license

### 2020-08-13 analysis/results.Rmd
paratext recognition oa articles

@@
 ---
-title: "results"
-author: "Najko Jahn"
-date: "7/7/2020"
 bibliography: references.bib
@@
 ```{r}
+# paratext pattern
+para <- c('^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
+```
+
+```{r}
 hybrid_md <- readr::read_csv(here::here("data", "hybrid_md.csv"))
@@
   inner_join(hybrid_md, by = c("doi")) %>% 
-  distinct()
+  distinct() %>%
+  filter(!grepl(para, title))
+
+flipped_jns_df <-
+  els_jns %>%
+    filter(issn %in% flipped_jns$issn)
 ```

### 2020-08-13 analysis/results.Rmd
variation

@@
 library(cowplot)
+library(ggeconodist)
 ```
@@
 Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio.
-Figure shows that the majority of journals supported the hybrid model.
+Figure shows that the majority of journals were hybrid journals.
 Overall, Elsevier published `r length(unique(els_jns$issn))` open access journals, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.
@@
 
-flipped_jns_df <-
-  els_jns %>%
-    filter(issn %in% flipped_jns$issn)
+flipped_jns_df <- els_jns %>%
+  filter(issn %in% flipped_jns$issn) %>%
+  select(oa_model, issn, year)
+
+hybrid_df <- left_join(hybrid_df, flipped_jns_df, by = c("issn", "issued_year" = "year")) %>%
+  filter(!oa_model %in% "Open Access") %>%
+  select(-oa_model)
 ```
@@
   ungroup() %>%
-  mutate(prop = oa_articles / articles * 100) %>% View()
+  mutate(prop = oa_articles / articles) 
 ```
 
-The open access share varied across Elsevier hybrid journals.
+```{r}
+perc_df <- hybrid_oa_per_jn %>%
+  group_by(issued_year) %>% 
+  summarize(median = median(prop),
+            p_10 = quantile(prop, 0.1),
+            p_90 = quantile(prop, 0.9)) %>%
+  ungroup()
+```
+
+The open access share varies across Elsevier hybrid journals.
 Since 2015, most hybrid journals have had a slow uptake rate of open access articles.
-In general, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume.
+Overall, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume.
+In 2019, the median open access share per hybrid journal was `r perc_df %>% filter(issued_year == 2019) %>% pull(median) * 100`%.
+In the same year, 90% of Elsevier hybrid journals had an open access uptake of `r perc_df %>% filter(issued_year == 2019) %>% pull(p_90) * 100`% or less.
+
+```{r, fig.cap='Open access uptake in Elsevier journals per year in percent, visualised as diminutive distribution chart removing outliers.
+Since 2015, most hybrid journals have had a slow uptake rate of open access articles.
+In general, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume.'}
+a_plot <- perc_df %>%
+ggplot() +
+  geom_econodist(aes(x = factor(issued_year), ymin = p_10, median = median, ymax = p_90),
+         stat = "identity", show.legend = TRUE) +
+  scale_y_continuous(position = "right", 
+                     labels = scales::percent_format(accuracy = 1L),
+                    # expand = expansion(mult = c(0, 0.15)), 
+                     limits = c(0,max(perc_df$p_90) + 0.01)) +
+  coord_flip() +
+  labs(
+    x = NULL, y = NULL
+   # title = "What is the uptake of hybrid open access among Elsevier journals?",
+  #  subtitle = "Proportion of immediate open access articles published in subscription-based journals\nJan 2015 to July 2019",
+   # caption = "Data source: Elsevier B.V. / Crossref"
+  ) +
+  theme_minimal(base_family = "Roboto") +
+  theme(panel.grid.minor = element_blank()) +
+  theme(axis.ticks = element_blank()) +
+  theme(panel.grid.major.y = element_blank()) +
+  theme(panel.border = element_blank())
+
+grid.newpage()
+left_align(a_plot, c("subtitle", "title", "caption")) %>% 
+  add_econodist_legend(econodist_legend_grob(family = "Roboto"), below = "subtitle") %>% 
+  grid.draw()  
+```
 

### 2020-08-21 analysis/results.Rmd
description variation
improve descriptive summary table

--- 2020-08-13 analysis/results.Rmd
@@
 output: github_document
+always_allow_html: true
 ---
@@
 library(cowplot)
+library(ggeconodist)
+library(gt)
 ```
@@
   group_by(issued_year) %>%
-  summarise(articles = sum(articles),
-            journals = n_distinct(issn))
+  summarise(journals = n_distinct(issn),
+            articles_n = sum(articles),
+            articles_mean = mean(articles),
+            articles_sd = sd(articles))
 
 # oa
-ind_per_year <- hybrid_df %>%
+oa_per_jn_year <- hybrid_df %>%
   filter(oa_archive == FALSE) %>%
+  group_by(issued_year, issn) %>%
+  summarise(oa = n()) %>%
+  ungroup()
+oa_ind <- oa_per_jn_year %>%
   group_by(issued_year) %>%
-  summarise(oa_articles = n()) %>%
-  inner_join(els_yr, by = "issued_year") %>%
-  ungroup() %>%
-  mutate(prop = oa_articles / articles * 100)
-ind_summary_table <- ind_per_year %>% 
-  pivot_longer(oa_articles:prop, names_to = "indicator", values_to = "count") %>%
-  pivot_wider(names_from = issued_year, values_from = count) %>%
-  mutate(indicator = factor(indicator, levels = c("journals", "articles", "oa_articles", "prop"))) %>%
-  arrange(indicator) %>%
-  mutate(indicator = recode(indicator,
-                            oa_articles = "Hybrid OA Articles", 
-                            articles = "All Articles", 
-                            journals = "Hybrid journals with at least one OA article", 
-                            prop = "Percentage"))
-`Total` <- c(
-  hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow(),
-  sum(ind_summary_table[2,2:6]),
-  sum(ind_summary_table[3,2:6]),
-  sum(ind_per_year$oa_articles) / sum(ind_per_year$articles) * 100)
-uptake_table <- bind_cols(ind_summary_table, Total = `Total`) %>%
-  rename(` ` = indicator) %>%
-  knitr::kable(digits = 1)
+  summarise(oa_n = sum(oa),
+            oa_mean = mean(oa),
+            oa_sd = sd(oa)) 
+# oa uptake 
+oa_uptake_ind <- els_pub_volume %>%
+  filter(issn %in% eligible_jns$issn) %>%
+  inner_join(oa_per_jn_year, by = c("issn", "issued_year")) %>%
+  mutate(prop = oa / articles) %>%
+  group_by(issued_year) %>%
+  summarise(oa_uptake_mean = mean(prop),
+            oa_uptake_sd = sd(prop))
+# oa uptake all 
+oa_uptake_15_19 <- els_pub_volume %>%
+    filter(issn %in% eligible_jns$issn) %>%
+    inner_join(oa_per_jn_year, by = c("issn", "issued_year")) %>%
+    group_by(issn) %>%
+    summarise(oa = sum(oa),
+              articles = sum(articles)) %>%
+    mutate(prop = oa / articles)
+
+# all 
+total_df <- tibble::tibble(
+  issued_year = "2015-19",
+  journals = hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow(),
+  articles_n = els_pub_volume %>% filter(issn %in% eligible_jns$issn) %>% pull(articles) %>% sum(),
+  articles_mean = els_pub_volume %>% filter(issn %in% eligible_jns$issn) %>% group_by(issn) %>% summarise(n = sum(articles)) %>% pull(n) %>% mean(),
+  articles_sd = els_pub_volume %>% filter(issn %in% eligible_jns$issn) %>% group_by(issn) %>% summarise(n = sum(articles)) %>% pull(n) %>% sd(),
+  oa_n = hybrid_df %>% filter(oa_archive == FALSE) %>% nrow(),
+  oa_mean = hybrid_df %>% filter(oa_archive == FALSE) %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% mean(),
+  oa_sd = hybrid_df %>% filter(oa_archive == FALSE) %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% sd(),
+  oa_uptake = oa_n / articles_n,
+  oa_uptake_mean = mean(oa_uptake_15_19$prop),
+  oa_uptake_sd = sd(oa_uptake_15_19$prop)
+         )
+
+ind_df <- inner_join(els_yr, oa_ind, by = "issued_year") %>%
+  mutate(oa_uptake = oa_n / articles_n) %>%
+  inner_join(oa_uptake_ind, by = "issued_year") %>%
+  mutate(issued_year = as.character(issued_year)) %>%
+  bind_rows(total_df) %>%
+  mutate(oa_uptake = oa_uptake * 100) %>%
+  mutate(oa_uptake_mean = oa_uptake_mean * 100) %>%
+  mutate(oa_uptake_sd = oa_uptake_sd * 100)
+
+ind_summary_table <- ind_df %>% 
+  pivot_longer(journals:oa_uptake_sd, names_to = "indicator", values_to = "count") %>%
+  pivot_wider(names_from = issued_year, values_from = count)
 ```
 
-Overall, `r Total[1]` hybrid journals provided immediate open access to `r Total[3]` articles between 2015 and 2019.
-The total share of hybrid open access in the publication volume of Elsevier journals was `r Total[4]`%.
+Overall, `r total_df$journals` hybrid journals provided immediate open access to `r total_df$oa_n` articles between 2015 and 2019.
+The total share of hybrid open access in the publication volume of Elsevier journals was `r total_df$oa_uptake * 100`%.
 Table present the high-level findings by year, illustrating a moderate growth in hybrid open access.
-Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_articles)` (2015) to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_articles)` (2019).
-However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(prop)`% (2015) to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(prop)`% (2019), because the number of articles published also increased during this five-years period.
+Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r ind_df %>% filter(issued_year == "2015") %>% pull(oa_n)` (2015) to `r ind_df %>% filter(issued_year == "2019") %>% pull(oa_n)` (2019).
+However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r ind_df %>% filter(issued_year == 2015) %>% pull(oa_uptake)`% (2015) to `r ind_df %>% filter(issued_year == 2019) %>% pull(oa_uptake)`% (2019), because the number of articles published also increased during this five-years period.
 
 ```{r}
-uptake_table
+table_descriptives_ind <- tribble(
+  ~ind_group, ~ind, ~ind_name,
+  "Hybrid Journals with OA articles", "Total", "Total",
+  "Publication Volume", "articles_n", "Articles",
+  "Publication Volume", "articles_mean", "Mean",
+  "Publication Volume", "articles_sd", "SD",
+  "OA Volume", "oa_n", "Total",
+  "OA Volume", "oa_mean", "Mean",
+  "OA Volume", "oa_sd", "SD",
+  "OA Uptake", "oa_uptake", "Overall OA Uptake",
+  "OA Uptake", "oa_uptake_mean", "Mean",
+  "OA Uptake", "oa_uptake_sd", "SD",
+)
+inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
+  dplyr::select(-ind) %>%
+  gt(rowname_col = "ind_name", groupname_col = "ind_group") %>%
+  fmt_number(contains("20")) %>%
+  fmt_number(contains("20"), 
+             rows = c("Articles", "Total"),
+             decimals = 0) %>%
+  tab_style(cell_text(weight = "bold"),
+            locations = cells_row_groups()
+  )
 ```
+
+*variation*
 
@@
   ungroup() %>%
-  mutate(prop = oa_articles / articles * 100) %>% View()
+  mutate(prop = oa_articles / articles) 
 ```
 
-The open access share varied across Elsevier hybrid journals.
+```{r}
+perc_df <- hybrid_oa_per_jn %>%
+  group_by(issued_year) %>% 
+  summarize(median = median(prop),
+            p_10 = quantile(prop, 0.1),
+            p_90 = quantile(prop, 0.9)) %>%
+  ungroup()
+```
+
+We observed considerable variations across Elsevier hybrid journals relative to the share of open access articles published by year.
+Figure shows, in a diminutive chart, the range of open access uptake across Elsevier hybrid journals between 2015 and 2019.
+The dot represents the median by year, while the left and the right bars represents to 10th or the 90th percentile; the distance between them represent the distribution of open access rates of 80% of Elsevier journals, excluding extreme outliers at the top or bottom respectively from the visual presentation.
+Notably, uptake among above-average performing journals in terms of published open access articles increased over the years.
+Nevertheless, the proportion of open access articles published by journals that operate above average can still be described as low; in 2019, 90% of Elsevier hybrid journals had an open access uptake of `r perc_df %>% filter(issued_year == 2019) %>% pull(p_90) * 100`% or less.
+
+```{r, fig.cap='Open access uptake in Elsevier hybrid journals per year in percent, visualised as diminutive distribution chart removing outliers at the top or bottom respectively.
 Since 2015, most hybrid journals have had a slow uptake rate of open access articles.
 In general, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume.
+However, many journals operating above average in terms of published open access articles were able to increase uptake rates over the years.'}
+a_plot <- perc_df %>%
+ggplot() +
+  geom_econodist(aes(x = factor(issued_year), ymin = p_10, median = median, ymax = p_90),
+         stat = "identity", show.legend = TRUE) +
+  scale_y_continuous(position = "right", 
+                     labels = scales::percent_format(accuracy = 1L),
+                    # expand = expansion(mult = c(0, 0.15)), 
+                     limits = c(0,max(perc_df$p_90) + 0.01)) +
+  coord_flip() +
+  labs(
+    x = NULL, y = NULL
+  ) +
+  theme_minimal(base_family = "Roboto") +
+  theme(panel.grid.minor = element_blank()) +
+  theme(axis.ticks = element_blank()) +
+  theme(panel.grid.major.y = element_blank()) +
+  theme(panel.border = element_blank())
 
-# license
+grid.newpage()
+left_align(a_plot, c("subtitle", "title", "caption")) %>% 
+  add_econodist_legend(econodist_legend_grob(family = "Roboto"), below = "subtitle") %>% 
+  grid.draw()  
+```
+
+alternative box plot presentation
+
+```{r}
+ggplot(hybrid_oa_per_jn, aes(factor(issued_year), prop)) + 
+  geom_boxplot(outlier.shape = NA) + 
+  coord_cartesian(ylim=c(0,0.15))
+
+# licenses used
+```

### 2020-08-23 analysis/results.Rmd
fix table

@@
 csl: chicago.csl
-output: github_document
+output: html_document
 always_allow_html: true
@@
 # pub volume eligble data
-els_yr <- els_pub_volume %>%
+hybrid_volume <- els_pub_volume %>%
   filter(issn %in% eligible_jns$issn) %>%
+  inner_join(hybrid_df, by = c("issn", "issued_year")) %>%
+  filter(oa_archive == FALSE)
+
+els_articles <- 
+  hybrid_volume %>%
+  distinct(issued_year, issn, articles) %>%
   group_by(issued_year) %>%
-  summarise(journals = n_distinct(issn),
+  summarise(
     articles_n = sum(articles),
             articles_mean = mean(articles),
-            articles_sd = sd(articles))
+            articles_sd = sd(articles)
+    )
+els_yr <- 
+  hybrid_volume %>%
+  group_by(issued_year) %>%
+  summarise(journals = n_distinct(issn)) %>%
+  inner_join(els_articles, by = "issued_year")
 
 # oa
-oa_per_jn_year <- hybrid_df %>%
-  filter(oa_archive == FALSE) %>%
+oa_per_jn_year <- hybrid_volume %>%
   group_by(issued_year, issn) %>%
-  summarise(oa = n()) %>%
-  ungroup()
+  summarise(oa = n(), .groups = "drop")
 oa_ind <- oa_per_jn_year %>%
@@
 # oa uptake 
-oa_uptake_ind <- els_pub_volume %>%
-  filter(issn %in% eligible_jns$issn) %>%
+oa_uptake_ind <- hybrid_volume %>% 
+  distinct(issued_year, issn, articles) %>%
   inner_join(oa_per_jn_year, by = c("issn", "issued_year")) %>%
@@
 # oa uptake all 
-oa_uptake_15_19 <- els_pub_volume %>%
-    filter(issn %in% eligible_jns$issn) %>%
+oa_uptake_15_19 <- hybrid_volume %>% 
+  distinct(issued_year, issn, articles) %>%
   inner_join(oa_per_jn_year, by = c("issn", "issued_year")) %>%
@@
   issued_year = "2015-19",
-  journals = hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow(),
-  articles_n = els_pub_volume %>% filter(issn %in% eligible_jns$issn) %>% pull(articles) %>% sum(),
-  articles_mean = els_pub_volume %>% filter(issn %in% eligible_jns$issn) %>% group_by(issn) %>% summarise(n = sum(articles)) %>% pull(n) %>% mean(),
-  articles_sd = els_pub_volume %>% filter(issn %in% eligible_jns$issn) %>% group_by(issn) %>% summarise(n = sum(articles)) %>% pull(n) %>% sd(),
-  oa_n = hybrid_df %>% filter(oa_archive == FALSE) %>% nrow(),
-  oa_mean = hybrid_df %>% filter(oa_archive == FALSE) %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% mean(),
-  oa_sd = hybrid_df %>% filter(oa_archive == FALSE) %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% sd(),
+  journals = hybrid_volume %>% distinct(issn) %>% nrow(),
+  articles_n = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% sum(),
+  articles_mean = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% mean(),
+  articles_sd = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% sd(),
+  oa_n = nrow(hybrid_volume),
+  oa_mean = hybrid_volume %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% mean(),
+  oa_sd = hybrid_volume %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% sd(),
   oa_uptake = oa_n / articles_n,
@@
   ~ind_group, ~ind, ~ind_name,
-  "Hybrid Journals with OA articles", "Total", "Total",
+  "Hybrid Journals with OA articles", "journals", "Journals",
   "Publication Volume", "articles_n", "Articles",
@@
   "Publication Volume", "articles_sd", "SD",
-  "OA Volume", "oa_n", "Total",
-  "OA Volume", "oa_mean", "Mean",
+  "OA Volume", "oa_n", "OA Articles",
+  "OA Volume", "oa_mean", "Mean per Journal",
   "OA Volume", "oa_sd", "SD",
-  "OA Uptake", "oa_uptake", "Overall OA Uptake",
-  "OA Uptake", "oa_uptake_mean", "Mean",
-  "OA Uptake", "oa_uptake_sd", "SD",
+  "OA Uptake (in%)", "oa_uptake", "OA Percentage",
+  "OA Uptake (in%)", "oa_uptake_mean", "Mean per Journal",
+  "OA Uptake (in%)", "oa_uptake_sd", "SD"
 )
@@
   gt(rowname_col = "ind_name", groupname_col = "ind_group") %>%
-  fmt_number(contains("20")) %>%
+  fmt_number(contains("20"), decimals = 2) %>%
   fmt_number(contains("20"), 
-             rows = c("Articles", "Total"),
+             rows = c("Journals", "Articles", "OA Articles"),
              decimals = 0) %>%

### 2020-09-06 analysis/006_license.Rmd
license

---
title: "OA Share"
output: github_document
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

### Connect to database

```{r}
# connect to google bg where we imported the json lines Unpaywall dump
library(tidyverse)
library(DBI)
library(bigrquery)
con <- dbConnect(
  bigrquery::bigquery(),
  project = "api-project-764811344545",
  dataset = "cr_dump_march_20"
)
```

### Get article-level metadata for all journals listed in Elseviers most current APC pricing list

```{sql, connection=con}
CREATE 
    OR REPLACE TABLE `api-project-764811344545.cr_dump_march_20.els_all_may_20` AS SELECT
        cr.*,
        els.oa_type 
    FROM
        `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` as cr 
    INNER JOIN
        `api-project-764811344545.cr_dump_march_20.elsevier_apc_list_30_May_20` as els 
            ON cr.issn = els.issn 
```

```{sql, connection=con}
CREATE 
    OR REPLACE TABLE `api-project-764811344545.cr_dump_march_20.els_hybrid_tagged` AS SELECT
        cr.*,
        els.oa_type 
    FROM
        `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` as cr 
    INNER JOIN
        `api-project-764811344545.cr_dump_march_20.elsevier_apc_list_30_May_20` as els 
            ON cr.issn = els.issn 
```

Aggregate

```{sql, connection=con, output.var = "license_agg"}
SELECT 
  REGEXP_REPLACE(license.URL, "http://|https://", "") as license_url,
  issued_year,
   oa_type,
  COUNT(DISTINCT(doi)) as articles
FROM `api-project-764811344545.cr_dump_march_20.els_all_may_20`, unnest(license) as license
WHERE issued_year > 2014 and issued_year < 2020 AND license.content_version = 'vor'
GROUP BY license_url, issued_year, oa_type
ORDER BY issued_year DESC
```

```{r}
license_by_year <- license_agg %>%
  mutate(license_code =
           case_when(
             grepl("/by/", license_url) ~ "cc-by",
             grepl("/by-nc", license_url) ~ "cc-by-nc",
             grepl("userlicense", license_url) ~ "els-user")
  ) %>%
  group_by(license_code, oa_type, issued_year) %>%
  summarise(n = sum(articles))
```
just immediate hybrid

```{r}
im_hybrid <- readr::read_csv(here::here("data", "hybrid_volume_eligible.csv"))
cc_hybrid <- im_hybrid %>%
  mutate(license_code = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
  group_by(license_code, issued_year) %>%
  summarise(hybrid_immediate_articles = n()) %>%
  mutate(immediate = "immediate",
         oa_type = "Hybrid") %>%
  inner_join(license_by_year, by = c("issued_year", "license_code", "oa_type")) %>%
  mutate(hybrid_delayed_articles = n - hybrid_immediate_articles) %>%
  select(-immediate, -oa_type) %>%
  pivot_longer(c(hybrid_immediate_articles, hybrid_delayed_articles), names_to = "type", values_to = "articles")
# user only open archive
els_user <- license_by_year %>%
  filter(license_code == "els-user") %>%
  mutate(type = "hybrid_delayed_articles") %>%
  group_by(issued_year, type, license_code) %>%
  summarise(articles = sum(n))
# full oa 
full_oa_df <- license_by_year %>%
  ungroup() %>%
  filter(license_code != "els-user", oa_type == "Open Access") %>%
  mutate(type = "full_oa") %>%
  rename(articles = n) %>%
  select(-oa_type)
bind_rows(cc_hybrid, els_user, full_oa_df) %>%
  ggplot(aes(issued_year, articles, fill = license_code, group = license_code)) +
  geom_area() +
  facet_wrap(~type)
```

### 2020-09-06 analysis/results.Rmd
license

@@
   filter(oa_archive == FALSE)
+# backup 
+write_csv(hybrid_volume, here::here("data", "hybrid_volume_eligible.csv"))
 
@@
             p_10 = quantile(prop, 0.1),
-            p_90 = quantile(prop, 0.9)) %>%
+            p_95 = quantile(prop, 0.95),
+            iqr = quantile(prop, 3/4) - quantile(prop, 1/4)) %>%
   ungroup()
@@
 
-We observed considerable variations across Elsevier hybrid journals relative to the share of open access articles published by year.
-Figure shows, in a diminutive chart, the range of open access uptake across Elsevier hybrid journals between 2015 and 2019.
-The dot represents the median by year, while the left and the right bars represents to 10th or the 90th percentile; the distance between them represent the distribution of open access rates of 80% of Elsevier journals, excluding extreme outliers at the top or bottom respectively from the visual presentation.
-Notably, uptake among above-average performing journals in terms of published open access articles increased over the years.
-Nevertheless, the proportion of open access articles published by journals that operate above average can still be described as low; in 2019, 90% of Elsevier hybrid journals had an open access uptake of `r perc_df %>% filter(issued_year == 2019) %>% pull(p_90) * 100`% or less.
-
-```{r, fig.cap='Open access uptake in Elsevier hybrid journals per year in percent, visualised as diminutive distribution chart removing outliers at the top or bottom respectively.
-Since 2015, most hybrid journals have had a slow uptake rate of open access articles.
-In general, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume.
-However, many journals operating above average in terms of published open access articles were able to increase uptake rates over the years.'}
-a_plot <- perc_df %>%
-ggplot() +
-  geom_econodist(aes(x = factor(issued_year), ymin = p_10, median = median, ymax = p_90),
-         stat = "identity", show.legend = TRUE) +
-  scale_y_continuous(position = "right", 
-                     labels = scales::percent_format(accuracy = 1L),
-                    # expand = expansion(mult = c(0, 0.15)), 
-                     limits = c(0,max(perc_df$p_90) + 0.01)) +
-  coord_flip() +
-  labs(
-    x = NULL, y = NULL
-  ) +
-  theme_minimal(base_family = "Roboto") +
-  theme(panel.grid.minor = element_blank()) +
-  theme(axis.ticks = element_blank()) +
-  theme(panel.grid.major.y = element_blank()) +
-  theme(panel.border = element_blank())
-
-grid.newpage()
-left_align(a_plot, c("subtitle", "title", "caption")) %>% 
-  add_econodist_legend(econodist_legend_grob(family = "Roboto"), below = "subtitle") %>% 
-  grid.draw()  
-```
-
-alternative box plot presentation
+Figure presents to which extent open access uptake rates varied among Elsevier hybrid journals between 2015 and 2019.
+The box-and-whiskers plot displays the distribution of the open access percentage per journal broken down by year.
+It visualizes the median open access proportion in the journal sample, the range of the majority of journals, and individual outliers represented as dots.
+Notably, upper quartiles and whiskers stretch farther over the years, indicating that the range of uptake among journals with an above-average proportion of open access articles increased over the years.
+Nevertheless, the prevalence of open access even among these journals can be described as relatively low; in 2019, 95% of Elsevier hybrid journals published `r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% or less open access articles.
 
@@
 ggplot(hybrid_oa_per_jn, aes(factor(issued_year), prop)) + 
-  geom_boxplot(outlier.shape = NA) + 
-  coord_cartesian(ylim=c(0,0.15))
-
+  geom_boxplot() + 
+  coord_cartesian(ylim=c(0,0.18)) +
+  theme_minimal_hgrid() +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 5L),
+      expand = expansion(mult = c(0, 0.05))) +
+  labs(x = "Publication Years",
+       y = "Open Access Uptake")
 # licenses used
 ```
+
+*Licenses used*
+
+```{r}
+# per year
+yrl_volume <- hybrid_oa_per_jn %>% 
+  group_by(issued_year) %>% 
+  summarise(articles = sum(articles))
+
+hybrid_license <- hybrid_volume %>%
+  filter(oa_archive == FALSE) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
+  group_by(cc_norm, issued_year) %>%
+  summarise(cc_articles = n())
+```
+
+Figure presents a breakdown by Creative Commons license variant used for hybrid open access by year.
+Elsevier provided commercial (cc-by) and a less permissive non-commercial license (cc-by-nc) for open access publishing in otherwise subscription-based journals.
+Nearly two-third of open access articles (`r  round(hybrid_license %>% group_by(cc_norm) %>% summarise(n = sum(cc_articles)) %>% mutate(prop = n / sum(n)) %>% filter(cc_norm == "cc-by-nc") %>% pull(prop) * 100,0)`%) can be re-used for non-commercial purposes.
+This proportion remained the same between 2015-19.
+
+```{r}
+hybrid_license %>%
+  ggplot(aes(factor(issued_year), cc_articles, fill = cc_norm, group = cc_norm)) +
+  geom_area(stat = "identity",
+            color = "white",
+            alpha = 0.8) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = function(x)
+        format(x, big.mark = ",", scientific = FALSE),
+      breaks = scales::pretty_breaks()
+  ) +
+  scale_fill_manual(values = 
+                      c(`cc-by` = "#0093c7",
+  `cc-by-nc` = "#B52141")) +
+  labs(x = "Publication year", y = "Total OA articles") +
+  theme_minimal_hgrid() +
+  theme(legend.position = "top",
+        legend.justification = "right") +
+  guides(fill = guide_legend("Creative Commons license variant"))
+````
+
+*Relationship to full and delayed OA?*

### 2020-09-07 analysis/006_license.Rmd
with paratext

@@
 
-```{sql, connection=con}
-CREATE 
-    OR REPLACE TABLE `api-project-764811344545.cr_dump_march_20.els_hybrid_tagged` AS SELECT
-        cr.*,
-        els.oa_type 
-    FROM
-        `api-project-764811344545.cr_dump_march_20.cr_08_20_expanded` as cr 
-    INNER JOIN
-        `api-project-764811344545.cr_dump_march_20.elsevier_apc_list_30_May_20` as els 
-            ON cr.issn = els.issn 
-```
-
 Aggregate
@@
 SELECT
-  REGEXP_REPLACE(license.URL, "http://|https://", "") as license_url,
+        REGEXP_REPLACE(license.URL,
+        "http://|https://",
+        "") as license_url,
         issued_year,
@@
         COUNT(DISTINCT(doi)) as articles 
-FROM `api-project-764811344545.cr_dump_march_20.els_all_may_20`, unnest(license) as license
-WHERE issued_year > 2014 and issued_year < 2020 AND license.content_version = 'vor'
-GROUP BY license_url, issued_year, oa_type
-ORDER BY issued_year DESC
+    FROM
+        `api-project-764811344545.cr_dump_march_20.els_all_may_20`,
+        unnest(license) as license 
+    WHERE
+        issued_year > 2014 
+        and issued_year < 2020 
+        AND license.content_version = 'vor'         
+        AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
+    GROUP BY
+        license_url,
+        issued_year,
+        oa_type 
+    ORDER BY
+        issued_year DESC
 ```
@@
   select(-oa_type)
-bind_rows(cc_hybrid, els_user, full_oa_df) %>%
-  ggplot(aes(issued_year, articles, fill = license_code, group = license_code)) +
+license_growth_df <- bind_rows(cc_hybrid, els_user, full_oa_df)
+
+ggplot(license_growth_df, aes(issued_year, articles, fill = license_code, group = license_code)) +
   geom_area() +
@@
 ```
+
+### relative
+
+overall article volume elsevier
+
+```{sql, connection=con, output.var="els_yearly"}
+SELECT  issued_year,
+        COUNT(DISTINCT(doi)) as all_articles 
+    FROM
+        `api-project-764811344545.cr_dump_march_20.els_all_may_20`
+    WHERE
+        issued_year > 2014 
+        and issued_year < 2020        
+        AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
+    GROUP BY
+        issued_year
+
+```
+
+```{r}
+year_per_oa_type_and_license <- inner_join(license_growth_df, els_yearly, by = "issued_year") %>%
+  mutate(prop = articles / all_articles) 
+# backup 
+write_csv(year_per_oa_type_and_license, here::here("data", "year_per_oa_type_and_license.csv"))
+```
+
+plots
+
+```{r}
+library(cowplot)
+p_1 <- year_per_oa_type_and_license %>%
+  mutate(type_plot = case_when(
+    type == "full_oa" ~ "Full OA",
+    type == "hybrid_immediate_articles" ~ "Hybrid OA",
+    type == "hybrid_delayed_articles" ~ "Delayed OA"
+  )) %>% 
+  mutate(type_plot = factor(type_plot, levels = c("Hybrid OA", "Full OA", "Delayed OA"))) %>%
+  ggplot(aes(gsub("20", "'", as.character(issued_year)), prop, fill = license_code, group = license_code)) +
+  geom_area(stat = "identity",
+            color = "white",
+            alpha = 0.8,
+            position = position_stack(reverse = T)) +
+  facet_grid(~type_plot) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.0005)),
+    labels = scales::percent_format(accuracy = 1L),
+    breaks = c(0, 0.02, 0.04, 0.06, 0.08)
+  ) +
+  scale_fill_manual(values = 
+                      c(`cc-by` = "#0093c7",
+                        `cc-by-nc` = "#B52141",
+                        `els-user` = "grey80")) +
+  labs(x = "Publication year", y = "OA Percentage") +
+  theme_minimal_hgrid() +
+  theme(legend.position = "top",
+        legend.justification = "right") +
+  guides(fill = guide_legend("Open content license"))
+p_1
+```
+
+overall
+
+```{r}
+p_2 <- year_per_oa_type_and_license %>%
+  mutate(type_plot = case_when(
+    type == "full_oa" ~ "Full OA",
+    type == "hybrid_immediate_articles" ~ "Hybrid OA",
+    type == "hybrid_delayed_articles" ~ "Delayed OA"
+  )) %>% 
+  mutate(type_plot = factor(type_plot, levels = c("Hybrid OA", "Full OA", "Delayed OA"))) %>%
+  group_by(type_plot) %>%
+  summarise(n = sum(articles)) %>%
+  ggplot(aes(x = "", y = n, fill = type_plot)) +
+  geom_bar(width = 1,
+           stat = "identity",
+           position = position_stack(reverse = TRUE),
+           color = "white") +
+  coord_flip() +
+  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE),
+                     limits = c(0, 400000)) +
+  scale_fill_brewer("OA Type", type = "qual", palette = "Set2") +
+  labs(x = NULL,
+       y = "Total OA Articles") +
+  theme_minimal_grid() +
+  theme(plot.margin = margin(30, 30, 30, 30)) +
+  theme(panel.grid.minor = element_blank()) +
+  theme(axis.ticks = element_blank()) +
+  theme(panel.grid.major.y = element_blank()) +
+  theme(panel.border = element_blank()) +
+  theme(legend.position = "bottom",
+        legend.justification = "center")
+p_2
+```
+
+```{r fig.width = 6, fig.height = 12}
+cowplot::plot_grid(
+  p_1, p_2,
+  labels = "AUTO", ncol = 1
+)
+```

### 2020-09-08 analysis/results.Rmd
improve table

@@
 Overall, `r total_df$journals` hybrid journals provided immediate open access to `r total_df$oa_n` articles between 2015 and 2019.
-The total share of hybrid open access in the publication volume of Elsevier journals was `r total_df$oa_uptake * 100`%.
+The total share of hybrid open access in the Articles published of Elsevier journals was `r total_df$oa_uptake * 100`%.
 Table present the high-level findings by year, illustrating a moderate growth in hybrid open access.
@@
   ~ind_group, ~ind, ~ind_name,
-  "Hybrid Journals with OA articles", "journals", "Journals",
-  "Publication Volume", "articles_n", "Articles",
-  "Publication Volume", "articles_mean", "Mean",
-  "Publication Volume", "articles_sd", "SD",
-  "OA Volume", "oa_n", "OA Articles",
-  "OA Volume", "oa_mean", "Mean per Journal",
-  "OA Volume", "oa_sd", "SD",
-  "OA Uptake (in%)", "oa_uptake", "OA Percentage",
-  "OA Uptake (in%)", "oa_uptake_mean", "Mean per Journal",
+  "Elsevier Hybrid Journals with OA article(s)", "journals", "Journals",
+  "Articles published", "articles_n", "Total",
+  "Articles published", "articles_mean", "Avg.",
+  "Articles published", "articles_sd", "SD",
+  "OA Articles published", "oa_n", "Total OA",
+  "OA Articles published", "oa_mean", "Avg.",
+  "OA Articles published", "oa_sd", "SD",
+  "OA Uptake (in%)", "oa_uptake", "Percentage",
+  "OA Uptake (in%)", "oa_uptake_mean", "Avg.",
   "OA Uptake (in%)", "oa_uptake_sd", "SD"
@@
   gt(rowname_col = "ind_name", groupname_col = "ind_group") %>%
-  fmt_number(contains("20"), decimals = 2) %>%
+  fmt_number(contains("20"), decimals = 1) %>%
   fmt_number(contains("20"), 
-             rows = c("Journals", "Articles", "OA Articles"),
+             rows = c("Total", "Total OA", "Journals"),
              decimals = 0) %>%
-  tab_style(cell_text(weight = "bold"),
-            locations = cells_row_groups()
+  tab_style(
+    style = cell_fill(
+      color = "grey90"
+    ),
+    locations = list(
+      cells_body(
+        columns = "2015-19"
+      ),
+      cells_column_labels(
+        columns = "2015-19"
+      )
+    )
+  ) %>%
+  tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = list(
+      cells_column_labels(everything())
+    )
+  )  %>%
+   tab_style(
+    style = cell_text(color = "black", style = "italic"),
+    locations = list(
+      cells_row_groups())
+    ) %>%
+  cols_width(
+    2 ~ px(125),
+    3:7 ~ px(80),
+    8 ~ px(110)
+  ) %>%
+  tab_options(
+    row_group.border.top.width = px(3),
+    row_group.border.top.color = "white",
+    row_group.border.bottom.color = "black",
+    table_body.hlines.color = "white",
+    table.border.top.color = "white",
+    table.border.top.width = px(3),
+    table.border.bottom.color = "white",
+    table.border.bottom.width = px(3),
+    column_labels.border.bottom.color = "black",
+    column_labels.border.bottom.width = px(2),
+    stub.border.color = "white"
   )

### 2020-09-08 analysis/006_license.Rmd
improve license analysis

@@
   theme(legend.position = "bottom",
-        legend.justification = "center")
+        legend.justification = "right")
 p_2
@@
 ```{r fig.width = 6, fig.height = 12}
-cowplot::plot_grid(
+p <- cowplot::plot_grid(
   p_1, p_2,
@@
 )
+ggsave(here::here("figure", "license_portfolio.png"), p, dpi = 300)
 ```

### 2020-09-08 analysis/results.Rmd
improve license analysis

@@
 *Relationship to full and delayed OA?*
+
+```{r}
+license_df <- readr::read_csv(here::here("data", "year_per_oa_type_and_license.csv"))
+summary_licenses <- 
+  license_df %>% 
+  group_by(type) %>% 
+  summarise(n = sum(articles)) %>%
+  mutate(prop = n / sum(n))
+```
+
+Next, we present a breakdown of open content licensed articles by open access type relative to Elsevier's journal productivity (Figure).
+In total, Elsevier provided open access to `r sum(summary_licenses$n)` articles at he time of our study, corresponding to an open access share of `r round(sum(summary_licenses$n) / license_df %>% distinct(all_articles) %>% pull(all_articles) %>% sum() * 100, 0)`%.
+Hybrid open access had the lowest proportion (`r round(filter(summary_licenses, type == "hybrid_immediate_articles") %>% pull(prop) * 100)`), while over half of open access articles were made freely available after an embargo period (`r round(filter(summary_licenses, type == "hybrid_delayed_articles") %>% pull(prop) * 100, 0)`%) (Figure A).
+`r round(filter(summary_licenses, type == "full_oa") %>% pull(prop) * 100)`% of open access articles appeared in full open access journals.
+
+```{r}
+knitr::include_graphics(here::here("figure", "license_portfolio.png"))
+```
+Figure B presents a breakdown of open content licensed articles by open access type and year relative to the total publication volume published in Elsevier journals between 2015-19.
+In addition to Creative Commons licenses, Elsevier also published delayed open access under an Elsevier user license (els-user), which prohibits re-use for commercial purposes.
+The license is mainly used by journals participating in Elsevier's open archive program comprising around 140 journals that make provide free access to published articles after an embargo period ranging between 12 months and 48 months.
+Figure reveals that about half of open access articles were made available under the Elsevier user license.
+The minority of open access articles can be re-used for commercial purposes (`r round(license_df %>% group_by(license_code) %>% summarise(n = sum(articles)) %>% mutate(prop = n / sum(n)) %>% filter(license_code == "cc-by") %>% pull(prop) * 100, 0)`% of open access articles under an CC-BY license).
+Hybrid open access had the highest proportion of openly available articles under a CC-BY licenses.

### 2020-09-09 analysis/005_article_volume.Rmd
exclude articles published in supplements

@@
 
+### Article volume regular issues excluding paratext
+
+```{sql, connection = "con", output.var = "article_volume_regular"}
+    SELECT
+        issued_year,
+        issn,
+        count(distinct(doi)) as articles          
+    FROM
+        `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`           
+    WHERE
+        issued_year > 2014          
+        and issued_year < 2020              
+        AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
+        AND NOT regexp_contains(page, '^S')     
+    GROUP BY
+        issued_year,
+        issn
+```
+
+How many articles by year?
+
+```{r}
+article_volume_regular %>%
+  group_by(issued_year) %>%
+  summarise(articles = sum(articles)) 
+```
+
+backup
+
+```{r}
+write_csv(article_volume_para, here::here("data", "article_volume_para_regular.csv"))
+```
+
 ### Article volume with at least one reference per journal and year

### 2020-09-09 analysis/006_license.Rmd
exclude articles published in supplements

@@
         AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
+        AND NOT regexp_contains(page, '^S')
     GROUP BY
@@
         AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
+        AND NOT regexp_contains(page, '^S') 
     GROUP BY

### 2020-09-09 analysis/results.Rmd
exclude articles published in supplements

@@
   inner_join(hybrid_df, by = c("issn", "issued_year")) %>%
-  filter(oa_archive == FALSE)
+  filter(oa_archive == FALSE) %>%
+  filter(!grepl("^S", page))
 # backup 
@@
 Hybrid open access had the highest proportion of openly available articles under a CC-BY licenses.
+
+## invoicing
+
+```{r}
+sponsored_articles <- hybrid_volume %>%
+  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
+  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
+                                  `FundingBody` = "Agreement",
+                                  `ElsevierWaived` = "Fee Waived")) %>%
+  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
+  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
+  count(issued_year, fct_source)
+all_articles <- hybrid_df %>%
+  group_by(issued_year) %>%
+  summarise(n = n())
+
+ggplot(sponsored_articles, aes(x = gsub("20", "", issued_year), y = n)) +
+  geom_bar(
+    data = all_articles,
+    aes(fill = "All hybrid open access articles"),
+    color = "transparent",
+    stat = "identity"
+  ) +
+  geom_bar(aes(fill = "by invoicing type"), color = "transparent", stat = "identity") +
+  facet_wrap( ~ fct_source, nrow = 1) +
+  scale_fill_manual(values = c("#b3b3b3a0", "#56B4E9"), name = "") +
+  labs(x = NULL, y = NULL) + 
+  scale_y_continuous(labels = scales::number_format(big.mark = " ")) +
+  theme_minimal(base_family="Roboto", base_size = 12) +
+  theme(panel.grid.minor=element_blank()) +
+  theme(axis.ticks=element_blank()) +
+  theme(panel.grid.major.x=element_blank()) +
+  theme(panel.border=element_blank()) +
+  theme(legend.position="top", legend.justification = "right")
+```

### 2020-09-09 analysis/006_license.Rmd
figures

@@
   scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE),
-                     limits = c(0, 400000)) +
+                     limits = c(0, 350000)) +
   scale_fill_brewer("OA Type", type = "qual", palette = "Set2") +
@@
 
-```{r fig.width = 6, fig.height = 12}
+```{r}
 p <- cowplot::plot_grid(
@@
 )
-ggsave(here::here("figure", "license_portfolio.png"), p, dpi = 300)
+ggsave(here::here("figure", "license_portfolio.png"), p, dpi = 300, width = 6, height = 6)
 ```

### 2020-09-09 analysis/results.Rmd
figures

--- 2020-09-08 analysis/results.Rmd
@@
             journals = n_distinct(issn))
-hybrid_oa_per_jn <- hybrid_df %>%
-  filter(oa_archive == FALSE) %>%
+hybrid_oa_per_jn <- hybrid_volume %>%
   group_by(issued_year, issn) %>%
@@
 ```
+
 Figure B presents a breakdown of open content licensed articles by open access type and year relative to the total publication volume published in Elsevier journals between 2015-19.
@@
 Hybrid open access had the highest proportion of openly available articles under a CC-BY licenses.
+
+## invoicing
+
+```{r}
+sponsored_articles <- hybrid_volume %>%
+  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
+  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
+                                  `FundingBody` = "Agreement",
+                                  `ElsevierWaived` = "Fee Waived")) %>%
+  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
+  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
+  count(issued_year, fct_source)
+all_articles <- hybrid_volume %>%
+  group_by(issued_year) %>%
+  summarise(n = n())
+
+ggplot(sponsored_articles, aes(x = gsub("20", "'", issued_year), y = n)) +
+  geom_bar(
+    data = all_articles,
+    aes(fill = "All hybrid OA"),
+    color = "transparent",
+    stat = "identity"
+  ) +
+  geom_bar(aes(fill = "by invoicing type"), color = "transparent", stat = "identity") +
+  facet_wrap( ~ fct_source, nrow = 1) +
+  scale_fill_manual(values = c("#b3b3b3a0", "#56B4E9"), name = "") +
+  labs(x = NULL, y = NULL) + 
+  scale_y_continuous(labels = scales::number_format(big.mark = " ")) +
+  theme_minimal(base_family="Roboto", base_size = 12) +
+  theme(panel.grid.minor=element_blank()) +
+  theme(axis.ticks=element_blank()) +
+  theme(panel.grid.major.x=element_blank()) +
+  theme(panel.border=element_blank()) +
+  theme(legend.position="top", legend.justification = "right")
+```

### 2020-09-11 analysis/007_invoice_exploration.Rmd
invoices

---
title: "invoice analysis"
output: github_document
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "fg", digits = 2))
    } else{
      return(x)
    }
  }
)
```

```{r}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
```

```{r}
# load eligible oa articles
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_volume_eligible.csv"))

# aggregate data
sponsored_articles <- hybrid_volume %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  count(issued_year, fct_source)
all_articles <- hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(n = n())
```

In most cases, Elsevier sent invoices for hybrid open access publication fees to individual authors (`r round(hybrid_volume %>% filter(oa_sponsor_type == "Author") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%). For around `r round(hybrid_volume %>% filter(oa_sponsor_type == "FundingBody") %>% nrow() / nrow(hybrid_volume) * 100, 0)`% of articles, the publisher directly billed funders and research organisations. The proportion of open access articles where Elsevier granted publication fee waivers was `r round(hybrid_volume %>% filter(oa_sponsor_type == "ElsevierWaived") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%.

Figure \@ref(fig:invoiceoverview) shows the annual development per invoicing type. Each type is visualised separately as parts of the total. The grey areas show the overall yearly distribution of open access articles in hybrid journals, and the proportion of each invoicing type is highlighted in blue, The figure reveals a general growth of open access articles in hybrid journals. It illustrates that this development was mainly driven by billing individual authors, while central invoicing stagnated. Also, the amount of fee-waived articles remained more or less constant from 2015 to date.

```{r invoiceoverview, fig.cap="Development of fee-based open access publishing in Elsevier hybrid journals by invoicing type. Colored bars represent the invoice recipient, or whether the fee was waived. Grey bars show the total number of hybrid open access articles published in Elsevier journals from 2015 to date."}
# plot
ggplot(sponsored_articles, aes(x = gsub("20", "'", issued_year), y = n)) +
  geom_bar(
    data = all_articles,
    aes(fill = "All hybrid OA"),
    color = "transparent",
    stat = "identity"
  ) +
  geom_bar(aes(fill = "by invoicing type"), color = "transparent", stat = "identity") +
  facet_wrap( ~ fct_source, nrow = 1) +
  scale_fill_manual(values = c("#b3b3b3a0", "#56B4E9"), name = "") +
  labs(x = NULL, y = NULL) + 
  scale_y_continuous(labels = scales::number_format(big.mark = " "),
                      expand = expansion(mult = c(0, 0.05))) + theme_minimal_hgrid() +
  theme(legend.position="top", legend.justification = "right")
```

```{r}
cc_invoice <- hybrid_volume %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
   mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  count(issued_year, fct_source, cc_norm)

cc_invoice_type <- cc_invoice %>% 
  group_by(fct_source, cc_norm) %>% 
  summarise(n = sum(n))
cc_norm_all <- cc_invoice_type %>%
  ungroup() %>%
  group_by(cc_norm) %>%
  summarise(all = sum(n))
cc_type_all <- cc_invoice_type %>%
  ungroup() %>%
  group_by(fct_source) %>%
  summarise(all = sum(n))
license_count <- inner_join(cc_invoice_type, cc_norm_all, by =  "cc_norm") %>% mutate(prop = n / all)
license_summary_count <- inner_join(cc_invoice_type, cc_type_all, by =  "fct_source") %>% mutate(prop = n / all)
```

In terms of re-use, the type of invoicing is strongly related to whether the open content license permits commercial adaption or not (see figure). Although most open access articles published in Elsevier's hybrid journals were billed to authors directly, centrally invoiced articles accounted for the largest share of articles under an CC BY licenses (`r round(filter(license_count, fct_source == "Agreement", cc_norm == "cc-by") %>% pull(prop) * 100, 0)`%). Consequently, `r round(filter(license_summary_count, fct_source == "Agreement", cc_norm == "cc-by") %>% pull(prop) * 100, 0)`% of centrally invoiced articles permits commercial adaption.

```{r}
ggplot(cc_invoice, aes(factor(issued_year), n, fill = fct_source, group = fct_source)) +
  geom_area(stat = "identity",
            color = "white",
            alpha = 0.8,
            position = position_stack(reverse = T)) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = function(x)
        format(x, big.mark = ",", scientific = FALSE),
      breaks = scales::pretty_breaks()
  ) +
  facet_grid(~cc_norm) +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "Publication year", y = "Total OA articles") +
  theme_minimal_hgrid() +
  theme(legend.position = "top",
        legend.justification = "right") +
  guides(fill = guide_legend("Invoicing"))
```

```{r}
# invoicing analysis
library(googlesheets4)
tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 2)
```

```{r}
hybrid_invoice <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>%
  rename(yrl_volume = articles) %>%
  left_join(tt, by = c("oa_sponsor_name"))

sponsor_country <- hybrid_invoice %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  group_by(country_factor) %>%
  summarise(n = n_distinct(doi)) %>%
  mutate(prop = n / sum(n))
```

Next, we examined funders and research organisations that were invoiced directly by Elsevier. In total, we were able to identify `r length(unique(hybrid_invoice$oa_sponsor_name))` sponsoring bodies. Figure presents a breakdown by country, highlighting the dominating role of a few countries. Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "Multinational") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%). In the latter case, invoices for individual open access articles were not send to research funders, but to the Dutch VSNU, a consortium  of Universities in the Netherlands that has negotiated agreements with Elsevier since 2016. The prominent role of Norway can be also explained by a national deal, which started in 2019. US-based open access sponsors originated mainly from the Bill & Melinda Gates Foundation (n = `r hybrid_invoice %>% filter(oa_sponsor_name == "Bill & Melinda Gates Foundation") %>% distinct(doi) %>% nrow()`).

```{r}
sponsor_country %>%
  mutate(country_factor = fct_reorder(country_factor, n)) %>%
  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
) %>%
  ggplot(aes(country_factor, n)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  coord_flip() +
  theme_minimal_vgrid() +
  labs(y = "Number of invoiced OA articles", x = NULL)
```

### 2020-09-11 analysis/results.Rmd
invoices

@@
 ggplot(hybrid_oa_per_jn, aes(factor(issued_year), prop)) + 
-  geom_boxplot() + 
+  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + 
   coord_cartesian(ylim=c(0,0.18)) +
@@
             color = "white",
-            alpha = 0.8) +
+            alpha = 0.8,
+            position = position_stack(reverse = T)) +
   scale_y_continuous(
@@
 
-## invoicing
+### How many payments for open access articles in hybrid journals were invoiced to authors, to funders, and how many APCs were waived?
 
 ```{r}
+# load eligible oa articles
+hybrid_volume <- readr::read_csv(here::here("data", "hybrid_volume_eligible.csv"))
+
+# aggregate data
 sponsored_articles <- hybrid_volume %>%
@@
   summarise(n = n())
+```
 
+In most cases, Elsevier sent invoices for hybrid open access publication fees to individual authors (`r round(hybrid_volume %>% filter(oa_sponsor_type == "Author") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%).
+For around `r round(hybrid_volume %>% filter(oa_sponsor_type == "FundingBody") %>% nrow() / nrow(hybrid_volume) * 100, 0)`% of articles, the publisher directly billed funders and research organisations.
+The proportion of open access articles where Elsevier granted publication fee waivers was `r round(hybrid_volume %>% filter(oa_sponsor_type == "ElsevierWaived") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%.
+
+Figure \@ref(fig:invoiceoverview) shows the annual development per invoicing type.
+Each type is visualised separately as parts of the total.
+The grey areas show the overall yearly distribution of open access articles in hybrid journals, and the proportion of each invoicing type is highlighted in blue, The figure reveals a general growth of open access articles in hybrid journals.
+It illustrates that this development was mainly driven by billing individual authors, while central invoicing stagnated.
+Also, the amount of fee-waived articles remained more or less constant from 2015 to date.
+
+```{r invoiceoverview, fig.cap="Development of fee-based open access publishing in Elsevier hybrid journals by invoicing type.
+Colored bars represent the invoice recipient, or whether the fee was waived.
+Grey bars show the total number of hybrid open access articles published in Elsevier journals from 2015 to date."}
+# plot
 ggplot(sponsored_articles, aes(x = gsub("20", "'", issued_year), y = n)) +
@@
   labs(x = NULL, y = NULL) + 
-  scale_y_continuous(labels = scales::number_format(big.mark = " ")) +
-  theme_minimal(base_family="Roboto", base_size = 12) +
-  theme(panel.grid.minor=element_blank()) +
-  theme(axis.ticks=element_blank()) +
-  theme(panel.grid.major.x=element_blank()) +
-  theme(panel.border=element_blank()) +
+  scale_y_continuous(labels = scales::number_format(big.mark = " "),
+                      expand = expansion(mult = c(0, 0.05))) + theme_minimal_hgrid() +
   theme(legend.position="top", legend.justification = "right")
 ```
+
+```{r}
+cc_invoice <- hybrid_volume %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
+   mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
+  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
+                                  `FundingBody` = "Agreement",
+                                  `ElsevierWaived` = "Fee Waived")) %>%
+  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
+  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
+  count(issued_year, fct_source, cc_norm)
+
+cc_invoice_type <- cc_invoice %>% 
+  group_by(fct_source, cc_norm) %>% 
+  summarise(n = sum(n))
+cc_norm_all <- cc_invoice_type %>%
+  ungroup() %>%
+  group_by(cc_norm) %>%
+  summarise(all = sum(n))
+cc_type_all <- cc_invoice_type %>%
+  ungroup() %>%
+  group_by(fct_source) %>%
+  summarise(all = sum(n))
+license_count <- inner_join(cc_invoice_type, cc_norm_all, by =  "cc_norm") %>% mutate(prop = n / all)
+license_summary_count <- inner_join(cc_invoice_type, cc_type_all, by =  "fct_source") %>% mutate(prop = n / all)
+```
+
+In terms of re-use, the type of invoicing is strongly related to whether the open content license permits commercial adaption or not (see figure).
+Although most open access articles published in Elsevier's hybrid journals were billed to authors directly, centrally invoiced articles accounted for the largest share of articles under an CC BY licenses (`r round(filter(license_count, fct_source == "Agreement", cc_norm == "cc-by") %>% pull(prop) * 100, 0)`%).
+Consequently, `r round(filter(license_summary_count, fct_source == "Agreement", cc_norm == "cc-by") %>% pull(prop) * 100, 0)`% of centrally invoiced articles permits commercial adaption.
+
+```{r}
+ggplot(cc_invoice, aes(factor(issued_year), n, fill = fct_source, group = fct_source)) +
+  geom_area(stat = "identity",
+            color = "white",
+            alpha = 0.8,
+            position = position_stack(reverse = T)) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = function(x)
+        format(x, big.mark = ",", scientific = FALSE),
+      breaks = scales::pretty_breaks()
+  ) +
+  facet_grid(~cc_norm) +
+  scale_fill_brewer(palette = "Set1") +
+  labs(x = "Publication year", y = "Total OA articles") +
+  theme_minimal_hgrid() +
+  theme(legend.position = "top",
+        legend.justification = "right") +
+  guides(fill = guide_legend("Invoicing"))
+```
+
+```{r}
+# invoicing analysis
+library(googlesheets4)
+tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 2)
+```
+
+```{r}
+hybrid_invoice <- hybrid_volume %>%
+  filter(oa_sponsor_type == "FundingBody") %>%
+  rename(yrl_volume = articles) %>%
+  left_join(tt, by = c("oa_sponsor_name"))
+
+sponsor_country <- hybrid_invoice %>%
+  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
+  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
+  group_by(country_factor) %>%
+  summarise(n = n_distinct(doi)) %>%
+  mutate(prop = n / sum(n))
+```
+
+Next, we examined funders and research organisations that were invoiced directly by Elsevier.
+In total, we were able to identify `r length(unique(hybrid_invoice$oa_sponsor_name))` sponsoring bodies.
+Figure presents a breakdown by country, highlighting the dominating role of a few countries.
+Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "Multinational") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%).
+In the latter case, invoices for individual open access articles were not send to research funders, but to the Dutch VSNU, a consortium  of Universities in the Netherlands that has negotiated agreements with Elsevier since 2016.
+The prominent role of Norway can be also explained by a national deal, which started in 2019.
+US-based open access sponsors originated mainly from the Bill & Melinda Gates Foundation (n = `r hybrid_invoice %>% filter(oa_sponsor_name == "Bill & Melinda Gates Foundation") %>% distinct(doi) %>% nrow()`).
+
+```{r}
+sponsor_country %>%
+  mutate(country_factor = fct_reorder(country_factor, n)) %>%
+  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
+) %>%
+  ggplot(aes(country_factor, n)) +
+  geom_bar(stat = "identity", fill = "#56B4E9") +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  coord_flip() +
+  theme_minimal_vgrid() +
+  labs(y = "Number of invoiced OA articles", x = NULL)
+```

### 2020-09-13 analysis/007_invoice_exploration.Rmd
inviocing

@@
 title: "invoice analysis"
-output: github_document
+output: html_document
 ---
@@
 
-In most cases, Elsevier sent invoices for hybrid open access publication fees to individual authors (`r round(hybrid_volume %>% filter(oa_sponsor_type == "Author") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%).
-For around `r round(hybrid_volume %>% filter(oa_sponsor_type == "FundingBody") %>% nrow() / nrow(hybrid_volume) * 100, 0)`% of articles, the publisher directly billed funders and research organisations.
-The proportion of open access articles where Elsevier granted publication fee waivers was `r round(hybrid_volume %>% filter(oa_sponsor_type == "ElsevierWaived") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%.
-
-Figure \@ref(fig:invoiceoverview) shows the annual development per invoicing type.
+To examine the funding structure for open access articles, we text-mined spending metadata from open access full-texts.
+Figure @ref(fig:invoiceoverview) shows the annual development per invoicing type.
 Each type is visualised separately as parts of the total.
-The grey areas show the overall yearly distribution of open access articles in hybrid journals, and the proportion of each invoicing type is highlighted in blue, The figure reveals a general growth of open access articles in hybrid journals.
-It illustrates that this development was mainly driven by billing individual authors, while central invoicing stagnated.
-Also, the amount of fee-waived articles remained more or less constant from 2015 to date.
+The grey areas show the overall yearly distribution of open access articles in hybrid journals, and the proportion of each invoicing type is highlighted in blue.
+The figure illustrates a general growth of open access articles in hybrid journals.
+This development was mainly driven by billing individual authors, while central invoicing stagnated.
+Also, the number of fee-waived articles remained more or less constant from 2015 to date.
+In most cases, Elsevier charged individual authors (`r round(hybrid_volume %>% filter(oa_sponsor_type == "Author") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%).
+Invoices addressed to funders and research organizations accounted for `r round(hybrid_volume %>% filter(oa_sponsor_type == "FundingBody") %>% nrow() / nrow(hybrid_volume) * 100, 0)`% of articles.
+The proportion of open access articles with publication fee waivers was `r round(hybrid_volume %>% filter(oa_sponsor_type == "ElsevierWaived") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%.
 
@@
   theme(legend.position="top", legend.justification = "right")
+```
+
+```{r}
+# invoicing analysis
+library(googlesheets4)
+tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 2)
+```
+
+```{r}
+hybrid_invoice <- hybrid_volume %>%
+  filter(oa_sponsor_type == "FundingBody") %>%
+  rename(yrl_volume = articles) %>%
+  left_join(tt, by = c("oa_sponsor_name")) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc"))
+
+sponsor_country <- hybrid_invoice %>%
+  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
+  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
+  group_by(country_factor) %>%
+  summarise(n = n_distinct(doi)) %>%
+  mutate(prop = n / sum(n))
+```
+
+Figure presents a breakdown by country, highlighting the dominating role of a few countries.
+Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "Multinational") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%).
+In the latter case, invoices for individual open access articles were not send to research funders, but to the Dutch VSNU, a consortium  of Universities in the Netherlands that has negotiated agreements with Elsevier since 2016.
+The prominent role of Norway can be also explained by a national deal, which started in 2019.
+US-based open access sponsors originated mainly from the Bill & Melinda Gates Foundation (n = `r hybrid_invoice %>% filter(oa_sponsor_name == "Bill & Melinda Gates Foundation") %>% distinct(doi) %>% nrow()`).
+
+```{r}
+sponsor_country %>%
+  mutate(country_factor = fct_reorder(country_factor, n)) %>%
+  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
+) %>%
+  ggplot(aes(country_factor, n)) +
+  geom_bar(stat = "identity", fill = "#56B4E9") +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  coord_flip() +
+  theme_minimal_vgrid() +
+  labs(y = "Number of invoiced OA articles", x = NULL)
 ```
@@
 
-In terms of re-use, the type of invoicing is strongly related to whether the open content license permits commercial adaption or not (see figure).
+In terms of licensing, the type of invoicing is strongly related to whether the open content license permits commercial adaption or not (see figure).
 Although most open access articles published in Elsevier's hybrid journals were billed to authors directly, centrally invoiced articles accounted for the largest share of articles under an CC BY licenses (`r round(filter(license_count, fct_source == "Agreement", cc_norm == "cc-by") %>% pull(prop) * 100, 0)`%).
-Consequently, `r round(filter(license_summary_count, fct_source == "Agreement", cc_norm == "cc-by") %>% pull(prop) * 100, 0)`% of centrally invoiced articles permits commercial adaption.
 
@@
 
-```{r}
-# invoicing analysis
-library(googlesheets4)
-tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 2)
+Examining funders and research organisations that were invoiced directly by  Elsevier, strongly suggests that a few funders with a strict CC-BY policy have contributed to the dominance of CC-BY license.
+Table presents the Top 10 out of `r length(unique(hybrid_invoice$oa_sponsor_name))` sponsoring bodies in terms of funded open access articles between 2015.19.
+Together, these funders accounted for around 80% of central payments.
+The table furthermore shows the number of distinct journals and the share of CC-BY licensed articles.
+Notably, UK-based research funders and the Bill & Melinda Gates Foundation sponsored mainly articles that were published under a CC-BY license.
+On the contrary, articles funded by the Dutch VSNU, the European Research Council and Norway Institutes had a much lower proportion.
+In fact, the Dutch and Norwegian agreements allowed authors to decide whether to publish under an CC-BY or CC-BY-ND license, while the UK-based funders had a CC-BY licensing policy.
+
+```{r opapc}
+# openapc
+u <-
+  "https://raw.githubusercontent.com/OpenAPC/openapc-de/master/data/apc_de.csv"
+o_apc <- readr::read_csv(u)
+#'
+#' We also would like to add data from transformative aggrements, which is also
+#' collected by the Open APC Initiative.
+#' The transformative agreements data-set does not include pricing information.
+#'
+oa_trans <-
+  readr::read_csv(
+    "https://raw.githubusercontent.com/OpenAPC/openapc-de/master/data/transformative_agreements/transformative_agreements.csv"
+  )
+oapc <- oa_trans %>%
+  bind_rows(o_apc) 
+hybrid_invoice <- hybrid_invoice %>%
+  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
+
 ```
@@
 ```{r}
-hybrid_invoice <- hybrid_volume %>%
-  filter(oa_sponsor_type == "FundingBody") %>%
-  rename(yrl_volume = articles) %>%
-  left_join(tt, by = c("oa_sponsor_name"))
+hybrid_invoice_tt <- hybrid_invoice %>%
+  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
+  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
+  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
+hybrid_invoice_agg_general <- hybrid_invoice_tt %>%
+ group_by(oa_sponsor) %>%
+  summarise(journals = n_distinct(issn),
+            articles = n_distinct(doi)) %>%
+  mutate(prop = articles / sum(articles))
+hybrid_invoice_agg_cc <- hybrid_invoice_tt %>%
+  ungroup() %>%
+  filter(cc_norm == "cc-by") %>%
+  group_by(oa_sponsor) %>%
+  summarise(cc_by = n_distinct(doi)) 
+hybrid_invoice_agg_oapc <- hybrid_invoice_tt %>%
+  filter(in_open_apc == TRUE) %>%
+  group_by(oa_sponsor) %>%
+  summarise(in_open_apc =  n_distinct(doi))
+ind_sponsor_df <- left_join(hybrid_invoice_agg_general, hybrid_invoice_agg_cc, by = "oa_sponsor") %>%
+  left_join(hybrid_invoice_agg_oapc, by = "oa_sponsor") %>%
+  mutate(cc_by = cc_by / articles,
+         in_open_apc = ifelse(is.na(in_open_apc), 0, in_open_apc),
+         in_open_apc = in_open_apc / articles) %>% 
+  mutate(oa_sponsor = fct_reorder(oa_sponsor, articles)) %>%
+  mutate(oa_sponsor = forcats::fct_relevel(oa_sponsor, "Other", after = 0L)) %>% arrange(desc(oa_sponsor))
+# overall
+overall_ind_sponsor_df <- tibble(oa_sponsor = "All", journals = length(unique(hybrid_invoice$issn)), articles = length(unique(hybrid_invoice$doi)), prop = 1, cc_by = filter(hybrid_invoice, cc_norm == "cc-by") %>% nrow() / length(unique(hybrid_invoice$doi)), in_open_apc = filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi)))
 
-sponsor_country <- hybrid_invoice %>%
-  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
-  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
-  group_by(country_factor) %>%
-  summarise(n = n_distinct(doi)) %>%
-  mutate(prop = n / sum(n))
+ind_sponsor_df %>%
+  bind_rows(overall_ind_sponsor_df) %>%
+  gt::gt() %>%
+  gt::cols_label(oa_sponsor = "OA Sponsor",
+    journals = "Journals",
+    articles = "Articles",
+    prop = "%",
+    cc_by = "CC-BY",
+    in_open_apc = "OAPC") %>%
+  tab_footnote(
+     cells_body(
+     columns = vars(oa_sponsor),
+     rows = oa_sponsor %in% c("VSNU", "Norway Institutes")
+     ),
+    footnote = "Transformative Agreements") %>%
+  tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = list(
+      cells_column_labels(everything())
+    )) %>%
+  fmt_number(
+    columns = vars(journals, articles),
+    decimals = 0
+  ) %>%
+  fmt_percent(
+    columns = vars(prop, cc_by, in_open_apc),
+    decimals = 0)  %>%
+  cols_width(
+    2:6 ~ px(80)) %>%
+    gt::cols_align(
+    align = "right",
+    columns = 2:6
+  ) %>%
+    gt::cols_align(
+    align = "left",
+    columns = 1
+  ) %>%
+  data_color(
+    columns =5:6,
+    colors = scales::col_numeric(
+      # custom defined values - notice that order matters!
+      palette = c("#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977","#b9257a","#7c1d6f"),
+      domain = NULL
+    )
+  ) %>%
+  tab_spanner(
+    label = md("**Compliance**"),
+    columns = 5:6
+  ) %>% 
+  tab_style(
+    style = cell_borders(
+      sides = "top",
+      color = "black",
+      weight = px(2)
+    ),
+    locations = cells_body(
+      columns = everything(),
+      rows = oa_sponsor == "All"
+    )
+  ) %>%
+   tab_options(
+    column_labels.border.top.color = "white",
+    column_labels.border.top.width = px(3),
+    column_labels.border.bottom.color = "black",
+    table_body.hlines.color = "white",
+    table.border.bottom.color = "white",
+    table.border.bottom.width = px(3)
+      )
 ```
 
-Next, we examined funders and research organisations that were invoiced directly by Elsevier.
-In total, we were able to identify `r length(unique(hybrid_invoice$oa_sponsor_name))` sponsoring bodies.
-Figure presents a breakdown by country, highlighting the dominating role of a few countries.
-Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "Multinational") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%).
-In the latter case, invoices for individual open access articles were not send to research funders, but to the Dutch VSNU, a consortium  of Universities in the Netherlands that has negotiated agreements with Elsevier since 2016.
-The prominent role of Norway can be also explained by a national deal, which started in 2019.
-US-based open access sponsors originated mainly from the Bill & Melinda Gates Foundation (n = `r hybrid_invoice %>% filter(oa_sponsor_name == "Bill & Melinda Gates Foundation") %>% distinct(doi) %>% nrow()`).
+```{r}
+o_apc_ind <- hybrid_volume %>% 
+  filter(oa_sponsor_type != "FundingBody") %>%
+   mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi)) %>%
+  group_by(in_open_apc) %>%
+  summarise(n = n_distinct(doi))
+```
 
-```{r}
-sponsor_country %>%
-  mutate(country_factor = fct_reorder(country_factor, n)) %>%
-  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
-) %>%
-  ggplot(aes(country_factor, n)) +
-  geom_bar(stat = "identity", fill = "#56B4E9") +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
-  coord_flip() +
-  theme_minimal_vgrid() +
-  labs(y = "Number of invoiced OA articles", x = NULL)
-```
+As another variable, the table highlights the proportion of public disclosure of funding through the Open APC initiative.
+The Open APC Initiative disclosed more than one-third of centrally invoiced articles (`r round( filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi)) * 100, 0)`%), but also publication fees that were billed to authors directly (n = `r filter(o_apc_ind, in_open_apc == TRUE) %>% pull(n)`).
+This indicates that authors were able to make use of other types of institutional support structures for publication fees such as publication funds.
+
+x

### 2020-09-13 analysis/results.Rmd
inviocing

@@
 
-## Results
-
 ### Uptake
@@
 
-#### How many Elsevier journals supported the hybrid model?
+#### Elseviers open access journal portfolio
+
+*How many Elsevier journals supported the hybrid model?*
 
@@
 
-Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio.
-Figure shows that the majority of journals were hybrid journals.
-Overall, Elsevier published `r length(unique(els_jns$issn))` open access journals, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.
-
-Elsevier's journal portfolio changed over the period of observation.
+Drawing on web-archived Elsevier APC pricing lists  [@Matthias_2020], we determined the evolution of Elsevier's open access journal portfolio.
+Figure shows that the majority of journals were hybrid journals: Elsevier published a total of `r length(unique(els_jns$issn))` open access journals, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.
+Elsevier's journal portfolio  hardly changed.
 The number of hybrid journals grew by  `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$n) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$n)` journals between 2015 and May 2020, while its proportion relative to Elsevier's overall journal portfolio slightly decreased from `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$prop, 2) * 100`% to `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$prop, 2) * 100`%.
@@
 
-```{r}
+```{r fig.cap = "Longitudinal development of Elsevier's journal portfolio by open access business model.
+The figure distinguishes between fully open access journals that provide immediate open access to all articles and the hybrid model.
+Data obtained from Elsevier APC pricing lists (Matthias 2020)."}
 ggplot(els_jns_count,
@@
 
-#### How many of hybrid journals flipped to fully open access?
+*How many of hybrid journals flipped to fully open access?*
 
@@
 
-Following the approach of our earlier study [@Matthias_2019], we started with APC pricing list information to determine the number of open access business model flips between 2015-2020.
+Following the approach of our earlier study [@Matthias_2019], we draw on historic APC pricing lists to determine the number of open access business model flips since 2015.
 Although the transition from hybrid to a fully open access business model has drawn most attention, also the reverse can be observed.
@@
 
-#### What is the uptake of open access in Elsevier's current hybrid journal portfolio?
+#### Open Access Uptake
 
-*Number and percentage of journals with at least one article*
+*What is the uptake of open access in Elsevier's current hybrid journal portfolio?*
 
@@
 They account for about 2.5% of DOIs registered by Crossref.
+Moreover, we focused on articles published in regular issues.
 
 [^1]: <https://support.unpaywall.org/support/solutions/articles/44001894783-what-does-is-paratext-mean-in-the-api->
-
-On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model.
-Of those, `r hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio.
-At the same time, `r 100 - round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of hybrid journals did not published an open access article.
 
@@
 ```
-
-*Number and percentage of immediate open access articles per hybrid journal*
 
@@
 
-Overall, `r total_df$journals` hybrid journals provided immediate open access to `r total_df$oa_n` articles between 2015 and 2019.
-The total share of hybrid open access in the Articles published of Elsevier journals was `r total_df$oa_uptake * 100`%.
+On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model.
+Of those, `r hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio.
+These journals published a total of `r total_df$oa_n` immediate open access articles between 2015 and 2019, corresponding to a share of `r total_df$oa_uptake * 100`%.
 Table present the high-level findings by year, illustrating a moderate growth in hybrid open access.
-Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r ind_df %>% filter(issued_year == "2015") %>% pull(oa_n)` (2015) to `r ind_df %>% filter(issued_year == "2019") %>% pull(oa_n)` (2019).
-However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r ind_df %>% filter(issued_year == 2015) %>% pull(oa_uptake)`% (2015) to `r ind_df %>% filter(issued_year == 2019) %>% pull(oa_uptake)`% (2019), because the number of articles published also increased during this five-years period.
+Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r ind_df %>% filter(issued_year == "2015") %>% pull(oa_n)` in 2015 to `r ind_df %>% filter(issued_year == "2019") %>% pull(oa_n)` in 2019.
+However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r ind_df %>% filter(issued_year == 2015) %>% pull(oa_uptake)`% to `r ind_df %>% filter(issued_year == 2019) %>% pull(oa_uptake)`%, because the number of articles published also increased during this five-years period.
 
@@
 
-*variation*
-
 ```{r}
@@
 
-Figure presents to which extent open access uptake rates varied among Elsevier hybrid journals between 2015 and 2019.
+Figure presents to which extent the open access uptake rates varied among Elsevier hybrid journals between 2015 and 2019.
 The box-and-whiskers plot displays the distribution of the open access percentage per journal broken down by year.
@@
 
-```{r}
+```{r, fig.cap = "Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 18%."}
 ggplot(hybrid_oa_per_jn, aes(factor(issued_year), prop)) + 
@@
 ```
-
-*Licenses used*
 
@@
 
+*Under which open content licenses are the open access articles distributed?*
+
 Figure presents a breakdown by Creative Commons license variant used for hybrid open access by year.
 Elsevier provided commercial (cc-by) and a less permissive non-commercial license (cc-by-nc) for open access publishing in otherwise subscription-based journals.
-Nearly two-third of open access articles (`r  round(hybrid_license %>% group_by(cc_norm) %>% summarise(n = sum(cc_articles)) %>% mutate(prop = n / sum(n)) %>% filter(cc_norm == "cc-by-nc") %>% pull(prop) * 100,0)`%) can be re-used for non-commercial purposes.
+`r  round(hybrid_license %>% group_by(cc_norm) %>% summarise(n = sum(cc_articles)) %>% mutate(prop = n / sum(n)) %>% filter(cc_norm == "cc-by-nc") %>% pull(prop) * 100,0)` percent of open access articles can be re-used for non-commercial purposes.
 This proportion remained the same between 2015-19.
 
-```{r}
+```{r, fig.cap = "Immediate open access articles in hybrid journals by open content license and year.
+The blue area represent articles permitting commercial adaption under a Creative Commons CC-BY license, while the purple area displays articles under a less permissive Commons CC-BY-ND license."}
 hybrid_license %>%
@@
   guides(fill = guide_legend("Creative Commons license variant"))
-````
+```
 
-*Relationship to full and delayed OA?*
+*How does hybrid open access compare to full and delayed open access?*
 
@@
 
-Next, we present a breakdown of open content licensed articles by open access type relative to Elsevier's journal productivity (Figure).
+Next, we present a breakdown of open content licensed articles by open access type relative to Elsevier's journal productivity (ses Figure).
 In total, Elsevier provided open access to `r sum(summary_licenses$n)` articles at he time of our study, corresponding to an open access share of `r round(sum(summary_licenses$n) / license_df %>% distinct(all_articles) %>% pull(all_articles) %>% sum() * 100, 0)`%.
@@
 
-```{r}
+```{r, fig.cap='Open Access in Elsevier journals in terms of article volume between 2015-2019, showing a breakdown by business type, distinguishing between immediate open access provision in otherwise subscription-based journals ("Hybrid OA"), articles in fully open access journals and delayed open access under the Elsevier Open Archive program.
+Figure A presents a breakdown by open content license.
+Y-Axis represents the proportion of articles relative to the overall article output in percent.
+Figure B presents the number of articles for each opena ccess types across the five-years period 2015-19.'}
+
 knitr::include_graphics(here::here("figure", "license_portfolio.png"))
@@
 In addition to Creative Commons licenses, Elsevier also published delayed open access under an Elsevier user license (els-user), which prohibits re-use for commercial purposes.
-The license is mainly used by journals participating in Elsevier's open archive program comprising around 140 journals that make provide free access to published articles after an embargo period ranging between 12 months and 48 months.
+The license is mainly used by journals participating in Elsevier's open archive program comprising around 140 journals, providing free access to scholarly articles after an embargo period that range between 12 months and 48 months.
 Figure reveals that about half of open access articles were made available under the Elsevier user license.
@@
 
-### How many payments for open access articles in hybrid journals were invoiced to authors, to funders, and how many APCs were waived?
+### Invoicing
 
@@
 
-In most cases, Elsevier sent invoices for hybrid open access publication fees to individual authors (`r round(hybrid_volume %>% filter(oa_sponsor_type == "Author") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%).
-For around `r round(hybrid_volume %>% filter(oa_sponsor_type == "FundingBody") %>% nrow() / nrow(hybrid_volume) * 100, 0)`% of articles, the publisher directly billed funders and research organisations.
-The proportion of open access articles where Elsevier granted publication fee waivers was `r round(hybrid_volume %>% filter(oa_sponsor_type == "ElsevierWaived") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%.
-
-Figure \@ref(fig:invoiceoverview) shows the annual development per invoicing type.
+To examine the funding structure for open access articles, we text-mined spending metadata from open access full-texts.
+Figure @ref(fig:invoiceoverview) shows the annual development per invoicing type.
 Each type is visualised separately as parts of the total.
-The grey areas show the overall yearly distribution of open access articles in hybrid journals, and the proportion of each invoicing type is highlighted in blue, The figure reveals a general growth of open access articles in hybrid journals.
-It illustrates that this development was mainly driven by billing individual authors, while central invoicing stagnated.
-Also, the amount of fee-waived articles remained more or less constant from 2015 to date.
+The grey areas show the overall yearly distribution of open access articles in hybrid journals, and the proportion of each invoicing type is highlighted in blue.
+The figure illustrates a general growth of open access articles in hybrid journals.
+This development was mainly driven by billing authors individually ("Author"), while central invoicing ("Agreement") stagnated.
+Also, the number of fee-waived articles ("Fee Waived") somehow constant from 2015 to date.
+In most cases, Elsevier charged individual authors (`r round(hybrid_volume %>% filter(oa_sponsor_type == "Author") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%.
+Invoices addressed to funders and research organizations accounted for `r round(hybrid_volume %>% filter(oa_sponsor_type == "FundingBody") %>% nrow() / nrow(hybrid_volume) * 100, 0)`% of articles.
+The proportion of open access articles with publication fee waivers was `r round(hybrid_volume %>% filter(oa_sponsor_type == "ElsevierWaived") %>% nrow() / nrow(hybrid_volume) * 100, 0)`%.
 
@@
   theme(legend.position="top", legend.justification = "right")
+```
+
+```{r}
+# invoicing analysis
+library(googlesheets4)
+tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 2)
+```
+
+```{r}
+hybrid_invoice <- hybrid_volume %>%
+  filter(oa_sponsor_type == "FundingBody") %>%
+  rename(yrl_volume = articles) %>%
+  left_join(tt, by = c("oa_sponsor_name")) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc"))
+
+sponsor_country <- hybrid_invoice %>%
+  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
+  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
+  group_by(country_factor) %>%
+  summarise(n = n_distinct(doi)) %>%
+  mutate(prop = n / sum(n))
+```
+
+Figure presents a breakdown by country, highlighting the dominating role of a few countries.
+Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "Multinational") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%).
+In the latter case, invoices for individual open access articles were not send to research funders, but to the Dutch VSNU, a consortium  of Universities in the Netherlands that has negotiated agreements with Elsevier since 2016.
+The prominent role of Norway can be also explained by a national deal, which started in 2019.
+US-based open access sponsors originated mainly from the Bill & Melinda Gates Foundation (n = `r hybrid_invoice %>% filter(oa_sponsor_name == "Bill & Melinda Gates Foundation") %>% distinct(doi) %>% nrow()`).
+
+```{r fig.cap  = 'Invoiced sponsors of open access articles in Elsevier hybrid journals by country (2015-19)'}
+sponsor_country %>%
+  mutate(country_factor = fct_reorder(country_factor, n)) %>%
+  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
+) %>%
+  ggplot(aes(country_factor, n)) +
+  geom_bar(stat = "identity", fill = "#56B4E9") +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  coord_flip() +
+  theme_minimal_vgrid() +
+  labs(y = "Number of invoiced OA articles", x = NULL)
 ```
@@
 
-In terms of re-use, the type of invoicing is strongly related to whether the open content license permits commercial adaption or not (see figure).
+In terms of licensing, the type of invoicing is strongly related to whether the open content license permits commercial adaption or not (see figure).
 Although most open access articles published in Elsevier's hybrid journals were billed to authors directly, centrally invoiced articles accounted for the largest share of articles under an CC BY licenses (`r round(filter(license_count, fct_source == "Agreement", cc_norm == "cc-by") %>% pull(prop) * 100, 0)`%).
-Consequently, `r round(filter(license_summary_count, fct_source == "Agreement", cc_norm == "cc-by") %>% pull(prop) * 100, 0)`% of centrally invoiced articles permits commercial adaption.
 
-```{r}
+```{r, fig.cap  = 'Invoicing for open access articles in hybrid journals by open content license, 2015-2019.'}
 ggplot(cc_invoice, aes(factor(issued_year), n, fill = fct_source, group = fct_source)) +
@@
 
-```{r}
-# invoicing analysis
-library(googlesheets4)
-tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 2)
+Examining funders and research organisations that were invoiced directly by  Elsevier, strongly suggests that a few funders with a strict CC-BY policy have contributed to the dominance of CC-BY license.
+Table presents the Top 10 out of `r length(unique(hybrid_invoice$oa_sponsor_name))` sponsoring bodies in terms of funded open access articles between 2015-19.
+Together, these funders accounted for around 80% of central payments.
+The table furthermore shows the number of distinct journals and the share of CC-BY licensed articles.
+Notably, UK-based research funders and the Bill & Melinda Gates Foundation sponsored mainly articles that were published under a CC-BY license.
+On the contrary, articles funded by the Dutch VSNU, the European Research Council and Norway Institutes had a much lower proportion.
+In fact, the Dutch and Norwegian agreements allowed authors to decide whether to publish under an CC-BY or CC-BY-ND license, while the UK-based funders had a CC-BY licensing policy.
+
+```{r opapc}
+# openapc
+u <-
+  "https://raw.githubusercontent.com/OpenAPC/openapc-de/master/data/apc_de.csv"
+o_apc <- readr::read_csv(u)
+#'
+#' We also would like to add data from transformative aggrements, which is also
+#' collected by the Open APC Initiative.
+#' The transformative agreements data-set does not include pricing information.
+#'
+oa_trans <-
+  readr::read_csv(
+    "https://raw.githubusercontent.com/OpenAPC/openapc-de/master/data/transformative_agreements/transformative_agreements.csv"
+  )
+oapc <- oa_trans %>%
+  bind_rows(o_apc) 
+hybrid_invoice <- hybrid_invoice %>%
+  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
+
 ```
@@
 ```{r}
-hybrid_invoice <- hybrid_volume %>%
-  filter(oa_sponsor_type == "FundingBody") %>%
-  rename(yrl_volume = articles) %>%
-  left_join(tt, by = c("oa_sponsor_name"))
+hybrid_invoice_tt <- hybrid_invoice %>%
+  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
+  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
+  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
+hybrid_invoice_agg_general <- hybrid_invoice_tt %>%
+ group_by(oa_sponsor) %>%
+  summarise(journals = n_distinct(issn),
+            articles = n_distinct(doi)) %>%
+  mutate(prop = articles / sum(articles))
+hybrid_invoice_agg_cc <- hybrid_invoice_tt %>%
+  ungroup() %>%
+  filter(cc_norm == "cc-by") %>%
+  group_by(oa_sponsor) %>%
+  summarise(cc_by = n_distinct(doi)) 
+hybrid_invoice_agg_oapc <- hybrid_invoice_tt %>%
+  filter(in_open_apc == TRUE) %>%
+  group_by(oa_sponsor) %>%
+  summarise(in_open_apc =  n_distinct(doi))
+ind_sponsor_df <- left_join(hybrid_invoice_agg_general, hybrid_invoice_agg_cc, by = "oa_sponsor") %>%
+  left_join(hybrid_invoice_agg_oapc, by = "oa_sponsor") %>%
+  mutate(cc_by = cc_by / articles,
+         in_open_apc = ifelse(is.na(in_open_apc), 0, in_open_apc),
+         in_open_apc = in_open_apc / articles) %>% 
+  mutate(oa_sponsor = fct_reorder(oa_sponsor, articles)) %>%
+  mutate(oa_sponsor = forcats::fct_relevel(oa_sponsor, "Other", after = 0L)) %>% arrange(desc(oa_sponsor))
+# overall
+overall_ind_sponsor_df <- tibble(oa_sponsor = "All", journals = length(unique(hybrid_invoice$issn)), articles = length(unique(hybrid_invoice$doi)), prop = 1, cc_by = filter(hybrid_invoice, cc_norm == "cc-by") %>% nrow() / length(unique(hybrid_invoice$doi)), in_open_apc = filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi)))
 
-sponsor_country <- hybrid_invoice %>%
-  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
-  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
-  group_by(country_factor) %>%
-  summarise(n = n_distinct(doi)) %>%
-  mutate(prop = n / sum(n))
+ind_sponsor_df %>%
+  bind_rows(overall_ind_sponsor_df) %>%
+  gt::gt() %>%
+  gt::cols_label(oa_sponsor = "OA Sponsor",
+    journals = "Journals",
+    articles = "Articles",
+    prop = "%",
+    cc_by = "CC-BY",
+    in_open_apc = "OAPC") %>%
+  tab_footnote(
+     cells_body(
+     columns = vars(oa_sponsor),
+     rows = oa_sponsor %in% c("VSNU", "Norway Institutes")
+     ),
+    footnote = "Transformative Agreements") %>%
+  tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = list(
+      cells_column_labels(everything())
+    )) %>%
+  fmt_number(
+    columns = vars(journals, articles),
+    decimals = 0
+  ) %>%
+  fmt_percent(
+    columns = vars(prop, cc_by, in_open_apc),
+    decimals = 0)  %>%
+  cols_width(
+    2:6 ~ px(80)) %>%
+    gt::cols_align(
+    align = "right",
+    columns = 2:6
+  ) %>%
+    gt::cols_align(
+    align = "left",
+    columns = 1
+  ) %>%
+  data_color(
+    columns =5:6,
+    colors = scales::col_numeric(
+      # custom defined values - notice that order matters!
+      palette = c("#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977","#b9257a","#7c1d6f"),
+      domain = NULL
+    )
+  ) %>%
+  tab_spanner(
+    label = md("**Compliance**"),
+    columns = 5:6
+  ) %>% 
+  tab_style(
+    style = cell_borders(
+      sides = "top",
+      color = "black",
+      weight = px(2)
+    ),
+    locations = cells_body(
+      columns = everything(),
+      rows = oa_sponsor == "All"
+    )
+  ) %>%
+   tab_options(
+    column_labels.border.top.color = "white",
+    column_labels.border.top.width = px(3),
+    column_labels.border.bottom.color = "black",
+    table_body.hlines.color = "white",
+    table.border.bottom.color = "white",
+    table.border.bottom.width = px(3)
+      )
 ```
 
-Next, we examined funders and research organisations that were invoiced directly by Elsevier.
-In total, we were able to identify `r length(unique(hybrid_invoice$oa_sponsor_name))` sponsoring bodies.
-Figure presents a breakdown by country, highlighting the dominating role of a few countries.
-Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "Multinational") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%).
-In the latter case, invoices for individual open access articles were not send to research funders, but to the Dutch VSNU, a consortium  of Universities in the Netherlands that has negotiated agreements with Elsevier since 2016.
-The prominent role of Norway can be also explained by a national deal, which started in 2019.
-US-based open access sponsors originated mainly from the Bill & Melinda Gates Foundation (n = `r hybrid_invoice %>% filter(oa_sponsor_name == "Bill & Melinda Gates Foundation") %>% distinct(doi) %>% nrow()`).
+```{r}
+o_apc_ind <- hybrid_volume %>% 
+  filter(oa_sponsor_type != "FundingBody") %>%
+   mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi)) %>%
+  group_by(in_open_apc) %>%
+  summarise(n = n_distinct(doi))
+```
 
-```{r}
-sponsor_country %>%
-  mutate(country_factor = fct_reorder(country_factor, n)) %>%
-  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
-) %>%
-  ggplot(aes(country_factor, n)) +
-  geom_bar(stat = "identity", fill = "#56B4E9") +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
-  coord_flip() +
-  theme_minimal_vgrid() +
-  labs(y = "Number of invoiced OA articles", x = NULL)
-```
+As another variable, the table highlights the proportion of public disclosure of funding through the Open APC initiative.
+The Open APC Initiative disclosed more than one-third of centrally invoiced articles (`r round( filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi)) * 100, 0)`%), but also publication fees that were billed to authors directly (n = `r filter(o_apc_ind, in_open_apc == TRUE) %>% pull(n)`).
+This indicates that authors were able to make use of other types of institutional support structures for publication fees such as publication funds.

### 2020-09-21 analysis/008_subject_exploration.Rmd
subject exploration

---
output: github_document
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "fg", digits = 2))
    } else{
      return(x)
    }
  }
)
```

```{r}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
```

```{r}
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_volume_eligible.csv"))
```

```{r}
tt <- hybrid_volume %>%
  group_by(issn, container_title, articles, issued_year) %>%
  summarise(n = n_distinct(doi)) %>%
  mutate(prop = n / articles)
```

read in scopus journal indicators

```{r}
scopus_19 <- readxl::read_excel("data/scopus_jn.xlsx", sheet = 2,
                                .name_repair = make_clean_names)
```

```{r}
scopus_norm <- scopus_19 %>%
  gather(print_issn, e_issn, key = "issn_tpye", value = "issn") %>%
  # trailing zero's missing in Excel spreadsheet
  mutate(
    issn = ifelse(nchar(issn) == 5, paste0("000", issn), issn),
    issn = ifelse(nchar(issn) == 6, paste0("00", issn), issn),
    issn = ifelse(nchar(issn) == 7, paste0("0", issn), issn)) %>%
  # missing hyphen
  mutate(issn = map_chr(issn, function(x) paste(c(substr(x, 1, 4), substr(x, 5, 8)), collapse = "-")))
```

missing journals (needs to be validated)

```{r}
hybrid_volume %>% 
  filter(!issn %in% scopus_norm$issn) %>% 
  distinct(issn, container_title)
```

backup

```{r}
cite_score_19 <- scopus_norm %>% 
  select(issn, contains("subject"), percentile, snip, sjr)
write_csv(cite_score_19, here::here("data", "cite_score_19.csv"))
```

```{r}
ajsc_mapped <- read_csv(here::here("data", "asjc_mapped.csv")) %>%
  mutate(top_level_code = as.character(top_level_code))
hybrid_subject <- left_join(tt, cite_score_19, by = c("issn" = "issn")) %>%
  mutate(top_level_code = substr(as.character(scopus_asjc_code_sub_subject_area), 1, 2)) %>%
  inner_join(ajsc_mapped, by = "top_level_code")
# backup subject liste
hybrid_subject %>% 
  ungroup() %>% 
  distinct(issn, container_title, subject_area, top_level = description) %>%
  write_csv(here::here("data", "jn_subjects.csv"))
```

### Correlation test

```{r}
tt <- hybrid_subject %>% 
  filter(issued_year == "2019", !is.na(percentile)
         #, prop > 0.01 & prop < 0.15) 
         )%>%
  distinct(issn, prop, percentile, snip, sjr)
```

spearman

```{r}
cor.test(tt$prop, tt$percentile, method = "spearman")
```

plot
```{r}
ggplot(tt, aes(prop, percentile)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
  coord_cartesian(xlim = c(0.01, 0.5), ylim = c(0,100))
```

- cell system, global food security

```{r}
hybrid_subject %>%
  filter(issued_year == "2019", !is.na(percentile)) %>% ggplot(aes(prop, percentile)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
  facet_wrap(~top_level_code) +
  coord_cartesian(xlim = c(0.01, 0.5), ylim = c(0,100))
```

## OA proportion per subject

```{r}
# sort by median per asjc subject group
subject_all_years <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(n),
            n = sum(articles)) %>%
  ungroup() %>%
  group_by(description) %>%
  filter(subject_area != "Multidisciplinary")
#  mutate(ajsc_top = fct_reorder(description, n / n_oa, .fun = median, .desc = TRUE)) 
 # filter(issued_year == "2019", !is.na(percentile)) %>%
  ggplot(subject_all_years, aes(fct_reorder(description, n / n_oa, .fun = median, .desc = TRUE), n_oa / n)) + 
  geom_boxplot(aes(color = subject_area)) +
  coord_flip() +
  scale_colour_manual(values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701"))

```

# nur

### 2020-09-23 analysis/008_subject_exploration.Rmd
tmp save

@@
 ```{r}
-tt <- hybrid_volume %>%
+ind_per_jn <- hybrid_volume %>%
   group_by(issn, container_title, articles, issued_year) %>%
@@
 ```{r}
-scopus_19 <- readxl::read_excel("data/scopus_jn.xlsx", sheet = 2,
+scopus_19 <- readxl::read_excel(here::here("data", "scopus_jn.xlsx"), sheet = 2,
                                 .name_repair = make_clean_names)
 ```
+
+normalize scopus jn indicators
 
@@
   mutate(top_level_code = as.character(top_level_code))
-hybrid_subject <- left_join(tt, cite_score_19, by = c("issn" = "issn")) %>%
+hybrid_subject <- left_join(ind_per_jn, cite_score_19, by = c("issn" = "issn")) %>%
   mutate(top_level_code = substr(as.character(scopus_asjc_code_sub_subject_area), 1, 2)) %>%
@@
 ```{r}
-tt <- hybrid_subject %>% 
+hybrid_subject_19 <- hybrid_subject %>% 
   filter(issued_year == "2019", !is.na(percentile)
@@
 ```{r}
-cor.test(tt$prop, tt$percentile, method = "spearman")
+cor.test(hybrid_subject_19$prop, hybrid_subject_19$percentile, method = "spearman")
 ```
@@
 ```{r}
-ggplot(tt, aes(prop, percentile)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
+ggplot(hybrid_subject_19, aes(prop, percentile)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
   coord_cartesian(xlim = c(0.01, 0.5), ylim = c(0,100))
@@
 hybrid_subject %>%
-  filter(issued_year == "2019", !is.na(percentile)) %>% ggplot(aes(prop, percentile)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
+  filter(issued_year == "2019", !is.na(percentile)) %>% 
+  ggplot(aes(prop, percentile)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
   facet_wrap(~top_level_code) +
@@
   filter(subject_area != "Multidisciplinary")
-#  mutate(ajsc_top = fct_reorder(description, n / n_oa, .fun = median, .desc = TRUE)) 
- # filter(issued_year == "2019", !is.na(percentile)) %>%
+
 ggplot(subject_all_years, aes(fct_reorder(description, n / n_oa, .fun = median, .desc = TRUE), n_oa / n)) + 
   geom_boxplot(aes(color = subject_area)) +
+  geom_jitter(alpha = .05, aes(color = subject_area)) +
   coord_flip() +
-  scale_colour_manual(values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701"))
+  coord_cartesian()
+  scale_colour_manual(values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701")) +
+  scale_y_continuous(
+    labels = scales::percent_format(accuracy = 5L),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
+  theme_minimal_vgrid()
+```
+- Comptes Rendus https://www.library.illinois.edu/mtx/2020/05/01/comptes-rendus-de-lacademie-des-sciences-important-publishing-and-access-changes/
 
+### by funder
+
+```{r}
+jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
+
+hybrid_volume_subjects <- hybrid_volume %>%
+  group_by(oa_sponsor_type, articles, issued_year, issn) %>% 
+  summarise(n_oa = n_distinct(doi)) %>% 
+  left_join(jn_subjects, by = "issn") %>%
+  ungroup()
+
+hybrid_volume_subjects %>%
+ # filter(container_title == "Journal of the American College of Cardiology") %>%
+  group_by(top_level, articles, issued_year) %>%
+  summarise(n_oa = sum(n_oa)) %>%
+  ungroup() %>%
+  group_by(top_level) %>%
+  summarise(n_oa = sum(n_oa),
+            n = sum(articles), .groups = "drop") %>%
+  mutate(prop = n_oa / n) %>%
+  arrange(desc(prop))
+
+hybrid_volume %>%
+  left_join(jn_subjects, by = "issn") %>%
+  group_by(oa_sponsor_type, issn, issued_year, top_level, articles) %>%
+  summarise(n_oa = n_distinct(doi)) %>%
+  group_by(oa_sponsor_type, issued_year, top_level) %>%
+  summarise(n_oa = sum(n_oa),
+            n = sum(articles)) %>%
+  ungroup() %>%
+  group_by(top_level, oa_sponsor_type) %>%
+  summarise(n_oa = sum(n_oa, na.rm = TRUE),
+            n = sum(n, na.rm = TRUE)) %>%
+  mutate(prop = n_oa / n) %>% View()
+
+hybrid_volume %>%
+  left_join(jn_subjects, by = "issn") %>%
+  group_by(issn, issued_year, top_level, article) %>%
+  summarise(n_oa = n_distinct(doi)) %>%
+  group_by(issued_year, top_level) %>%
+  summarise(nn = sum(n_oa))
 ```
-
-# nur

### 2020-09-24 analysis/results.Rmd
re-render post
fix pub volume 15-19

@@
   distinct(issn)
-# pub volume eligble data
+# pub volume eligible data
 hybrid_volume <- els_pub_volume %>%
@@
     mutate(prop = oa / articles) 
+# article volume 2015-19
+volume_15_19 <- els_pub_volume %>%
+  # elgible journals
+  filter(issn %in% eligible_jns$issn) %>%
+  summarise(n = sum(articles),
+            mean_articles = mean(articles),
+            sd = sd(articles))
 
@@
   journals = hybrid_volume %>% distinct(issn) %>% nrow(),
-  articles_n = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% sum(),
-  articles_mean = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% mean(),
-  articles_sd = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% sd(),
+  articles_n = volume_15_19$n,
+  articles_mean = volume_15_19$mean_articles,
+  articles_sd = volume_15_19$sd,
   oa_n = nrow(hybrid_volume),
@@
 )
-inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
+gt_1 <- inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
   dplyr::select(-ind) %>%
@@
   )
+gt_1
+#gt::gtsave(gt_1, here::here("figure", "gt_overview.png"), dpi = 300)
 ```

### 2020-09-30 analysis/008_subject_exploration.Rmd
subject by oa funder

@@
   message = FALSE,
-  echo = FALSE,
+  echo = TRUE,
   fig.width = 6,
@@
   inner_join(ajsc_mapped, by = "top_level_code")
+# back up ind
+ write_csv(hybrid_subject, here::here("data", "jn_scopus_ind_subjects.csv"))
 # backup subject liste
@@
 ```{r}
-cor.test(hybrid_subject_19$prop, hybrid_subject_19$percentile, method = "spearman")
+cor.test(hybrid_subject_19$prop, hybrid_subject_19$snip, method = "spearman")
 ```
@@
 ```{r}
-ggplot(hybrid_subject_19, aes(prop, percentile)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
-  coord_cartesian(xlim = c(0.01, 0.5), ylim = c(0,100))
+ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
+  coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
+  labs(x = "OA Uptake Rate", y = "SNIP") +
+  theme_cowplot()
 ```
@@
   filter(issued_year == "2019", !is.na(percentile)) %>% 
-  ggplot(aes(prop, percentile)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
-  facet_wrap(~top_level_code) +
-  coord_cartesian(xlim = c(0.01, 0.5), ylim = c(0,100))
+  ggplot(aes(prop, snip)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
+  facet_wrap(~top_level_code, scales = "free") +
+  theme_minimal()
 ```
@@
 
+median_all <- subject_all_years %>% 
+  ungroup() %>% 
+  distinct(issn, .keep_all = TRUE) %>% 
+  mutate(prop = n_oa /n) %>% 
+  .$prop %>% 
+  median()
+
 ggplot(subject_all_years, aes(fct_reorder(description, n / n_oa, .fun = median, .desc = TRUE), n_oa / n)) + 
   geom_boxplot(aes(color = subject_area)) +
-  geom_jitter(alpha = .05, aes(color = subject_area)) +
-  coord_flip() +
-  coord_cartesian()
-  scale_colour_manual(values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701")) +
+  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
+  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
+  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701")) +
   scale_y_continuous(
@@
   ) +
+  coord_flip(ylim=c(0,0.2)) +
   labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
-  theme_minimal_vgrid()
+  theme_minimal_vgrid(font_size = 10) +
+  theme(legend.position = "bottom") +
+  guides(color = guide_legend(ncol = 2))
 ```
+
 - Comptes Rendus https://www.library.illinois.edu/mtx/2020/05/01/comptes-rendus-de-lacademie-des-sciences-important-publishing-and-access-changes/
@@
 
-hybrid_volume %>%
+# hybrid_volume %>%
+#   left_join(jn_subjects, by = "issn") %>%
+#   group_by(oa_sponsor_type, issn, issued_year, top_level, articles) %>%
+#   summarise(n_oa = n_distinct(doi)) %>%
+#   group_by(oa_sponsor_type, issued_year, top_level) %>%
+#   summarise(n_oa = sum(n_oa),
+#             n = sum(articles)) %>%
+#   ungroup() %>%
+#   group_by(top_level, oa_sponsor_type) %>%
+#   summarise(n_oa = sum(n_oa, na.rm = TRUE),
+#             n = sum(n, na.rm = TRUE)) %>%
+#   mutate(prop = n_oa / n) %>% View()
+# 
+# 
+# hybrid_volume %>%
+#   left_join(jn_subjects, by = "issn") %>%
+#   group_by(issn, issued_year, top_level, article) %>%
+#   summarise(n_oa = n_distinct(doi)) %>%
+#   group_by(issued_year, top_level) %>%
+#   summarise(nn = sum(n_oa))
+```
+
+```{r}
+hybrid_volume_jn <- hybrid_volume %>%
+  group_by(issn) %>%
+  summarise(n_oa = n_distinct(doi))
+
+share_per_invoicing_type <- hybrid_volume %>%
+  group_by(oa_sponsor_type, issn) %>%
+  summarise(n_oa_type = n_distinct(doi)) %>%
+  left_join(hybrid_volume_jn, by = "issn") %>%
+  mutate(prop = n_oa_type / n_oa) %>%
   left_join(jn_subjects, by = "issn") %>%
-  group_by(oa_sponsor_type, issn, issued_year, top_level, articles) %>%
-  summarise(n_oa = n_distinct(doi)) %>%
-  group_by(oa_sponsor_type, issued_year, top_level) %>%
-  summarise(n_oa = sum(n_oa),
-            n = sum(articles)) %>%
+  ungroup()
+share_per_invoicing_type %>%
+  filter(oa_sponsor_type == c("FundingBody")) %>%
+  ggplot(aes(fct_reorder(top_level, prop, .fun = median, .desc = TRUE), prop)) +
+  geom_boxplot() +
+  coord_flip()
+```
+
+```{r}
+funder_shares <- hybrid_volume %>%
+   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
+  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
+                                  `FundingBody` = "Agreement",
+                                  `ElsevierWaived` = "Fee Waived")) %>%
+  group_by(oa_sponsor_type, issn) %>%
+  summarise(n_oa_type = n_distinct(doi)) %>%
+  left_join(hybrid_volume_jn, by = "issn") %>%
+  mutate(prop = n_oa_type / n_oa) %>%
+  left_join(jn_subjects, by = "issn") %>%
   ungroup() %>%
   group_by(top_level, oa_sponsor_type) %>%
-  summarise(n_oa = sum(n_oa, na.rm = TRUE),
-            n = sum(n, na.rm = TRUE)) %>%
-  mutate(prop = n_oa / n) %>% View()
+  summarise(oa_type = sum(n_oa_type)) %>%
+  mutate(prop = oa_type / sum(oa_type))
+# overall props
+oa_sponsor_type_all <- hybrid_volume %>% 
+  #exclude multidisciplinary
+  filter(issn != "2095-9273") %>%
+  count(oa_sponsor_type) %>% 
+  mutate(prop = n / sum(n))
 
-hybrid_volume %>%
+funder_shares %>% 
+  filter(oa_sponsor_type %in% c("Author", "Agreement"), top_level != "Multidisciplinary") %>%
+  pivot_wider(id_cols = top_level, names_from = oa_sponsor_type, values_from = prop) %>%
+  filter(!is.na(top_level)) %>%
+  ggplot() +
+   ggalt::geom_dumbbell(
+    aes(y = reorder(top_level, Author), x = Agreement, xend = Author),
+    colour = "grey20",
+    colour_xend = "#EDAE49",
+    colour_x = "#30638E",
+    size_x = 1.5,
+    alpha = 0.9,
+    size_xend = 1.5,
+    show.legend = TRUE
+   ) +
+  geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "Author") %>% .$prop,  colour = "#EDAE49", linetype ="dashed", size = 1) +
+    geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "FundingBody") %>% .$prop,  colour = "#30638E", linetype ="dashed", size = 1) +
+  scale_x_continuous(
+    labels = scales::percent_format(accuracy = 5L),
+    expand = expansion(mult = c(0, 0.05)),
+    limits = c(0,1)
+  ) +
+  labs(y = NULL, x = "Percentage Invoicing Type") +
+  theme_minimal_vgrid(font_size = 10)
+```
+
+issues:
+
+- high share of elseveir waived articles due to ome issue published in Nuclear and Particle Physics Proceedings: "37th International Conference on High Energy Physics (ICHEP)
+(ICHEP 2014)2-9 July 2014, Valencia, Spain" Do we want to exclude this journal?
+
+table representation
+
+```{r}
+subject_all <- funder_shares %>%
+  group_by(top_level) %>%
+  summarise(oa = sum(oa_type))
+funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level")
+# gt
+funder_overview_subject %>%
+  pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
+  ungroup() %>% 
+  # remove na and multidisciplinary 
+  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
+  relocate(Agreement, .after = Author) %>%
+  arrange(desc(Author)) %>%
+  gt::gt(rowname_col = "top_level") %>%
+  gt::cols_label(oa = "OA Articles") %>%
+  fmt_percent(
+    columns = vars(Author, Agreement, `Fee Waived`, Other),
+    decimals = 0) %>%
+  cols_width(
+    2:6 ~ px(80)) %>%
+  cols_align(
+    align = "right",
+    columns = 2:6
+  ) %>%
+  cols_width(
+    2 ~ px(120)
+  ) %>%
+    gt::cols_align(
+    align = "left",
+    columns = 1
+  ) %>%
+  fmt_number(
+    columns = vars(oa),
+    decimals = 0
+  ) %>%
+  data_color(
+    columns =3:6,
+    colors = scales::col_numeric(
+      # custom defined values - notice that order matters!
+      palette = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977","#b9257a","#7c1d6f"),
+      domain = NULL
+    )
+  ) %>%
+  tab_spanner(
+    label = md("**OA Sponsor Type**"),
+    columns = 3:6
+  ) %>%
+    tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = list(
+      cells_column_labels(everything())
+    )) %>%
+   tab_options(
+    column_labels.border.top.color = "white",
+    column_labels.border.top.width = px(3),
+    column_labels.border.bottom.color = "black",
+    table_body.hlines.color = "white",
+    table.border.bottom.color = "white",
+    table.border.bottom.width = px(3)
+      )
+```
+ correlation funder / oa share
+
+```{r}
+funder_jn_volume <- hybrid_volume %>%
+  group_by(oa_sponsor_type, issn) %>%
+  summarise(n_oa_type = n_distinct(doi)) %>%
+  left_join(hybrid_volume_jn, by = "issn") %>%
+  mutate(prop = n_oa_type / n_oa) %>%
+ # left_join(jn_subjects, by = "issn") %>%
+  ungroup() %>%
+  filter(oa_sponsor_type == "FundingBody")
+hybrid_funder_shares <- hybrid_volume %>%
+  group_by(issn) %>%
+  summarise(all = sum(articles)) %>%
+  inner_join(funder_jn_volume, by = "issn") %>%
+  mutate(oa_prop = n_oa / all)
+
+  ggplot(hybrid_funder_shares, aes(oa_prop, prop)) +
+  geom_point(alpha = .3) + 
+  geom_smooth(method = "lm") +
+  theme_cowplot()
+```
+
+```{r}
+cor.test(hybrid_funder_shares$prop, hybrid_funder_shares$oa_prop, method = "spearman")
+```
+
+oa sponsor name subject
+
+```{r}
+test_a <- hybrid_volume %>%
+  filter(oa_sponsor_type == "FundingBody") %>%
+  group_by(oa_sponsor_name, issn) %>%
+  summarise(n_oa_type = n_distinct(doi)) %>%
+  left_join(hybrid_volume_jn, by = "issn") %>%
+  mutate(prop = n_oa_type / n_oa) %>%
   left_join(jn_subjects, by = "issn") %>%
-  group_by(issn, issued_year, top_level, article) %>%
-  summarise(n_oa = n_distinct(doi)) %>%
-  group_by(issued_year, top_level) %>%
-  summarise(nn = sum(n_oa))
+  ungroup()  %>%
+  group_by(top_level, oa_sponsor_name) %>%
+  summarise(oa = sum(n_oa_type))
 ```
+
+```{r}
+tt <- test_a %>%
+  pivot_wider(names_from = "oa_sponsor_name", values_from = "oa")
+wilcox.test(tt$`Bill & Melinda Gates Foundation`, tt$VSNU, paired = TRUE)
+```

### 2020-10-01 analysis/008_subject_exploration.Rmd
co-occurence matrix funder/subject

@@
 
-ggplot(subject_all_years, aes(fct_reorder(description, n / n_oa, .fun = median, .desc = TRUE), n_oa / n)) + 
+ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
   geom_boxplot(aes(color = subject_area)) +
@@
 ```
+
+## fields per funder
+
+```{r}
+field_funder <- hybrid_volume %>%
+  filter(oa_sponsor_type == "FundingBody") %>% 
+  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
+   mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
+  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
+  left_join(jn_subjects, by = "issn") %>%
+  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
+  group_by(oa_sponsor, top_level) %>%
+  summarise(n_oa_type = n_distinct(doi)) %>%
+  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
+  mutate(oa_sponsor = fct_reorder(oa_sponsor, n_oa_type))
+ggplot(field_funder, aes(x = oa_sponsor, 
+                         y = reorder(top_level, desc(top_level)), 
+                         colour = prop, size = n_oa_type)) + 
+  geom_point() +
+  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
+                      option = "magma",
+                      name = "Percentage per funder",
+                      breaks =  scales::extended_breaks()) +
+  scale_x_discrete(position = "top") +
+  labs(x = NULL, y = NULL) +
+  theme_minimal_grid() +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank()) +
+  theme(legend.position = "top", legend.justification = "right") +
+  theme(legend.direction = "horizontal", legend.box = "vertical") +
+  theme(strip.text = element_text(face="bold"))
+```
+
  correlation funder / oa share

### 2020-10-11 analysis/005_article_volume.Rmd
exclude corrections and withdrawn articles; fix sql to include articles without page numbers

--- 2020-07-07 analysis/005_article_volume.Rmd
@@
 write_csv(article_volume, here::here("data", "article_volume.csv"))
+```
+
+### Article volume excluding paratext
+
+```{sql, connection = "con", output.var = "article_volume_para"}
+SELECT
+        issued_year,
+        issn,
+        count(distinct(doi)) as articles     
+    FROM
+        `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`      
+    WHERE
+        issued_year > 2014 
+        and issued_year < 2020     
+    AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page|^Correction$|^Corrections to|^Corrections$|^Withdrawn')
+    GROUP BY
+        issued_year,
+        issn
+```
+
+How many articles by year?
+
+```{r}
+article_volume_para %>%
+  group_by(issued_year) %>%
+  summarise(articles = sum(articles)) %>%
+  inner_join(els_vol, by = "issued_year") %>%
+  mutate(para = n - articles) %>%
+  mutate(para / n * 100)
+```
+backup
+
+```{r}
+write_csv(article_volume_para, here::here("data", "article_volume_para.csv"))
+```
+
+### Article volume regular issues excluding paratext
+
+```{sql, connection = "con", output.var = "article_volume_regular"}
+    SELECT
+        issued_year,
+        issn,
+        count(distinct(doi)) as articles          
+    FROM
+        `api-project-764811344545.cr_dump_march_20.els_hybrid_cr`           
+    WHERE
+        issued_year > 2014          
+        and issued_year < 2020              
+        AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page|^Correction$|^Corrections to|^Corrections$|^Withdrawn')
+         AND (NOT regexp_contains(page, '^S') OR page is NULL)
+    GROUP BY
+        issued_year,
+        issn
+```
+
+How many articles by year?
+
+```{r}
+article_volume_regular %>%
+  group_by(issued_year) %>%
+  summarise(articles = sum(articles)) 
+```
+
+backup
+
+```{r}
+write_csv(article_volume_regular, here::here("data", "article_volume_para_regular.csv"))
 ```

### 2020-10-11 data-raw/hybrid_jns_articles.R

@@ -44,7 +44,7 @@ invoice_df <-  readr::read_csv(here::here("data", "invoice_df.csv"))
 #' matching table between Crossref and invoicing metadata
 els_tdm <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
 #' patterns used to exclude paratext
-para <- c('^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
+para <- c('^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page|^Correction$|^Corrections to|^Corrections$|^Withdrawn')
 #'
 #' join tables
 hybrid_df <- els_tdm %>% 

### 2020-10-11 analysis/006_license.Rmd
update license data

@@
 
+```{r}
+hybrid_articles <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
+```
+
 Aggregate
@@
         AND license.content_version = 'vor'         
-        AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
-        AND NOT regexp_contains(page, '^S')
+        AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page|^Correction$|^Corrections to|^Corrections$|^Withdrawn')
+        AND (NOT regexp_contains(page, '^S') OR page is NULL)
     GROUP BY
@@
              grepl("/by/", license_url) ~ "cc-by",
-             grepl("/by-nc", license_url) ~ "cc-by-nc",
+             grepl("/by-nc", license_url) ~ "cc-by-nc-nd",
              grepl("userlicense", license_url) ~ "els-user")
@@
 ```{r}
-im_hybrid <- readr::read_csv(here::here("data", "hybrid_volume_eligible.csv"))
+im_hybrid <- hybrid_articles <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
+
 cc_hybrid <- im_hybrid %>%
-  mutate(license_code = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
+  mutate(license_code = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
   group_by(license_code, issued_year) %>%
@@
         and issued_year < 2020        
-        AND NOT     regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
-        AND NOT regexp_contains(page, '^S') 
+        and not    regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page|^Correction$|^Corrections to|^Corrections$|^Withdrawn')
+        and (not regexp_contains(page, '^S') or page is NULL)
     GROUP BY
@@
   scale_fill_manual(values = 
-                      c(`cc-by` = "#0093c7",
-                        `cc-by-nc` = "#B52141",
+                      c(`cc-by` = "#B52141",
+                        `cc-by-nc-nd` = "#0093c7",
                         `els-user` = "grey80")) +

### 2020-10-11 analysis/results.Rmd
re-render results

--- 2020-09-13 analysis/results.Rmd
@@
 library(cowplot)
-library(ggeconodist)
 library(gt)
@@
 
-```{r}
-els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
-```
-
-```{r}
-# paratext pattern
-para <- c('^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
-```
-
-```{r}
-hybrid_md <- readr::read_csv(here::here("data", "hybrid_md.csv"))
-els_tdm <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
-invoice_df <-  readr::read_csv(here::here("data", "invoice_df.csv"))
-
-hybrid_df <- els_tdm %>% 
-  select(doi, URL) %>% 
-  inner_join(invoice_df, by = c("URL" = "tdm_url")) %>% 
-  select(-URL) %>% 
-  inner_join(hybrid_md, by = c("doi")) %>% 
-  distinct() %>%
-  filter(!grepl(para, title))
-
-flipped_jns_df <- els_jns %>%
-  filter(issn %in% flipped_jns$issn) %>%
-  select(oa_model, issn, year)
-
-hybrid_df <- left_join(hybrid_df, flipped_jns_df, by = c("issn", "issued_year" = "year")) %>%
-  filter(!oa_model %in% "Open Access") %>%
-  select(-oa_model)
-```
-
 To determine the number of Elsevier hybrid journals with at least one open access article, we focused on the most current version of Elsevier's APC pricing list, lastly updated on 23 May 2020 [@Matthias_2020].
@@
 
-```{r}
-eligible_jns <- hybrid_df %>% 
-  filter(oa_archive == FALSE) %>% 
-  distinct(issn)
-```
-
-```{r}
-# required data spreadsheets
-# pub volume
-els_pub_volume <- readr::read_csv(
-  here::here("data", "article_volume_para.csv")
-  ) 
-# eligble journals 
-eligible_jns <- hybrid_df %>% 
-  filter(oa_archive == FALSE) %>% 
-  distinct(issn)
-# pub volume eligble data
-hybrid_volume <- els_pub_volume %>%
-  filter(issn %in% eligible_jns$issn) %>%
-  inner_join(hybrid_df, by = c("issn", "issued_year")) %>%
-  filter(oa_archive == FALSE) %>%
-  filter(!grepl("^S", page))
-# backup 
-write_csv(hybrid_volume, here::here("data", "hybrid_volume_eligible.csv"))
-
-els_articles <- 
-  hybrid_volume %>%
-  distinct(issued_year, issn, articles) %>%
+```{r summary_statistics_per_year}
+hybrid_oa_volume <- readr::read_csv(here::here("data", "hybrid_oa_volume.csv"))
+# journal and article metrics **per year**
+hybrid_oa_per_year <- hybrid_oa_volume %>%
   group_by(issued_year) %>%
+  # distinct jns with at least one article per year
+  filter(oa > 0) %>%
   summarise(
+      # distinct jns with at least one article per year
+    n_journals = n(),
+    # article metrics
     articles_n = sum(articles),
     articles_mean = mean(articles),
-            articles_sd = sd(articles)
-    )
-els_yr <- 
-  hybrid_volume %>%
-  group_by(issued_year) %>%
-  summarise(journals = n_distinct(issn)) %>%
-  inner_join(els_articles, by = "issued_year")
-
-# oa
-oa_per_jn_year <- hybrid_volume %>%
-  group_by(issued_year, issn) %>%
-  summarise(oa = n(), .groups = "drop")
-oa_ind <- oa_per_jn_year %>%
-  group_by(issued_year) %>%
-  summarise(oa_n = sum(oa),
+    articles_sd = sd(articles),
+    # oa articles metrics
+    oa_n = sum(oa),
     oa_mean = mean(oa),
-            oa_sd = sd(oa)) 
-# oa uptake 
-oa_uptake_ind <- hybrid_volume %>% 
-  distinct(issued_year, issn, articles) %>%
-  inner_join(oa_per_jn_year, by = c("issn", "issued_year")) %>%
+    oa_sd = sd(oa)
+    ) %>%
+  mutate(oa_uptake = oa_n / articles_n)
+# oa uptake per year
+oa_uptake_per_year <- hybrid_oa_volume %>% 
+  filter(oa > 0) %>%
   mutate(prop = oa / articles) %>%
@@
             oa_uptake_sd = sd(prop))
-# oa uptake all 
-oa_uptake_15_19 <- hybrid_volume %>% 
-  distinct(issued_year, issn, articles) %>%
-  inner_join(oa_per_jn_year, by = c("issn", "issued_year")) %>%
+# merge indicators
+ind_per_year <- inner_join(hybrid_oa_per_year, oa_uptake_per_year, by = "issued_year")
+```
+
+```{r summary_statistics_2015_19}
+# five-years article volume
+all_jn_articles <- hybrid_oa_volume %>%
   group_by(issn) %>%
-    summarise(oa = sum(oa),
-              articles = sum(articles)) %>%
-    mutate(prop = oa / articles)
+  summarise(
+    # articles
+    articles_p_jn = sum(articles)
+    ) %>%
+  ungroup() %>%
+  summarise(
+    n_journals = n(),
+    articles_n = sum(articles_p_jn),
+    articles_mean = mean(articles_p_jn),
+    articles_sd = sd(articles_p_jn)
+  )
+# five-years OA article volume
+all_oa_articles <- 
+  hybrid_oa_volume %>%
+  group_by(issn) %>%
+  summarise(
+    # articles
+    oa_p_jn = sum(oa)
+    ) %>%
+  ungroup() %>%
+  summarise(
+    oa_n = sum(oa_p_jn),
+    oa_mean = mean(oa_p_jn),
+    oa_sd = sd(oa_p_jn)
+  )
+# five-years OA uptake rate
+oa_uptake <- 
+  hybrid_oa_volume %>%
+  group_by(issn) %>%
+  summarise(
+    # articles
+    articles_p_jn = sum(articles),
+    oa_p_jn = sum(oa)
+    ) %>%
+  mutate(prop = oa_p_jn / articles_p_jn) %>%
+  ungroup() %>%
+  summarise(oa_uptake = sum(oa_p_jn) / sum(articles_p_jn),
+            oa_uptake_mean = mean(prop),
+            oa_uptake_sd = sd(prop)
+  )
+# bringing it all together
+five_years_ind <- bind_cols(issued_year = "2015-19", all_jn_articles, all_oa_articles, oa_uptake)
 
-# all 
-total_df <- tibble::tibble(
-  issued_year = "2015-19",
-  journals = hybrid_volume %>% distinct(issn) %>% nrow(),
-  articles_n = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% sum(),
-  articles_mean = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% mean(),
-  articles_sd = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% sd(),
-  oa_n = nrow(hybrid_volume),
-  oa_mean = hybrid_volume %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% mean(),
-  oa_sd = hybrid_volume %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% sd(),
-  oa_uptake = oa_n / articles_n,
-  oa_uptake_mean = mean(oa_uptake_15_19$prop),
-  oa_uptake_sd = sd(oa_uptake_15_19$prop)
-         )
-
-ind_df <- inner_join(els_yr, oa_ind, by = "issued_year") %>%
-  mutate(oa_uptake = oa_n / articles_n) %>%
-  inner_join(oa_uptake_ind, by = "issued_year") %>%
+ind_summary_table <- ind_per_year %>%
   mutate(issued_year = as.character(issued_year)) %>% 
-  bind_rows(total_df) %>%
-  mutate(oa_uptake = oa_uptake * 100) %>%
-  mutate(oa_uptake_mean = oa_uptake_mean * 100) %>%
-  mutate(oa_uptake_sd = oa_uptake_sd * 100)
-
-ind_summary_table <- ind_df %>% 
-  pivot_longer(journals:oa_uptake_sd, names_to = "indicator", values_to = "count") %>%
+  bind_rows(five_years_ind) %>%
+  mutate(across(contains("uptake")) * 100) %>%
+  pivot_longer(n_journals:oa_uptake_sd, names_to = "indicator", values_to = "count") %>%
   pivot_wider(names_from = issued_year, values_from = count)
@@
 On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model.
-Of those, `r hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio.
-These journals published a total of `r total_df$oa_n` immediate open access articles between 2015 and 2019, corresponding to a share of `r total_df$oa_uptake * 100`%.
+Of those, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_oa_volume %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio.
+These journals published a total of `r sum(hybrid_oa_volume$oa)` immediate open access articles between 2015 and 2019, corresponding to a share of `r round(sum(hybrid_oa_volume$oa) / sum(hybrid_oa_volume$articles) * 100, 1)`%.
 Table present the high-level findings by year, illustrating a moderate growth in hybrid open access.
-Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r ind_df %>% filter(issued_year == "2015") %>% pull(oa_n)` in 2015 to `r ind_df %>% filter(issued_year == "2019") %>% pull(oa_n)` in 2019.
-However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r ind_df %>% filter(issued_year == 2015) %>% pull(oa_uptake)`% to `r ind_df %>% filter(issued_year == 2019) %>% pull(oa_uptake)`%, because the number of articles published also increased during this five-years period.
+Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r hybrid_oa_volume %>% filter(issued_year == "2015") %>% pull(oa) %>% sum()` in 2015 to `r hybrid_oa_volume %>% filter(issued_year == "2019") %>% pull(oa) %>% sum()` in 2019.
+However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r round(ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake) * 100, 1)`% to `r round(ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake) *100, 1)`%, because the number of articles published also increased during this five-years period.
 
@@
   ~ind_group, ~ind, ~ind_name,
-  "Elsevier Hybrid Journals with OA article(s)", "journals", "Journals",
+  "Elsevier Hybrid Journals with OA article(s)", "n_journals", "Journals",
   "Articles published", "articles_n", "Total",
@@
   )
+gt_1
 ```
@@
 ```{r}
-els_jn_pub_volume <- els_pub_volume %>%
-  filter(issn %in% eligible_jns$issn) %>%
-  group_by(issued_year, issn) %>%
-  summarise(articles = sum(articles),
-            journals = n_distinct(issn))
-hybrid_oa_per_jn <- hybrid_volume %>%
-  group_by(issued_year, issn) %>%
-  summarise(oa_articles = n())%>%
-  inner_join(els_jn_pub_volume, by = c("issued_year", "issn")) %>%
-  ungroup() %>%
-  mutate(prop = oa_articles / articles) 
-```
-
-```{r}
-perc_df <- hybrid_oa_per_jn %>%
+perc_df <- hybrid_oa_volume %>%
+  mutate(prop = oa / articles) %>%
+  filter(oa > 0) %>%
   group_by(issued_year) %>% 
@@
 ```{r, fig.cap = "Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 18%."}
-ggplot(hybrid_oa_per_jn, aes(factor(issued_year), prop)) + 
+hybrid_oa_volume %>%
+  mutate(prop = oa / articles) %>%
+  filter(oa > 0) %>%
+ggplot(aes(factor(issued_year), prop)) + 
   geom_boxplot(outlier.size = 1, outlier.alpha = .5) + 
-  coord_cartesian(ylim=c(0,0.18)) +
+  coord_cartesian(ylim=c(0,0.2)) +
   theme_minimal_hgrid() +
@@
        y = "Open Access Uptake")
-# licenses used
 ```
@@
 # per year
-yrl_volume <- hybrid_oa_per_jn %>% 
-  group_by(issued_year) %>% 
-  summarise(articles = sum(articles))
+hybrid_articles <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
 
-hybrid_license <- hybrid_volume %>%
-  filter(oa_archive == FALSE) %>%
-  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
+hybrid_license <- hybrid_articles %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
   group_by(cc_norm, issued_year) %>%
@@
 Elsevier provided commercial (cc-by) and a less permissive non-commercial license (cc-by-nc) for open access publishing in otherwise subscription-based journals.
-`r  round(hybrid_license %>% group_by(cc_norm) %>% summarise(n = sum(cc_articles)) %>% mutate(prop = n / sum(n)) %>% filter(cc_norm == "cc-by-nc") %>% pull(prop) * 100,0)` percent of open access articles can be re-used for non-commercial purposes.
+`r  round(hybrid_license %>% group_by(cc_norm) %>% summarise(n = sum(cc_articles)) %>% mutate(prop = n / sum(n)) %>% filter(cc_norm == "cc-by-nc-nd") %>% pull(prop) * 100,0)` percent of open access articles can be re-used for non-commercial purposes.
 This proportion remained the same between 2015-19.
@@
 ```{r, fig.cap = "Immediate open access articles in hybrid journals by open content license and year.
-The blue area represent articles permitting commercial adaption under a Creative Commons CC-BY license, while the purple area displays articles under a less permissive Commons CC-BY-ND license."}
+The blue area represent articles permitting commercial adaption under a Creative Commons CC-BY license, while the purple area displays articles under a less permissive Commons CC-BY-ND-NC license."}
 hybrid_license %>%
@@
   scale_fill_manual(values = 
-                      c(`cc-by` = "#0093c7",
-  `cc-by-nc` = "#B52141")) +
+                      c(`cc-by` = "#B52141",
+                        `cc-by-nc-nd` = "#0093c7")) +
   labs(x = "Publication year", y = "Total OA articles") +
@@
 # load eligible oa articles
-hybrid_volume <- readr::read_csv(here::here("data", "hybrid_volume_eligible.csv"))
+hybrid_volume <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
 
@@
   filter(oa_sponsor_type == "FundingBody") %>%
-  rename(yrl_volume = articles) %>%
+ # rename(yrl_volume = articles) %>%
   left_join(tt, by = c("oa_sponsor_name")) %>%
-  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc"))
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd"))
 
@@
   geom_bar(stat = "identity", fill = "#56B4E9") +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::number_format(big.mark = " ")) +
   coord_flip() +
@@
   mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
-  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
   mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%

### 2020-10-12 analysis/008_subject_exploration.Rmd
update to new datasets

@@
 ```{r}
-hybrid_volume <- readr::read_csv(here::here("data", "hybrid_volume_eligible.csv"))
-```
-
-```{r}
-ind_per_jn <- hybrid_volume %>%
-  group_by(issn, container_title, articles, issued_year) %>%
-  summarise(n = n_distinct(doi)) %>%
-  mutate(prop = n / articles)
+hybrid_oa_volume <- readr::read_csv(here::here("data", "hybrid_oa_volume.csv"))
 ```
@@
 ```{r}
-hybrid_volume %>% 
+hybrid_oa_volume %>% 
   filter(!issn %in% scopus_norm$issn) %>% 
-  distinct(issn, container_title)
+  distinct(issn)
 ```
@@
   mutate(top_level_code = as.character(top_level_code))
-hybrid_subject <- left_join(ind_per_jn, cite_score_19, by = c("issn" = "issn")) %>%
+hybrid_subject <- left_join(hybrid_oa_volume, cite_score_19, by = c("issn" = "issn")) %>%
   mutate(top_level_code = substr(as.character(scopus_asjc_code_sub_subject_area), 1, 2)) %>%
@@
   ungroup() %>% 
-  distinct(issn, container_title, subject_area, top_level = description) %>%
+  distinct(issn, subject_area, top_level = description) %>%
   write_csv(here::here("data", "jn_subjects.csv"))
@@
 hybrid_subject_19 <- hybrid_subject %>% 
+  mutate(prop = oa /articles) %>%
   filter(issued_year == "2019", !is.na(percentile)
@@
 hybrid_subject %>%
+  mutate(prop = oa /articles) %>%
   filter(issued_year == "2019", !is.na(percentile)) %>% 
@@
   group_by(issn, description, subject_area) %>% 
-  summarise(n_oa = sum(n),
+  summarise(n_oa = sum(oa),
             n = sum(articles)) %>%
@@
 ```{r}
-funder_shares <- hybrid_volume %>%
+hybrid_jn_volume <- hybrid_oa_volume %>%
+  group_by(issn) %>% 
+  summarise(n_oa = sum(oa))
+
+hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
+funder_shares <- hybrid_df %>%
    mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
@@
   summarise(n_oa_type = n_distinct(doi)) %>%
-  left_join(hybrid_volume_jn, by = "issn") %>%
+  left_join(hybrid_jn_volume, by = "issn") %>%
   mutate(prop = n_oa_type / n_oa) %>%
@@
 # overall props
-oa_sponsor_type_all <- hybrid_volume %>% 
+oa_sponsor_type_all <- hybrid_df %>% 
   #exclude multidisciplinary
@@
 ```{r}
-field_funder <- hybrid_volume %>%
+field_funder <- hybrid_df %>%
   filter(oa_sponsor_type == "FundingBody") %>% 
@@
                       breaks =  scales::extended_breaks()) +
-  scale_x_discrete(position = "top") +
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 6), position = "top") +
   labs(x = NULL, y = NULL) +
-  theme_minimal_grid() +
+  theme_minimal_grid(font_size = 10) +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
@@
   theme(legend.position = "top", legend.justification = "right") +
-  theme(legend.direction = "horizontal", legend.box = "vertical") +
-  theme(strip.text = element_text(face="bold"))
+  theme(legend.direction = "horizontal", legend.box = "vertical") 
 ```

### 2020-10-22 analysis/007_invoice_exploration.Rmd
subject exploration
cc licenses per sponsorship type

@@
 title: "invoice analysis"
-output: html_document
+output: github_document
 ---
@@
   mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
-  count(issued_year, fct_source)
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
+  count(issued_year, fct_source, cc_norm)
 all_articles <- hybrid_volume %>%
@@
 # plot
-ggplot(sponsored_articles, aes(x = gsub("20", "'", issued_year), y = n)) +
+ggplot(sponsored_articles, aes(x = gsub("20", "'", issued_year), y = n), alpha = 0.8) +
   geom_bar(
     data = all_articles,
-    aes(fill = "All hybrid OA"),
     color = "transparent",
-    stat = "identity"
+    stat = "identity",
+    fill = "#b3b3b3a0"
   ) +
-  geom_bar(aes(fill = "by invoicing type"), color = "transparent", stat = "identity") +
+  geom_bar(aes(fill = cc_norm), color = "white", stat = "identity", alpha = 0.8, size = 0.25,
+           position = position_stack(reverse = T)) +
   facet_wrap( ~ fct_source, nrow = 1) +
-  scale_fill_manual(values = c("#b3b3b3a0", "#56B4E9"), name = "") +
+  scale_fill_manual(values = 
+                      c(`cc-by` = "#B52141",
+                        `cc-by-nc-nd` = "#0093c7"),
+                    name = "CC license variant") +
   labs(x = NULL, y = NULL) + 

### 2020-10-22 analysis/results.Rmd
cc licenses per sponsorship type

@@
 csl: chicago.csl
-output: html_document
+output: 
+  github_document
 always_allow_html: true

### 2020-10-25 analysis/009_pricing_exploration.Rmd
apc / funding type

---
title: "pricing differences"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
```

read invoicing data

```{r}
hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
```

apc price list

```{r}
els_jn <- readr::read_csv(here::here("data", "els_historic_jns.csv"))
```

APC development hybrid oa portflio

```{r}
els_jn %>%
  filter(oa_model == "Hybrid") %>%
  ggplot(aes(year, apc, group = year)) +
  geom_boxplot()
```

By invoicing

```{r}
my_df <- left_join(hybrid_df, els_jn, by = c("issn", "issued_year" = "year")) %>%
    mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  filter(!fct_source %in% c("Other", NA))

```

```{r}
ggplot(my_df, aes(factor(issued_year), apc,  color = fct_source)) +
  geom_boxplot()
```

All

```{r}
ggplot(my_df, aes(factor(fct_source), apc)) +
  geom_boxplot()
```

### 2020-10-25 analysis/008_subject_exploration.Rmd
update results

@@
 ```{r}
-jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
-
-hybrid_volume_subjects <- hybrid_volume %>%
-  group_by(oa_sponsor_type, articles, issued_year, issn) %>% 
-  summarise(n_oa = n_distinct(doi)) %>% 
-  left_join(jn_subjects, by = "issn") %>%
-  ungroup()
-
-hybrid_volume_subjects %>%
- # filter(container_title == "Journal of the American College of Cardiology") %>%
-  group_by(top_level, articles, issued_year) %>%
-  summarise(n_oa = sum(n_oa)) %>%
-  ungroup() %>%
-  group_by(top_level) %>%
-  summarise(n_oa = sum(n_oa),
-            n = sum(articles), .groups = "drop") %>%
-  mutate(prop = n_oa / n) %>%
-  arrange(desc(prop))
+# jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
+# 
+# hybrid_volume_subjects <- hybrid_volume %>%
+#   group_by(oa_sponsor_type, articles, issued_year, issn) %>% 
+#   summarise(n_oa = n_distinct(doi)) %>% 
+#   left_join(jn_subjects, by = "issn") %>%
+#   ungroup()
+# 
+# hybrid_volume_subjects %>%
+#  # filter(container_title == "Journal of the American College of Cardiology") %>%
+#   group_by(top_level, articles, issued_year) %>%
+#   summarise(n_oa = sum(n_oa)) %>%
+#   ungroup() %>%
+#   group_by(top_level) %>%
+#   summarise(n_oa = sum(n_oa),
+#             n = sum(articles), .groups = "drop") %>%
+#   mutate(prop = n_oa / n) %>%
+#   arrange(desc(prop))
+#
 
@@
 - high share of elseveir waived articles due to ome issue published in Nuclear and Particle Physics Proceedings: "37th International Conference on High Energy Physics (ICHEP)
-(ICHEP 2014)2-9 July 2014, Valencia, Spain" Do we want to exclude this journal?
+(ICHEP 2014) 2-9 July 2014, Valencia, Spain" Do we want to exclude this journal?
 
@@
 ```{r}
+#short funder labels
+funder_codes <- tribble(
+  ~oa_sponsor, ~oa_sponsor_short,
+  "Engineering and Physical Sciences Research Council", "EPSRC",
+  "VSNU", "VSNU",
+  "Wellcome Trust", "Wellcome",
+  "European Research Council", "ERC",
+  "Medical Research Council", "MRC",
+  "Natural Environment Research Council", "NERC",
+  "Biotechnology and Biological Sciences Research Council", "BBSRC",
+  "Bill & Melinda Gates Foundation", "Gates",
+  "Economic and Social Research Council", "ESRC",
+  "Norway Institutes", "NO",
+  "Other", "Other")
+```
+
+```{r}
 field_funder <- hybrid_df %>%
   filter(oa_sponsor_type == "FundingBody") %>% 
-  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
+  left_join(funder_codes, by = c("oa_sponsor_name"="oa_sponsor")) %>% 
+  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_short), "Other", oa_sponsor_short)) %>%
   mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
@@
   mutate(prop = n_oa_type / sum(n_oa_type)) %>%
-  mutate(oa_sponsor = fct_reorder(oa_sponsor, n_oa_type))
-ggplot(field_funder, aes(x = oa_sponsor, 
+  mutate(oa_sponsor = fct_relevel(oa_sponsor, "Other", after = Inf))
+ggplot(field_funder, aes(x = factor(oa_sponsor), 
                          y = reorder(top_level, desc(top_level)), 

### 2020-10-25 analysis/results.Rmd
update results

@@
 
-On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model.
+On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model, without reverse-flips: `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() - filter(flipped_jns_ind, oa_model != "Open Access")$n`.
 Of those, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_oa_volume %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio.
@@
 This indicates that authors were able to make use of other types of institutional support structures for publication fees such as publication funds.
+
+### Field cultures
+
+```{r}
+hybrid_subject <- readr::read_csv(here::here("data", "jn_scopus_ind_subjects.csv"))
+```
+
+```{r}
+subject_all_years <- hybrid_subject %>% 
+  ungroup() %>% 
+  group_by(issn, description, subject_area) %>% 
+  summarise(n_oa = sum(oa),
+            n = sum(articles)) %>%
+  ungroup() %>%
+  group_by(description) %>%
+  filter(subject_area != "Multidisciplinary")
+
+median_all <- subject_all_years %>% 
+  ungroup() %>% 
+  distinct(issn, .keep_all = TRUE) %>% 
+  mutate(prop = n_oa /n) %>% 
+  .$prop %>% 
+  median()
+# rank
+oa_per_subject_rank <- hybrid_subject %>% 
+  ungroup() %>% 
+  group_by(issn, description, subject_area) %>% 
+  summarise(n_oa = sum(oa),
+            n = sum(articles)) %>%
+  mutate(prop = n_oa / n) %>%
+  ungroup() %>% 
+  group_by(description) %>%
+  summarise(median_prop = median(prop)) %>%
+  arrange(desc(median_prop))
+
+ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
+  geom_boxplot(aes(color = subject_area)) +
+  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
+  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
+  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701")) +
+  scale_y_continuous(
+    labels = scales::percent_format(accuracy = 5L),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  coord_flip(ylim=c(0,0.2)) +
+  labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
+  theme_minimal_vgrid(font_size = 10) +
+  theme(legend.position = "bottom") +
+  guides(color = guide_legend(ncol = 2))
+```
+
+We used the Scopus’ All Sciences Journal Classification (ASJC) to examine if and to what extent open access uptake rates varied by subject.
+Figure presents the distribution of the journal’s five-years open access uptake rates by subject fields in comparison to the overall average (Mdn = `r round(median_all * 100, 1)`.
+The box plot reveals large variations between 27 journal subject fields (excluding one journal belonging to the area Multidisciplinary), with the highest OA rates in Immunology and Microbiology (Mdn = `r round(filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop) * 100,1)`), Environmental Science (Mdn = `r round(filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop) * 100,1)`) and Neuroscience  (Mdn = `r round(filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop) * 100,1)`), Environmental Science (Mdn = `r round(filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop) * 100,1)`) , and the lowest in Chemistry (Mdn = `r round(filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop) * 100,1)`), Materials Science (Mdn = `r round(filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop) * 100,1)`) and Dentistry (Mdn = `r round(filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop) * 100,1)`).
+Figure also highlights variations across subject areas.
+Notably, fields belonging to the subject areas Health Sciences and Physical Sciences show a lower median proportion of open access articles, whereas most Life Sciences and Social Sciences fields ranked above average.
+
+```{r}
+hybrid_jn_volume <- hybrid_oa_volume %>%
+  group_by(issn) %>% 
+  summarise(n_oa = sum(oa))
+
+jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
+
+hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
+funder_shares <- hybrid_df %>%
+   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
+  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
+                                  `FundingBody` = "Agreement",
+                                  `ElsevierWaived` = "Fee Waived")) %>%
+  group_by(oa_sponsor_type, issn) %>%
+  summarise(n_oa_type = n_distinct(doi)) %>%
+  left_join(hybrid_jn_volume, by = "issn") %>%
+  mutate(prop = n_oa_type / n_oa) %>%
+  left_join(jn_subjects, by = "issn") %>%
+  ungroup() %>%
+  group_by(top_level, oa_sponsor_type) %>%
+  summarise(oa_type = sum(n_oa_type)) %>%
+  mutate(prop = oa_type / sum(oa_type))
+# overall props
+oa_sponsor_type_all <- hybrid_df %>% 
+  #exclude multidisciplinary
+  filter(issn != "2095-9273") %>%
+  count(oa_sponsor_type) %>% 
+  mutate(prop = n / sum(n))
+
+funder_shares %>% 
+  filter(oa_sponsor_type %in% c("Author", "Agreement"), top_level != "Multidisciplinary") %>%
+  pivot_wider(id_cols = top_level, names_from = oa_sponsor_type, values_from = prop) %>%
+  filter(!is.na(top_level)) %>%
+  ggplot() +
+   ggalt::geom_dumbbell(
+    aes(y = reorder(top_level, Author), x = Agreement, xend = Author),
+    colour = "grey20",
+    colour_xend = "#EDAE49",
+    colour_x = "#30638E",
+    size_x = 1.5,
+    alpha = 0.9,
+    size_xend = 1.5,
+    show.legend = TRUE
+   ) +
+  geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "Author") %>% .$prop,  colour = "#EDAE49", linetype ="dashed", size = 1) +
+    geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "FundingBody") %>% .$prop,  colour = "#30638E", linetype ="dashed", size = 1) +
+  scale_x_continuous(
+    labels = scales::percent_format(accuracy = 5L),
+    expand = expansion(mult = c(0, 0.05)),
+    limits = c(0,1)
+  ) +
+  labs(y = NULL, x = "Percentage Invoicing Type") +
+  theme_minimal_vgrid(font_size = 10)
+```
+
+```{r}
+subject_all <- funder_shares %>%
+  group_by(top_level) %>%
+  summarise(oa = sum(oa_type))
+funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level")
+# gt
+funder_overview_subject %>%
+  pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
+  ungroup() %>% 
+  # remove na and multidisciplinary 
+  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
+  relocate(Agreement, .after = Author) %>%
+  arrange(desc(Author)) %>%
+  gt::gt(rowname_col = "top_level") %>%
+  gt::cols_label(oa = "OA Articles") %>%
+  fmt_percent(
+    columns = vars(Author, Agreement, `Fee Waived`, Other),
+    decimals = 0) %>%
+  cols_width(
+    2:6 ~ px(80)) %>%
+  cols_align(
+    align = "right",
+    columns = 2:6
+  ) %>%
+  cols_width(
+    2 ~ px(120)
+  ) %>%
+    gt::cols_align(
+    align = "left",
+    columns = 1
+  ) %>%
+  fmt_number(
+    columns = vars(oa),
+    decimals = 0
+  ) %>%
+  data_color(
+    columns =3:6,
+    colors = scales::col_numeric(
+      # custom defined values - notice that order matters!
+      palette = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977","#b9257a","#7c1d6f"),
+      domain = NULL
+    )
+  ) %>%
+  tab_spanner(
+    label = md("**OA Sponsor Type**"),
+    columns = 3:6
+  ) %>%
+    tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = list(
+      cells_column_labels(everything())
+    )) %>%
+   tab_options(
+    column_labels.border.top.color = "white",
+    column_labels.border.top.width = px(3),
+    column_labels.border.bottom.color = "black",
+    table_body.hlines.color = "white",
+    table.border.bottom.color = "white",
+    table.border.bottom.width = px(3)
+      )
+```
+
+funder by subject
+
+```{r}
+#short funder labels
+funder_codes <- tribble(
+  ~oa_sponsor, ~oa_sponsor_short,
+  "Engineering and Physical Sciences Research Council", "EPSRC",
+  "VSNU", "VSNU",
+  "Wellcome Trust", "Wellcome",
+  "European Research Council", "ERC",
+  "Medical Research Council", "MRC",
+  "Natural Environment Research Council", "NERC",
+  "Biotechnology and Biological Sciences Research Council", "BBSRC",
+  "Bill & Melinda Gates Foundation", "Gates",
+  "Economic and Social Research Council", "ESRC",
+  "Norway Institutes", "NO",
+  "Other", "Other")
+```
+
+```{r}
+field_funder <- hybrid_df %>%
+  filter(oa_sponsor_type == "FundingBody") %>% 
+  left_join(funder_codes, by = c("oa_sponsor_name"="oa_sponsor")) %>% 
+  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_short), "Other", oa_sponsor_short)) %>%
+  mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
+  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
+  left_join(jn_subjects, by = "issn") %>%
+  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
+  group_by(oa_sponsor, top_level) %>%
+  summarise(n_oa_type = n_distinct(doi)) %>%
+  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
+  mutate(oa_sponsor = fct_relevel(oa_sponsor, "Other", after = Inf))
+ggplot(field_funder, aes(x = factor(oa_sponsor), 
+                         y = reorder(top_level, desc(top_level)), 
+                         colour = prop, size = n_oa_type)) + 
+  geom_point() +
+  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
+                      option = "magma",
+                      name = "Percentage per funder",
+                      breaks =  scales::extended_breaks()) +
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 6), position = "top") +
+  labs(x = NULL, y = NULL) +
+  theme_minimal_grid(font_size = 10) +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank()) +
+  theme(legend.position = "top", legend.justification = "right") +
+  theme(legend.direction = "horizontal", legend.box = "vertical") 
+```

### 2020-10-27 analysis/results.Rmd
facetted subject view

@@
 
+```{r}
+p_subject <- ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
+  geom_boxplot(aes(color = subject_area)) +
+  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
+  facet_wrap(~subject_area, scales = "free_y") +
+  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
+  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701"),
+                      guide=FALSE) +
+  scale_y_continuous(
+    labels = scales::percent_format(accuracy = 5L),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  coord_flip(ylim=c(0,0.2)) +
+  labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
+  theme_minimal_vgrid(font_size = 10) 
+ggsave(here::here("figure", "subject_by_area.png"), p_subject, dpi = 300, width = 8, height = 6)
+p_subject
+```
+
 We used the Scopus’ All Sciences Journal Classification (ASJC) to examine if and to what extent open access uptake rates varied by subject.
@@
   filter(oa_sponsor_type == "FundingBody") %>%
-  left_join(funder_codes, by = c("oa_sponsor_name"="oa_sponsor")) %>% 
+  left_join(funder_codes, by = c("oa_sponsor_name" = "oa_sponsor")) %>%
   mutate(oa_sponsor = ifelse(is.na(oa_sponsor_short), "Other", oa_sponsor_short)) %>%
@@
   mutate(oa_sponsor = fct_relevel(oa_sponsor, "Other", after = Inf))
-ggplot(field_funder, aes(x = factor(oa_sponsor), 
+# plot
+pp <- ggplot(field_funder,
+       aes(
+         x = factor(oa_sponsor),
          y = reorder(top_level, desc(top_level)),
-                         colour = prop, size = n_oa_type)) + 
+         colour = prop,
+         size = n_oa_type
+       )) +
   geom_point() +
-  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
+  scale_color_viridis_c(
+    labels = scales::percent_format(accuracy = 5L),
     option = "magma",
     name = "Percentage per funder",
-                      breaks =  scales::extended_breaks()) +
-  scale_x_discrete(labels = function(x) str_wrap(x, width = 6), position = "top") +
+    breaks =  scales::extended_breaks(),
+    begin = 0.15,
+    end = 0.85
+  ) +
+  scale_size("OA Articles") +
+  scale_x_discrete(position = "top") +
   labs(x = NULL, y = NULL) +
   theme_minimal_grid(font_size = 10) +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank()) +
+  theme(
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    panel.background = element_blank()
+  ) +
   theme(legend.position = "top", legend.justification = "right") +
   theme(legend.direction = "horizontal", legend.box = "vertical") 
+ggsave(here::here("figure", "funder_subject.png"), pp, dpi = 300, width = 8, height = 6)
 ```

### 2020-11-18 analysis/results.Rmd
funder charaterusutucs added

@@
 library(gt)
+library(janitor)
 ```
@@
 sponsor_country <- hybrid_invoice %>%
+  # manual changes
+  mutate(Country = ifelse(grepl("European|CERN", oa_sponsor_name), "EU", Country)) %>%
+  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
   mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
   mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
-  group_by(country_factor) %>%
+  mutate(country_factor = fct_infreq(country_factor)) %>%
+  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf)) %>%
+  group_by(country_factor, `Primary research focus`) %>%
   summarise(n = n_distinct(doi)) %>%
-  mutate(prop = n / sum(n))
+  ungroup() %>%
+  mutate(prop = round(n / sum(n), 2))
+
+sponsor_country %>% 
+  ungroup() %>%
+  pivot_wider(id_cols = c("country_factor"), names_from = `Primary research focus`, values_from = prop) %>%
+  janitor::adorn_totals(c("col", "row")) %>%
+  rename(`Country/Subject Area` = country_factor) %>%
+  knitr::kable()
 ```
@@
 Figure presents a breakdown by country, highlighting the dominating role of a few countries.
-Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "Multinational") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%).
+Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "EU") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%).
 In the latter case, invoices for individual open access articles were not send to research funders, but to the Dutch VSNU, a consortium  of Universities in the Netherlands that has negotiated agreements with Elsevier since 2016.
@@
 
-```{r fig.cap  = 'Invoiced sponsors of open access articles in Elsevier hybrid journals by country (2015-19)'}
+```{r fig.cap  = 'Invoiced sponsors of open access articles in Elsevier hybrid journals by country (2015-19)', eval = FALSE}
 sponsor_country %>%

### 2020-11-18 analysis/results.Rmd
add cor test snip / prop

@@
 ```
+
+### citation
+
+```{r}
+hybrid_subject_19 <- hybrid_subject %>% 
+  mutate(prop = oa /articles) %>%
+  filter(issued_year == "2019", !is.na(percentile)
+         , prop != 0) %>%
+  distinct(issn, prop, percentile, snip, sjr)
+```
+
+spearman
+
+```{r}
+cor.test(hybrid_subject_19$prop, hybrid_subject_19$percentile, method = "spearman")
+```
+
+```{r}
+ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = 0.2) + 
+  geom_smooth(method = "loess", se = TRUE) +
+  coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
+  labs(x = "OA Uptake Rate", y = "SNIP") +
+  theme_cowplot()
+```

### 2020-12-09 analysis/006_license.Rmd
check for mirror journals

@@
 ```
+
+### License for Mirror journals
+
+```{sql, connection=con, output.var="els_yearly_mirror"}
+SELECT  issued_year,
+        container_title,
+        COUNT(DISTINCT(doi)) as all_articles 
+    FROM
+        `api-project-764811344545.cr_dump_march_20.els_all_may_20`
+    WHERE
+        issued_year > 2014 
+        and issued_year < 2020        
+        and not    regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page|^Correction$|^Corrections to|^Corrections$|^Withdrawn')
+        and (not regexp_contains(page, '^S') or page is NULL)
+        and regexp_contains(container_title, '^[a-z|A-Z|0-9]+[^I]\\s? X{1}$')
+    GROUP BY
+        issued_year,
+        container_title
+```
+
+```{r}
+els_yearly_mirror
+```
+
+```{r}
+sum(els_yearly_mirror$all_articles)
+```

### 2020-12-09 analysis/006_license.Rmd
export data
bugfix regex

--- 2020-10-11 analysis/006_license.Rmd
@@
 ```
+
+### License for Mirror journals
+
+```{sql, connection=con, output.var="els_yearly_mirror"}
+SELECT  issued_year,
+        container_title,
+        COUNT(DISTINCT(doi)) as all_articles 
+    FROM
+        `api-project-764811344545.cr_dump_march_20.els_all_may_20`
+    WHERE
+        issued_year > 2014 
+        and issued_year < 2020        
+        and not    regexp_contains(title,'^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page|^Correction$|^Corrections to|^Corrections$|^Withdrawn')
+        and (not regexp_contains(page, '^S') or page is NULL)
+        and regexp_contains(container_title, ' X$')
+    GROUP BY
+        issued_year,
+        container_title
+```
+
+```{r}
+els_yearly_mirror
+```
+
+```{r}
+sum(els_yearly_mirror$all_articles)
+length(unique(els_yearly_mirror$container_title))
+write_csv(els_yearly_mirror, here::here("data", "x_journals_volume.csv"))
+```

### 2020-12-10 analysis/results.Rmd
summary tables + gt export

@@
 
-```{r fig.cap = "Longitudinal development of Elsevier's journal portfolio by open access business model.
+```{r number_of_els_journals, fig.cap = "Longitudinal development of Elsevier's journal portfolio by open access business model.
 The figure distinguishes between fully open access journals that provide immediate open access to all articles and the hybrid model.
@@
     name = NULL,
-    labels = c("Fully", "Hybrid")
+    labels = c("Fully OA", "Hybrid OA")
   ) +
-  labs(x = "Year", y = "Elsevier Journals") +
+  labs(x = "Year", y = "Number of Elsevier Journals") +
   scale_y_continuous(
@@
         legend.justification = "right") +
-  guides(fill = guide_legend("OA Business Model"))
+  guides(fill = guide_legend("Business Model"))
 ```
@@
   ~ind_group, ~ind, ~ind_name,
-  "Elsevier Hybrid Journals with OA article(s)", "n_journals", "Journals",
   "Articles published", "articles_n", "Total",
@@
   "Articles published", "articles_sd", "SD",
-  "OA Articles published", "oa_n", "Total OA",
-  "OA Articles published", "oa_mean", "Avg.",
-  "OA Articles published", "oa_sd", "SD",
+  "of which OA", "oa_n", "Total ",
+  "of which OA", "oa_mean", "Avg.",
+  "of which OA", "oa_sd", "SD",
   "OA Uptake (in%)", "oa_uptake", "Percentage",
   "OA Uptake (in%)", "oa_uptake_mean", "Avg.",
-  "OA Uptake (in%)", "oa_uptake_sd", "SD"
+  "OA Uptake (in%)", "oa_uptake_sd", "SD",
+  "Elsevier Hybrid Journals with OA article(s)", "n_journals", "Journals"
 )
@@
   fmt_number(contains("20"), 
-             rows = c("Total", "Total OA", "Journals"),
+             rows = c("Total", "Total ", "Journals"),
              decimals = 0) %>%
@@
   group_by(cc_norm, issued_year) %>%
-  summarise(cc_articles = n())
+  summarise(cc_articles = n()) %>%
+  ungroup() %>%
+  group_by(issued_year) %>%
+  mutate(prop = cc_articles /sum(cc_articles))
+hybrid_license
 ```
@@
 Elsevier provided commercial (cc-by) and a less permissive non-commercial license (cc-by-nc) for open access publishing in otherwise subscription-based journals.
-`r  round(hybrid_license %>% group_by(cc_norm) %>% summarise(n = sum(cc_articles)) %>% mutate(prop = n / sum(n)) %>% filter(cc_norm == "cc-by-nc-nd") %>% pull(prop) * 100,0)` percent of open access articles can be re-used for non-commercial purposes.
+`r  round(hybrid_license %>% group_by(cc_norm) %>% summarise(n = sum(cc_articles)) %>% mutate(prop = n / sum(n)) %>% filter(cc_norm == "cc-by-nc-nd") %>% pull(prop) * 100,0)` percent of open access articles can be re-used for non-commercial purposes (n = `r hybrid_license %>% filter(cc_norm == "cc-by-nc-nd") %>% pull(cc_articles) %>% sum()`).
 This proportion remained the same between 2015-19.
@@
   mutate(prop = n / sum(n))
+summary_licenses
+
+license_df %>%
+  group_by(type, license_code) %>%
+  summarise(n = sum(articles)) %>%
+  mutate(prop = n / sum(n))
+```
+
+```{r mirror}
+mirror_jn <- readr::read_csv(here::here("data", "x_journals_volume.csv"))
 ```
@@
 `r round(filter(summary_licenses, type == "full_oa") %>% pull(prop) * 100)`% of open access articles appeared in full open access journals.
+
+Of those,
 
@@
 
+### wavied
+
+```{r}
+hybrid_volume %>% 
+  filter(oa_sponsor_type == "ElsevierWaived") %>%
+  count(oa_sponsor_name, sort = TRUE) %>%
+  mutate(prop = n / sum(n))
+```
 ```{r}
@@
 
+```{r}
+cc_invoice %>%
+  group_by(fct_source) %>%
+  summarise(nn = sum(n)) %>%
+  mutate(prop = nn / sum(nn))
+```
 In terms of licensing, the type of invoicing is strongly related to whether the open content license permits commercial adaption or not (see figure).
-Although most open access articles published in Elsevier's hybrid journals were billed to authors directly, centrally invoiced articles accounted for the largest share of articles under an CC BY licenses (`r round(filter(license_count, fct_source == "Agreement", cc_norm == "cc-by") %>% pull(prop) * 100, 0)`%).
+Although most open access articles published in Elsevier's hybrid journals were billed to authors directly, centrally invoiced articles accounted for the largest share of articles under an CC BY licenses.
+
+```{r}
+cc_invoice %>%
+  filter(fct_source == "Agreement") %>%
+  group_by(cc_norm) %>%
+  summarise(nn = sum(n)) %>%
+  mutate(prop = nn / sum(nn))
+```
 
@@
 
-ind_sponsor_df %>%
+gt_sponsor <-ind_sponsor_df %>%
   bind_rows(overall_ind_sponsor_df) %>%
   gt::gt() %>%
-  gt::cols_label(oa_sponsor = "OA Sponsor",
+  gt::cols_label(oa_sponsor = "Hybrid OA Sponsor",
     journals = "Journals",
@@
       )
+gt_sponsor
+gt::gtsave(gt_sponsor, here::here("figure","gt_sponsor.png"))
 ```
@@
 # gt
-funder_overview_subject %>%
+gt_subject_all <- funder_overview_subject %>%
   pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
@@
   gt::gt(rowname_col = "top_level") %>%
-  gt::cols_label(oa = "OA Articles") %>%
+  gt::cols_label(oa = "Hybrid OA Articles") %>%
   fmt_percent(
@@
   tab_spanner(
-    label = md("**OA Sponsor Type**"),
+    label = md("**Hybrid OA Sponsor Type**"),
     columns = 3:6
@@
       )
+gt::gtsave(gt_subject_all, here::here("figure","gt_subject_all.png"))
+gt_subject_all
 ```

### 2020-12-17 analysis/008_subject_exploration.Rmd
re order, first overall volume, then closer look into licenses

@@
   filter(issued_year == "2019", !is.na(percentile)
-         #, prop > 0.01 & prop < 0.15) 
-         )%>%
+         , prop != 0) %>%
   distinct(issn, prop, percentile, snip, sjr)
@@
 ```{r}
-cor.test(hybrid_subject_19$prop, hybrid_subject_19$snip, method = "spearman")
+cor.test(hybrid_subject_19$prop, hybrid_subject_19$percentile, method = "spearman")
 ```
@@
 ```{r}
-ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
-  coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
+ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = 0.2) + 
+  geom_smooth(method = "loess", se = TRUE) +
+ # coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
   labs(x = "OA Uptake Rate", y = "SNIP") +
@@
 ```{r}
-hybrid_volume_jn <- hybrid_volume %>%
-  group_by(issn) %>%
-  summarise(n_oa = n_distinct(doi))
-
-share_per_invoicing_type <- hybrid_volume %>%
-  group_by(oa_sponsor_type, issn) %>%
-  summarise(n_oa_type = n_distinct(doi)) %>%
-  left_join(hybrid_volume_jn, by = "issn") %>%
-  mutate(prop = n_oa_type / n_oa) %>%
-  left_join(jn_subjects, by = "issn") %>%
-  ungroup()
-share_per_invoicing_type %>%
-  filter(oa_sponsor_type == c("FundingBody")) %>%
-  ggplot(aes(fct_reorder(top_level, prop, .fun = median, .desc = TRUE), prop)) +
-  geom_boxplot() +
-  coord_flip()
+# hybrid_volume_jn <- hybrid_volume %>%
+#   group_by(issn) %>%
+#   summarise(n_oa = n_distinct(doi))
+# 
+# share_per_invoicing_type <- hybrid_volume %>%
+#   group_by(oa_sponsor_type, issn) %>%
+#   summarise(n_oa_type = n_distinct(doi)) %>%
+#   left_join(hybrid_volume_jn, by = "issn") %>%
+#   mutate(prop = n_oa_type / n_oa) %>%
+#   left_join(jn_subjects, by = "issn") %>%
+#   ungroup()
+# share_per_invoicing_type %>%
+#   filter(oa_sponsor_type == c("FundingBody")) %>%
+#   ggplot(aes(fct_reorder(top_level, prop, .fun = median, .desc = TRUE), prop)) +
+#   geom_boxplot() +
+#   coord_flip()
 ```

### 2020-12-17 analysis/006_license.Rmd
update license

@@
 p <- cowplot::plot_grid(
-  p_1, p_2,
+  p_2, p_1,
   labels = "AUTO", ncol = 1

### 2020-12-17 analysis/results.Rmd
tmp save

@@
   ~ind_group, ~ind, ~ind_name,
-  "Articles published", "articles_n", "Total",
-  "Articles published", "articles_mean", "Avg.",
-  "Articles published", "articles_sd", "SD",
-  "of which OA", "oa_n", "Total ",
-  "of which OA", "oa_mean", "Avg.",
-  "of which OA", "oa_sd", "SD",
-  "OA Uptake (in%)", "oa_uptake", "Percentage",
-  "OA Uptake (in%)", "oa_uptake_mean", "Avg.",
-  "OA Uptake (in%)", "oa_uptake_sd", "SD",
-  "Elsevier Hybrid Journals with OA article(s)", "n_journals", "Journals"
+    "Elsevier Hybrid Journals with ≥ 1 OA article", "n_journals", "All (n)",
+  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_n", "All (n) ",
+  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_mean", "Avg.",
+  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_sd", "SD",
+  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_n", "OA (n)",
+  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_mean", "Avg.",
+  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_sd", "SD",
+  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake", "Total (%)",
+  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_mean", "Avg.",
+  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_sd", "SD"
 )
@@
   fmt_number(contains("20"), 
-             rows = c("Total", "Total ", "Journals"),
+             rows = c("All (n)", "All (n) ", "OA (n)"),
              decimals = 0) %>%
@@
   ) %>%
+  tab_style(
+    style = 
+      cell_text(weight = "bold"),
+    locations = cells_stub(c("All (n)", "All (n) ", "OA (n)", "Total (%)"))) %>%
   tab_style(
@@
 gt_1
+gt::gtsave(gt_1, here::here("figure","gt_overview.png"))
+
 ```
@@
       expand = expansion(mult = c(0, 0.05))) +
-  labs(x = "Publication Years",
-       y = "Open Access Uptake")
+  labs(x = "Publication Year",
+       y = "Hybrid OA Uptake")
 ```
@@
   mutate(prop = n / sum(n))
+all_oa <- license_df %>% 
+  group_by(type) %>% 
+  summarise(n = sum(articles)) %>%
+  pull(n) %>%
+  sum()
+all_oa
 ```
@@
   summarise(n = n())
+```
+
+```{r}
+sponsored_articles %>% 
+  group_by(fct_source) %>% 
+  summarise(nn = sum(n)) %>% 
+  mutate(cumsum = cumsum(nn)) %>% 
+  mutate(prop = nn / sum(nn) * 100) %>%
+  mutate(cumprop = cumsum(prop)) %>%
+  adorn_totals("row") %>%
+  knitr::kable(digits = 1)
 ```
@@
                     name = "CC license variant") +
-  labs(x = NULL, y = NULL) + 
-  scale_y_continuous(labels = scales::number_format(big.mark = " "),
-                      expand = expansion(mult = c(0, 0.05))) + theme_minimal_hgrid() +
+  labs(x = "Publication year", y = "Hybrid OA Articles") + 
+  scale_y_continuous( 
+                     labels = scales::number_format(big.mark = " "),
+                      expand = expansion(mult = c(0, 0.05))) +
+  theme_minimal_hgrid() +
   theme(legend.position="top", legend.justification = "right")
@@
 
+```{r}
+hybrid_invoice <-hybrid_volume %>%
+  filter(oa_sponsor_type == "FundingBody") %>%
+ # rename(yrl_volume = articles) %>%
+  left_join(tt, by = c("oa_sponsor_name")) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd"))
+
+sponsor_country_t <- hybrid_invoice %>%
+  # manual changes
+  mutate(Country = ifelse(grepl("European|CERN", oa_sponsor_name), "EU", Country)) %>%
+  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
+  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
+  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
+  mutate(country_factor = fct_infreq(country_factor)) %>%
+  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))
+
+sponsor_country_t %>%
+  count(issued_year, Sector, country_factor, cc_norm) %>%
+  ggplot(aes(issued_year, n, fill = Sector)) + 
+  geom_bar(stat = "identity") +
+  facet_wrap(~country_factor)
+```
+
 Figure presents a breakdown by country, highlighting the dominating role of a few countries.
@@
 US-based open access sponsors originated mainly from the Bill & Melinda Gates Foundation (n = `r hybrid_invoice %>% filter(oa_sponsor_name == "Bill & Melinda Gates Foundation") %>% distinct(doi) %>% nrow()`).
+
+```{r}
+sponsor_country %>%
+  group_by(country_factor) %>%
+  summarise(nn = sum(n)) %>%
+  mutate(prop = nn / sum(nn))
+```
 
@@
   coord_flip(ylim=c(0,0.2)) +
-  labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
+  labs(x = NULL, y = "Five-Years Hybrid OA Uptake Rate per Journal") +
   theme_minimal_vgrid(font_size = 10) 
@@
   coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
-  labs(x = "OA Uptake Rate", y = "SNIP") +
+  labs(x = "Hybrid OA Uptake", y = "SNIP") +
   theme_cowplot()

### 2020-12-20 analysis/results.Rmd
sector analysis

@@
 ```{r}
+els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv")) 
 els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
-els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv")) 
+
 els_jns_count <- els_jns %>%
@@
   mutate(oa_model_fct = fct_rev(as_factor(oa_model))) 
+els_jns_count
 ```
@@
 all_oa
+# all articles
+license_df %>%
+  distinct(all_articles, issued_year) %>%
+  pull(all_articles) %>%
+  sum()
 ```
@@
 library(googlesheets4)
-tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 2)
+tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 3)
 ```
@@
   mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd"))
-
+```
+
+```{r license_count}
+sponsor_df <- hybrid_invoice %>% 
+  filter(!is.na(Country)) %>% 
+  # manual changes
+  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "EU", Country)) %>%
+  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
+  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
+  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
+  mutate(country_factor = fct_infreq(country_factor)) %>%
+  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))
+sponsor_df %>%
+  group_by(country_factor, cc_norm) %>% 
+  summarise(n = n_distinct(doi)) %>%
+  ggplot(aes(fct_rev(country_factor), n, fill = cc_norm)) +
+  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
+            color = "white", alpha = 0.8, size = 0.25) +
+  coord_flip() +
+  scale_fill_manual(values = 
+                      c(`cc-by` = "#B52141",
+                        `cc-by-nc-nd` = "#0093c7"),
+                    name = "CC license variant") +
+  labs(x = "OA Sponsor by Country", y = "Hybrid OA Articles (Agreement)") +
+  scale_y_continuous(labels = scales::number_format(big.mark = " "),
+                      expand = expansion(mult = c(0, 0.05)),
+                     breaks = seq(0,15000, 2000)) +
+  theme_minimal_vgrid() +
+  theme(legend.position="top", legend.justification = "right")
+```
+
+```{r}
 sponsor_country <- hybrid_invoice %>%
+  filter(!is.na(`Primary research focus`)) %>%
   # manual changes
-  mutate(Country = ifelse(grepl("European|CERN", oa_sponsor_name), "EU", Country)) %>%
+  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "EU", Country)) %>%
   mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
@@
   mutate(prop = round(n / sum(n), 2))
-
+```
+
+```{r}
 sponsor_country %>%
@@
   pivot_wider(id_cols = c("country_factor"), names_from = `Primary research focus`, values_from = prop) %>%
+  mutate_all(~replace(., is.na(.), 0)) %>%
   janitor::adorn_totals(c("col", "row")) %>%
@@
   # manual changes
-  mutate(Country = ifelse(grepl("European|CERN", oa_sponsor_name), "EU", Country)) %>%
+  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "EU", Country)) %>%
   mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
@@
 
-sponsor_country_t %>%
-  count(issued_year, Sector, country_factor, cc_norm) %>%
-  ggplot(aes(issued_year, n, fill = Sector)) + 
-  geom_bar(stat = "identity") +
-  facet_wrap(~country_factor)
+sponsor_df %>%
+  count(issued_year, Sector, country_factor) %>%
+  mutate(sector_level = factor(Sector)) %>%
+  mutate(sector_level = fct_relevel(sector_level, "government", "higher education", "private non-profit", "other")) %>%
+  ggplot(aes(gsub("20", "'", issued_year), n, fill = sector_level)) + 
+  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
+            color = "white", alpha = 0.8, size = 0.25) +
+  facet_wrap(~country_factor) +
+  scale_fill_manual("Sector", values = c("#65BADA", "#00A757", "#E5BA52", "grey50")) +
+  scale_y_continuous(labels = scales::number_format(big.mark = " "),
+                      expand = expansion(mult = c(0, 0.05))) +
+  theme_minimal_hgrid() +
+  theme(legend.position="top", legend.justification = "right") +
+  guides(fill = guide_legend(nrow = 2)) +
+  labs(x = "Publication year", y = "Hybrid OA Articles (Agreement)")
 ```
@@
 ```
+
 ```{r}

### 2021-01-17 analysis/008_subject_exploration.Rmd
improve issn matching

--- 2020-10-12 analysis/008_subject_exploration.Rmd
@@
 scopus_norm <- scopus_19 %>%
-  gather(print_issn, e_issn, key = "issn_tpye", value = "issn") %>%
+  separate(print_issn, c("print_issn", "print_issn_extra")) %>% 
+  gather(print_issn, e_issn, print_issn_extra, key = "issn_tpye", value = "issn") %>%
   # trailing zero's missing in Excel spreadsheet
@@
 ```
-
-### Correlation test
-
-```{r}
-hybrid_subject_19 <- hybrid_subject %>% 
-  mutate(prop = oa /articles) %>%
-  filter(issued_year == "2019", !is.na(percentile)
-         #, prop > 0.01 & prop < 0.15) 
-         )%>%
-  distinct(issn, prop, percentile, snip, sjr)
-```
-
-spearman
-
-```{r}
-cor.test(hybrid_subject_19$prop, hybrid_subject_19$snip, method = "spearman")
-```
-
-plot
-```{r}
-ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
-  coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
-  labs(x = "OA Uptake Rate", y = "SNIP") +
-  theme_cowplot()
-```
-
-- cell system, global food security
-
-```{r}
-hybrid_subject %>%
-  mutate(prop = oa /articles) %>%
-  filter(issued_year == "2019", !is.na(percentile)) %>% 
-  ggplot(aes(prop, snip)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
-  facet_wrap(~top_level_code, scales = "free") +
-  theme_minimal()
-```
-
-## OA proportion per subject
-
-```{r}
-# sort by median per asjc subject group
-subject_all_years <- hybrid_subject %>% 
-  ungroup() %>% 
-  group_by(issn, description, subject_area) %>% 
-  summarise(n_oa = sum(oa),
-            n = sum(articles)) %>%
-  ungroup() %>%
-  group_by(description) %>%
-  filter(subject_area != "Multidisciplinary")
-
-median_all <- subject_all_years %>% 
-  ungroup() %>% 
-  distinct(issn, .keep_all = TRUE) %>% 
-  mutate(prop = n_oa /n) %>% 
-  .$prop %>% 
-  median()
-
-ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
-  geom_boxplot(aes(color = subject_area)) +
-  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
-  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
-  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701")) +
-  scale_y_continuous(
-    labels = scales::percent_format(accuracy = 5L),
-    expand = expansion(mult = c(0, 0.05))
-  ) +
-  coord_flip(ylim=c(0,0.2)) +
-  labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
-  theme_minimal_vgrid(font_size = 10) +
-  theme(legend.position = "bottom") +
-  guides(color = guide_legend(ncol = 2))
-```
-
-- Comptes Rendus https://www.library.illinois.edu/mtx/2020/05/01/comptes-rendus-de-lacademie-des-sciences-important-publishing-and-access-changes/
-
-### by funder
-
-```{r}
-jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
-
-hybrid_volume_subjects <- hybrid_volume %>%
-  group_by(oa_sponsor_type, articles, issued_year, issn) %>% 
-  summarise(n_oa = n_distinct(doi)) %>% 
-  left_join(jn_subjects, by = "issn") %>%
-  ungroup()
-
-hybrid_volume_subjects %>%
- # filter(container_title == "Journal of the American College of Cardiology") %>%
-  group_by(top_level, articles, issued_year) %>%
-  summarise(n_oa = sum(n_oa)) %>%
-  ungroup() %>%
-  group_by(top_level) %>%
-  summarise(n_oa = sum(n_oa),
-            n = sum(articles), .groups = "drop") %>%
-  mutate(prop = n_oa / n) %>%
-  arrange(desc(prop))
-
-# hybrid_volume %>%
-#   left_join(jn_subjects, by = "issn") %>%
-#   group_by(oa_sponsor_type, issn, issued_year, top_level, articles) %>%
-#   summarise(n_oa = n_distinct(doi)) %>%
-#   group_by(oa_sponsor_type, issued_year, top_level) %>%
-#   summarise(n_oa = sum(n_oa),
-#             n = sum(articles)) %>%
-#   ungroup() %>%
-#   group_by(top_level, oa_sponsor_type) %>%
-#   summarise(n_oa = sum(n_oa, na.rm = TRUE),
-#             n = sum(n, na.rm = TRUE)) %>%
-#   mutate(prop = n_oa / n) %>% View()
-# 
-# 
-# hybrid_volume %>%
-#   left_join(jn_subjects, by = "issn") %>%
-#   group_by(issn, issued_year, top_level, article) %>%
-#   summarise(n_oa = n_distinct(doi)) %>%
-#   group_by(issued_year, top_level) %>%
-#   summarise(nn = sum(n_oa))
-```
-
-```{r}
-hybrid_volume_jn <- hybrid_volume %>%
-  group_by(issn) %>%
-  summarise(n_oa = n_distinct(doi))
-
-share_per_invoicing_type <- hybrid_volume %>%
-  group_by(oa_sponsor_type, issn) %>%
-  summarise(n_oa_type = n_distinct(doi)) %>%
-  left_join(hybrid_volume_jn, by = "issn") %>%
-  mutate(prop = n_oa_type / n_oa) %>%
-  left_join(jn_subjects, by = "issn") %>%
-  ungroup()
-share_per_invoicing_type %>%
-  filter(oa_sponsor_type == c("FundingBody")) %>%
-  ggplot(aes(fct_reorder(top_level, prop, .fun = median, .desc = TRUE), prop)) +
-  geom_boxplot() +
-  coord_flip()
-```
-
-```{r}
-hybrid_jn_volume <- hybrid_oa_volume %>%
-  group_by(issn) %>% 
-  summarise(n_oa = sum(oa))
-
-hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
-funder_shares <- hybrid_df %>%
-   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
-  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
-                                  `FundingBody` = "Agreement",
-                                  `ElsevierWaived` = "Fee Waived")) %>%
-  group_by(oa_sponsor_type, issn) %>%
-  summarise(n_oa_type = n_distinct(doi)) %>%
-  left_join(hybrid_jn_volume, by = "issn") %>%
-  mutate(prop = n_oa_type / n_oa) %>%
-  left_join(jn_subjects, by = "issn") %>%
-  ungroup() %>%
-  group_by(top_level, oa_sponsor_type) %>%
-  summarise(oa_type = sum(n_oa_type)) %>%
-  mutate(prop = oa_type / sum(oa_type))
-# overall props
-oa_sponsor_type_all <- hybrid_df %>% 
-  #exclude multidisciplinary
-  filter(issn != "2095-9273") %>%
-  count(oa_sponsor_type) %>% 
-  mutate(prop = n / sum(n))
-
-funder_shares %>% 
-  filter(oa_sponsor_type %in% c("Author", "Agreement"), top_level != "Multidisciplinary") %>%
-  pivot_wider(id_cols = top_level, names_from = oa_sponsor_type, values_from = prop) %>%
-  filter(!is.na(top_level)) %>%
-  ggplot() +
-   ggalt::geom_dumbbell(
-    aes(y = reorder(top_level, Author), x = Agreement, xend = Author),
-    colour = "grey20",
-    colour_xend = "#EDAE49",
-    colour_x = "#30638E",
-    size_x = 1.5,
-    alpha = 0.9,
-    size_xend = 1.5,
-    show.legend = TRUE
-   ) +
-  geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "Author") %>% .$prop,  colour = "#EDAE49", linetype ="dashed", size = 1) +
-    geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "FundingBody") %>% .$prop,  colour = "#30638E", linetype ="dashed", size = 1) +
-  scale_x_continuous(
-    labels = scales::percent_format(accuracy = 5L),
-    expand = expansion(mult = c(0, 0.05)),
-    limits = c(0,1)
-  ) +
-  labs(y = NULL, x = "Percentage Invoicing Type") +
-  theme_minimal_vgrid(font_size = 10)
-```
-
-issues:
-
-- high share of elseveir waived articles due to ome issue published in Nuclear and Particle Physics Proceedings: "37th International Conference on High Energy Physics (ICHEP)
-(ICHEP 2014)2-9 July 2014, Valencia, Spain" Do we want to exclude this journal?
-
-table representation
-
-```{r}
-subject_all <- funder_shares %>%
-  group_by(top_level) %>%
-  summarise(oa = sum(oa_type))
-funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level")
-# gt
-funder_overview_subject %>%
-  pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
-  ungroup() %>% 
-  # remove na and multidisciplinary 
-  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
-  relocate(Agreement, .after = Author) %>%
-  arrange(desc(Author)) %>%
-  gt::gt(rowname_col = "top_level") %>%
-  gt::cols_label(oa = "OA Articles") %>%
-  fmt_percent(
-    columns = vars(Author, Agreement, `Fee Waived`, Other),
-    decimals = 0) %>%
-  cols_width(
-    2:6 ~ px(80)) %>%
-  cols_align(
-    align = "right",
-    columns = 2:6
-  ) %>%
-  cols_width(
-    2 ~ px(120)
-  ) %>%
-    gt::cols_align(
-    align = "left",
-    columns = 1
-  ) %>%
-  fmt_number(
-    columns = vars(oa),
-    decimals = 0
-  ) %>%
-  data_color(
-    columns =3:6,
-    colors = scales::col_numeric(
-      # custom defined values - notice that order matters!
-      palette = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977","#b9257a","#7c1d6f"),
-      domain = NULL
-    )
-  ) %>%
-  tab_spanner(
-    label = md("**OA Sponsor Type**"),
-    columns = 3:6
-  ) %>%
-    tab_style(
-    style = cell_text(color = "black", weight = "bold"),
-    locations = list(
-      cells_column_labels(everything())
-    )) %>%
-   tab_options(
-    column_labels.border.top.color = "white",
-    column_labels.border.top.width = px(3),
-    column_labels.border.bottom.color = "black",
-    table_body.hlines.color = "white",
-    table.border.bottom.color = "white",
-    table.border.bottom.width = px(3)
-      )
-```
-
-## fields per funder
-
-```{r}
-field_funder <- hybrid_df %>%
-  filter(oa_sponsor_type == "FundingBody") %>% 
-  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
-   mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
-  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
-  left_join(jn_subjects, by = "issn") %>%
-  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
-  group_by(oa_sponsor, top_level) %>%
-  summarise(n_oa_type = n_distinct(doi)) %>%
-  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
-  mutate(oa_sponsor = fct_reorder(oa_sponsor, n_oa_type))
-ggplot(field_funder, aes(x = oa_sponsor, 
-                         y = reorder(top_level, desc(top_level)), 
-                         colour = prop, size = n_oa_type)) + 
-  geom_point() +
-  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
-                      option = "magma",
-                      name = "Percentage per funder",
-                      breaks =  scales::extended_breaks()) +
-  scale_x_discrete(labels = function(x) str_wrap(x, width = 6), position = "top") +
-  labs(x = NULL, y = NULL) +
-  theme_minimal_grid(font_size = 10) +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank()) +
-  theme(legend.position = "top", legend.justification = "right") +
-  theme(legend.direction = "horizontal", legend.box = "vertical") 
-```
-
- correlation funder / oa share
-
-```{r}
-funder_jn_volume <- hybrid_volume %>%
-  group_by(oa_sponsor_type, issn) %>%
-  summarise(n_oa_type = n_distinct(doi)) %>%
-  left_join(hybrid_volume_jn, by = "issn") %>%
-  mutate(prop = n_oa_type / n_oa) %>%
- # left_join(jn_subjects, by = "issn") %>%
-  ungroup() %>%
-  filter(oa_sponsor_type == "FundingBody")
-hybrid_funder_shares <- hybrid_volume %>%
-  group_by(issn) %>%
-  summarise(all = sum(articles)) %>%
-  inner_join(funder_jn_volume, by = "issn") %>%
-  mutate(oa_prop = n_oa / all)
-
-  ggplot(hybrid_funder_shares, aes(oa_prop, prop)) +
-  geom_point(alpha = .3) + 
-  geom_smooth(method = "lm") +
-  theme_cowplot()
-```
-
-```{r}
-cor.test(hybrid_funder_shares$prop, hybrid_funder_shares$oa_prop, method = "spearman")
-```
-
-oa sponsor name subject
-
-```{r}
-test_a <- hybrid_volume %>%
-  filter(oa_sponsor_type == "FundingBody") %>%
-  group_by(oa_sponsor_name, issn) %>%
-  summarise(n_oa_type = n_distinct(doi)) %>%
-  left_join(hybrid_volume_jn, by = "issn") %>%
-  mutate(prop = n_oa_type / n_oa) %>%
-  left_join(jn_subjects, by = "issn") %>%
-  ungroup()  %>%
-  group_by(top_level, oa_sponsor_name) %>%
-  summarise(oa = sum(n_oa_type))
-```
-
-```{r}
-tt <- test_a %>%
-  pivot_wider(names_from = "oa_sponsor_name", values_from = "oa")
-wilcox.test(tt$`Bill & Melinda Gates Foundation`, tt$VSNU, paired = TRUE)
-```

### 2021-01-23 analysis/006_license.Rmd
apa label

@@
            case_when(
-             grepl("/by/", license_url) ~ "cc-by",
-             grepl("/by-nc", license_url) ~ "cc-by-nc-nd",
-             grepl("userlicense", license_url) ~ "els-user")
+             grepl("/by/", license_url) ~ "CC-BY",
+             grepl("/by-nc", license_url) ~ "CC-BY-NC-ND",
+             grepl("userlicense", license_url) ~ "Els-User")
   ) %>%
@@
 cc_hybrid <- im_hybrid %>%
-  mutate(license_code = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
+  mutate(license_code = ifelse(grepl("/by/", URL), "CC-BY", "CC-BY-NC-ND")) %>%
   group_by(license_code, issued_year) %>%
@@
 els_user <- license_by_year %>%
-  filter(license_code == "els-user") %>%
+  filter(license_code == "Els-User") %>%
   mutate(type = "hybrid_delayed_articles") %>%
@@
   ungroup() %>%
-  filter(license_code != "els-user", oa_type == "Open Access") %>%
+  filter(license_code != "Els-User", oa_type == "Open Access") %>%
   mutate(type = "full_oa") %>%
@@
   scale_fill_manual(values = 
-                      c(`cc-by` = "#B52141",
-                        `cc-by-nc-nd` = "#0093c7",
-                        `els-user` = "grey80")) +
-  labs(x = "Publication year", y = "OA Percentage") +
+                      c(`CC-BY` = "#B52141",
+                        `CC-BY-NC-ND` = "#0093c7",
+                        `Els-User` = "grey80")) +
+  labs(x = "Publication Year", y = "OA Percentage") +
   theme_minimal_hgrid() +

### 2021-01-23 analysis/manuscript.Rmd
refactor results

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: no
    theme: readable
csl: apa.csl
bibliography: references.bib
---

```{r knitr_options, , echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300,
  fig.pos = "H", 
  out.extra = ""
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "f", digits = 0))
    } else{
      return(x)
    }
  }
)
```

```{r libraries}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
library(kableExtra)
```

# Results{.unnumbered}

This section first presents the results of our analysis of hybrid OA uptake in Elsevier's journal portfolio with a view to licensing, disciplinary differences, and citation impact. Then, we present a descriptive analysis of Elsevier's invoicing data, highlighting licensing and disciplinary differences for different sponsor types, institutional sectors, and individual funders.

```{r data}
# journal-level data
els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv")) 
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
# article-level data
hybrid_oa_volume <- readr::read_csv(here::here("data", "hybrid_oa_volume.csv"))
```
## Open Access Uptake in Hybrid Journals{.unnumbered}

### Journal-level analysis{.unnumbered}

```{r}
# journals by year and oa model
els_jns_count <- els_jns %>%
  group_by(year, oa_model) %>%
  count(year) %>%
  group_by(year) %>%
  mutate(prop = n / sum(n)) %>%
  mutate(oa_model_fct = fct_rev(as_factor(oa_model)))
# overall count
 els_overall <- els_jns %>% 
  group_by(oa_model) %>% 
  summarise(n = n_distinct(issn))
```

Figure \ref{fig:number_of_els_journals} shows that in Elsevier's OA portfolio of fully OA and hybrid OA journals, the latter has outweighed the former consistently throughout the five years, with only little variation overall. In 2015, Elsevier published `r els_jns_count %>% filter(year == 2015, oa_model == "Open Access") %>% pull(n)` fully OA journals and `r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(n)` hybrid journals (`r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(prop)*100`%). Since then, Elsevier's hybrid journals increased in number by `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n)) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% pull(n))` journals (n=`r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` in 2020) and slightly decreased in proportion to Elsevier's overall OA journal portfolio by 1% (`r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(prop), 2) * 100`%; `r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` out of `r els_jns_count %>% filter(year == 2020) %>% pull(n) %>% sum()` journals).

```{r number_of_els_journals, fig.cap = "Longitudinal development of Elsevier's journal portfolio by open access business model. The figure distinguishes between fully open access journals that provide immediate open access to all articles and the hybrid model. Data obtained from Elsevier APC pricing lists (Matthias 2020)."}
ggplot(els_jns_count,
       aes(as.character(year), n,  fill = oa_model_fct, group = oa_model_fct)) +
  geom_area(alpha = 0.8,
            colour = "white") +
  scale_fill_manual(
    values = c("#cccccca0", "#56b4e9"),
    name = NULL,
    labels = c("Fully OA", "Hybrid OA")
  ) +
  labs(x = "Year", y = "Number of Elsevier Journals") +
  scale_y_continuous(
    labels = scales::number_format(big.mark = ","),
    expand = expansion(mult = c(0, 0.05)),
    breaks =  scales::extended_breaks()
  ) +
  theme_minimal_hgrid() +
  theme(legend.position = "top",
        legend.justification = "right") +
  guides(fill = guide_legend("Business Model"))
```

```{r, rev_flipped}
flipped_jns <- els_jns %>%
  distinct(issn, oa_model) %>%
  group_by(issn) %>%
  filter(n() > 1)
flipped_jns_ind <- els_jns %>%
  filter(issn %in% flipped_jns$issn) %>%
  group_by(issn) %>%
  slice(which.max(year)) %>%
  ungroup() %>%
  count(oa_model)
```

We identified `r filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` OA journals that reverse flipped--transitioned from fully OA to hybrid [@Matthias_2019]. We excluded these journals from the following article-level analysis; if we had not, OA articles published in an OA journal before it reverse-flipped would have been falsely attributed to the hybrid model. Likewise, we identified and excluded journals flipping from hybrid to fully OA in the same timeframe (n=`r filter(flipped_jns_ind, oa_model == "Open Access") %>% pull(n)`). Our final sample for the article-level analysis consisted of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` unique hybrid journals.

### Article-level analysis{.unnumbered}

```{r hybrid_oa_per_year}
# journal and article metrics **per year**
hybrid_oa_per_year <- hybrid_oa_volume %>%
  group_by(issued_year) %>%
  # distinct jns with at least one article per year
  filter(oa > 0) %>%
  summarise(
      # distinct jns with at least one article per year
    n_journals = n(),
    # article metrics
    articles_n = sum(articles),
    articles_mean = mean(articles),
    articles_sd = sd(articles),
    # oa articles metrics
    oa_n = sum(oa),
    oa_mean = mean(oa),
    oa_sd = sd(oa)
    ) %>%
  mutate(oa_uptake = oa_n / articles_n)
# oa uptake per year
oa_uptake_per_year <- hybrid_oa_volume %>% 
  filter(oa > 0) %>%
  mutate(prop = oa / articles) %>%
  group_by(issued_year) %>%
  summarise(oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop))
# merge indicators
ind_per_year <- inner_join(hybrid_oa_per_year, oa_uptake_per_year, by = "issued_year")

# five-years article volume
all_jn_articles <- hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles)
    ) %>%
  ungroup() %>%
  summarise(
    n_journals = n(),
    articles_n = sum(articles_p_jn),
    articles_mean = mean(articles_p_jn),
    articles_sd = sd(articles_p_jn)
  )
# five-years OA article volume
all_oa_articles <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    oa_p_jn = sum(oa)
    ) %>%
  ungroup() %>%
  summarise(
    oa_n = sum(oa_p_jn),
    oa_mean = mean(oa_p_jn),
    oa_sd = sd(oa_p_jn)
  )
# five-years OA uptake rate
oa_uptake <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles),
    oa_p_jn = sum(oa)
    ) %>%
  mutate(prop = oa_p_jn / articles_p_jn) %>%
  ungroup() %>%
  summarise(oa_uptake = sum(oa_p_jn) / sum(articles_p_jn),
            oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop)
  )
# bringing it all together
five_years_ind <- bind_cols(issued_year = "2015-19", all_jn_articles, all_oa_articles, oa_uptake)

ind_summary_table <- ind_per_year %>%
  mutate(issued_year = as.character(issued_year)) %>% 
  bind_rows(five_years_ind) %>%
  mutate(across(contains("uptake")) * 100) %>%
  pivot_longer(n_journals:oa_uptake_sd, names_to = "indicator", values_to = "count") %>%
  pivot_wider(names_from = issued_year, values_from = count)
```

Between 2015 and 2019, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` (`r  hybrid_oa_volume %>% distinct(issn) %>% nrow() / (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n))*100`%) hybrid journals published at least one OA article. Together these journals published `r five_years_ind %>% pull(oa_n)` OA articles, representing `r five_years_ind %>% pull(oa_uptake)*100`% of their total article volume (n=`r five_years_ind %>% pull(articles_n)`). Table \ref{tab:overview} presents these findings for each year and the aggregated five-year period between 2015-19, illustrating a moderate growth in hybrid OA over time. The number of hybrid journals with at least one OA article increased each year, as did the number of OA articles, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` 10,672 in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019. However, since the total number of articles published in these journals also increased over time, the relative share of OA articles in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.

```{r overview}
table_descriptives_ind <- tribble(
  ~ind_group, ~ind, ~` `,
    "Elsevier Hybrid Journals with ≥ 1 OA article", "n_journals", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_n", "Total ",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_sd", "SD",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_n", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_sd", "SD",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake", "Total (%)",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_mean", "Avg.",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_sd", "SD"
)
inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
  dplyr::select(-ind, -ind_group) %>%
  kableExtra::kbl(booktabs = T, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
                  caption = "Overview of hybrid open access uptake across Elsevier hybrid journals with at least one OA article 2015-2019.",
                  position = "H"
) %>%
  kable_styling() %>%
  pack_rows("Elsevier Hybrid Journals with OA article", 1, 1) %>%
  pack_rows("Articles", 2, 4) %>%
  pack_rows("OA Articles", 5, 7) %>%
  pack_rows("OA Uptake", 8, 10)
```

```{r}
perc_df <- hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
  group_by(issued_year) %>% 
  summarize(median = median(prop),
            p_10 = quantile(prop, 0.1),
            p_95 = quantile(prop, 0.95),
            iqr = quantile(prop, 3/4) - quantile(prop, 1/4)) %>%
  ungroup()
```

Figure \ref{fig:boxuptake} presents the hybrid OA uptake by year, showing considerable variation among Elsevier's hybrid journals. The box-and-whiskers plot visualizes the median hybrid OA uptake (horizontal line inside the box), the range of the majority of journals (the box), and individual outliers (dots). Notably, upper quartiles and whiskers stretch farther over the years, indicating that the range of uptake among journals with an above-average proportion of OA articles increased over the years. Nevertheless, the prevalence of hybrid OA is relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.

```{r boxuptake, fig.cap = "Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 20\\%."}
hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
ggplot(aes(factor(issued_year), prop)) + 
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + 
  coord_cartesian(ylim=c(0,0.2)) +
  theme_minimal_hgrid() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L),
      expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Publication Year",
       y = "Hybrid OA Uptake")
```

### Share of the total volume published in OA{.unnumbered}

```{r oa_volume}
# see analysis/006_license.Rmd how the data was obtained

license_df <- readr::read_csv(here::here("data", "year_per_oa_type_and_license.csv"))
# all oa
all_oa <- license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  pull(n) %>%
  sum()
# all articles
all_els <- license_df %>%
  distinct(all_articles, issued_year) %>%
  pull(all_articles) %>%
  sum()
# summary by type
summary_oa_type <- 
  license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n))
# mirror journals
mirror_jn <- readr::read_csv(here::here("data", "x_journals_volume.csv"))
```

During the same period, Elsevier provided OA to `r all_oa` articles, corresponding to a share of `r all_oa / all_els * 100` of the total article volume (n=`r all_els`). The OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind fully OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`, `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r length(unique(mirror_jn$container_title))` mirror journals that published `r sum(mirror_jn$all_articles)` articles, and Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) both in number and share of OA articles (see Figure \ref{fig:oa_volume_fig}-A).

```{r oa_volume_fig, fig.cap="Open Access in Elsevier journals in terms of article volume between 2015-2019. Figure A presents the share of open access (OA) articles relative to Elsevier's overall article output for each OA type from 2015-2019 and by open content license. Figure B presents the aggregated OA article volume of Elsevier journals by OA type."}
# see analysis/006_license.Rmd how the figure was created

knitr::include_graphics(here::here("figure", "license_portfolio.png"))
```

### License prevalence{.unnumbered}

```{r cc_by_by_type}
cc_by_by_type <- license_df %>%
  group_by(type, license_code) %>%
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n)) %>%
  filter(license_code == "CC-BY")
```

As far as we observed, Elsevier licensed OA articles in hybrid journals under two possible Creative Commons (CC) licenses—CC-BY or CC-BY-NC-ND. CC-BY allows others to distribute, remix, adapt, and build upon the licensed work, including for commercial purposes, as long as the original author is credited. In contrast, CC-BY-NC-ND is less permissive as it prohibits commercial reuse and derivatives. In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC-BY-NC-ND license marginally increased and maintained a greater share than that of CC-BY (see Figure \ref{fig:oa_volume_fig}-B). Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid OA had the highest number and proportion of openly available articles licensed under CC-BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to fully OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`, `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program, Open Archive (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) . Most of these delayed OA articles were provided under an Elsevier user license (Els-User), prohibiting reuse for commercial purposes.

### Hybrid OA uptake across subjects{.unnumbered}

```{r subject_data}
# Load hybrid article-level dataset and journal classification
# see
jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
# article level data set, see data-raw/hybrid_jns_articles.R
hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
subject_hybrid <- left_join(hybrid_df, jn_subjects, by = "issn") %>%
  select(doi, issn, top_level, subject_area)

# Number of Journals (full and fractional counting)
subject_area_all_jn <- subject_hybrid %>% 
  group_by(issn, subject_area) %>%
  summarise(n = n()) %>% 
  mutate(fractions = n / sum(n)) %>% 
  ungroup() %>% 
  group_by(subject_area) %>% 
  summarise(journals_n = n_distinct(issn),
            journals_frac = sum(fractions)) %>% 
  arrange(desc(journals_frac))
# Number of articles (full and fractional counting)
subject_area_all_article <- subject_hybrid %>% 
  group_by(doi, subject_area) %>%
  summarise(n = n()) %>% 
  mutate(fractions = n / sum(n)) %>% 
  ungroup() %>% 
  group_by(subject_area) %>% 
  summarise(articles_n = n_distinct(doi),
            articles_frac = sum(fractions)) %>% 
  arrange(desc(articles_frac))
```

Next, we present disciplinary variations in hybrid OA. Table \ref{tab:subject_area_table} presents high-level findings by AJSC subject area. In the five-years period studied, the Physical Science (`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and Health Sciences (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) were the most prevalent in terms of hybrid journals with at least one OA article published.  In terms of article volume, most OA articles were published in hybrid journals belonging to the Life Sciences (`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles), while Social Sciences  (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles) journals published the lowest number of OA articles. However, it must be noted that there is a large overlap between the journals’ subject areas. Therefore, Table also assigns hybrid journals and OA articles fractionally to a subject area in addition to full counting when a journal belongs to more than one field. In the following, we apply full counting, meaning that journals and articles associated with more than one subject in the AJSC classification were counted once for each area or field.

```{r subject_area_table}
# join
inner_join(subject_area_all_jn, subject_area_all_article, by = "subject_area") %>%
  arrange(subject_area) %>%
  select(1, journals_n, articles_n, journals_frac, articles_frac) %>%
  filter(subject_area != "Multidisciplinary", !is.na(subject_area)) %>% 
  kbl(booktabs = TRUE, col.names = c("Subject area", "Journals", "OA articles", "Journals", "OA articles"),
      caption = "Subject area overview using full and fractional counting. Table reports the number of Elsevier hybrid journals with at least one OA article, as well as the number of OA articles published in them between 2015-2019. Not shown is one journal coded as Multidisciplinary that published 61 OA articles. Eleven journals with 37 OA articles could not be matched to the Scopus Source Title list used to obtain ASJC codes.") %>%
  add_header_above(c(" " = 1, "Full counting" = 2, "Fractional counting" = 2)) %>%
  kable_styling() 
```

```{r subject_ind}
# see 008_subject_exploration
hybrid_subject <- readr::read_csv(here::here("data", "jn_scopus_ind_subjects.csv"))
```

```{r}
subject_all_years <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  ungroup() %>%
  group_by(description) %>%
  filter(subject_area != "Multidisciplinary")

median_all <- subject_all_years %>% 
  ungroup() %>% 
  distinct(issn, .keep_all = TRUE) %>% 
  mutate(prop = n_oa /n) %>% 
  .$prop %>% 
  median()
# rank
oa_per_subject_rank <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  mutate(prop = n_oa / n) %>%
  ungroup() %>% 
  group_by(description) %>%
  summarise(median_prop = median(prop)) %>%
  arrange(desc(median_prop))
```

Figure \ref{fig:oa_sub_uptake} presents the journals' five-year OA uptake grouped by subject area and field and in comparison to the overall average (Mdn= `r median_all * 100`; excluding one journal coded as Multidisciplinary). The box plot reveals some variation between the subject areas. Notably, with the exception of Environmental Science and Earth and Planetary Sciences, journals from the Physical Sciences show a lower median proportion of hybrid OA articles, whereas most Life Sciences and Social Sciences journals ranked above average. Figure X also shows large variations among the 26 subject fields, with the highest OA rates in Immunology and Microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), Environmental Science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and Neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), and the lowest in Chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), Materials Science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and Dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`).

```{r oa_sub_uptake, fig.cap='Open access (OA) uptake in Elsevier hybrid journals in percent grouped by subject area and field between 2015-19. Visualized as a box-and-whiskers plot limited to an OA share of 20\\%; the red line represents the overall median (2.5), jittered dots represent individual journals. Subject fields are ordered based on the median OA uptake.', out.width="90%", fig.width=9}
p_subject <- ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
  geom_boxplot(aes(color = subject_area)) +
  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
  facet_wrap(~subject_area, scales = "free_y") +
  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701"),
                      guide=FALSE) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05))
  ) +
  coord_flip(ylim=c(0,0.2)) +
  labs(x = NULL, y = "Five-Year Hybrid OA Uptake Rate per Journal") +
  theme_minimal_vgrid() 
p_subject
```

### Citation impact{.unnumbered}

```{r}
hybrid_subject_19 <- hybrid_subject %>% 
  mutate(prop = oa /articles) %>%
  filter(issued_year == "2019", !is.na(percentile)
         , prop != 0) %>%
  distinct(issn, prop, percentile, snip, sjr)

cor_test_res <- cor.test(hybrid_subject_19$prop, hybrid_subject_19$percentile, method = "spearman") 
```

Furthermore, we were interested in the relationship between a journal's field-specific impact and its OA uptake. Spearman's rho correlation coefficient was used to assess the relationship between the 2019 Source Normalized Impact per Paper (SNIP) value calculated by Scopus and the journal's OA uptake that year. Considering only journals with at least one OA article, we found a weak positive correlation between journal impact and OA uptake ($r_s$=`r as.character(round(cor_test_res$estimate,4))`, $p<0.001$), suggesting that the relationship is not very strong, based on our data (see Figure \ref{fig:sniptest}).

```{r sniptest, fig.cap="Source Normalized Impact per Paper (SNIP) versus open access (OA) uptake 2019. Each dot represents a hybrid journal with at least one published OA article in 2019. There is only a weak, positive correlation between the two variables ($r_s$=0.1679, $p<0.001$)"}
ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = 0.2) + 
  geom_smooth(method = "loess", se = TRUE) +
  coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
  labs(x = "Hybrid OA Uptake", y = "SNIP") +
  theme_cowplot()
```

## Invoicing for Hybrid Open Access{.unnumbered}

### Invoice types

```{r}
# load eligible oa articles
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_articles.csv"))

# aggregate data
sponsored_articles <- hybrid_volume %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
  count(issued_year, fct_source, cc_norm)
all_articles <- hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(n = n())
sponsored_art <- sponsored_articles %>% 
  group_by(fct_source) %>% 
  summarise(nn = sum(n)) %>% 
  #mutate(cumsum = cumsum(nn)) %>% 
  mutate(prop = nn / sum(nn) * 100) %>%
 # mutate(cumprop = cumsum(prop)) %>%
  adorn_totals("row")
```

Table \ref{tab:invoicing_overview} shows that authors (`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) were by far the most frequent invoicing type for hybrid OA, followed by funding bodies/institutional agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%). Interestingly, we also found a small number of cases where hybrid APCs were waived (`r filter(sponsored_art, fct_source == "Fee Waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee Waived") %>% pull(prop)`%).

```{r invoicing_overview}
sponsored_art %>% 
  kbl(booktabs = TRUE, col.names = c("Invoicing type", "Hybrid OA articles", "Percentage"),
      caption = "Types of invoicing for hybrid OA in Elsevier journals between 2015-19", digits = 1,
      format.args = list(big.mark = ",")) %>%
  kable_styling()
```

Figure X illustrates that over the years, Elsevier has increasingly and, to a greater extent, invoiced hybrid OA APCs to authors directly rather than to funders or consortia ("Agreement"). The share of fee-waived articles remained relatively stable at around 6%. However, we found different types of waivers. Around 52% of fee waivers were linked to an OA sponsor, indicating a third-party paid the APC, whereas 48% of Elsevier waived articles did not disclose an OA sponsor. For instance, 856 waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors.  In terms of licensing, Figure X reveals that the type of invoicing was strongly related to the open content license variant. When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=32,086; 77%), whereas most articles with agreed payment were licensed under the more permissive CC BY license (n=18,331; 76%).

### 2021-01-23 analysis/006_license.Rmd
tmp save

@@
         legend.justification = "right") +
-  guides(fill = guide_legend("Open content license"))
+  guides(fill = guide_legend("Open Content License"))
 p_1

### 2021-01-23 analysis/results.Rmd
tmp save

@@
   theme(legend.position="top", legend.justification = "right")
+```
+```{r}
+sponsored_articles %>% 
+  group_by(fct_source, cc_norm) %>%
+  summarise(n = sum(n)) %>%
+  mutate(prop = n / sum(n))
 ```
@@
   theme(legend.position="top", legend.justification = "right")
+```
+
+```{r}
+sponsor_df %>%
+  group_by(country_factor) %>% 
+  summarise(n = n_distinct(doi)) %>%
+  mutate(prop = n /sum(n))
 ```

### 2021-01-25 analysis/006_license.Rmd
kableextra with conditional color :-)

@@
 )
-ggsave(here::here("figure", "license_portfolio.png"), p, dpi = 300, width = 6, height = 6)
+ggsave(here::here("figure", "license_portfolio.png"), p, dpi = 300, width = 6.5, height = 6)
 ```

### 2021-01-25 analysis/manuscript.Rmd
kableextra with conditional color :-)

@@
 library(kableExtra)
+library(stringr)
 ```
@@
 This section first presents the results of our analysis of hybrid OA uptake in Elsevier's journal portfolio with a view to licensing, disciplinary differences, and citation impact.
-Then, we present a descriptive analysis of Elsevier's invoicing data, highlighting licensing and disciplinary differences for different sponsor types, institutional sectors, and individual funders.
+Then, we present a descriptive analysis of Elsevier's invoicing data, highlighting licensing and disciplinary differences for different invoicing channels, institutional sectors, and individual invoice recipients.
 
@@
 
-### Journal-level analysis{.unnumbered}
+### Journal-Level Analysis{.unnumbered}
 
@@
 
-Figure \ref{fig:number_of_els_journals} shows that in Elsevier's OA portfolio of fully OA and hybrid OA journals, the latter has outweighed the former consistently throughout the five years, with only little variation overall.
+```{r, rev_flipped}
+flipped_jns <- els_jns %>%
+  distinct(issn, oa_model) %>%
+  group_by(issn) %>%
+  filter(n() > 1)
+flipped_jns_ind <- els_jns %>%
+  filter(issn %in% flipped_jns$issn) %>%
+  group_by(issn) %>%
+  slice(which.max(year)) %>%
+  ungroup() %>%
+  count(oa_model)
+```
+
+Figure \ref{fig:number_of_els_journals} visualizes  Elsevier's OA portfolio throughout the five years from 2015-2019, and shows that hybrid journals consistently outweighed fully OA journals, with only little variation overall.
 In 2015, Elsevier published `r els_jns_count %>% filter(year == 2015, oa_model == "Open Access") %>% pull(n)` fully OA journals and `r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(n)` hybrid journals (`r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(prop)*100`%).
-Since then, Elsevier's hybrid journals increased in number by `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n)) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% pull(n))` journals (n=`r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` in 2020) and slightly decreased in proportion to Elsevier's overall OA journal portfolio by 1% (`r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(prop), 2) * 100`%; `r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` out of `r els_jns_count %>% filter(year == 2020) %>% pull(n) %>% sum()` journals).
+Since then, hybrid journals grew in number by `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n)) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% pull(n))` journals (n=`r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` in 2020), resulting in a slightly decreased share of Elsevier's overall OA journal portfolio by 1% (`r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(prop), 2) * 100`%; `r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` out of `r els_jns_count %>% filter(year == 2020) %>% pull(n) %>% sum()` journals).
+Of these, `r filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` journals transitioned from fully OA to hybrid.
+For the article-level analysis, we excluded these reverse-flip journals, focussing on `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` unique hybrid journals without OA business model change for the remaining results.
 
@@
 
-```{r, rev_flipped}
-flipped_jns <- els_jns %>%
-  distinct(issn, oa_model) %>%
-  group_by(issn) %>%
-  filter(n() > 1)
-flipped_jns_ind <- els_jns %>%
-  filter(issn %in% flipped_jns$issn) %>%
-  group_by(issn) %>%
-  slice(which.max(year)) %>%
-  ungroup() %>%
-  count(oa_model)
-```
-
-We identified `r filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` OA journals that reverse flipped--transitioned from fully OA to hybrid [@Matthias_2019].
-We excluded these journals from the following article-level analysis; if we had not, OA articles published in an OA journal before it reverse-flipped would have been falsely attributed to the hybrid model.
-Likewise, we identified and excluded journals flipping from hybrid to fully OA in the same timeframe (n=`r filter(flipped_jns_ind, oa_model == "Open Access") %>% pull(n)`).
-Our final sample for the article-level analysis consisted of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` unique hybrid journals.
-
-### Article-level analysis{.unnumbered}
+### Article-Level Analysis{.unnumbered}
 
@@
 Between 2015 and 2019, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` (`r  hybrid_oa_volume %>% distinct(issn) %>% nrow() / (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n))*100`%) hybrid journals published at least one OA article.
-Together these journals published `r five_years_ind %>% pull(oa_n)` OA articles, representing `r five_years_ind %>% pull(oa_uptake)*100`% of their total article volume (n=`r five_years_ind %>% pull(articles_n)`).
-Table \ref{tab:overview} presents these findings for each year and the aggregated five-year period between 2015-19, illustrating a moderate growth in hybrid OA over time.
-The number of hybrid journals with at least one OA article increased each year, as did the number of OA articles, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` 10,672 in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019.
-However, since the total number of articles published in these journals also increased over time, the relative share of OA articles in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.
+Together these journals published `r five_years_ind %>% pull(oa_n)` OA articles, which represents `r five_years_ind %>% pull(oa_uptake)*100`% of their total article volume (n=`r five_years_ind %>% pull(articles_n)`).
+Table \ref{tab:overview} presents these findings for each year and the aggregated five-year period, illustrating a moderate growth in hybrid OA over time.
+With each year, the number of hybrid journals with at least one OA article increased, as did the number of OA articles, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019.
+However, since the total number of articles published in these journals also grew over time, the relative share of OA articles in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.
 
@@
   dplyr::select(-ind, -ind_group) %>%
-  kableExtra::kbl(booktabs = T, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
+  kableExtra::kbl(booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
                   caption = "Overview of hybrid open access uptake across Elsevier hybrid journals with at least one OA article 2015-2019.",
@@
   kable_styling() %>%
-  pack_rows("Elsevier Hybrid Journals with OA article", 1, 1) %>%
-  pack_rows("Articles", 2, 4) %>%
-  pack_rows("OA Articles", 5, 7) %>%
-  pack_rows("OA Uptake", 8, 10)
+  pack_rows("Elsevier Hybrid Journals with ≥ 1 OA article", 1, 1) %>%
+  pack_rows("Articles in Elsevier Hybrid Journals with ≥ 1 OA Article", 2, 4) %>%
+  pack_rows("OA Articles in Elsevier Hybrid Journals with ≥ 1 OA Article", 5, 7) %>%
+  pack_rows("OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA Article", 8, 10)
 ```
@@
 The box-and-whiskers plot visualizes the median hybrid OA uptake (horizontal line inside the box), the range of the majority of journals (the box), and individual outliers (dots).
-Notably, upper quartiles and whiskers stretch farther over the years, indicating that the range of uptake among journals with an above-average proportion of OA articles increased over the years.
-Nevertheless, the prevalence of hybrid OA is relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.
+Notably, the upper quartiles and whiskers stretch farther over the years, which indicates that the range of uptake among journals with an above-average proportion of OA articles increased over the years.
+Nevertheless, the prevalence of hybrid OA remained relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.
 
@@
 
-### Share of the total volume published in OA{.unnumbered}
+### Share of the Total Volume published in OA{.unnumbered}
 
@@
 
-During the same period, Elsevier provided OA to `r all_oa` articles, corresponding to a share of `r all_oa / all_els * 100` of the total article volume (n=`r all_els`).
-The OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind fully OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`, `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r length(unique(mirror_jn$container_title))` mirror journals that published `r sum(mirror_jn$all_articles)` articles, and Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) both in number and share of OA articles (see Figure \ref{fig:oa_volume_fig}-A).
+During the same period, Elsevier published `r all_oa` OA articles in total (i.e., in hybrid and fully OA journals), representing `r all_oa / all_els * 100` of the total article volume (n=`r all_els`).
+Looking at Figure \ref{fig:oa_volume_fig}, it is apparant that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind that of Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and fully OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`, `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.
 
 ```{r oa_volume_fig, fig.cap="Open Access in Elsevier journals in terms of article volume between 2015-2019.
-Figure A presents the share of open access (OA) articles relative to Elsevier's overall article output for each OA type from 2015-2019 and by open content license.
-Figure B presents the aggregated OA article volume of Elsevier journals by OA type."}
+Figure A presents the aggregated OA article volume of Elsevier journals from 2015-2019 by OA type.
+Figure B presents the share of open access (OA) articles relative to Elsevier's overall article output by OA type and open content license from 2015-2019."}
 # see analysis/006_license.Rmd how the figure was created
@@
 
-### License prevalence{.unnumbered}
+### License Prevalence{.unnumbered}
 
@@
 
-As far as we observed, Elsevier licensed OA articles in hybrid journals under two possible Creative Commons (CC) licenses—CC-BY or CC-BY-NC-ND.
+As far as we observed, OA articles in Elsevier hybrid journals were published under two possible CC licenses--CC-BY or CC-BY-NC-ND.
 CC-BY allows others to distribute, remix, adapt, and build upon the licensed work, including for commercial purposes, as long as the original author is credited.
@@
 In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC-BY-NC-ND license marginally increased and maintained a greater share than that of CC-BY (see Figure \ref{fig:oa_volume_fig}-B).
-Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid OA had the highest number and proportion of openly available articles licensed under CC-BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to fully OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`, `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program, Open Archive (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) .
-Most of these delayed OA articles were provided under an Elsevier user license (Els-User), prohibiting reuse for commercial purposes.
+Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC-BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to fully OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`, `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) .
+Most  delayed OA articles were provided under an Elsevier user license (Els-User), which prohibits reuse for commercial purposes.
 
-### Hybrid OA uptake across subjects{.unnumbered}
+### Hybrid OA Uptake across Subjects{.unnumbered}
 
@@
 Next, we present disciplinary variations in hybrid OA.
-Table \ref{tab:subject_area_table} presents high-level findings by AJSC subject area.
-In the five-years period studied, the Physical Science (`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and Health Sciences (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) were the most prevalent in terms of hybrid journals with at least one OA article published.
-In terms of article volume, most OA articles were published in hybrid journals belonging to the Life Sciences (`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles), while Social Sciences  (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles) journals published the lowest number of OA articles.
-However, it must be noted that there is a large overlap between the journals’ subject areas.
-Therefore, Table also assigns hybrid journals and OA articles fractionally to a subject area in addition to full counting when a journal belongs to more than one field.
-In the following, we apply full counting, meaning that journals and articles associated with more than one subject in the AJSC classification were counted once for each area or field.
+Table \ref{tab:subject_area_table} presents the high-level findings by AJSC subject area.
+In the five-year period we studied, the physical sciences (`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and health sciences (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) were the most prevalent in terms of hybrid journals with at least one OA article.
+In terms of article volume, most OA articles were published in hybrid journals belonging to the life sciences (`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles), while social sciences  (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles) journals published the lowest number of OA articles.
+However, it is important to note that there is a large overlap between the journals’ subject areas.
+Therefore, Table  \ref{tab:subject_area_table} also assigns hybrid journals and OA articles fractionally to a subject area and compares it to the full counting, where journals assigned more than one subject in the AJSC classification were counted once for each field.
+For the remaining results, we only present the full counting.
 
@@
   filter(subject_area != "Multidisciplinary", !is.na(subject_area)) %>% 
+  mutate(subject_area = stringr::str_to_sentence(subject_area)) %>%
   kbl(booktabs = TRUE, col.names = c("Subject area", "Journals", "OA articles", "Journals", "OA articles"),
       caption = "Subject area overview using full and fractional counting.
-Table reports the number of Elsevier hybrid journals with at least one OA article, as well as the number of OA articles published in them between 2015-2019.
-Not shown is one journal coded as Multidisciplinary that published 61 OA articles.
-Eleven journals with 37 OA articles could not be matched to the Scopus Source Title list used to obtain ASJC codes.") %>%
+Table reports the number of Elsevier hybrid journals with at least one OA article, and OA articles published from 2015-2019.
+One multidisciplinary with 61 OA articles and 11 journals with 37 OA articles that  could not be matched to the Scopus Source Title list used to obtain ASJC codes are not included.",
+      format.args = list(big.mark = ",")) %>%
   add_header_above(c(" " = 1, "Full counting" = 2, "Fractional counting" = 2)) %>%
@@
 
-Figure \ref{fig:oa_sub_uptake} presents the journals' five-year OA uptake grouped by subject area and field and in comparison to the overall average (Mdn= `r median_all * 100`; excluding one journal coded as Multidisciplinary).
+Figure \ref{fig:oa_sub_uptake} presents the journals' five-year OA uptake grouped by subject area and field and in comparison to the overall median (Mdn= `r median_all * 100`; excluding one journal coded as multidisciplinary).
 The box plot reveals some variation between the subject areas.
-Notably, with the exception of Environmental Science and Earth and Planetary Sciences, journals from the Physical Sciences show a lower median proportion of hybrid OA articles, whereas most Life Sciences and Social Sciences journals ranked above average.
-Figure X also shows large variations among the 26 subject fields, with the highest OA rates in Immunology and Microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), Environmental Science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and Neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), and the lowest in Chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), Materials Science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and Dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`).
+Notably, with the exception of environmental science and earth and planetary sciences, journals from the physical sciences show a lower median proportion of hybrid OA articles, whereas most life sciences and social sciences journals ranked above average.
+Figure X also shows large variations among the 26 subject fields, with the highest OA rates in immunology and microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), environmental science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), whilst  chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), materials science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`) recorded the lowest OA rate.
 
@@
 
-### Citation impact{.unnumbered}
+### Citation Impact{.unnumbered}
 
@@
 
-Furthermore, we were interested in the relationship between a journal's field-specific impact and its OA uptake.
-Spearman's rho correlation coefficient was used to assess the relationship between the 2019 Source Normalized Impact per Paper (SNIP) value calculated by Scopus and the journal's OA uptake that year.
-Considering only journals with at least one OA article, we found a weak positive correlation between journal impact and OA uptake ($r_s$=`r as.character(round(cor_test_res$estimate,4))`, $p<0.001$), suggesting that the relationship is not very strong, based on our data (see Figure \ref{fig:sniptest}).
+Furthermore, we were interested in the relationship between the journals' field-specific impact and their OA uptake.
+Spearman's rho correlation coefficient was used to assess the relationship between the 2019 Source Normalized Impact per Paper (SNIP) value calculated by Scopus and the journals' OA uptake that year.
+Considering only journals with at least one OA article, we found a weak but positive correlation between journal impact and OA uptake ($r_s$=`r as.character(round(cor_test_res$estimate,4))`, $p<0.001$), suggesting that the relationship is not very strong, based on our data (see Figure \ref{fig:sniptest}).
 
@@
 
-### Invoice types
+### Invoice Channels{.unnumbered}
 
@@
                                   `FundingBody` = "Agreement",
-                                  `ElsevierWaived` = "Fee Waived")) %>%
+                                  `ElsevierWaived` = "Fee waived")) %>%
   mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
   mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
-  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC-BY", "CC-BY-NC-ND")) %>%
   count(issued_year, fct_source, cc_norm)
@@
 
-Table \ref{tab:invoicing_overview} shows that authors (`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) were by far the most frequent invoicing type for hybrid OA, followed by funding bodies/institutional agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%).
-Interestingly, we also found a small number of cases where hybrid APCs were waived (`r filter(sponsored_art, fct_source == "Fee Waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee Waived") %>% pull(prop)`%).
+As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APC were most often invoiced to authors (n=`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) and to a lesser extent part of agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%).
+Interestingly, we also found a small number of cases where hybrid APCs were waived (n=`r filter(sponsored_art, fct_source == "Fee waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee waived") %>% pull(prop)`%).
 
@@
 
-Figure X illustrates that over the years, Elsevier has increasingly and, to a greater extent, invoiced hybrid OA APCs to authors directly rather than to funders or consortia ("Agreement").
-The share of fee-waived articles remained relatively stable at around 6%.
-However, we found different types of waivers.
-Around 52% of fee waivers were linked to an OA sponsor, indicating a third-party paid the APC, whereas 48% of Elsevier waived articles did not disclose an OA sponsor.
-For instance, 856 waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors.
-In terms of licensing, Figure X reveals that the type of invoicing was strongly related to the open content license variant.
-When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=32,086; 77%), whereas most articles with agreed payment were licensed under the more permissive CC BY license (n=18,331; 76%).
+```{r els_waived}
+els_waived_df <- hybrid_volume %>% 
+  filter(oa_sponsor_type == "ElsevierWaived") 
+```
+
+Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has increasingly invoiced hybrid OA APCs directly to authors compared to research funders or academic consortia ("Agreement").
+The share of fee-waived articles remained relatively stable, but we found different types of waivers.
+Around `r filter(els_waived_df, !is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)* 100`% of fee waivers were linked to an OA sponsor, which indicates that a third party paid the APC.
+For instance, `r filter(els_waived_df, oa_sponsor_name == "Académie des Sciences") %>% distinct(doi) %>% nrow()` waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors.
+The remaining `r filter(els_waived_df, is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)*100`% of waived articles did not disclose any OA sponsor.
+
+```{r invoiceoverview, fig.cap="Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoicing type.
+Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant.
+Grey bars show the total number of hybrid OA articles published in Elsevier journals from 2015-2019."}
+# plot
+sponsored_articles %>%
+  mutate(fct_source = stringr::str_to_title(fct_source)) %>%
+  ggplot(aes(x = gsub("20", "'", issued_year), y = n), alpha = 0.8) +
+  geom_bar(
+    data = all_articles,
+    color = "transparent",
+    stat = "identity",
+    fill = "#b3b3b3a0"
+  ) +
+  geom_bar(aes(fill = cc_norm), color = "white", stat = "identity", alpha = 0.8, size = 0.25,
+           position = position_stack(reverse = T)) +
+  facet_wrap( ~ fct_source, nrow = 1) +
+  scale_fill_manual(values = 
+                      c(`CC-BY` = "#B52141",
+                        `CC-BY-NC-ND` = "#0093c7"),
+                    name = "CC License Variant") +
+  labs(x = "Publication Year", y = "Hybrid OA Articles") + 
+  scale_y_continuous( 
+                     labels = scales::number_format(big.mark = " "),
+                      expand = expansion(mult = c(0, 0.05))) +
+  theme_minimal_hgrid() +
+  theme(legend.position="top", legend.justification = "right")
+```
+
+Moreover, Figure \ref{fig:invoiceoverview} relates the type of invoicing to the open content license variant.
+When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=`r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC-BY-NC-ND") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC-BY-NC-ND") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Author") %>% pull(n) %>% sum()* 100`%), whereas most articles billed as part of agreements were licensed under the more permissive CC BY license (n=`r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC-BY") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC-BY") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Agreement") %>% pull(n) %>% sum() * 100`%).
+
+### Invoicing variations across disciplines{.unnumbered}
+
+We also note large differences in OA invoicing among subject fields.
+Table X shows the number of OA articles by subject field by invoicing type.
+For articles in Nursing, Decision Sciences, and Pharmacology, Toxicology and Pharmaceutics, Elsevier predominantly invoiced authors, whereas most Energy and Chemical Engineering articles were invoiced through central agreements.
+Likewise, the majority of articles in Materials Science, Chemistry and Physics and Astronomy were not invoiced to authors but facilitated through central agreements or waived.
+The large share of waived APCs in Physics and Astronomy can be attributed to a single 2015 issue of Nuclear and Particle Physics Proceedings, which offered sponsored OA.
+
+```{r, results='asis'}
+hybrid_jn_volume <- hybrid_oa_volume %>%
+  group_by(issn) %>% 
+  summarise(n_oa = sum(oa))
+
+jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
+
+hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
+funder_shares <- hybrid_df %>%
+   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
+  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
+                                  `FundingBody` = "Agreement",
+                                  `ElsevierWaived` = "Fee Waived")) %>%
+  group_by(oa_sponsor_type, issn) %>%
+  summarise(n_oa_type = n_distinct(doi)) %>%
+  left_join(hybrid_jn_volume, by = "issn") %>%
+  mutate(prop = n_oa_type / n_oa) %>%
+  left_join(jn_subjects, by = "issn") %>%
+  ungroup() %>%
+  group_by(top_level, oa_sponsor_type) %>%
+  summarise(oa_type = sum(n_oa_type)) %>%
+  mutate(prop = oa_type / sum(oa_type))
+# overall props
+oa_sponsor_type_all <- hybrid_df %>% 
+  #exclude multidisciplinary
+  filter(issn != "2095-9273") %>%
+  count(oa_sponsor_type) %>% 
+  mutate(prop = n / sum(n))
+subject_all <- funder_shares %>%
+  group_by(top_level) %>%
+  summarise(oa = sum(oa_type))
+
+funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level") %>%
+  tidyr::pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
+  ungroup() %>% 
+  # remove na and multidisciplinary 
+  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
+  relocate(Agreement, .after = Author) %>%
+  arrange(desc(Author))
+
+make_pal <- scales::col_numeric(
+  palette  = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977"),
+  na.color = "#808080",
+  domain = NULL,
+  alpha = FALSE,
+  reverse = FALSE
+)
+
+funder_overview_subject %>%
+  mutate_at(3:6, function(x){
+    round(x*100,0)
+  }) %>%
+  mutate_at(3:6, function(x){
+    cell_spec(x, background = make_pal(x))}) %>%
+  kbl(booktabs = TRUE, "latex", format.args = list(big.mark = ","), digits = 1, escape = FALSE,
+                  caption = "Invoicing variations across disciplines 2015-2019.",
+                  position = "H",
+      col.names = linebreak(c("Subject", "OA articles", "Author", "Agreement", "Fee waived", "Other")),
+       linesep = "",
+      align=c("l|rrrrr")) %>%
+  add_header_above(c(" " = 2, "Invoicing channel" = 4)) %>%
+  kable_styling(latex_options = c("scale_down"))
+```

### 2021-01-26 analysis/manuscript.Rmd
full draft results

@@
 hybrid_oa_volume <- readr::read_csv(here::here("data", "hybrid_oa_volume.csv"))
+# invoicing analysis
+curated_invoicing_df <- readr::read_csv(here::here("data", "curated_invoicing_data.csv"))
+# open apc data
+o_apc <- readr::read_csv(here::here("data", "o_apc.csv"))
 ```
+
 ## Open Access Uptake in Hybrid Journals{.unnumbered}
@@
 
-```{r invoiceoverview, fig.cap="Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoicing type.
-Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant.
+```{r invoiceoverview, fig.cap="Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoicing channel Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant.
 Grey bars show the total number of hybrid OA articles published in Elsevier journals from 2015-2019."}
@@
   mutate(fct_source = stringr::str_to_title(fct_source)) %>%
+  mutate(fct_source = ordered(fct_source, levels = c("Author", "Agreement", "Fee Waived", "Other"))) %>%
   ggplot(aes(x = gsub("20", "'", issued_year), y = n), alpha = 0.8) +
@@
   scale_y_continuous( 
-                     labels = scales::number_format(big.mark = " "),
+                     labels = scales::number_format(big.mark = ","),
                       expand = expansion(mult = c(0, 0.05))) +
@@
 We also note large differences in OA invoicing among subject fields.
-Table X shows the number of OA articles by subject field by invoicing type.
+Table \ref{tab:invocing_by_subject} shows the number of OA articles by subject field by invoicing type.
 For articles in Nursing, Decision Sciences, and Pharmacology, Toxicology and Pharmaceutics, Elsevier predominantly invoiced authors, whereas most Energy and Chemical Engineering articles were invoiced through central agreements.
@@
 
-```{r, results='asis'}
+```{r invocing_by_subject, results='asis'}
 hybrid_jn_volume <- hybrid_oa_volume %>%
@@
                   position = "H",
-      col.names = linebreak(c("Subject", "OA articles", "Author", "Agreement", "Fee waived", "Other")),
+      col.names = c("Subject", "OA articles", "Author", "Agreement", "Fee waived", "Other"),
        linesep = "",
-      align=c("l|rrrrr")) %>%
-  add_header_above(c(" " = 2, "Invoicing channel" = 4)) %>%
+      align = "lrrrrr") %>%
+  add_header_above(c(" " = 2, "Invoicing channel (in %)" = 4)) %>%
   kable_styling(latex_options = c("scale_down"))
 ```
+### Breakdown by OA sponsor{.unnumbered}
+
+Table X presents the top ten of 62 sponsoring bodies in our sample.
+Together, these ten funders accounted for around 80% of centrally invoiced APCs.
+The table also highlights the number of distinct journals and the share of CC -BY licensed articles.
+Notably, UK-based research funders and the US-based Bill & Melinda Gates Foundation mainly sponsored articles published under a CC -BY license.
+In contrast, articles funded by VSNU, the European Research Council, and Norway Institutes had a much lower proportion of CC -BY.
+
+Figure shows the yearly distribution of central invoiced articles by year and institutional sector.
+While UK-based open access invoices were mainly addressed to discipline-specific governmental and non-profit research funders,  the situation is different in the Netherlands and Norway where national consortia of institutions belonging to the sector of higher education (e.g.
+universities) had transformative agreements with Elsevier.
+In 2019, Elsevier also launched similar agreements in other countries with lower publication output including Hungary and Poland.
+Besides, we found that invoiced funding bodies from the UK and the US mainly represent discipline-specific funders, while the remaining funding bodies from other countries are focused on a broad variety of disciplines (Table).
+
+### Invoicing variations across countries{.unnumbered}
+
+```{r}
+hybrid_invoice <- hybrid_volume %>%
+  filter(oa_sponsor_type == "FundingBody") %>%
+  left_join(curated_invoicing_df, by = c("oa_sponsor_name")) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC-BY", "CC-BY-NC-ND"))
+```
+
+```{r license_count}
+sponsor_df <- hybrid_invoice %>% 
+#  filter(!is.na(Country)) %>% 
+  # manual changes
+  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "European Union", Country)) %>%
+  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
+  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
+  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
+  mutate(country_factor = fct_infreq(country_factor)) %>%
+  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))
+```
+
+Elsevier’s data offer more insight into publishing fees invoiced as part of OA agreements.
+Overall, we identified 62 sponsoring institutions that received invoices for hybrid publishing fees as part of an agreement with Elsevier.
+We found that, by a large margin, most invoices were linked to UK-based research funders and institutions (n=`r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%), followed by the Netherlands (n=`r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%) and the European Union (n=`r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%).
+Figure \ref{fig:invoice_sponsor_country} provides a closer look at the geographical distribution by license prevalence, highlighting the dominance of CC -BY licensed articles invoiced to sponsoring institutions in the UK, the United States, and, to a much lesser extent, the Netherlands.
+In contrast, the majority of articles invoiced to sponsoring institutions from Norway or representing the EU were published under a non-commercial license.
+
+```{r invoice_sponsor_country, fig.cap="Number of centrally invoiced OA articles in Elsevier hybrid journals by country of recipient.
+Colored bars represent the Creative Commons (CC) license variant under which the article was published between 2015 and 2019."}
+sponsor_df %>%
+  group_by(country_factor, cc_norm) %>% 
+  summarise(n = n_distinct(doi)) %>%
+  ggplot(aes(fct_rev(country_factor), n, fill = cc_norm)) +
+  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
+            color = "white", alpha = 0.8, size = 0.25) +
+  coord_flip() +
+  scale_fill_manual(values = 
+                      c(`CC-BY` = "#B52141",
+                        `CC-BY-NC-ND` = "#0093c7"),
+                    name = "CC License Variant") +
+  labs(x = "OA Sponsor by Country", y = "Hybrid OA Articles (Agreement)") +
+  scale_y_continuous(labels = scales::number_format(big.mark = ","),
+                      expand = expansion(mult = c(0, 0.05)),
+                     breaks = seq(0,15000, 2000)) +
+  theme_minimal_vgrid() +
+  theme(legend.position="top", legend.justification = "right")
+```
+
+Figure \ref{fig:sponsor_sector} shows the yearly distribution of central invoiced articles by year and institutional sector.
+While UK-based open access invoices were mainly addressed to discipline-specific governmental and non-profit research funders,  the situation is different in the Netherlands and Norway where national consortia of institutions belonging to the sector of higher education (e.g.
+universities) had transformative agreements with Elsevier.
+In 2019, Elsevier also launched similar agreements in other countries with lower publication output including Hungary and Poland.
+Besides, we found that invoiced funding bodies from the UK and the US mainly represent discipline-specific funders, while the remaining funding bodies from other countries are focused on a broad variety of disciplines (Table).
+
+```{r sponsor_sector, fig.cap=' Development of centrally invoiced OA articles in hybrid journals by country. Colored bars represent the institutional sector OA funding bodies belong to.'}
+sponsor_df %>%
+  filter(!is.na(Sector)) %>%
+  count(issued_year, Sector, country_factor) %>%
+  mutate(Sector = stringr::str_to_title(Sector)) %>% 
+  mutate(sector_level = factor(Sector)) %>%
+  mutate(sector_level = fct_relevel(sector_level, "Government", "Higher Education", "Private Non-Profit", "Other")) %>%
+  ggplot(aes(gsub("20", "'", issued_year), n, fill = sector_level)) + 
+  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
+            color = "white", alpha = 0.8, size = 0.25) +
+  facet_wrap(~country_factor) +
+  scale_fill_manual("Sector", values = c("#65BADA", "#00A757", "#E5BA52", "grey50")) +
+  scale_y_continuous(labels = scales::number_format(big.mark = " "),
+                      expand = expansion(mult = c(0, 0.05))) +
+  theme_minimal_hgrid() +
+  theme(legend.position="top", legend.justification = "right",
+        panel.spacing = unit(2, "lines")) +
+  guides(fill = guide_legend(nrow = 2)) +
+  labs(x = "Publication Year", y = "Hybrid OA Articles (Agreement)")
+```
+
+```{r}
+sponsor_country <- hybrid_invoice %>%
+  filter(!is.na(`Primary research focus`)) %>%
+  # manual changes
+  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "European Union", Country)) %>%
+  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
+  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
+  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
+  mutate(country_factor = fct_infreq(country_factor)) %>%
+  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf)) %>%
+  mutate(`Primary research focus` = ifelse(`Primary research focus` == "SSH", "SSH", stringr::str_to_sentence(`Primary research focus`))) %>%
+  group_by(country_factor, `Primary research focus`) %>%
+  summarise(n = n_distinct(doi)) %>%
+  ungroup() %>%
+  mutate(prop = round(n / sum(n), 2) * 100)
+```
+
+```{r}
+sponsor_country %>%
+  ungroup() %>%
+  pivot_wider(id_cols = c("country_factor"), names_from = `Primary research focus`, values_from = prop) %>%
+  mutate_all(~replace(., is.na(.), 0)) %>%
+  janitor::adorn_totals(c("col", "row")) %>%
+  mutate_if(is.numeric, ~paste0(., "%")) %>%
+  rename(`Country/Subject area` = country_factor) %>%
+  kableExtra::kbl(booktabs = TRUE,
+                  caption = " Relative subject profile of centrally invoiced OA funding bodies.") %>%
+  kable_styling()
+```
+
+### Breakdown by OA sponsor{.unnumbered}
+
+```{r open_apc_check}
+hybrid_invoice <- hybrid_invoice %>%
+  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
+```
+
+Table \ref{tab:sponsor_top_10} presents the top ten of `r hybrid_invoice %>% filter(!is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` sponsoring bodies in our sample.
+Together, these ten funders accounted for around 80% of centrally invoiced APCs.
+The table also highlights the number of distinct journals and the share of CC -BY licensed articles.
+Notably, UK-based research funders and the US-based Bill & Melinda Gates Foundation mainly sponsored articles published under a CC BY license.
+In contrast, articles funded by VSNU, the European Research Council, and Norway Institutes had a much lower proportion of CC BY.
+
+```{r, sponsor_top_10}
+hybrid_invoice_tt <- hybrid_invoice %>%
+  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_cleaned), "Other", oa_sponsor_cleaned)) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC-BY", "CC-BY-NC-ND")) %>%
+  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
+  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
+hybrid_invoice_agg_general <- hybrid_invoice_tt %>%
+  group_by(oa_sponsor) %>%
+  summarise(journals = n_distinct(issn),
+            articles = n_distinct(doi)) %>%
+  mutate(prop = articles / sum(articles))
+hybrid_invoice_agg_cc <- hybrid_invoice_tt %>%
+  ungroup() %>%
+  filter(cc_norm == "CC-BY") %>%
+  group_by(oa_sponsor) %>%
+  summarise(cc_by = n_distinct(doi))
+hybrid_invoice_agg_oapc <- hybrid_invoice_tt %>%
+  filter(in_open_apc == TRUE) %>%
+  group_by(oa_sponsor) %>%
+  summarise(in_open_apc =  n_distinct(doi))
+ind_sponsor_df <-
+  left_join(hybrid_invoice_agg_general, hybrid_invoice_agg_cc, by = "oa_sponsor") %>%
+  left_join(hybrid_invoice_agg_oapc, by = "oa_sponsor") %>%
+  mutate(
+    cc_by = cc_by / articles,
+    in_open_apc = ifelse(is.na(in_open_apc), 0, in_open_apc),
+    in_open_apc = in_open_apc / articles
+  ) %>%
+  mutate(oa_sponsor = fct_reorder(oa_sponsor, articles)) %>%
+  mutate(oa_sponsor = forcats::fct_relevel(oa_sponsor, "Other", after = 0L)) %>% arrange(desc(oa_sponsor))
+# overall
+overall_ind_sponsor_df <-
+  tibble(
+    oa_sponsor = "All",
+    journals = length(unique(hybrid_invoice$issn)),
+    articles = length(unique(hybrid_invoice$doi)),
+    prop = 1,
+    cc_by = filter(hybrid_invoice, cc_norm == "CC-BY") %>% nrow() / length(unique(hybrid_invoice$doi)),
+    in_open_apc = filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi))
+  )
+
+ind_sponsor_df %>%
+  bind_rows(overall_ind_sponsor_df) %>%
+  mutate(oa_sponsor = gsub("&", "and", oa_sponsor)) %>%
+  mutate_at(4:6, function(x){
+    round(x*100,0)
+  }) %>%
+  mutate_at(5:6, function(x){
+    cell_spec(x, background = make_pal(x))}) %>%
+  kbl("latex", booktabs = TRUE, escape = FALSE,
+      col.names = c("Hybrid OA sponsors", "Journals", "Articles", "\\%", "CC BY", "OAPC"),
+      align = c("lrrrrr"),
+      linesep = "",
+      caption = "Top 10 Elsevier hybrid OA sponsors by centrally invoiced articles 2015-2019.
+The table furthermore presents the number of Elsevier hybrid journals with at least one invoiced article and two indicators of compliance: proportion of articles provided under a CC BY license and where spending was disclosed through the Open APC Initiative (OAPC).
+VSNU and Norway Institutes paid for OA through transformative agreements according to the ESAC Transformative Agreement Registry.") %>%
+  row_spec(11:12, hline_after = TRUE) %>%
+  add_header_above(c(" " = 4, "Compliance (in %)" = 2)) %>%
+   kable_styling(latex_options = c("scale_down"))
+```
+
+The table also highlights the proportion of APC funding publicly disclosed through the Open APC initiative, suggesting that funders with a large CC -BY share were also more willing to share their spending publicly.
+On the other hand, APC funding activities through transformative agreements were not publicly shared.
+
+Figure X shows the subject profile of OA funders relative to the journals’ subject fields.
+As expected, the figure shows that subject-specific funders tended to focus on a limited set of subject fields.
+In contrast, national consortia, such as VSNU and Norwegian Institutes (NO), financed OA articles  for a broader range of fields.
+However, within these consortia, the subject distribution also varied.
+VSNU OA spending, for example, had a strong focus on articles relating to Medicine, Environmental Science, and Energy, presumably reflecting disciplines with generally strong publication output at Dutch universities.
+Norwegian universities mostly published hybrid OA in Engineering, Environmental Science and Agricultural and Biological Sciences journals.
+
+```{r}
+#short funder labels
+funder_codes <- tribble(
+  ~oa_sponsor, ~oa_sponsor_short,
+  "Engineering and Physical Sciences Research Council", "EPSRC",
+  "VSNU", "VSNU",
+  "Wellcome Trust", "Wellcome",
+  "European Research Council", "ERC",
+  "Medical Research Council", "MRC",
+  "Natural Environment Research Council", "NERC",
+  "Biotechnology and Biological Sciences Research Council", "BBSRC",
+  "Bill & Melinda Gates Foundation", "Gates",
+  "Economic and Social Research Council", "ESRC",
+  "Norway Institutes", "NO",
+  "Other", "Other")
+```
+
+```{r field_funder_matrix, out.width="90%", fig.width=8, fig.height=12}
+field_funder <- hybrid_df %>%
+  filter(oa_sponsor_type == "FundingBody") %>%
+  left_join(funder_codes, by = c("oa_sponsor_name" = "oa_sponsor")) %>%
+  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_short), "Other", oa_sponsor_short)) %>%
+  mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
+  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
+  left_join(jn_subjects, by = "issn") %>%
+  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
+  group_by(oa_sponsor, top_level) %>%
+  summarise(n_oa_type = n_distinct(doi)) %>%
+  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
+  mutate(oa_sponsor = fct_relevel(oa_sponsor, "Other", after = Inf))
+# plot
+pp <- ggplot(field_funder,
+       aes(
+         x = factor(oa_sponsor),
+         y = reorder(top_level, desc(top_level)),
+         colour = prop,
+         size = n_oa_type
+       )) +
+  geom_point() +
+  scale_color_viridis_c(
+    labels = scales::percent_format(accuracy = 5L),
+    option = "magma",
+    name = "Percentage per Funder",
+    breaks =  scales::extended_breaks(),
+    begin = 0.15,
+    end = 0.85
+  ) +
+  scale_size("OA Articles") +
+  scale_x_discrete(position = "top") +
+  labs(x = NULL, y = NULL) +
+  theme_minimal_grid(font_size = 12) +
+  theme(
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    panel.background = element_blank()
+  ) +
+  theme(legend.position = "top", legend.justification = "right") +
+  theme(legend.direction = "horizontal", 
+        legend.box = "horizontal")
+ggsave(here::here("figure", "funder_subject.png"), pp, dpi = 300, width = 9.5, height = 9)
+```
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.95\textwidth]{../figure/funder_subject.png}
+
+}
+
+\caption{Top 10 Elsevier hybrid OA institutional invoice recipients in terms of articles published by journal subject.
+EPSRC, Engineering and Physical Sciences Research Council; VSNU, VSNU; Wellcome, Wellcome Trust; ERC, European Research Council; ERC, European Research Council; MRC, Medical Research Council; NERC, Natural Environment Research Council; BBSRC, Biotechnology and Biological Sciences Research Council; Gates, Bill and Melinda Gates Foundation; ESRC, Economic and Social Research Council; NO, Norway Institutes.
+}\label{fig:field_funder_matrix}
+\end{figure}

### 2021-01-26 data-raw/oapc_data.R

@@ -1,0 +1,16 @@
+# openapc https://github.com/OpenAPC/openapc-de/releases/tag/v4.8.4-3-0
+u <-
+  "https://raw.githubusercontent.com/OpenAPC/openapc-de/3d2f9ba66e6dab29856cb3702a9ce8ae4da0d09b/data/apc_de.csv"
+o_apc <- readr::read_csv(u)
+#'
+#' We also would like to add data from transformative aggrements, which is also
+#' collected by the Open APC Initiative.
+#' The transformative agreements data-set does not include pricing information.
+#'
+oa_trans <-
+  readr::read_csv(
+    "https://raw.githubusercontent.com/OpenAPC/openapc-de/3d2f9ba66e6dab29856cb3702a9ce8ae4da0d09b/data/transformative_agreements/transformative_agreements.csv"
+  )
+oapc <- oa_trans %>%
+  bind_rows(o_apc) 
+write_csv(oapc, here::here("data", "o_apc.csv"))

### 2021-01-27 analysis/006_license.Rmd
final results :-)

@@
            case_when(
-             grepl("/by/", license_url) ~ "CC-BY",
-             grepl("/by-nc", license_url) ~ "CC-BY-NC-ND",
+             grepl("/by/", license_url) ~ "CC BY",
+             grepl("/by-nc", license_url) ~ "CC BY-NC-ND",
              grepl("userlicense", license_url) ~ "Els-User")
@@
 cc_hybrid <- im_hybrid %>%
-  mutate(license_code = ifelse(grepl("/by/", URL), "CC-BY", "CC-BY-NC-ND")) %>%
+  mutate(license_code = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
   group_by(license_code, issued_year) %>%
@@
   scale_fill_manual(values = 
-                      c(`CC-BY` = "#B52141",
-                        `CC-BY-NC-ND` = "#0093c7",
+                      c(`CC BY` = "#B52141",
+                        `CC BY-NC-ND` = "#0093c7",
                         `Els-User` = "grey80")) +

### 2021-01-27 analysis/manuscript.Rmd
final results :-)

@@
     if (is.numeric(x)) {
-      return(formatC(x, big.mark = ",", format = "f", digits = 0))
+      return(formatC(x, big.mark = ",", format = "f", digits = 1, drop0trailing = TRUE))
     } else{
@@
 
-## Open Access Uptake in Hybrid Journals{.unnumbered}
+## Uptake of Hybrid Open Access{.unnumbered}
 
-### Journal-Level Analysis{.unnumbered}
+### Overview{.unnumbered}
 
@@
 ```
-
-Figure \ref{fig:number_of_els_journals} visualizes  Elsevier's OA portfolio throughout the five years from 2015-2019, and shows that hybrid journals consistently outweighed fully OA journals, with only little variation overall.
-In 2015, Elsevier published `r els_jns_count %>% filter(year == 2015, oa_model == "Open Access") %>% pull(n)` fully OA journals and `r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(n)` hybrid journals (`r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(prop)*100`%).
-Since then, hybrid journals grew in number by `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n)) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% pull(n))` journals (n=`r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` in 2020), resulting in a slightly decreased share of Elsevier's overall OA journal portfolio by 1% (`r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(prop), 2) * 100`%; `r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` out of `r els_jns_count %>% filter(year == 2020) %>% pull(n) %>% sum()` journals).
-Of these, `r filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` journals transitioned from fully OA to hybrid.
-For the article-level analysis, we excluded these reverse-flip journals, focussing on `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` unique hybrid journals without OA business model change for the remaining results.
-
-```{r number_of_els_journals, fig.cap = "Longitudinal development of Elsevier's journal portfolio by open access business model.
-The figure distinguishes between fully open access journals that provide immediate open access to all articles and the hybrid model.
-Data obtained from Elsevier APC pricing lists (Matthias 2020)."}
-ggplot(els_jns_count,
-       aes(as.character(year), n,  fill = oa_model_fct, group = oa_model_fct)) +
-  geom_area(alpha = 0.8,
-            colour = "white") +
-  scale_fill_manual(
-    values = c("#cccccca0", "#56b4e9"),
-    name = NULL,
-    labels = c("Fully OA", "Hybrid OA")
-  ) +
-  labs(x = "Year", y = "Number of Elsevier Journals") +
-  scale_y_continuous(
-    labels = scales::number_format(big.mark = ","),
-    expand = expansion(mult = c(0, 0.05)),
-    breaks =  scales::extended_breaks()
-  ) +
-  theme_minimal_hgrid() +
-  theme(legend.position = "top",
-        legend.justification = "right") +
-  guides(fill = guide_legend("Business Model"))
-```
-
-### Article-Level Analysis{.unnumbered}
-
 ```{r hybrid_oa_per_year}
@@
 Table \ref{tab:overview} presents these findings for each year and the aggregated five-year period, illustrating a moderate growth in hybrid OA over time.
-With each year, the number of hybrid journals with at least one OA article increased, as did the number of OA articles, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019.
+With each year, the number of hybrid journals with at least one OA article increased, as did the number of immediate OA articles in these journals, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019.
 However, since the total number of articles published in these journals also grew over time, the relative share of OA articles in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.
@@
   kableExtra::kbl(booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
-                  caption = "Overview of hybrid open access uptake across Elsevier hybrid journals with at least one OA article 2015-2019.",
+                  caption = "Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.",
                   position = "H"
@@
 
-During the same period, Elsevier published `r all_oa` OA articles in total (i.e., in hybrid and fully OA journals), representing `r all_oa / all_els * 100` of the total article volume (n=`r all_els`).
-Looking at Figure \ref{fig:oa_volume_fig}, it is apparant that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind that of Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and fully OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`, `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.
+During the same period, Elsevier published `r all_oa` OA articles in total (i.e., in hybrid and full OA journals), representing `r all_oa / all_els * 100`% of the total article volume (n=`r all_els`).
+Looking at Figure \ref{fig:oa_volume_fig}, it is apparent that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind that of Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and full OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`, `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.
 
-```{r oa_volume_fig, fig.cap="Open Access in Elsevier journals in terms of article volume between 2015-2019.
+```{r oa_volume_fig, fig.cap="Open Access (OA) in Elsevier journals in terms of article volume between 2015-2019.
 Figure A presents the aggregated OA article volume of Elsevier journals from 2015-2019 by OA type.
-Figure B presents the share of open access (OA) articles relative to Elsevier's overall article output by OA type and open content license from 2015-2019."}
+Figure B presents the share of OA articles relative to Elsevier's overall article output by OA type and open content license from 2015-2019."}
 # see analysis/006_license.Rmd how the figure was created
@@
   mutate(prop = n / sum(n)) %>%
-  filter(license_code == "CC-BY")
+  filter(license_code == "CC BY")
 ```
 
-As far as we observed, OA articles in Elsevier hybrid journals were published under two possible CC licenses--CC-BY or CC-BY-NC-ND.
-CC-BY allows others to distribute, remix, adapt, and build upon the licensed work, including for commercial purposes, as long as the original author is credited.
-In contrast, CC-BY-NC-ND is less permissive as it prohibits commercial reuse and derivatives.
-In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC-BY-NC-ND license marginally increased and maintained a greater share than that of CC-BY (see Figure \ref{fig:oa_volume_fig}-B).
-Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC-BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to fully OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`, `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) .
+As far as we observed, OA articles in Elsevier hybrid journals were published under two possible CC licenses--CC BY or CC BY-NC-ND.
+CC BY allows others to distribute, remix, adapt, and build upon the licensed work, including for commercial purposes, as long as the original author is credited.
+In contrast, CC BY-NC-ND is less permissive as it prohibits commercial reuse and derivatives.
+In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC BY-NC-ND license marginally increased and maintained a greater share than that of CC BY (see Figure \ref{fig:oa_volume_fig}-B).
+Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to full OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`, `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) .
 Most  delayed OA articles were provided under an Elsevier user license (Els-User), which prohibits reuse for commercial purposes.
 
-### Hybrid OA Uptake across Subjects{.unnumbered}
+### Subject Area and Field{.unnumbered}
 
@@
       caption = "Subject area overview using full and fractional counting.
-Table reports the number of Elsevier hybrid journals with at least one OA article, and OA articles published from 2015-2019.
+Table reports the number of Elsevier hybrid journals with at least one open access (OA) article, and OA articles published from 2015-2019.
 One multidisciplinary with 61 OA articles and 11 journals with 37 OA articles that  could not be matched to the Scopus Source Title list used to obtain ASJC codes are not included.",
@@
   mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
-  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC-BY", "CC-BY-NC-ND")) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
   count(issued_year, fct_source, cc_norm)
@@
 
-As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APC were most often invoiced to authors (n=`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) and to a lesser extent part of agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%).
+As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APC were most often invoiced to authors (n=`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) and to a lesser extent as part of agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%).
 Interestingly, we also found a small number of cases where hybrid APCs were waived (n=`r filter(sponsored_art, fct_source == "Fee waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee waived") %>% pull(prop)`%).
@@
 sponsored_art %>% 
-  kbl(booktabs = TRUE, col.names = c("Invoicing type", "Hybrid OA articles", "Percentage"),
-      caption = "Types of invoicing for hybrid OA in Elsevier journals between 2015-19", digits = 1,
+  kbl(booktabs = TRUE, col.names = c("Invoicing channel", "Hybrid OA articles (n)", "Percentage"),
+      caption = "Invoice channels for hybrid OA in Elsevier journals between 2015-19.", digits = 1,
       format.args = list(big.mark = ",")) %>%
-  kable_styling()
+  kable_styling() %>%
+  row_spec(4, hline_after = TRUE)
 ```
@@
 The remaining `r filter(els_waived_df, is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)*100`% of waived articles did not disclose any OA sponsor.
+Moreover, Figure \ref{fig:invoiceoverview} compares the invoicing channels based on open content license variants.
+When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=`r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Author") %>% pull(n) %>% sum()* 100`%), whereas most articles billed as part of agreements were licensed under the more permissive CC BY license (n=`r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Agreement") %>% pull(n) %>% sum() * 100`%).
 
-```{r invoiceoverview, fig.cap="Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoicing channel Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant.
+```{r invoiceoverview, fig.cap="Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoice channel Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant.
 Grey bars show the total number of hybrid OA articles published in Elsevier journals from 2015-2019."}
@@
   scale_fill_manual(values = 
-                      c(`CC-BY` = "#B52141",
-                        `CC-BY-NC-ND` = "#0093c7"),
+                      c(`CC BY` = "#B52141",
+                        `CC BY-NC-ND` = "#0093c7"),
                     name = "CC License Variant") +
@@
 
-Moreover, Figure \ref{fig:invoiceoverview} relates the type of invoicing to the open content license variant.
-When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=`r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC-BY-NC-ND") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC-BY-NC-ND") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Author") %>% pull(n) %>% sum()* 100`%), whereas most articles billed as part of agreements were licensed under the more permissive CC BY license (n=`r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC-BY") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC-BY") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Agreement") %>% pull(n) %>% sum() * 100`%).
-
-### Invoicing variations across disciplines{.unnumbered}
-
-We also note large differences in OA invoicing among subject fields.
-Table \ref{tab:invocing_by_subject} shows the number of OA articles by subject field by invoicing type.
-For articles in Nursing, Decision Sciences, and Pharmacology, Toxicology and Pharmaceutics, Elsevier predominantly invoiced authors, whereas most Energy and Chemical Engineering articles were invoiced through central agreements.
-Likewise, the majority of articles in Materials Science, Chemistry and Physics and Astronomy were not invoiced to authors but facilitated through central agreements or waived.
-The large share of waived APCs in Physics and Astronomy can be attributed to a single 2015 issue of Nuclear and Particle Physics Proceedings, which offered sponsored OA.
+We also observed large differences in OA invoicing among subject fields.
+Table \ref{tab:invocing_by_subject} shows the number of OA articles by subject field and invoice channel.
+For articles in nursing, decision dciences, and pharmacology, toxicology and pharmaceutics, Elsevier predominantly invoiced authors, whereas most energy and chemical engineering articles were invoiced through agreements.
+Likewise, the majority of articles in materials science, chemistry and physics and astronomy were not invoiced to authors but facilitated through agreements or waived.
+The large share of waived APCs in physics and astronomy can be attributed to a single 2015 issue of Nuclear and Particle Physics Proceedings, which offered sponsored OA.
 
@@
   kbl(booktabs = TRUE, "latex", format.args = list(big.mark = ","), digits = 1, escape = FALSE,
-                  caption = "Invoicing variations across disciplines 2015-2019.",
+                  caption = "Invoice channels by discipline 2015-2019.",
                   position = "H",
@@
       align = "lrrrrr") %>%
-  add_header_above(c(" " = 2, "Invoicing channel (in %)" = 4)) %>%
+  add_header_above(c(" " = 2, "Invoice channel (in %)" = 4)) %>%
   kable_styling(latex_options = c("scale_down"))
 ```
-### Breakdown by OA sponsor{.unnumbered}
 
-Table X presents the top ten of 62 sponsoring bodies in our sample.
-Together, these ten funders accounted for around 80% of centrally invoiced APCs.
-The table also highlights the number of distinct journals and the share of CC -BY licensed articles.
-Notably, UK-based research funders and the US-based Bill & Melinda Gates Foundation mainly sponsored articles published under a CC -BY license.
-In contrast, articles funded by VSNU, the European Research Council, and Norway Institutes had a much lower proportion of CC -BY.
-
-Figure shows the yearly distribution of central invoiced articles by year and institutional sector.
-While UK-based open access invoices were mainly addressed to discipline-specific governmental and non-profit research funders,  the situation is different in the Netherlands and Norway where national consortia of institutions belonging to the sector of higher education (e.g.
-universities) had transformative agreements with Elsevier.
-In 2019, Elsevier also launched similar agreements in other countries with lower publication output including Hungary and Poland.
-Besides, we found that invoiced funding bodies from the UK and the US mainly represent discipline-specific funders, while the remaining funding bodies from other countries are focused on a broad variety of disciplines (Table).
-
-### Invoicing variations across countries{.unnumbered}
+### Invoice Recipients{.unnumbered}
 
@@
   left_join(curated_invoicing_df, by = c("oa_sponsor_name")) %>%
-  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC-BY", "CC-BY-NC-ND"))
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND"))
 ```
@@
 
-Elsevier’s data offer more insight into publishing fees invoiced as part of OA agreements.
-Overall, we identified 62 sponsoring institutions that received invoices for hybrid publishing fees as part of an agreement with Elsevier.
-We found that, by a large margin, most invoices were linked to UK-based research funders and institutions (n=`r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%), followed by the Netherlands (n=`r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%) and the European Union (n=`r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%).
+Elsevier’s data offer more insight into invoicing.
+Overall, we identified 62 sponsoring institutions and organizations that received invoices for hybrid publishing fees as part of an agreement with Elsevier.
+We found that, by a large margin, most invoices were issued to UK-based research funders and institutions (n=`r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%), followed by the Netherlands (n=`r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%) and the European Union (n=`r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%; EU).
 Figure \ref{fig:invoice_sponsor_country} provides a closer look at the geographical distribution by license prevalence, highlighting the dominance of CC -BY licensed articles invoiced to sponsoring institutions in the UK, the United States, and, to a much lesser extent, the Netherlands.
@@
 In 2019, Elsevier also launched similar agreements in other countries with lower publication output including Hungary and Poland.
-Besides, we found that invoiced funding bodies from the UK and the US mainly represent discipline-specific funders, while the remaining funding bodies from other countries are focused on a broad variety of disciplines (Table).
+Besides, we found that invoiced funding bodies from the UK and the US mainly represent discipline-specific funders, while the remaining funding bodies from other countries are focused on a broad variety of disciplines (see Table \ref{tab:subject_profile}).
 
@@
 
-```{r}
+```{r subject_profile}
 sponsor_country %>%
@@
   kableExtra::kbl(booktabs = TRUE,
-                  caption = " Relative subject profile of centrally invoiced OA funding bodies.") %>%
-  kable_styling()
+                  caption = " Relative subject profile of centrally invoiced OA funding bodies.", align = c("l|rrrrr|r")) %>%
+    row_spec(6:7, hline_after = TRUE) %>%
+   kable_styling(latex_options = c("scale_down"))
 ```
 
-### Breakdown by OA sponsor{.unnumbered}
+### Institutional ranking{.unnumbered}
 
@@
 
-Table \ref{tab:sponsor_top_10} presents the top ten of `r hybrid_invoice %>% filter(!is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` sponsoring bodies in our sample.
-Together, these ten funders accounted for around 80% of centrally invoiced APCs.
-The table also highlights the number of distinct journals and the share of CC -BY licensed articles.
-Notably, UK-based research funders and the US-based Bill & Melinda Gates Foundation mainly sponsored articles published under a CC BY license.
+Table \ref{tab:sponsor_top_10} presents the top ten of `r hybrid_invoice %>% filter(!is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` invoiced institutions and organizations in our sample.
+Together, they accounted for around 80% of centrally invoiced APCs.
+The table also highlights the number of distinct journals and the share of CC BY licensed articles.
+Notably, UK-based research funders and the US-based Bill and Melinda Gates Foundation mainly sponsored articles published under a CC BY license.
 In contrast, articles funded by VSNU, the European Research Council, and Norway Institutes had a much lower proportion of CC BY.
@@
   mutate(oa_sponsor = ifelse(is.na(oa_sponsor_cleaned), "Other", oa_sponsor_cleaned)) %>%
-  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC-BY", "CC-BY-NC-ND")) %>%
+  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
   mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
@@
   ungroup() %>%
-  filter(cc_norm == "CC-BY") %>%
+  filter(cc_norm == "CC BY") %>%
   group_by(oa_sponsor) %>%
@@
     prop = 1,
-    cc_by = filter(hybrid_invoice, cc_norm == "CC-BY") %>% nrow() / length(unique(hybrid_invoice$doi)),
+    cc_by = filter(hybrid_invoice, cc_norm == "CC BY") %>% nrow() / length(unique(hybrid_invoice$doi)),
     in_open_apc = filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi))
@@
 
-The table also highlights the proportion of APC funding publicly disclosed through the Open APC initiative, suggesting that funders with a large CC -BY share were also more willing to share their spending publicly.
+The table also highlights the proportion of APC funding publicly disclosed through the Open APC initiative, suggesting that funders with a large CC BY share were also more willing to share their spending publicly.
 On the other hand, APC funding activities through transformative agreements were not publicly shared.
 
-Figure X shows the subject profile of OA funders relative to the journals’ subject fields.
+Figure \ref{fig:field_funder_matrix} shows the subject profile of OA funders relative to the journals’ subject fields.
 As expected, the figure shows that subject-specific funders tended to focus on a limited set of subject fields.

### 2021-01-27 analysis/manuscript.Rmd
tmp save

@@
 ```
+
+# Background
+
+The ability to identify and discover hybrid OA articles plays a fundamental role in monitoring hybrid OA, yet time and again, previous studies have underscored the difficulties involved in this (cf.
+[@Akbaritabar_2019; @van_Leeuwen_2018; @Laakso_2016; @Emery_2013; @Bj_rk_2012].
+Insights into the volume of hybrid OA and the various levels and sources of APC funding could create a better understanding of the costs involved in scholarly publishing and, thus, strengthen institutions' bargaining power in negotiations with publishers, and promote greater equity in scholarly publishing.
+In the following section, we examine how previous research has tackled the challenge of identifying hybrid OA content and what these studies taught us about the uptake and funding sources of hybrid OA.
+
+## Prevalence / uptake studies
+
+Since there is no standardized way to measure the uptake of hybrid OA, estimates can vary greatly depending on how uptake is operationalized-—for example, through the number of hybrid journals with at least one OA article, the average number of OA articles per hybrid journal, or the share of OA articles compared to subscription articles.
+
+In the first article-level study on hybrid OA, @Laakso_2016 reported that between 2007 and 2013, the five largest publishers--Elsevier, Sage, Springer, Taylor & Francis, and Wiley--recorded growing numbers of hybrid journals that coincided with a twenty-fold increase in hybrid OA articles (from 666 to 13,994).
+As a result, hybrid journals with at least one OA article more than doubled in number (1,082 in 2009, 2,714 in 2013) but fell by 31% relative to the number of hybrid journals in total.
+Although Laakso and Björk's (2016) presented a comprehensive overview tracking the growth of hybrid OA over time and across multiple publishers, their exploratory approach is unsuitable for repeated use as it required a considerable amount of manual data collection and cleaning.
+Minimizing such manual tasks, @Nelson_2017 narrowed their focus to one publisher--the American Chemical Society (ACS)--and requested a list of all hybrid OA articles published from 2006-2011 directly from the ACS.
+To assess the hybrid OA uptake, the authors aggregated and compared the number of OA (n=814) and subscription articles (n=27,621), finding that over the five years, 2.9% of ACS articles were published as hybrid OA.
+
+Contrary to the previous two studies, @Kirkman_2018 took a research funder as a starting point, reporting a 26% share of hybrid OA for articles funded by the Australian National Health and Medical Research Council from 2012-2014 (816 of 3,190).
+Kirkman collected research articles using the Funding Agency and Funding Text search fields on Web of Science, and classified articles as OA when they were freely available on the publisher website, and further distinguished hybrid OA articles based on journal-level information from the publisher website, such as hybrid OA journal and APC lists.
+Another approach is presented by @P_l_nen_2020, who used current research information system (CRIS) data—institutional publication data—to study the extent of OA among Finnish research publications from 2016-2017 (n=34,507 research articles).
+As part of the analysis, the authors identified hybrid OA articles through a dedicated OA status metadata field that has been mandatory to report in Finland since 2016 and found a 7% share of hybrid OA.
+However, the drawbacks of Kirkman's (2018) and Pölönen et al.'s (2020) approaches are that such data sources might not be openly available or, regarding CRIS data, might not be available at all since not all countries maintain comprehensive institutional publication data.
+
+Other studies assessed the share of (hybrid) OA for the entire corpus of scholarly research articles.
+For instance, using research articles with an assigned DOI and indexed in the Web of Science Core Collection (n=2,610,305) as a benchmark, @Mart_n_Mart_n_2018 assessed how many research articles from 2009 and 2014 are freely available through Google Scholar.
+After identifying OA articles through licensing information from Crossref, the authors estimated the share of hybrid OA at around 0.5% in 2009 and 1.5% in 2014.
+In a similar vein, @Piwowar_2018 drew on Crossref to identify content licenses.
+The authors followed a data-driven approach based on Unpaywall data, which is openly available and contained 66 million Crossref DOIs at the time, to determine how much of the scholarly literature is OA.
+Based on a random sample of 100,000 Crossref DOIs, the authors found 3.6% of hybrid OA, while pointing towards a growing trend in recent years that recorded a hybrid OA share of 9.4% of all articles published in 2015.
+While Piwowar et al.'s (2018) and Martín-Martín et al.'s (2018) methods are well-suited for large-scale and repeated analyses of the prevalence of hybrid OA, their main limitation is that they lack information about publishing costs.
+
+Moreover, Unpaywall's hybrid OA detection method has been refined several times since its launch in 2017, reflecting growing albeit incomprehensive license information in Crossref metadata.
+To overcome these shortcomings, the developers extended previous web-scraping approaches to better distinguish between immediate and delayed OA based on licenses at the date of observation @Piwowar_2019.
+Relatedly, in October 2020, Elsevier's license for delayed OA content was added to Unpaywall's list of license patterns [@Unpaywall_oa_license].
+Finally, at the journal-level, a more fine-grained identification of full OA journals using OA article information was developed to better distinguish them from subscription journals including hybrid journals [@Unpaywall_types].
+Because of these ongoing changes due to non-standardized data, uptake studies reporting hybrid OA uptake based on Unpaywall data cannot easily be compared with each other.
+
+## Financial studies
+
+Some countries like Sweden and the United Kingdom (UK) have negotiated national publishing agreements that allow authors to publish hybrid OA at a discounted fee or for free.
+These agreements can serve as the basis for studies on the uptake of hybrid OA or cost-benefit analyses (cf.
+@Olsson1271866; @Lawson_2018) but are often limited in scope as they tend to focus on a single country and rely on data that seldom is publicly available.
+Hence, while being valuable for specific use-cases, such studies can neither provide a comprehensive understanding of how the scholarly publishing system functions beyond national boundaries nor facilitate collective sharing of financial information about hybrid OA.
+
+Pinfield et al.'s (2016) analysis of APC payment records provided by 23 UK higher education institutions revealed a sharp increase in central payments from 2007-2013, which was largely attributed to the introduction of block grants by Research Councils UK (RCUK) and non-compliance sanctions by the Wellcome Trust.
+Moreover, the study showed that OA fees were paid almost exclusively through RCUK and Wellcome block grants (92%), and only a small number of APCs were paid through internal funding (7%).
+In contrast, a recent Springer Nature survey found that authors draw on a range of funding sources to cover OA fees, such as dedicated institutional OA funds, block grants, OA agreements, or research grants [@Monaghan_2020].
+Most hybrid OA authors were supported through dedicated institutional OA funds (43%, excluding block grants) and OA agreements with Springer Nature (34%).
+The differences between these two studies may be due to changes in policy and funding arrangements—in particular, the introduction of OA agreements—since @Pinfield_2016, or due to Monaghan et al.'s (2020) more regionally diverse sample, indicating regional differences in APC funding.
+
+Regional and policy differences in APC payments also came to light in a study by @Jahn_2016 that analyzed APC records from 30 German higher education and research institutions crowd-sourced by the Open APC initiative, the Austrian Science Fund (FWF), Jisc, and the Wellcome Trust.
+In particular, the study revealed large differences in the amount of hybrid OA funded from 2014-2015.
+Whilst hybrid OA accounted for less than 1% of APCs paid by German institutions (23 of 3,846), the three non-German research funders recorded a hybrid share of75% (11,533 of 15,779).
+According to @Jahn_2016, this could point towards differences in science policy, such as hybrid OA being supported by the three non-German research funders but not by Germany's largest national funder (DFG).
+Another possibility is that German hybrid OA fees were paid from budgets not reported to the Open APC initiative, from where the authors acquired the German APC data.
+Among these unreported funds are research grants and research unit budgets, which author surveys identified  as funding sources for hybrid OA fees [@Monaghan_2020; @Graaf_2017].
+As such, Jahn and Tullney's (2016) findings could reflect the complexities and potential limitations of institutional OA spending data that Pinfield et al.
+(2016) and Monaghan et al.
+(2020) attributed to incomplete or missing records.
+
+As this brief review has shown, the most comprehensive studies approach hybrid OA funding from an institutional perspective that focuses on research institutions or funders and draws on internal records or crowd-sourced data like from the Open APC initiative.
+The advantages of Open APC data are that they reflect the actual APCs paid rather than list prices; they are not limited to certain publishers or journals; they are provided in a standardized format regardless of the publisher.
+On the other hand, the drawbacks include the limited number of participating institutions that are almost exclusively based in Europe and the dependency on participating organizations for data integrity.
+Moreover, since Open APC data does not include waivers or free APCs, it cannot present a complete picture of all publishing activities and expenses.
+This, however, would pose less of an issue when analyzing just hybrid OA.
+
+## Data-driven workflows for hybrid OA
+
+In September 2018, coalition S presented a set of principles known as Plan S and announced.
+that from 2021 onwards, signatories would no longer fund publications in hybrid OA journals.
+This principle includes so-called "mirror" and sister journals—hybrid journals with fully OA counterparts or split into their fully OA and subscription parts—that have launched since Plan S was announced but makes an exception for hybrid journals with transformative agreements [@Frantsv_g_2019].
+While the scale and means of these transformative agreements may differ, many address the lack of transparency surrounding hybrid OA by making publisher-provided metadata a defining element of publishing workflows between institutions and publishers [@Geschuhn_2017; @Marques_2019]
+
+Elsevier, one of the world's largest scholarly publishers, provides article-level metadata to Crossref, including license information, which can be used to identify OA articles, and links to HTML and XML full-text versions for text and data mining purposes.
+Surprisingly, besides the full-text article, the XML files also contain invoicing information and the article's OA status.
+Table \ref{tab:els_xml} summarizes the key metadata elements, which comprise the type of entity to which the APC was invoiced (`openaccessSponsorType`) and the name of that entity if available (`openaccessSponsorName`).
+Moreover, the article-level metadata indicates if OA was provided through Elsevier's Open Archive program (`openArchiveArticle`), which enables free access to participating subscription journals after an embargo period (i.e., delayed OA; @Els_Archive).
+As we have been unable to identify similar efforts by other large publishers (e.g., Springer Nature), this increased transparency through openly available bibliometric and financial data at the article level seems unique to Elsevier.
+
+\begin{table}[H]
+\caption{\label{tab:els_xml}Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.}
+\centering
+\begin{tabular}[t]{lp{10cm}}
+\toprule
+XML Node & Description\\
+\midrule
+\texttt{openaccessArticle} & Indicates if the article is open
+access. \\
+\texttt{openaccessSponsorType} & Invoice recipient type.\\
+\texttt{openaccessSponsorName} & Entity that received Elsevier's APC
+invoice.\\
+\texttt{openArchiveArticle} & Indicates if open access was provided through
+Elsevier's Open Archive programme.\tabularnewline \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Table 1: Overview of metadata provided in XML-based Elsevier full-texts representing information about open access status and invoicing.
+A sample XML is available here: https://api.elsevier.com/content/article/PII:S0169409X18301479?httpAccept=text/xml
+
+In this study, we utilize Elsevier's publicly available metadata to analyze the uptake and invoicing of their hybrid OA publishing option.
+In particular, we use this novel methodology to answer the following questions:
+
+- What was the uptake of Elsevier's hybrid OA publishing option between 2015 and 2019?
+
+- Through which channels were hybrid APCs invoiced, and who were the recipients?
+
+# Methodology
+
+For this study, we collected data relating to Elsevier's hybrid OA option by drawing on multiple freely available data sources.
+We identified Elsevier hybrid journals through Elsevier’s APC list and supplemented our sample with Crossref metadata and text-mined invoicing data to investigate the financing of immediate OA articles provided under a Creative Commons license in subscription-based journals.
+Figure \ref{fig:workflow} visualizes the automated workflow we used to collect data from Elsevier and Crossref.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.95\textwidth]{../figure/els_flow.pdf}
+
+}
+
+\caption{Data Analytics Workflow to obtain article-level OA and invoicing data using freely available data sources provided by Elsevier and Crossref.
+}\label{fig:workflow}
+\end{figure}
+
+First, we obtained hybrid journals through an Elsevier APC price list from May 2020 provided as captured by Matthias (2020).
+We only included hybrid journals that were published under the hybrid model throughout the entire period of our analysis.
+Hence, we excluded journals that flipped from OA to hybrid and from hybrid to OA.
+This approach allowed us to ensure that the article-level differences we observed were not due changes at the journal-level (i.e., increases or decreases in number).
+Overall, we identified 1,970 unique hybrid journals.
+
+Next, we used an openly available Crossref database snapshot [@Crossref_2020], which contains all Crossref records registered until March 2020, to calculate the journals' combined article volume for the five-year period from 2015-2019.
+We only included articles published in regular issues aside from supplements containing conference contributions like meeting abstracts, indicated by non-numeric pagination.
+We excluded non-scholarly journal content, such as the table of contents and lists of reviewers, following Unpaywall's paratext recognition approach [@Unpaywall_para], which we expanded to include patterns indicating corrections.
+Furthermore, we categorized the articles by subject according to the All Science Journal Classification code (ASJC) and added citation indicators from the Scopus Source Title list (version June 2020) based on the journal they were published in.
+
+Crossref metadata also contains license information that indicates if and under which terms the publisher has made articles openly available on their website [@Hendricks_2020].
+We identified OA articles through Creative Commons (CC) licenses and then downloaded the XML version of all CC-licensed articles published in a hybrid journal.
+From the XML files, we determined whether or not the articles were immediate OA using the XML node `openArchiveArticle` and measured the uptake of hybrid OA through three bibliometric indicators [@Nelson_2017; @Laakso_2016]-—the number of hybrid journals with at least one OA article, the number of hybrid OA articles, and the number of hybrid OA articles relative to the number of closed-accessed articles in hybrid journals with at least one OA article.
+
+Moreover, we obtained APC invoicing metadata from the XML (see Table 1), which enabled us to identify the invoice channels and recipients of hybrid OA APCs.
+Based on the `openaccessSponsorType` node in the XML files, we distinguish between four invoice channels, including invoices billed to the corresponding author, as part of publishing agreements with funding bodies (hereinafter "agreements"; cf.
+@Els_Agreements), fee waivers (e.g., in "cases of genuine need" or due to society or university sponsorships, cf.
+ref), and other types not specified by Elsevier.
+Finally, when hybrid APCs were invoiced as part of OA agreements, we identified invoice recipients through the `openaccessSponsorName` node.
+
+We manually classified invoice recipients based on their institutional sectors, countries, and primary research areas.
+Following the OECD's Frascati Manual [@Frascati, p.
+91], we coded for four sectors, business enterprise, government, higher education, and private non-profit.
+The government sector consists of federal, state, and municipal governments, plus non-profit institutions controlled by government units [@Frascati, p.
+100].
+The higher education sector comprises institutions that perform research and development and provide formal tertiary education services, such as universities or research institutes [@Frascati, p.
+101].
+Legal or social entities that produce goods and services without generating profit, such as learned societies or charities, constitute the private non-profit sector [@Frascati, p.
+288].
+Due to the low article volume, we combined invoice recipients Elsevier listed as "authors" and "third-party sponsor", and those we had coded as business enterprises [@Frascati, p.
+97], into "Others".
+Moreover, we categorized invoice recipients according to the country or countries representing their scope of funding and based on the following primary research areas health sciences, life sciences, physical sciences and mathematics, social sciences and humanities, unknown, and broad (i.e., multiple research areas).
+
+It is important to note that we could not find any official comprehensive documentation of Elsevier's XML schema , at the time of writing.
+However, our conceptualization of the variables and interpretation of the data was informed by a comprehensive review of Elsevier's websites about OA, including a list of funding agreements and details about Elsevier's Open Archive program,and Elsevier’s article submission process in the context of dedicated OA agreements [@lund; @vsnu; @Els_videos].
+For validation, we matched the invoicing data provided by Elsevier to data from  the Open APC initiative.
+
+Throughout the mostly automated data gathering and analysis process, we used tools from the Tidyverse [@tidyverse] for the R programming language [@r].
+To allow for efficient data manipulation and retrieval, we imported the Crossref dump to Google BigQuery, applying the rcrossref [@rcrossref] parsers to extract relevant metadata fields.
+We interfaced Google BigQuery with SQL with the bigrquery client [@bigrquery].
+We used crminer [@crminer] to obtain the XML files from Elsevier and the xml2 package [@xml2] to parse relevant metadata.
+To ensure our results are reproducible, we provided all code for obtaining the data and performing the analyses with detailed documentation as RMarkdown notebooks on GitHub (<https://github.com/njahn82/elsevier_hybrid_invoicing>).
 
@@
 During the same period, Elsevier published `r all_oa` OA articles in total (i.e., in hybrid and full OA journals), representing `r all_oa / all_els * 100`% of the total article volume (n=`r all_els`).
-Looking at Figure \ref{fig:oa_volume_fig}, it is apparent that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind that of Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and full OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`, `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.
+Looking at Figure \ref{fig:oa_volume_fig}, it is apparent that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind that of Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and full OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`; `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.
 
@@
 In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC BY-NC-ND license marginally increased and maintained a greater share than that of CC BY (see Figure \ref{fig:oa_volume_fig}-B).
-Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to full OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`, `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) .
+Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`; `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to full OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`; `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`; `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) .
 Most  delayed OA articles were provided under an Elsevier user license (Els-User), which prohibits reuse for commercial purposes.
@@
 Figure \ref{fig:sponsor_sector} shows the yearly distribution of central invoiced articles by year and institutional sector.
-While UK-based open access invoices were mainly addressed to discipline-specific governmental and non-profit research funders,  the situation is different in the Netherlands and Norway where national consortia of institutions belonging to the sector of higher education (e.g.
-universities) had transformative agreements with Elsevier.
-In 2019, Elsevier also launched similar agreements in other countries with lower publication output including Hungary and Poland.
-Besides, we found that invoiced funding bodies from the UK and the US mainly represent discipline-specific funders, while the remaining funding bodies from other countries are focused on a broad variety of disciplines (see Table \ref{tab:subject_profile}).
+While UK-based OA invoices were mainly addressed to discipline-specific governmental and non-profit research funders, invoices to the Netherlands and Norway were issued to national academic consortia representing the higher education sector.
+In 2019, Elsevier also launched similar agreements in countries with lower publication output including Hungary and Poland.
+Besides, we found that invoiced funding bodies from the UK and the US mainly represented discipline-specific funders, while the invoiced funding bodies from other countries were focused on a broad variety of disciplines (see Table \ref{tab:subject_profile}).
 
@@
 
-Table \ref{tab:sponsor_top_10} presents the top ten of `r hybrid_invoice %>% filter(!is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` invoiced institutions and organizations in our sample.
+Table \ref{tab:sponsor_top_10} presents the top ten of `r hybrid_invoice %>% filter(!is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` sponsoring bodies in our sample.
 Together, they accounted for around 80% of centrally invoiced APCs.
@@
       linesep = "",
-      caption = "Top 10 Elsevier hybrid OA sponsors by centrally invoiced articles 2015-2019.
-The table furthermore presents the number of Elsevier hybrid journals with at least one invoiced article and two indicators of compliance: proportion of articles provided under a CC BY license and where spending was disclosed through the Open APC Initiative (OAPC).
-VSNU and Norway Institutes paid for OA through transformative agreements according to the ESAC Transformative Agreement Registry.") %>%
+      caption = " Top 10 Elsevier hybrid OA sponsors by centrally invoiced articles 2015-2019.
+The table furthermore presents the number of Elsevier hybrid journals with at least one invoiced article and two indicators of compliance: proportion of  CC BY licensed articles  and disclosed spending through the Open APC Initiative (OAPC.") %>%
   row_spec(11:12, hline_after = TRUE) %>%
@@
 
-The table also highlights the proportion of APC funding publicly disclosed through the Open APC initiative, suggesting that funders with a large CC BY share were also more willing to share their spending publicly.
-On the other hand, APC funding activities through transformative agreements were not publicly shared.
+The table also highlights the proportion of APC funding publicly disclosed through the Open APC initiative, showing higher disclosure rates for  funders with a large CC BY share.
+On the other hand, APCs that were invoiced as part of transformative agreements were not publicly shared.
 
-Figure \ref{fig:field_funder_matrix} shows the subject profile of OA funders relative to the journals’ subject fields.
-As expected, the figure shows that subject-specific funders tended to focus on a limited set of subject fields.
+Figure \ref{fig:field_funder_matrix} shows the subject profile of OA sponsors based on to the journals’ subject fields of centrally invoiced articles.
+As expected, the figure shows that subject-specific funders, such as the British Engineering and Physical Sciences Research Council (EPSRC), tended to focus on a limited set of subject fields.
 In contrast, national consortia, such as VSNU and Norwegian Institutes (NO), financed OA articles  for a broader range of fields.
 However, within these consortia, the subject distribution also varied.
-VSNU OA spending, for example, had a strong focus on articles relating to Medicine, Environmental Science, and Energy, presumably reflecting disciplines with generally strong publication output at Dutch universities.
-Norwegian universities mostly published hybrid OA in Engineering, Environmental Science and Agricultural and Biological Sciences journals.
+VSNU OA spending, for example, had a strong focus on articles relating to medicine, environmental science, and energy, presumably reflecting disciplines with a generally strong publication output at Dutch universities.
+Norwegian universities mostly published hybrid OA in engineering, environmental science and agricultural and biological sciences journals.
 

### 2021-01-28 analysis/manuscript.Rmd
introduction

@@
 ```
+# Introduction
+
+The rise of open access (OA) to journal articles has added complexity to scholarly publishing, particularly concerning transparency of economic dimensions.
+Financial transparency in journal publishing has long been lacking since Big Deal subscription contracts between academic institutions and large academic publishers usually include confidentiality clauses that prohibit publicly sharing information about the scope and pricing of such agreements [@Frazier_2001; @Bergstrom_2014; @Larivi_re_2015].
+Further, the distributed process of paying for accessing and publishing scholarly articles, which can involve several actors at the same time, including authors, research organizations, libraries, and research funders, adds to the opacity of the financial flows of scholarly publishing [@Lawson_2016].
+
+The problem with this lack of transparency becomes more apparent in the case of hybrid OA.
+In this OA business model, individual articles can be made openly available for a fee, also known as article-processing charge (APC), while the journal as a whole remains subscription access.
+Hence, hybrid OA was originally introduced as a low-risk transitional model that allows journals to gradually convert to full OA and reduce subscription costs as the OA uptake increases [@Prosser_2003].
+Many large subscription publishers have since incorporated this OA model, and while the uptake among authors increased only slowly during the first decade [@Bj_rk_2012], it has gradually gained popularity in recent years.
+This development was driven by the implementation of OA policies by research institutions and funders, the allocation of OA funds, and formal agreements between publishers and institutions that allow affiliated authors to publish OA free-of-charge.
+However, there have been concerns that the hybrid model allows for double-dipping—receiving two payments for one article, the APC and subscription fees [@Shieber_2009; @Prosser_2015].
+While publishers assure they combine subscription fees and APCs to adjust pricing [@Mittermaier_2015], without transparency around the uptake of hybrid OA and revenue streams, such claims are impossible to evaluate.
+
+With the recent introduction of transformative agreements, an evolving term for contracts that shift library spendings for subscriptions to OA [@Borrego_2020; @Hinchliffe_2019], the demand for publisher-provided data has even increased.
+If publishers would enable transparency about the OA uptake in hybrid journals and APC invoicing, academic libraries could improve their assessments and adjustments of contracts with publishers [@antelmann_2017].
+Likewise, research organizations and funders could improve compliance monitoring with OA mandates, while ensuring that the different actors were not charged multiple times [@Larivi_re_2018].
+However, as formal bibliometrics and financial studies have noted, the lack of publicly available and standardized data about hybrid OA publishing is considerable [@Laakso_2016; @Lawson_2015; @Pinfield_2016]
+
+This current study investigates the extent to which publisher-provided data can be utilized to assess hybrid OA.
+The study takes Elsevier, the largest scholarly journal publisher in terms of article volume [@Larivi_re_2015], as the starting point for an empirical analysis of hybrid OA uptake and invoicing.
+Our approach is based on openly available data sources and brings together scholarly data from Crossref and machine-readable APC invoicing data from Elsevier OA full-texts, which enables us to analyze the development of Elsevier’s hybrid journal portfolio from 2015-19, and to compare it with that of Elsevier’s full and delayed OA program.
+Further, we investigate whether hybrid publishing fees were waived or charged to the authors or academic institutions or research funders with Elsevier OA agreements.
+In particular, we identify academic consortia and research funders that were directly invoiced by Elsevier, and compare their hybrid OA rates by license and subject.
+As such, our findings have implications for research and OA policy and highlight the potential of publisher-provided data for large-scale, instantaneous analysis of hybrid OA uptake and invoicing.
 
@@
 
-Since there is no standardized way to measure the uptake of hybrid OA, estimates can vary greatly depending on how uptake is operationalized-—for example, through the number of hybrid journals with at least one OA article, the average number of OA articles per hybrid journal, or the share of OA articles compared to subscription articles.
+Since there is no standardized way to measure the uptake of hybrid OA, estimates can vary greatly depending on how uptake is operationalized--for example, through the number of hybrid journals with at least one OA article, the average number of OA articles per hybrid journal, or the share of OA articles compared to subscription articles.
 

### 2021-01-28 analysis/manuscript.tex
introduction

\documentclass[a4paper,man,floatsintext,longtable,noextraspace,12pt]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{hyperref}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

% bibliography
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\renewcommand{\baselinestretch}{1.5}
\setlength\parindent{24pt}

\title{Your APA6-Style Manuscript}
\shorttitle{}
\author{You}
\affiliation{Somewhere}

\abstract{Your abstract here.}

\begin{document}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The rise of open access (OA) to journal articles has added complexity to
scholarly publishing, particularly concerning transparency of economic
dimensions. Financial transparency in journal publishing has long been
lacking since Big Deal subscription contracts between academic
institutions and large academic publishers usually include
confidentiality clauses that prohibit publicly sharing information about
the scope and pricing of such agreements (Bergstrom et al., 2014;
Frazier, 2001; Larivière et al., 2015). Further, the distributed process
of paying for accessing and publishing scholarly articles, which can
involve several actors at the same time, including authors, research
organizations, libraries, and research funders, adds to the opacity of
the financial flows of scholarly publishing (Lawson et al., 2016).

The problem with this lack of transparency becomes more apparent in the
case of hybrid OA. In this OA business model, individual articles can be
made openly available for a fee, also known as article-processing charge
(APC), while the journal as a whole remains subscription access. Hence,
hybrid OA was originally introduced as a low-risk transitional model
that allows journals to gradually convert to full OA and reduce
subscription costs as the OA uptake increases (Prosser, 2003). Many
large subscription publishers have since incorporated this OA model, and
while the uptake among authors increased only slowly during the first
decade (Björk, 2012), it has gradually gained popularity in recent
years. This development was driven by the implementation of OA policies
by research institutions and funders, the allocation of OA funds, and
formal agreements between publishers and institutions that allow
affiliated authors to publish OA free-of-charge. However, there have
been concerns that the hybrid model allows for
double-dipping---receiving two payments for one article, the APC and
subscription fees (Prosser, 2015; Shieber, 2009). While publishers
assure they combine subscription fees and APCs to adjust pricing
(Mittermaier, 2015), without transparency around the uptake of hybrid OA
and revenue streams, such claims are impossible to evaluate.

With the recent introduction of transformative agreements, an evolving
term for contracts to shift library spendings for subscriptions to OA
(Borrego et al., 2020; Hinchliffe, 2019), the demand for
publisher-provided data has even increased. If publishers would enable
transparency about the OA uptake in hybrid journals and APC invoicing,
academic libraries could improve their assessments and adjustments of
contracts with publishers (Antelman, 2017). Likewise, research
organizations and funders could improve compliance monitoring with OA
mandates, while ensuring that the different actors were not charged
multiple times (Larivière \& Sugimoto, 2018). However, as formal
bibliometrics and financial studies have noted, the lack of publicly
available and standardized data about hybrid OA publishing is
considerable (Laakso \& Björk, 2016; Lawson, 2015; Pinfield et al.,
2016)

This current study investigates the extent to which publisher-provided
data can be utilized to assess hybrid OA. The study takes Elsevier, the
largest scholarly journal publisher in terms of article volume
(Larivière et al., 2015), as the starting point for an empirical
analysis of hybrid OA uptake and invoicing. Our approach is based on
openly available data sources and brings together scholarly data from
Crossref and machine-readable APC invoicing data from Elsevier OA
full-texts, which enables us to analyze the development of Elsevier's
hybrid journal portfolio from 2015-19, and to compare it with that of
Elsevier's full and delayed OA program. Further, we investigate whether
hybrid publishing fees were waived or charged to the authors or academic
institutions or research funders with Elsevier OA agreements. In
particular, we identify academic consortia and research funders that
were directly invoiced by Elsevier, and compare their hybrid OA rates by
license and subject. As such, our findings have implications for
research and OA policy and highlight the potential of publisher-provided
data for large-scale, instantaneous analysis of hybrid OA uptake and
invoicing.

\hypertarget{background}{%
\section{Background}\label{background}}

The ability to identify and discover hybrid OA articles plays a
fundamental role in monitoring hybrid OA, yet time and again, previous
studies have underscored the difficulties involved in this
(cf.~(Akbaritabar \& Stahlschmidt, 2019; Björk, 2012; Emery, 2013;
Laakso \& Björk, 2016; Leeuwen et al., 2018). Insights into the volume
of hybrid OA and the various levels and sources of APC funding could
create a better understanding of the costs involved in scholarly
publishing and, thus, strengthen institutions' bargaining power in
negotiations with publishers, and promote greater equity in scholarly
publishing. In the following section, we examine how previous research
has tackled the challenge of identifying hybrid OA content and what
these studies taught us about the uptake and funding sources of hybrid
OA.

\hypertarget{prevalence-uptake-studies}{%
\subsection{Prevalence / uptake
studies}\label{prevalence-uptake-studies}}

Since there is no standardized way to measure the uptake of hybrid OA,
estimates can vary greatly depending on how uptake is
operationalized--for example, through the number of hybrid journals with
at least one OA article, the average number of OA articles per hybrid
journal, or the share of OA articles compared to subscription articles.

In the first article-level study on hybrid OA, Laakso \& Björk (2016)
reported that between 2007 and 2013, the five largest
publishers--Elsevier, Sage, Springer, Taylor \& Francis, and
Wiley--recorded growing numbers of hybrid journals that coincided with a
twenty-fold increase in hybrid OA articles (from 666 to 13,994). As a
result, hybrid journals with at least one OA article more than doubled
in number (1,082 in 2009, 2,714 in 2013) but fell by 31\% relative to
the number of hybrid journals in total. Although Laakso and Björk's
(2016) presented a comprehensive overview tracking the growth of hybrid
OA over time and across multiple publishers, their exploratory approach
is unsuitable for repeated use as it required a considerable amount of
manual data collection and cleaning. Minimizing such manual tasks,
Nelson \& Eggett (2017) narrowed their focus to one publisher--the
American Chemical Society (ACS)--and requested a list of all hybrid OA
articles published from 2006-2011 directly from the ACS. To assess the
hybrid OA uptake, the authors aggregated and compared the number of OA
(n=814) and subscription articles (n=27,621), finding that over the five
years, 2.9\% of ACS articles were published as hybrid OA.

Contrary to the previous two studies, Kirkman (2018) took a research
funder as a starting point, reporting a 26\% share of hybrid OA for
articles funded by the Australian National Health and Medical Research
Council from 2012-2014 (816 of 3,190). Kirkman collected research
articles using the Funding Agency and Funding Text search fields on Web
of Science, and classified articles as OA when they were freely
available on the publisher website, and further distinguished hybrid OA
articles based on journal-level information from the publisher website,
such as hybrid OA journal and APC lists. Another approach is presented
by Pölönen et al. (2020), who used current research information system
(CRIS) data---institutional publication data---to study the extent of OA
among Finnish research publications from 2016-2017 (n=34,507 research
articles). As part of the analysis, the authors identified hybrid OA
articles through a dedicated OA status metadata field that has been
mandatory to report in Finland since 2016 and found a 7\% share of
hybrid OA. However, the drawbacks of Kirkman's (2018) and Pölönen et
al.'s (2020) approaches are that such data sources might not be openly
available or, regarding CRIS data, might not be available at all since
not all countries maintain comprehensive institutional publication data.

Other studies assessed the share of (hybrid) OA for the entire corpus of
scholarly research articles. For instance, using research articles with
an assigned DOI and indexed in the Web of Science Core Collection
(n=2,610,305) as a benchmark, Martín-Martín et al. (2018) assessed how
many research articles from 2009 and 2014 are freely available through
Google Scholar. After identifying OA articles through licensing
information from Crossref, the authors estimated the share of hybrid OA
at around 0.5\% in 2009 and 1.5\% in 2014. In a similar vein, Piwowar et
al. (2018) drew on Crossref to identify content licenses. The authors
followed a data-driven approach based on Unpaywall data, which is openly
available and contained 66 million Crossref DOIs at the time, to
determine how much of the scholarly literature is OA. Based on a random
sample of 100,000 Crossref DOIs, the authors found 3.6\% of hybrid OA,
while pointing towards a growing trend in recent years that recorded a
hybrid OA share of 9.4\% of all articles published in 2015. While
Piwowar et al.'s (2018) and Martín-Martín et al.'s (2018) methods are
well-suited for large-scale and repeated analyses of the prevalence of
hybrid OA, their main limitation is that they lack information about
publishing costs.

Moreover, Unpaywall's hybrid OA detection method has been refined
several times since its launch in 2017, reflecting growing albeit
incomprehensive license information in Crossref metadata. To overcome
these shortcomings, the developers extended previous web-scraping
approaches to better distinguish between immediate and delayed OA based
on licenses at the date of observation Piwowar et al. (2019). Relatedly,
in October 2020, Elsevier's license for delayed OA content was added to
Unpaywall's list of license patterns (Unpaywall, n.d.-c). Finally, at
the journal-level, a more fine-grained identification of full OA
journals using OA article information was developed to better
distinguish them from subscription journals including hybrid journals
(Unpaywall, n.d.-b). Because of these ongoing changes due to
non-standardized data, uptake studies reporting hybrid OA uptake based
on Unpaywall data cannot easily be compared with each other.

\hypertarget{financial-studies}{%
\subsection{Financial studies}\label{financial-studies}}

Some countries like Sweden and the United Kingdom (UK) have negotiated
national publishing agreements that allow authors to publish hybrid OA
at a discounted fee or for free. These agreements can serve as the basis
for studies on the uptake of hybrid OA or cost-benefit analyses
(cf.~Olsson (2018); Lawson (2017)) but are often limited in scope as
they tend to focus on a single country and rely on data that seldom is
publicly available. Hence, while being valuable for specific use-cases,
such studies can neither provide a comprehensive understanding of how
the scholarly publishing system functions beyond national boundaries nor
facilitate collective sharing of financial information about hybrid OA.

Pinfield et al.'s (2016) analysis of APC payment records provided by 23
UK higher education institutions revealed a sharp increase in central
payments from 2007-2013, which was largely attributed to the
introduction of block grants by Research Councils UK (RCUK) and
non-compliance sanctions by the Wellcome Trust. Moreover, the study
showed that OA fees were paid almost exclusively through RCUK and
Wellcome block grants (92\%), and only a small number of APCs were paid
through internal funding (7\%). In contrast, a recent Springer Nature
survey found that authors draw on a range of funding sources to cover OA
fees, such as dedicated institutional OA funds, block grants, OA
agreements, or research grants (Monaghan et al., 2020). Most hybrid OA
authors were supported through dedicated institutional OA funds (43\%,
excluding block grants) and OA agreements with Springer Nature (34\%).
The differences between these two studies may be due to changes in
policy and funding arrangements---in particular, the introduction of OA
agreements---since Pinfield et al. (2016), or due to Monaghan et al.'s
(2020) more regionally diverse sample, indicating regional differences
in APC funding.

Regional and policy differences in APC payments also came to light in a
study by Jahn \& Tullney (2016) that analyzed APC records from 30 German
higher education and research institutions crowd-sourced by the Open APC
initiative, the Austrian Science Fund (FWF), Jisc, and the Wellcome
Trust. In particular, the study revealed large differences in the amount
of hybrid OA funded from 2014-2015. Whilst hybrid OA accounted for less
than 1\% of APCs paid by German institutions (23 of 3,846), the three
non-German research funders recorded a hybrid share of75\% (11,533 of
15,779). According to Jahn \& Tullney (2016), this could point towards
differences in science policy, such as hybrid OA being supported by the
three non-German research funders but not by Germany's largest national
funder (DFG). Another possibility is that German hybrid OA fees were
paid from budgets not reported to the Open APC initiative, from where
the authors acquired the German APC data. Among these unreported funds
are research grants and research unit budgets, which author surveys
identified as funding sources for hybrid OA fees (Graaf, 2017; Monaghan
et al., 2020). As such, Jahn and Tullney's (2016) findings could reflect
the complexities and potential limitations of institutional OA spending
data that Pinfield et al.~(2016) and Monaghan et al.~(2020) attributed
to incomplete or missing records.

As this brief review has shown, the most comprehensive studies approach
hybrid OA funding from an institutional perspective that focuses on
research institutions or funders and draws on internal records or
crowd-sourced data like from the Open APC initiative. The advantages of
Open APC data are that they reflect the actual APCs paid rather than
list prices; they are not limited to certain publishers or journals;
they are provided in a standardized format regardless of the publisher.
On the other hand, the drawbacks include the limited number of
participating institutions that are almost exclusively based in Europe
and the dependency on participating organizations for data integrity.
Moreover, since Open APC data does not include waivers or free APCs, it
cannot present a complete picture of all publishing activities and
expenses. This, however, would pose less of an issue when analyzing just
hybrid OA.

\hypertarget{data-driven-workflows-for-hybrid-oa}{%
\subsection{Data-driven workflows for hybrid
OA}\label{data-driven-workflows-for-hybrid-oa}}

In September 2018, coalition S presented a set of principles known as
Plan S and announced. that from 2021 onwards, signatories would no
longer fund publications in hybrid OA journals. This principle includes
so-called ``mirror'' and sister journals---hybrid journals with fully OA
counterparts or split into their fully OA and subscription parts---that
have launched since Plan S was announced but makes an exception for
hybrid journals with transformative agreements (Frantsvåg \& Strømme,
2019). While the scale and means of these transformative agreements may
differ, many address the lack of transparency surrounding hybrid OA by
making publisher-provided metadata a defining element of publishing
workflows between institutions and publishers (Geschuhn \& Stone, 2017;
Marques et al., 2019)

Elsevier, one of the world's largest scholarly publishers, provides
article-level metadata to Crossref, including license information, which
can be used to identify OA articles, and links to HTML and XML full-text
versions for text and data mining purposes. Surprisingly, besides the
full-text article, the XML files also contain invoicing information and
the article's OA status. Table \ref{tab:els_xml} summarizes the key
metadata elements, which comprise the type of entity to which the APC
was invoiced (\texttt{openaccessSponsorType}) and the name of that
entity if available (\texttt{openaccessSponsorName}). Moreover, the
article-level metadata indicates if OA was provided through Elsevier's
Open Archive program (\texttt{openArchiveArticle}), which enables free
access to participating subscription journals after an embargo period
(i.e., delayed OA; Elsevier (n.d.-b)). As we have been unable to
identify similar efforts by other large publishers (e.g., Springer
Nature), this increased transparency through openly available
bibliometric and financial data at the article level seems unique to
Elsevier.

\begin{table}[H]
\caption{\label{tab:els_xml}Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.}
\centering
\begin{tabular}[t]{lp{10cm}}
\toprule
XML Node & Description\\
\midrule
\texttt{openaccessArticle} & Indicates if the article is open
access. \\
\texttt{openaccessSponsorType} & Invoice recipient type.\\
\texttt{openaccessSponsorName} & Entity that received Elsevier's APC
invoice.\\
\texttt{openArchiveArticle} & Indicates if open access was provided through
Elsevier's Open Archive programme.\tabularnewline \\
\bottomrule
\end{tabular}
\end{table}

Table 1: Overview of metadata provided in XML-based Elsevier full-texts
representing information about open access status and invoicing. A
sample XML is available here:
\url{https://api.elsevier.com/content/article/PII:S0169409X18301479?httpAccept=text/xml}

In this study, we utilize Elsevier's publicly available metadata to
analyze the uptake and invoicing of their hybrid OA publishing option.
In particular, we use this novel methodology to answer the following
questions:

\begin{itemize}
\item
  What was the uptake of Elsevier's hybrid OA publishing option between
  2015 and 2019?
\item
  Through which channels were hybrid APCs invoiced, and who were the
  recipients?
\end{itemize}

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

For this study, we collected data relating to Elsevier's hybrid OA
option by drawing on multiple freely available data sources. We
identified Elsevier hybrid journals through Elsevier's APC list and
supplemented our sample with Crossref metadata and text-mined invoicing
data to investigate the financing of immediate OA articles provided
under a Creative Commons license in subscription-based journals. Figure
\ref{fig:workflow} visualizes the automated workflow we used to collect
data from Elsevier and Crossref.

\begin{figure}[H]

{\centering \includegraphics[width=0.95\textwidth]{../figure/els_flow.pdf}

}

\caption{Data Analytics Workflow to obtain article-level OA and invoicing data using freely available data sources provided by Elsevier and Crossref.
}\label{fig:workflow}
\end{figure}

First, we obtained hybrid journals through an Elsevier APC price list
from May 2020 provided as captured by Matthias (2020). We only included
hybrid journals that were published under the hybrid model throughout
the entire period of our analysis. Hence, we excluded journals that
flipped from OA to hybrid and from hybrid to OA. This approach allowed
us to ensure that the article-level differences we observed were not due
changes at the journal-level (i.e., increases or decreases in number).
Overall, we identified 1,970 unique hybrid journals.

Next, we used an openly available Crossref database snapshot (Crossref,
2020), which contains all Crossref records registered until March 2020,
to calculate the journals' combined article volume for the five-year
period from 2015-2019. We only included articles published in regular
issues aside from supplements containing conference contributions like
meeting abstracts, indicated by non-numeric pagination. We excluded
non-scholarly journal content, such as the table of contents and lists
of reviewers, following Unpaywall's paratext recognition approach
(Unpaywall, n.d.-a), which we expanded to include patterns indicating
corrections. Furthermore, we categorized the articles by subject
according to the All Science Journal Classification code (ASJC) and
added citation indicators from the Scopus Source Title list (version
June 2020) based on the journal they were published in.

Crossref metadata also contains license information that indicates if
and under which terms the publisher has made articles openly available
on their website (Hendricks et al., 2020). We identified OA articles
through Creative Commons (CC) licenses and then downloaded the XML
version of all CC-licensed articles published in a hybrid journal. From
the XML files, we determined whether or not the articles were immediate
OA using the XML node \texttt{openArchiveArticle} and measured the
uptake of hybrid OA through three bibliometric indicators (Laakso \&
Björk, 2016; Nelson \& Eggett, 2017)-\/---the number of hybrid journals
with at least one OA article, the number of hybrid OA articles, and the
number of hybrid OA articles relative to the number of closed-accessed
articles in hybrid journals with at least one OA article.

Moreover, we obtained APC invoicing metadata from the XML (see Table 1),
which enabled us to identify the invoice channels and recipients of
hybrid OA APCs. Based on the \texttt{openaccessSponsorType} node in the
XML files, we distinguish between four invoice channels, including
invoices billed to the corresponding author, as part of publishing
agreements with funding bodies (hereinafter ``agreements''; cf.~Elsevier
(n.d.-a)), fee waivers (e.g., in ``cases of genuine need'' or due to
society or university sponsorships, cf.~ref), and other types not
specified by Elsevier. Finally, when hybrid APCs were invoiced as part
of OA agreements, we identified invoice recipients through the
\texttt{openaccessSponsorName} node.

We manually classified invoice recipients based on their institutional
sectors, countries, and primary research areas. Following the OECD's
Frascati Manual (OECD, 2015, p. 91), we coded for four sectors, business
enterprise, government, higher education, and private non-profit. The
government sector consists of federal, state, and municipal governments,
plus non-profit institutions controlled by government units (OECD, 2015,
p. 100). The higher education sector comprises institutions that perform
research and development and provide formal tertiary education services,
such as universities or research institutes (OECD, 2015, p. 101). Legal
or social entities that produce goods and services without generating
profit, such as learned societies or charities, constitute the private
non-profit sector (OECD, 2015, p. 288). Due to the low article volume,
we combined invoice recipients Elsevier listed as ``authors'' and
``third-party sponsor'', and those we had coded as business enterprises
(OECD, 2015, p. 97), into ``Others''. Moreover, we categorized invoice
recipients according to the country or countries representing their
scope of funding and based on the following primary research areas
health sciences, life sciences, physical sciences and mathematics,
social sciences and humanities, unknown, and broad (i.e., multiple
research areas).

It is important to note that we could not find any official
comprehensive documentation of Elsevier's XML schema , at the time of
writing. However, our conceptualization of the variables and
interpretation of the data was informed by a comprehensive review of
Elsevier's websites about OA, including a list of funding agreements and
details about Elsevier's Open Archive program,and Elsevier's article
submission process in the context of dedicated OA agreements (Elsevier,
n.d.-d, n.d.-e, n.d.-c). For validation, we matched the invoicing data
provided by Elsevier to data from the Open APC initiative.

Throughout the mostly automated data gathering and analysis process, we
used tools from the Tidyverse (Wickham et al., 2019) for the R
programming language (R Core Team, 2020). To allow for efficient data
manipulation and retrieval, we imported the Crossref dump to Google
BigQuery, applying the rcrossref (Chamberlain et al., 2020) parsers to
extract relevant metadata fields. We interfaced Google BigQuery with SQL
with the bigrquery client (Wickham \& Bryan, 2020). We used crminer
(Chamberlain, 2020) to obtain the XML files from Elsevier and the xml2
package (Wickham et al., 2020) to parse relevant metadata. To ensure our
results are reproducible, we provided all code for obtaining the data
and performing the analyses with detailed documentation as RMarkdown
notebooks on GitHub
(\url{https://github.com/njahn82/elsevier_hybrid_invoicing}).

\hypertarget{results}{%
\section*{Results}\label{results}}
\addcontentsline{toc}{section}{Results}

This section first presents the results of our analysis of hybrid OA
uptake in Elsevier's journal portfolio with a view to licensing,
disciplinary differences, and citation impact. Then, we present a
descriptive analysis of Elsevier's invoicing data, highlighting
licensing and disciplinary differences for different invoicing channels,
institutional sectors, and individual invoice recipients.

\hypertarget{uptake-of-hybrid-open-access}{%
\subsection*{Uptake of Hybrid Open
Access}\label{uptake-of-hybrid-open-access}}
\addcontentsline{toc}{subsection}{Uptake of Hybrid Open Access}

\hypertarget{overview}{%
\subsubsection*{Overview}\label{overview}}
\addcontentsline{toc}{subsubsection}{Overview}

Between 2015 and 2019, 1,755 of 1,970 (89.1\%) hybrid journals published
at least one OA article. Together these journals published 71,643 OA
articles, which represents 3\% of their total article volume
(n=2,422,087). Table \ref{tab:overview} presents these findings for each
year and the aggregated five-year period, illustrating a moderate growth
in hybrid OA over time. With each year, the number of hybrid journals
with at least one OA article increased, as did the number of immediate
OA articles in these journals, which nearly doubled from 10,672 in 2015
to 19,311 in 2019. However, since the total number of articles published
in these journals also grew over time, the relative share of OA articles
in hybrid journals only increased slightly from 2.6\% to 3.7\%.

\begin{table}[H]

\caption{\label{tab:overview}Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & 2015 & 2016 & 2017 & 2018 & 2019 & 2015-19\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Elsevier Hybrid Journals with ≥ 1 OA article}}\\
\hspace{1em}Total & 1,317.0 & 1,364.0 & 1,401.0 & 1,501.0 & 1,600.0 & 1,755.0\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Articles in Elsevier Hybrid Journals with ≥ 1 OA Article}}\\
\hspace{1em}Total & 406,701.0 & 422,423.0 & 436,418.0 & 468,952.0 & 518,062.0 & 2,422,087.0\\
\hspace{1em}Avg. & 308.8 & 309.7 & 311.5 & 312.4 & 323.8 & 1,380.1\\
\hspace{1em}SD & 401.1 & 419.6 & 427.5 & 445.6 & 479.9 & 1,937.5\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{OA Articles in Elsevier Hybrid Journals with ≥ 1 OA Article}}\\
\hspace{1em}Total & 10,672.0 & 12,729.0 & 13,361.0 & 15,570.0 & 19,311.0 & 71,643.0\\
\hspace{1em}Avg. & 8.1 & 9.3 & 9.5 & 10.4 & 12.1 & 40.8\\
\hspace{1em}SD & 12.9 & 19.7 & 15.6 & 16.2 & 20.1 & 68.9\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA Article}}\\
\hspace{1em}Total (\%) & 2.6 & 3.0 & 3.1 & 3.3 & 3.7 & 3.0\\
\hspace{1em}Avg. & 3.7 & 4.3 & 4.4 & 4.7 & 5.4 & 3.8\\
\hspace{1em}SD & 4.5 & 5.9 & 5.1 & 4.9 & 5.7 & 4.1\\
\bottomrule
\end{tabular}
\end{table}

Figure \ref{fig:boxuptake} presents the hybrid OA uptake by year,
showing considerable variation among Elsevier's hybrid journals. The
box-and-whiskers plot visualizes the median hybrid OA uptake (horizontal
line inside the box), the range of the majority of journals (the box),
and individual outliers (dots). Notably, the upper quartiles and
whiskers stretch farther over the years, which indicates that the range
of uptake among journals with an above-average proportion of OA articles
increased over the years. Nevertheless, the prevalence of hybrid OA
remained relatively low; in 2019, 95\% of Elsevier hybrid journals
published ≤15.6\% of their articles OA.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{manuscript_files/figure-latex/boxuptake-1}

}

\caption{Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 20\%.}\label{fig:boxuptake}
\end{figure}

\hypertarget{share-of-the-total-volume-published-in-oa}{%
\subsubsection*{Share of the Total Volume published in
OA}\label{share-of-the-total-volume-published-in-oa}}
\addcontentsline{toc}{subsubsection}{Share of the Total Volume published
in OA}

During the same period, Elsevier published 328,601 OA articles in total
(i.e., in hybrid and full OA journals), representing 12.4\% of the total
article volume (n=2,643,474). Looking at Figure \ref{fig:oa_volume_fig},
it is apparent that the OA article volume of hybrid journals (n=71,643;
21.8\%) lagged behind that of Elsevier's Open Archive program
(n=163,643; 49.8\%) and full OA journals (n=93,315; 28.4\%), which
included 817 articles published in 38 mirror journals.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{/Users/najkojahn/Documents/papers/elsevier_hybrid_invoicing/figure/license_portfolio}

}

\caption{Open Access (OA) in Elsevier journals in terms of article volume between 2015-2019. Figure A presents the aggregated OA article volume of Elsevier journals from 2015-2019 by OA type. Figure B presents the share of OA articles relative to Elsevier's overall article output by OA type and open content license from 2015-2019.}\label{fig:oa_volume_fig}
\end{figure}

\hypertarget{license-prevalence}{%
\subsubsection*{License Prevalence}\label{license-prevalence}}
\addcontentsline{toc}{subsubsection}{License Prevalence}

As far as we observed, OA articles in Elsevier hybrid journals were
published under two possible CC licenses--CC BY or CC BY-NC-ND. CC BY
allows others to distribute, remix, adapt, and build upon the licensed
work, including for commercial purposes, as long as the original author
is credited. In contrast, CC BY-NC-ND is less permissive as it prohibits
commercial reuse and derivatives. In the five-year period from
2015-2019, the proportion of hybrid OA articles published under a CC
BY-NC-ND license marginally increased and maintained a greater share
than that of CC BY (see Figure \ref{fig:oa_volume_fig}-B).
Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid
journals had the highest number and proportion of openly available
articles licensed under CC BY (n=29,752; 41.5\%) compared to full OA
journals (n=17,293; 18.5\%) and Elsevier's delayed OA program (n=1,568;
1\%) . Most delayed OA articles were provided under an Elsevier user
license (Els-User), which prohibits reuse for commercial purposes.

\hypertarget{subject-area-and-field}{%
\subsubsection*{Subject Area and Field}\label{subject-area-and-field}}
\addcontentsline{toc}{subsubsection}{Subject Area and Field}

Next, we present disciplinary variations in hybrid OA. Table
\ref{tab:subject_area_table} presents the high-level findings by AJSC
subject area. In the five-year period we studied, the physical sciences
(712 hybrid journals, 29,584 OA articles) and health sciences (634
hybrid journals, 25,119 OA articles) were the most prevalent in terms of
hybrid journals with at least one OA article. In terms of article
volume, most OA articles were published in hybrid journals belonging to
the life sciences (538 hybrid journals, 31,383 OA articles), while
social sciences (372 hybrid journals, 11,204 articles) journals
published the lowest number of OA articles. However, it is important to
note that there is a large overlap between the journals' subject areas.
Therefore, Table \ref{tab:subject_area_table} also assigns hybrid
journals and OA articles fractionally to a subject area and compares it
to the full counting, where journals assigned more than one subject in
the AJSC classification were counted once for each field. For the
remaining results, we only present the full counting.

\begin{table}

\caption{\label{tab:subject_area_table}Subject area overview using full and fractional counting. Table reports the number of Elsevier hybrid journals with at least one open access (OA) article, and OA articles published from 2015-2019. One multidisciplinary with 61 OA articles and 11 journals with 37 OA articles that  could not be matched to the Scopus Source Title list used to obtain ASJC codes are not included.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Full counting} & \multicolumn{2}{c}{Fractional counting} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Subject area & Journals & OA articles & Journals & OA articles\\
\midrule
Health sciences & 634 & 25,119 & 506.6 & 18,415\\
Life sciences & 538 & 31,383 & 368.5 & 21,753\\
Physical sciences & 712 & 29,584 & 584.4 & 23,915\\
Social sciences & 372 & 11,204 & 283.5 & 7,462\\
\bottomrule
\end{tabular}
\end{table}

Figure \ref{fig:oa_sub_uptake} presents the journals' five-year OA
uptake grouped by subject area and field and in comparison to the
overall median (Mdn= 2.5; excluding one journal coded as
multidisciplinary). The box plot reveals some variation between the
subject areas. Notably, with the exception of environmental science and
earth and planetary sciences, journals from the physical sciences show a
lower median proportion of hybrid OA articles, whereas most life
sciences and social sciences journals ranked above average. Figure X
also shows large variations among the 26 subject fields, with the
highest OA rates in immunology and microbiology (Mdn=5.1), environmental
science (Mdn=4.8), and neuroscience (Mdn=4.3), whilst chemistry
(Mdn=1.4), materials science (Mdn=1.4), and dentistry (Mdn=0.6) recorded
the lowest OA rate.

\begin{figure}[H]

{\centering \includegraphics[width=0.9\linewidth,]{manuscript_files/figure-latex/oa_sub_uptake-1}

}

\caption{Open access (OA) uptake in Elsevier hybrid journals in percent grouped by subject area and field between 2015-19. Visualized as a box-and-whiskers plot limited to an OA share of 20\%; the red line represents the overall median (2.5), jittered dots represent individual journals. Subject fields are ordered based on the median OA uptake.}\label{fig:oa_sub_uptake}
\end{figure}

\hypertarget{citation-impact}{%
\subsubsection*{Citation Impact}\label{citation-impact}}
\addcontentsline{toc}{subsubsection}{Citation Impact}

Furthermore, we were interested in the relationship between the
journals' field-specific impact and their OA uptake. Spearman's rho
correlation coefficient was used to assess the relationship between the
2019 Source Normalized Impact per Paper (SNIP) value calculated by
Scopus and the journals' OA uptake that year. Considering only journals
with at least one OA article, we found a weak but positive correlation
between journal impact and OA uptake (\(r_s\)=0.1679, \(p<0.001\)),
suggesting that the relationship is not very strong, based on our data
(see Figure \ref{fig:sniptest}).

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{manuscript_files/figure-latex/sniptest-1}

}

\caption{Source Normalized Impact per Paper (SNIP) versus open access (OA) uptake 2019. Each dot represents a hybrid journal with at least one published OA article in 2019. There is only a weak, positive correlation between the two variables ($r_s$=0.1679, $p<0.001$)}\label{fig:sniptest}
\end{figure}

\hypertarget{invoicing-for-hybrid-open-access}{%
\subsection*{Invoicing for Hybrid Open
Access}\label{invoicing-for-hybrid-open-access}}
\addcontentsline{toc}{subsection}{Invoicing for Hybrid Open Access}

\hypertarget{invoice-channels}{%
\subsubsection*{Invoice Channels}\label{invoice-channels}}
\addcontentsline{toc}{subsubsection}{Invoice Channels}

As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APC
were most often invoiced to authors (n=41,725; 58.2\%) and to a lesser
extent as part of agreements (n=24,250; 33.8\%). Interestingly, we also
found a small number of cases where hybrid APCs were waived (n=4,345;
6.1\%).

\begin{table}

\caption{\label{tab:invoicing_overview}Invoice channels for hybrid OA in Elsevier journals between 2015-19.}
\centering
\begin{tabular}[t]{lrr}
\toprule
Invoicing channel & Hybrid OA articles (n) & Percentage\\
\midrule
Author & 41,725 & 58.2\\
Agreement & 24,250 & 33.8\\
Fee waived & 4,345 & 6.1\\
Other & 1,323 & 1.8\\
\midrule
Total & 71,643 & 100.0\\
\bottomrule
\end{tabular}
\end{table}

Figure \ref{fig:invoiceoverview} illustrates that over the years,
Elsevier has increasingly invoiced hybrid OA APCs directly to authors
compared to research funders or academic consortia (``Agreement''). The
share of fee-waived articles remained relatively stable, but we found
different types of waivers. Around 51.7\% of fee waivers were linked to
an OA sponsor, which indicates that a third party paid the APC. For
instance, 853 waivers were linked to the French Académie des Sciences,
presumably covering OA publication in its society journals for
affiliated authors. The remaining 48.3\% of waived articles did not
disclose any OA sponsor. Moreover, Figure \ref{fig:invoiceoverview}
compares the invoicing channels based on open content license variants.
When Elsevier invoiced authors directly, most OA articles in hybrid
journals were published under a non-commercial license (n=32,086;
76.9\%), whereas most articles billed as part of agreements were
licensed under the more permissive CC BY license (n=18,331; 75.6\%).

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{manuscript_files/figure-latex/invoiceoverview-1}

}

\caption{Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoice channel Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant. Grey bars show the total number of hybrid OA articles published in Elsevier journals from 2015-2019.}\label{fig:invoiceoverview}
\end{figure}

We also observed large differences in OA invoicing among subject fields.
Table \ref{tab:invocing_by_subject} shows the number of OA articles by
subject field and invoice channel. For articles in nursing, decision
dciences, and pharmacology, toxicology and pharmaceutics, Elsevier
predominantly invoiced authors, whereas most energy and chemical
engineering articles were invoiced through agreements. Likewise, the
majority of articles in materials science, chemistry and physics and
astronomy were not invoiced to authors but facilitated through
agreements or waived. The large share of waived APCs in physics and
astronomy can be attributed to a single 2015 issue of Nuclear and
Particle Physics Proceedings, which offered sponsored OA.

\begin{table}[H]

\caption{\label{tab:invocing_by_subject}Invoice channels by discipline 2015-2019.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrr}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{4}{c}{Invoice channel (in \%)} \\
\cmidrule(l{3pt}r{3pt}){3-6}
Subject & OA articles & Author & Agreement & Fee waived & Other\\
\midrule
Nursing & 1,504 & \cellcolor[HTML]{DC3977}{82} & \cellcolor[HTML]{FFFFFF}{16} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FCDE9C}{1}\\
Decision Sciences & 785 & \cellcolor[HTML]{E6576F}{72} & \cellcolor[HTML]{FCC088}{27} & \cellcolor[HTML]{FFFFFF}{0} & \cellcolor[HTML]{FCDE9C}{1}\\
Pharmacology, Toxicology and Pharmaceutics & 4,075 & \cellcolor[HTML]{EA646F}{69} & \cellcolor[HTML]{FCC088}{27} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FAA476}{2}\\
Business, Management and Accounting & 1,751 & \cellcolor[HTML]{EC686E}{68} & \cellcolor[HTML]{FBAF7D}{29} & \cellcolor[HTML]{FFF0D2}{2} & \cellcolor[HTML]{FCDE9C}{1}\\
Dentistry & 288 & \cellcolor[HTML]{EC686E}{68} & \cellcolor[HTML]{FFFAF1}{17} & \cellcolor[HTML]{F89774}{10} & \cellcolor[HTML]{DC3977}{5}\\
Health Professions & 674 & \cellcolor[HTML]{ED6D6E}{67} & \cellcolor[HTML]{FCD898}{24} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{DC3977}{5}\\
Medicine & 23,623 & \cellcolor[HTML]{F0756E}{65} & \cellcolor[HTML]{FCD092}{25} & \cellcolor[HTML]{FCC98E}{6} & \cellcolor[HTML]{E34F6F}{4}\\
Agricultural and Biological Sciences & 7,956 & \cellcolor[HTML]{F38170}{63} & \cellcolor[HTML]{F9A075}{31} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FCDE9C}{1}\\
Computer Science & 3,185 & \cellcolor[HTML]{F38170}{63} & \cellcolor[HTML]{F27F70}{36} & \cellcolor[HTML]{FFFFFF}{0} & \cellcolor[HTML]{FCDE9C}{1}\\
Economics, Econometrics and Finance & 2,071 & \cellcolor[HTML]{F48771}{62} & \cellcolor[HTML]{F79373}{33} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FFFFFF}{0}\\
Veterinary & 2,091 & \cellcolor[HTML]{F58D72}{61} & \cellcolor[HTML]{FCD092}{25} & \cellcolor[HTML]{F1766E}{13} & \cellcolor[HTML]{FCDE9C}{1}\\
Social Sciences & 6,306 & \cellcolor[HTML]{F79273}{60} & \cellcolor[HTML]{F27F70}{36} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FFFFFF}{0}\\
Earth and Planetary Sciences & 4,532 & \cellcolor[HTML]{F89874}{59} & \cellcolor[HTML]{F48671}{35} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FFFFFF}{0}\\
Biochemistry, Genetics and Molecular Biology & 15,903 & \cellcolor[HTML]{F89874}{59} & \cellcolor[HTML]{F48671}{35} & \cellcolor[HTML]{FDE1A5}{4} & \cellcolor[HTML]{FAA476}{2}\\
Immunology and Microbiology & 5,542 & \cellcolor[HTML]{F89874}{59} & \cellcolor[HTML]{F58D72}{34} & \cellcolor[HTML]{FCC98E}{6} & \cellcolor[HTML]{FAA476}{2}\\
Environmental Science & 9,000 & \cellcolor[HTML]{F99D75}{58} & \cellcolor[HTML]{EF726E}{38} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FCDE9C}{1}\\
Psychology & 3,087 & \cellcolor[HTML]{FAA376}{57} & \cellcolor[HTML]{EC686E}{40} & \cellcolor[HTML]{FFF0D2}{2} & \cellcolor[HTML]{FCDE9C}{1}\\
Neuroscience & 5,906 & \cellcolor[HTML]{FBAA7A}{56} & \cellcolor[HTML]{EC686E}{40} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FCDE9C}{1}\\
Arts and Humanities & 1,565 & \cellcolor[HTML]{FCB882}{54} & \cellcolor[HTML]{E4536F}{44} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FFFFFF}{0}\\
Mathematics & 2,105 & \cellcolor[HTML]{FCBF87}{53} & \cellcolor[HTML]{EF726E}{38} & \cellcolor[HTML]{FBAF7D}{8} & \cellcolor[HTML]{FFFFFF}{0}\\
Engineering & 8,405 & \cellcolor[HTML]{FCD395}{50} & \cellcolor[HTML]{E24C70}{46} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FCDE9C}{1}\\
Materials Science & 4,838 & \cellcolor[HTML]{FCDA99}{49} & \cellcolor[HTML]{E14971}{47} & \cellcolor[HTML]{FDE1A5}{4} & \cellcolor[HTML]{FFFFFF}{0}\\
Chemistry & 4,033 & \cellcolor[HTML]{FCE0A1}{48} & \cellcolor[HTML]{E24C70}{46} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FCDE9C}{1}\\
Energy & 4,286 & \cellcolor[HTML]{FEE7B8}{46} & \cellcolor[HTML]{DC3977}{52} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FCDE9C}{1}\\
Chemical Engineering & 2,977 & \cellcolor[HTML]{FFEBC4}{45} & \cellcolor[HTML]{E14971}{47} & \cellcolor[HTML]{FCBC85}{7} & \cellcolor[HTML]{FCDE9C}{1}\\
Physics and Astronomy & 5,859 & \cellcolor[HTML]{FFFFFF}{40} & \cellcolor[HTML]{EF726E}{38} & \cellcolor[HTML]{DC3977}{22} & \cellcolor[HTML]{FFFFFF}{0}\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{invoice-recipients}{%
\subsubsection*{Invoice Recipients}\label{invoice-recipients}}
\addcontentsline{toc}{subsubsection}{Invoice Recipients}

Elsevier's data offer more insight into invoicing. Overall, we
identified 62 sponsoring institutions and organizations that received
invoices for hybrid publishing fees as part of an agreement with
Elsevier. We found that, by a large margin, most invoices were issued to
UK-based research funders and institutions (n=14,344; 59.2\%), followed
by the Netherlands (n=2,835; 11.7\%) and the European Union (n=2,164;
8.9\%; EU). Figure \ref{fig:invoice_sponsor_country} provides a closer
look at the geographical distribution by license prevalence,
highlighting the dominance of CC -BY licensed articles invoiced to
sponsoring institutions in the UK, the United States, and, to a much
lesser extent, the Netherlands. In contrast, the majority of articles
invoiced to sponsoring institutions from Norway or representing the EU
were published under a non-commercial license.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{manuscript_files/figure-latex/invoice_sponsor_country-1}

}

\caption{Number of centrally invoiced OA articles in Elsevier hybrid journals by country of recipient. Colored bars represent the Creative Commons (CC) license variant under which the article was published between 2015 and 2019.}\label{fig:invoice_sponsor_country}
\end{figure}

Figure \ref{fig:sponsor_sector} shows the yearly distribution of central
invoiced articles by year and institutional sector. While UK-based OA
invoices were mainly addressed to discipline-specific governmental and
non-profit research funders, invoices to the Netherlands and Norway were
issued to national academic consortia representing the higher education
sector. In 2019, Elsevier also launched similar agreements in countries
with lower publication output including Hungary and Poland. Besides, we
found that invoiced funding bodies from the UK and the US mainly
represented discipline-specific funders, while the invoiced funding
bodies from other countries were focused on a broad variety of
disciplines (see Table \ref{tab:subject_profile}).

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth,]{manuscript_files/figure-latex/sponsor_sector-1}

}

\caption{ Development of centrally invoiced OA articles in hybrid journals by country. Colored bars represent the institutional sector OA funding bodies belong to.}\label{fig:sponsor_sector}
\end{figure}

\begin{table}

\caption{\label{tab:subject_profile} Relative subject profile of centrally invoiced OA funding bodies.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{l|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|r}
\toprule
Country/Subject area & Health science & Life sciences & Physical sciences & SSH & Broad & Total\\
\midrule
United Kingdom & 25\% & 11\% & 21\% & 5\% & 0\% & 62\%\\
The Netherlands & 0\% & 0\% & 0\% & 0\% & 12\% & 12\%\\
European Union & 0\% & 0\% & 0\% & 0\% & 9\% & 9\%\\
United States & 6\% & 0\% & 0\% & 0\% & 1\% & 7\%\\
Norway & 0\% & 0\% & 0\% & 0\% & 3\% & 3\%\\
\addlinespace
Other & 1\% & 0\% & 1\% & 0\% & 5\% & 7\%\\
\midrule
Total & 32\% & 11\% & 22\% & 5\% & 30\% & 100\%\\
\midrule
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{institutional-ranking}{%
\subsubsection*{Institutional ranking}\label{institutional-ranking}}
\addcontentsline{toc}{subsubsection}{Institutional ranking}

Table \ref{tab:sponsor_top_10} presents the top ten of 63 sponsoring
bodies in our sample. Together, they accounted for around 80\% of
centrally invoiced APCs. The table also highlights the number of
distinct journals and the share of CC BY licensed articles. Notably,
UK-based research funders and the US-based Bill and Melinda Gates
Foundation mainly sponsored articles published under a CC BY license. In
contrast, articles funded by VSNU, the European Research Council, and
Norway Institutes had a much lower proportion of CC BY.

\begin{table}

\caption{\label{tab:sponsor_top_10} Top 10 Elsevier hybrid OA sponsors by centrally invoiced articles 2015-2019.  The table furthermore presents the number of Elsevier hybrid journals with at least one invoiced article and two indicators of compliance: proportion of  CC BY licensed articles  and disclosed spending through the Open APC Initiative (OAPC.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrr}
\toprule
\multicolumn{4}{c}{ } & \multicolumn{2}{c}{Compliance (in \%)} \\
\cmidrule(l{3pt}r{3pt}){5-6}
Hybrid OA sponsors & Journals & Articles & \% & CC BY & OAPC\\
\midrule
Engineering and Physical Sciences Research Council & 619 & 4663 & 19 & \cellcolor[HTML]{DE3E75}{97} & \cellcolor[HTML]{E34E6F}{55}\\
VSNU & 357 & 2835 & 12 & \cellcolor[HTML]{FCC38A}{52} & \cellcolor[HTML]{FFFFFF}{0}\\
Wellcome Trust & 448 & 2506 & 10 & \cellcolor[HTML]{DD3D76}{98} & \cellcolor[HTML]{DD3D76}{66}\\
European Research Council & 541 & 1986 & 8 & \cellcolor[HTML]{FFFFFF}{32} & \cellcolor[HTML]{FFE9BD}{9}\\
Medical Research Council & 384 & 1922 & 8 & \cellcolor[HTML]{DF4274}{95} & \cellcolor[HTML]{E6596F}{51}\\
Natural Environment Research Council & 203 & 1357 & 6 & \cellcolor[HTML]{DE3E75}{97} & \cellcolor[HTML]{EC696E}{45}\\
Biotechnology and Biological Sciences Research Council & 309 & 1169 & 5 & \cellcolor[HTML]{E04573}{93} & \cellcolor[HTML]{F0736E}{41}\\
Bill and Melinda Gates Foundation & 254 & 1030 & 4 & \cellcolor[HTML]{DC3977}{100} & \cellcolor[HTML]{DC3977}{68}\\
Economic and Social Research Council & 245 & 922 & 4 & \cellcolor[HTML]{DE4075}{96} & \cellcolor[HTML]{E9616F}{48}\\
Norway Institutes & 374 & 694 & 3 & \cellcolor[HTML]{FFE9BD}{41} & \cellcolor[HTML]{FFFFFF}{0}\\
Other & 1012 & 5166 & 21 & \cellcolor[HTML]{FCBA84}{54} & \cellcolor[HTML]{FCBF87}{21}\\
\midrule
All & 1423 & 24250 & 100 & \cellcolor[HTML]{ED6C6E}{76} & \cellcolor[HTML]{F48671}{36}\\
\midrule
\bottomrule
\end{tabular}}
\end{table}

The table also highlights the proportion of APC funding publicly
disclosed through the Open APC initiative, showing higher disclosure
rates for funders with a large CC BY share. On the other hand, APCs that
were invoiced as part of transformative agreements were not publicly
shared.

Figure \ref{fig:field_funder_matrix} shows the subject profile of OA
sponsors based on to the journals' subject fields of centrally invoiced
articles. As expected, the figure shows that subject-specific funders,
such as the British Engineering and Physical Sciences Research Council
(EPSRC), tended to focus on a limited set of subject fields. In
contrast, national consortia, such as VSNU and Norwegian Institutes
(NO), financed OA articles for a broader range of fields. However,
within these consortia, the subject distribution also varied. VSNU OA
spending, for example, had a strong focus on articles relating to
medicine, environmental science, and energy, presumably reflecting
disciplines with a generally strong publication output at Dutch
universities. Norwegian universities mostly published hybrid OA in
engineering, environmental science and agricultural and biological
sciences journals.

\begin{figure}[H]

{\centering \includegraphics[width=0.95\textwidth]{../figure/funder_subject.png}

}

\caption{Top 10 Elsevier hybrid OA institutional invoice recipients in terms of articles published by journal subject. EPSRC, Engineering and Physical Sciences Research Council; VSNU, VSNU; Wellcome, Wellcome Trust; ERC, European Research Council; ERC, European Research Council; MRC, Medical Research Council; NERC, Natural Environment Research Council; BBSRC, Biotechnology and Biological Sciences Research Council; Gates, Bill and Melinda Gates Foundation; ESRC, Economic and Social Research Council; NO, Norway Institutes. }\label{fig:field_funder_matrix}
\end{figure}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Akbaritabar_2019}{}%
Akbaritabar, A., \& Stahlschmidt, S. (2019). \emph{Applying crossref and
unpaywall information to identify gold, hidden gold, hybrid and delayed
open access publications in the KB publication corpus}. SocArXiv.
\url{https://doi.org/10.31235/osf.io/sdzft}

\leavevmode\hypertarget{ref-antelmann_2017}{}%
Antelman, K. (2017). Leveraging the growth of open access in library
collection decision making. \emph{At the helm: leading transformation},
411--422.

\leavevmode\hypertarget{ref-Bergstrom_2014}{}%
Bergstrom, T. C., Courant, P. N., McAfee, R. P., \& Williams, M. A.
(2014). Evaluating big deal journal bundles. \emph{Proceedings of the
National Academy of Sciences}, \emph{111}(26), 9425--9430.
\url{https://doi.org/10.1073/pnas.1403006111}

\leavevmode\hypertarget{ref-Bj_rk_2012}{}%
Björk, B.-C. (2012). The hybrid model for open access publication of
scholarly articles: A failed experiment? \emph{Journal of the American
Society for Information Science and Technology}, \emph{63}(8),
1496--1504. \url{https://doi.org/10.1002/asi.22709}

\leavevmode\hypertarget{ref-Borrego_2020}{}%
Borrego, Á., Anglada, L., \& Abadal, E. (2020). Transformative
agreements: Do they pave the way to open access? \emph{Learned
Publishing}. \url{https://doi.org/10.1002/leap.1347}

\leavevmode\hypertarget{ref-crminer}{}%
Chamberlain, S. (2020). \emph{crminer: Fetch scholary full text from
CrossRef}. \url{https://CRAN.R-project.org/package=crminer}

\leavevmode\hypertarget{ref-rcrossref}{}%
Chamberlain, S., Zhu, H., Jahn, N., Boettiger, C., \& Ram, K. (2020).
\emph{rcrossref: Client for various CrossRef APIs}.
\url{https://CRAN.R-project.org/package=rcrossref}

\leavevmode\hypertarget{ref-Crossref_2020}{}%
Crossref. (2020). \emph{March 2020 public data file from crossref}.
Crossref. \url{https://doi.org/10.13003/83b2gp}

\leavevmode\hypertarget{ref-Els_Agreements}{}%
Elsevier. (n.d.-a). \emph{Agreements}.
\url{https://web.archive.org/web/20210127152729/https://www.elsevier.com/open-access/agreements}.

\leavevmode\hypertarget{ref-Els_Archive}{}%
Elsevier. (n.d.-b). \emph{Open archive}.
\url{http://web.archive.org/web/20210127201740/https://www.elsevier.com/open-access/open-archive}.

\leavevmode\hypertarget{ref-Els_videos}{}%
Elsevier. (n.d.-c). \emph{Publishing Journey videos: How do I complete
the Rights and Access form?}
\url{https://web.archive.org/web/20210127154144/https://service.elsevier.com/app/answers/detail/a_id/29789/supporthub/publishing/track/APN2ZgoIDv8a~RNiGvwa~yKgpv0qOS75Mv9e~zj~PP_X/}.

\leavevmode\hypertarget{ref-lund}{}%
Elsevier. (n.d.-d). \emph{Publishing options: BIBSAM institute
associated authors}.
\url{https://web.archive.org/web/20210127153520/https://www.ub.lu.se/en/sites/ub.lu.se.en/files/publicering_under_elsevier_avtalet.pdf}.

\leavevmode\hypertarget{ref-vsnu}{}%
Elsevier. (n.d.-e). \emph{Publishing options for: Dutch universities \&
institutes associated authors}.
\url{https://web.archive.org/web/20190530214833/https://www.openaccess.nl/sites/www.openaccess.nl/files/documenten/elsevier_-_vsnu_workflow_updated_18july_v2.pdf}.

\leavevmode\hypertarget{ref-Emery_2013}{}%
Emery, J. (2013). Mining for gold: Identifying the librarians' toolkit
for managing hybrid open access. \emph{Insights: The UKSG Journal},
\emph{26}(2), 115--119. \url{https://doi.org/10.1629/2048-7754.65}

\leavevmode\hypertarget{ref-Frantsv_g_2019}{}%
Frantsvåg, J. E., \& Strømme, T. E. (2019). Few open access journals are
compliant with plan s. \emph{Publications}, \emph{7}(2), 26.
\url{https://doi.org/10.3390/publications7020026}

\leavevmode\hypertarget{ref-Frazier_2001}{}%
Frazier, K. (2001). The librarians' dilemma: Contemplating the costs of
the "big deal". \emph{D-Lib Magazine}, \emph{7}(3).
\url{https://web.archive.org/web/20210125073550/https://librarytechnology.org/document/8950}

\leavevmode\hypertarget{ref-Geschuhn_2017}{}%
Geschuhn, K., \& Stone, G. (2017). It's the workflows, stupid! What is
required to make ``offsetting'' work for the open access transition.
\emph{Insights the UKSG Journal}, \emph{30}(3), 103--114.
\url{https://doi.org/10.1629/uksg.391}

\leavevmode\hypertarget{ref-Graaf_2017}{}%
Graaf, M. V. D. (2017). \emph{Paying for open access: The author's
perspective}. Zenodo. \url{https://doi.org/10.5281/ZENODO.438037}

\leavevmode\hypertarget{ref-Hendricks_2020}{}%
Hendricks, G., Tkaczyk, D., Lin, J., \& Feeney, P. (2020). Crossref: The
sustainable source of community-owned scholarly metadata.
\emph{Quantitative Science Studies}, \emph{1}(1), 414--427.
\url{https://doi.org/10.1162/qss_a_00022}

\leavevmode\hypertarget{ref-Hinchliffe_2019}{}%
Hinchliffe, L. J. (2019). \emph{Transformative agreements: A primer}.
\url{https://web.archive.org/web/20210128170342/https://scholarlykitchen.sspnet.org/2019/04/23/transformative-agreements/};
The Scholarly Kitchen.

\leavevmode\hypertarget{ref-Jahn_2016}{}%
Jahn, N., \& Tullney, M. (2016). A study of institutional spending on
open access publication fees in germany. \emph{PeerJ}, \emph{4}, e2323.
\url{https://doi.org/10.7717/peerj.2323}

\leavevmode\hypertarget{ref-Kirkman_2018}{}%
Kirkman, N. S. (2018). \emph{A study of open access publishing by nhmrc
grant recipients} {[}Curtin University{]}.
\url{http://hdl.handle.net/20.500.11937/77026}

\leavevmode\hypertarget{ref-Laakso_2016}{}%
Laakso, M., \& Björk, B.-C. (2016). Hybrid open access--a longitudinal
study. \emph{Journal of Informetrics}, \emph{10}(4), 919--932.
\url{https://doi.org/10.1016/j.joi.2016.08.002}

\leavevmode\hypertarget{ref-Larivi_re_2015}{}%
Larivière, V., Haustein, S., \& Mongeon, P. (2015). The oligopoly of
academic publishers in the digital era. \emph{PLOS ONE}, \emph{10}(6),
e0127502. \url{https://doi.org/10.1371/journal.pone.0127502}

\leavevmode\hypertarget{ref-Larivi_re_2018}{}%
Larivière, V., \& Sugimoto, C. R. (2018). Do authors comply when funders
enforce open access to research? \emph{Nature}, \emph{562}(7728),
483--486. \url{https://doi.org/10.1038/d41586-018-07101-w}

\leavevmode\hypertarget{ref-Lawson_2018}{}%
Lawson, S. (2017). \emph{Report on offset agreements: Evaluating current
jisc collections deals. Year 2 -- evaluating 2016 deals}. figshare.
\url{https://doi.org/10.6084/M9.FIGSHARE.5383861.V1}

\leavevmode\hypertarget{ref-Lawson_2015}{}%
Lawson, S. (2015). "Total cost of ownership" of scholarly communication:
Managing subscription and APC payments together. \emph{Learned
Publishing}, \emph{28}(1), 9--13. \url{https://doi.org/10.1087/20150103}

\leavevmode\hypertarget{ref-Lawson_2016}{}%
Lawson, S., Gray, J., \& Mauri, M. (2016). Opening the black box of
scholarly communication funding: A public data infrastructure for
financial flows in academic publishing. \emph{Open Library of
Humanities}, \emph{2}(1). \url{https://doi.org/10.16995/olh.72}

\leavevmode\hypertarget{ref-van_Leeuwen_2018}{}%
Leeuwen, T. N. van, Tatum, C., \& Wouters, P. F. (2018). Exploring
possibilities to use bibliometric data to monitor gold open access
publishing at the national level. \emph{Journal of the Association for
Information Science and Technology}, \emph{69}(9), 1161--1173.
\url{https://doi.org/10.1002/asi.24029}

\leavevmode\hypertarget{ref-Marques_2019}{}%
Marques, M., Woutersen-Windhouwer, S., \& Tuuliniemi, A. (2019).
Monitoring agreements with open access elements: Why article-level
metadata are important. \emph{Insights the UKSG Journal}, \emph{32}.
\url{https://doi.org/10.1629/uksg.489}

\leavevmode\hypertarget{ref-Mart_n_Mart_n_2018}{}%
Martín-Martín, A., Costas, R., Leeuwen, T. van, \& López-Cózar, E. D.
(2018). Evidence of open access of scientific publications in google
scholar: A large-scale analysis. \emph{Journal of Informetrics},
\emph{12}(3), 819--841. \url{https://doi.org/10.1016/j.joi.2018.06.012}

\leavevmode\hypertarget{ref-Mittermaier_2015}{}%
Mittermaier, B. (2015). Double dipping in hybrid open access -- chimera
or reality? \emph{ScienceOpen Research}.
\url{https://doi.org/10.14293/s2199-1006.1.sor-socsci.aowntu.v1}

\leavevmode\hypertarget{ref-Monaghan_2020}{}%
Monaghan, J., Lucraft, M., \& Allin, K. (2020). \emph{'APCs in the
wild': Could increased monitoring and consolidation of funding
accelerate the transition to open access?} figshare.
\url{https://doi.org/10.6084/M9.FIGSHARE.11988123.V4}

\leavevmode\hypertarget{ref-Nelson_2017}{}%
Nelson, G. M., \& Eggett, D. L. (2017). Citations, mandates, and money:
Author motivations to publish in chemistry hybrid open access journals.
\emph{Journal of the Association for Information Science and
Technology}, \emph{68}(10), 2501--2510.
\url{https://doi.org/10.1002/asi.23897}

\leavevmode\hypertarget{ref-Frascati}{}%
OECD. (2015). \emph{Frascati manual 2015: Guidelines for collecting and
reporting data on research and experimental development}. OECD
Publishing. \url{https://doi.org/10.1787/24132764}

\leavevmode\hypertarget{ref-Olsson1271866}{}%
Olsson, L. (2018). \emph{Evaluation of offset agreements -- report 4 :
Springer compact} (p. 20). Kungliga biblioteket.
\url{http://urn.kb.se/resolve?urn=urn\%3Anbn\%3Ase\%3Ahb\%3Adiva-15501}

\leavevmode\hypertarget{ref-Pinfield_2016}{}%
Pinfield, S., Salter, J., \& Bath, P. A. (2016). The "total cost of
publication" in a hybrid open-access environment: Institutional
approaches to funding journal article-processing charges in combination
with subscriptions. \emph{Journal of the Association for Information
Science and Technology}, \emph{67}(7), 1751--1766.
\url{https://doi.org/10.1002/asi.23446}

\leavevmode\hypertarget{ref-Piwowar_2018}{}%
Piwowar, H., Priem, J., Larivière, V., Alperin, J. P., Matthias, L.,
Norlander, B., Farley, A., West, J., \& Haustein, S. (2018). The state
of OA: A large-scale analysis of the prevalence and impact of open
access articles. \emph{PeerJ}, \emph{6}, e4375.
\url{https://doi.org/10.7717/peerj.4375}

\leavevmode\hypertarget{ref-Piwowar_2019}{}%
Piwowar, H., Priem, J., \& Orr, R. (2019). \emph{The future of OA: A
large-scale analysis projecting open access publication and readership}.
bioRxiv. \url{https://doi.org/10.1101/795310}

\leavevmode\hypertarget{ref-P_l_nen_2020}{}%
Pölönen, J., Laakso, M., Guns, R., Kulczycki, E., \& Sivertsen, G.
(2020). Open access at the national level: A comprehensive analysis of
publications by finnish researchers. \emph{Quantitative Science
Studies}, \emph{1}(4), 1396--1428.
\url{https://doi.org/10.1162/qss_a_00084}

\leavevmode\hypertarget{ref-Prosser_2015}{}%
Prosser, D. (2015). \emph{The costs of double dipping}.
\url{https://web.archive.org/web/20210128164930/https://www.rluk.ac.uk/the-costs-of-double-dipping/}.

\leavevmode\hypertarget{ref-Prosser_2003}{}%
Prosser, D. C. (2003). From here to there: A proposed mechanism for
transforming journals from closed to open access. \emph{Learned
Publishing}, \emph{16}(3), 163--166.
\url{https://doi.org/10.1087/095315103322110923}

\leavevmode\hypertarget{ref-r}{}%
R Core Team. (2020). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\leavevmode\hypertarget{ref-Shieber_2009}{}%
Shieber, S. M. (2009). Equity for open-access journal publishing.
\emph{PLoS Biology}, \emph{7}(8), e1000165.
\url{https://doi.org/10.1371/journal.pbio.1000165}

\leavevmode\hypertarget{ref-Unpaywall_para}{}%
Unpaywall. (n.d.-a). \emph{What does is\_paratext mean in the API?}
\url{https://web.archive.org/web/20201126131640/https://support.unpaywall.org/support/solutions/articles/44001894783}.

\leavevmode\hypertarget{ref-Unpaywall_types}{}%
Unpaywall. (n.d.-b). \emph{What do the types of oa\_status (green, gold,
hybrid, and bronze) mean?}
\url{http://web.archive.org/web/20210127192910/https://support.unpaywall.org/support/solutions/articles/44001777288}.

\leavevmode\hypertarget{ref-Unpaywall_oa_license}{}%
Unpaywall. (n.d.-c). \emph{What is an OA license?}
\url{http://web.archive.org/web/20210127192625/https://support.unpaywall.org/support/solutions/articles/44002063718-what-is-an-oa-license-}.

\leavevmode\hypertarget{ref-tidyverse}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
Welcome to the tidyverse. \emph{Journal of Open Source Software},
\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}

\leavevmode\hypertarget{ref-bigrquery}{}%
Wickham, H., \& Bryan, J. (2020). \emph{bigrquery: An interface to
Google's BigQuery API}.
\url{https://CRAN.R-project.org/package=bigrquery}

\leavevmode\hypertarget{ref-xml2}{}%
Wickham, H., Hester, J., \& Ooms, J. (2020). \emph{xml2: Parse XML}.
\url{https://CRAN.R-project.org/package=xml2}
\end{cslreferences}

\end{document}

### 2021-01-29 analysis/manuscript.Rmd
version pushed to overleaf for final polishing

@@
 Many large subscription publishers have since incorporated this OA model, and while the uptake among authors increased only slowly during the first decade [@Bj_rk_2012], it has gradually gained popularity in recent years.
-This development was driven by the implementation of OA policies by research institutions and funders, the allocation of OA funds, and formal agreements between publishers and institutions that allow affiliated authors to publish OA free-of-charge.
+This development was driven by the implementation of OA policies by research institutions and funders, the allocation of OA funds, and formal agreements between publishers and institutions that allow affiliated authors to publish OA free-of-charge [@Laakso_2016].
 However, there have been concerns that the hybrid model allows for double-dipping—receiving two payments for one article, the APC and subscription fees [@Shieber_2009; @Prosser_2015].
@@
 With the recent introduction of transformative agreements, an evolving term for contracts that shift library spendings for subscriptions to OA [@Borrego_2020; @Hinchliffe_2019], the demand for publisher-provided data has even increased.
-If publishers would enable transparency about the OA uptake in hybrid journals and APC invoicing, academic libraries could improve their assessments and adjustments of contracts with publishers [@antelmann_2017].
+If publishers would enable transparency about the OA uptake in hybrid journals and APC invoicing, academic libraries could improve their assessments and adjustments of contracts with publishers [@antelmann_2017; @Schimmer_2015].
 Likewise, research organizations and funders could improve compliance monitoring with OA mandates, while ensuring that the different actors were not charged multiple times [@Larivi_re_2018].
@@
 
-This current study investigates the extent to which publisher-provided data can be utilized to assess hybrid OA.
+This current study investigates the extent to which publisher-provided data can be utilized to research hybrid OA.
 The study takes Elsevier, the largest scholarly journal publisher in terms of article volume [@Larivi_re_2015], as the starting point for an empirical analysis of hybrid OA uptake and invoicing.
-Our approach is based on openly available data sources and brings together scholarly data from Crossref and machine-readable APC invoicing data from Elsevier OA full-texts, which enables us to analyze the development of Elsevier’s hybrid journal portfolio from 2015-19, and to compare it with that of Elsevier’s full and delayed OA program.
-Further, we investigate whether hybrid publishing fees were waived or charged to the authors or academic institutions or research funders with Elsevier OA agreements.
+Our approach is based on openly available data sources and brings together scholarly metadata from Crossref, a DOI registration agency for scholarly works, and machine-readable APC invoicing data from Elsevier OA full-texts.
+The so-compiled data enable us to analyze the development of Elsevier’s hybrid journal portfolio from 2015-19, and to compare it with that of Elsevier’s full and delayed OA program.
+Further, we investigate whether hybrid publishing fees were waived or charged.
 In particular, we identify academic consortia and research funders that were directly invoiced by Elsevier, and compare their hybrid OA rates by license and subject.
@@
 Based on a random sample of 100,000 Crossref DOIs, the authors found 3.6% of hybrid OA, while pointing towards a growing trend in recent years that recorded a hybrid OA share of 9.4% of all articles published in 2015.
-While Piwowar et al.'s (2018) and Martín-Martín et al.'s (2018) methods are well-suited for large-scale and repeated analyses of the prevalence of hybrid OA, their main limitation is that they lack information about publishing costs.
-
-Moreover, Unpaywall's hybrid OA detection method has been refined several times since its launch in 2017, reflecting growing albeit incomprehensive license information in Crossref metadata.
+
+Since its launch, Unpaywall has become widely used as an OA evidence source in bibliometric research and rankings[@Robinson_Garcia_2020; @Huang_2020].
+However, results cannot easily be compared with each other over time because of Unpaywall's ongoing changes due to non-standardized data to discover hybrid OA.
+These changes mainly reflect the growing albeit incomprehensive license information in Crossref metadata.
 To overcome these shortcomings, the developers extended previous web-scraping approaches to better distinguish between immediate and delayed OA based on licenses at the date of observation @Piwowar_2019.
 Relatedly, in October 2020, Elsevier's license for delayed OA content was added to Unpaywall's list of license patterns [@Unpaywall_oa_license].
-Finally, at the journal-level, a more fine-grained identification of full OA journals using OA article information was developed to better distinguish them from subscription journals including hybrid journals [@Unpaywall_types].
-Because of these ongoing changes due to non-standardized data, uptake studies reporting hybrid OA uptake based on Unpaywall data cannot easily be compared with each other.
+At the journal-level, a more fine-grained identification of full OA journals using OA article information was developed to better distinguish them from subscription journals including hybrid journals [@Unpaywall_types].
+With its servie unsub, Unpaywall recently launched a paid service, where libraries can compare hybrid OA uptake rates with internal financial and usage data to assess the value of journal subscriptions [@Chawla_2020].
 
@@
 The advantages of Open APC data are that they reflect the actual APCs paid rather than list prices; they are not limited to certain publishers or journals; they are provided in a standardized format regardless of the publisher.
-On the other hand, the drawbacks include the limited number of participating institutions that are almost exclusively based in Europe and the dependency on participating organizations for data integrity.
+On the other hand, the drawbacks include the limited number of participating institutions that are almost exclusively based in Europe and the dependency on participating organizations for data integrity [@Pieper_2018].
 Moreover, since Open APC data does not include waivers or free APCs, it cannot present a complete picture of all publishing activities and expenses.
@@
 
-## Data-driven workflows for hybrid OA
-
-In September 2018, coalition S presented a set of principles known as Plan S and announced.
-that from 2021 onwards, signatories would no longer fund publications in hybrid OA journals.
-This principle includes so-called "mirror" and sister journals—hybrid journals with fully OA counterparts or split into their fully OA and subscription parts—that have launched since Plan S was announced but makes an exception for hybrid journals with transformative agreements [@Frantsv_g_2019].
-While the scale and means of these transformative agreements may differ, many address the lack of transparency surrounding hybrid OA by making publisher-provided metadata a defining element of publishing workflows between institutions and publishers [@Geschuhn_2017; @Marques_2019]
-
-Elsevier, one of the world's largest scholarly publishers, provides article-level metadata to Crossref, including license information, which can be used to identify OA articles, and links to HTML and XML full-text versions for text and data mining purposes.
+## The role of publisher-provided data
+
+Uptake and financial studies have OA policy implications for hybrid OA.
+Most prominently, the cOAlition S, an international network of research funders, released its widely discussed Plan S.
+In September 2018, the cOAlition S announced that from 2021 onwards, signatories would no longer fund articles in hybrid OA journals.
+This principle includes so-called "mirror" and sister journals--hybrid journals with fully OA counterparts or split into their fully OA and subscription parts—that have launched since Plan S was announced but makes an exception for hybrid journals with transformative agreements [@Frantsv_g_2019].
+The Plan S refers to the Efficiency and Standards for Article Charges (ESAC) guidelines [@Geschuhn_2017] as critical component of transformative agreements, emphasizing the need for publisher-provided metadata to address the lack of transparency surrounding hybrid OA.
+However, studies confirm that publisher with such agreements lack essential data about hybrid OA uptake and invoicing, or are not able or willing to provide such information publicly and instantaneously [@Monaghan_2020; @Marques_2019], which, in turn, not just limit the monitoring of hybrid OA, but also research studies on its size and financial implications.
+
+Here, we focus on Elsevier.
+The publisher provides article-level metadata to Crossref, including license information, which can be used to identify OA articles, and links to HTML and XML full-text versions for text and data mining purposes.
 Surprisingly, besides the full-text article, the XML files also contain invoicing information and the article's OA status.
@@
 \begin{table}[H]
-\caption{\label{tab:els_xml}Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.}
+\caption{\label{tab:els_xml} Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019}
 \centering
@@
 \end{table}
-
-Table 1: Overview of metadata provided in XML-based Elsevier full-texts representing information about open access status and invoicing.
-A sample XML is available here: https://api.elsevier.com/content/article/PII:S0169409X18301479?httpAccept=text/xml
 
@@
 \end{figure}
+
+# Discussion
+
+Over the last five years, Elsevier recorded growth in the uptake of hybrid OA for each of our three indicators: The number of hybrid OA articles published per year doubled, the number of hybrid journals with at least one OA article grew by 21%, and the share of hybrid OA articles relative to closed-accessed articles in these journals increased from 2.6% to 3.7%.
+As @Laakso_2016, we found only a weak relationship between citation metrics like the SNIP and hybrid OA uptake and observed differences in the uptake of hybrid OA among journal subject areas.
+In particular, we found the highest count of hybrid OA articles in physical sciences journals (see Table \ref{tab:subject_area_table}).
+This was followed by the life sciences and health sciences, whereas the social sciences had the lowest count of hybrid articles.
+This order mostly reflects the disciplines' overall publication output.
+According to the Open Science Monitor by the European Commission, the physical sciences and mathematics publish the most articles, followed by the health sciences, life sciences, and social sciences and humanities.
+
+Such disciplinary differences in hybrid OA prevalence become more meaningful when considering the relative share of OA articles to closed-access articles.
+In line with previous research, we found that Elsevier journals from the life sciences and social sciences [@Jubb_2017; @Kramer_2018; @Laakso_2016] recorded greater than typical hybrid OA uptake, whereas physical sciences journals generally had a lower than typical uptake [@Kramer_2018; @Mart_n_Mart_n_2018; @Laakso_2016].
+In a systematic review of OA publishing patterns, @Severin_2020 linked such disciplinary differences to socio-cultural and technological factors that shape publishing cultures and practices.
+Among other aspects, the authors pointed to the availability of prestigious journals, the possibility to provide OA through alternative routes, and ease of access to funding sources to cover publishing fees.
+As such, looking at hybrid OA uptake in isolation can only provide limited insight, so that placing it within the broader context of disciplinary publishing and OA practices is necessary.
+For instance, since many branches of the physical sciences have established self-archiving practices and the technical infrastructure to support these (e.g., the arXiv; @Severin_2020, @Bj_rk_2010, @Laakso_2016), researchers can provide OA through repositories (i.e., green OA), and indeed, according to the Open Science Monitor the physical sciences provide OA primarily through the green route (31.2% total OA share, 24% OA if only considering green OA).
+Hence, the low share of hybrid OA in our sample might be because physical scientists perceive much less of a need for hybrid OA.
+Further, @Severin_2020 note project-based funding structures enable OA publishing in fully OA and hybrid journals as they allow for easy integration of publishing costs.
+Since these funding structures are typical for the life sciences (ibid.), we would expect higher rates of OA.
+Data from the Open Science Monitor seems to support this notion as the life sciences record particularly high shares of full OA articles and comparatively high  shares of hybrid OA in some fields, such as biology and agricultural science.
+While we expected the observed hybrid OA shares for the life sciences and physical sciences, we were surprised by the high share of hybrid OA belonging to the AJSC subject area social sciences, which also includes the humanities, since articles "social scientists have reported to face significant difficulties in access to grant funding for [publication charges], as most research in the social sciences is not done by means of project-specific funding" @Severin_2020 [p.
+15].
+However, previous research has found comparatively low awareness and uptake of green OA among researchers in the  field [@Bj_rk_2010; @Creaser_2010], which might encourage hybrid OA uptake.
+Another possible explanation,is that Elsevier has a number of high prestigious social science and humanities journals [@Larivi_re_2015], which means that our findings are not representative of the general population of SSH journals.
+Likewise, this limitation could be extended to our findings in general that whilst not representative of the population of hybrid OA publishing, they provide valuable insight into a key player of scholarly publishing.
+
+Finally, OA policies in favor of hybrid OA in combination with invoicing agreements with Elsevier might be a driver.
+For instance, we found that nearly half of articles in arts and humanities were facilitated through agreements, which was above-average.
+This is in line with previous research, suggesting that OA policies and funder mandates have a stronger effect on OA uptake than disciplinary publication cultures [@Huang_2020; @Larivi_re_2018].
+
+Our overall investigation of invoice channels and recipients of hybrid OA APCs found that the majority of Elsevier hybrid articles were invoiced to the author (n=41,725; 58.2%).
+However, it is important to note that although Elsevier's metadata classifies the sponsor type as "author", this does not necessarily mean that authors paid for APCs out of their own pocket but rather just that the APC was not invoiced through an agreement.
+Hence, it is possible that APCs invoiced to the author were covered by institutional funds or as part of research grants.
+For instance, a recent Springer Nature survey found that hybrid OA was predominantly supported through institutional and funder sources (71%), followed by OA agreements (34%), while only 6% were paid from personal funds or savings [@Monaghan_2020].
+We also observed notable differences in licensing among invoice types.
+The majority of hybrid OA articles, regardless of the invoice type, were licensed under the more restrictive CC BY-NC-ND license.
+There is a lack of dedicated studies on licence selection, but indications so far suggest that when given a choice authors tend to select more restrictive variants for their published works [@Fraser_2020; @Rowley_2017; @Van_Noorden_2013].
+On the other hand, when hybrid articles were invoiced through agreements, most articles were licensed under the more liberal CC BY license.
+As several funding bodies mandate CC BY licenses, including the UK research funders that account for 62% of our "agreement" subsample, this result is perhaps not surprising but suggests the effectiveness of such agreements.
+
+Based on our findings, around a third of the articles were invoiced directly through OA agreements (e.g., research funders, national library consortia).
+Strikingly, sponsors based in the UK accounted for more than half of these , and a quarter was charged to sponsors from other European countries.
+This finding reiterates reports from previous studies that the UK's OA profile differs markedly from other countries (cf.
+ref, ref) and points to the effects of science policy.
+To increase OA to research publications, the UK implemented centralized APC funding and embraced hybrid OA as a transition model.
+Within five years of the publication of the Finch Report in 2012, the UK recorded an 18% increase in immediate OA, coupled with a rise in hybrid OA from 2.7% in 2012 to 15.4% of all articles in 2016 [@Jubb_2017], which can be attributed to the availability of APC funding from Research Councils UK (RCUK).
+Indeed, RCUK block grants have been the largest single source of OA funds in the UK—the Wellcome Trust, and more recently transformative agreements [@Tickel_2018; @Jubb_2017].
+However, since then publishing expenditures of UK universities have been rising rapidly (ref), so that several universities stopped supporting hybrid OA in late 2018-2019 [@oxford_2019; @walker_2019; @Castro_2019; @birmingham].
+This development might explain the slight decrease in hybrid articles invoiced to the UK our data showed around that time (see Figure \ref{fig:invoice_sponsor_country}).
+
+An essential part in the absence of transparency in scholarly publishing and especially in hybrid OA has been the lack of transparency in publishers’ article management.
+This has prohibited consumers to get a comprehensive and comparative overview of their publishing activities and the associated costs and, thus, limited their negotiating power.
+However, certain initiatives in Europe like the ESAC Initiative have begun to press for improved and more transparent workflows for transformative agreements [@Geschuhn_2017].
+In particular, the ESAC Workflow Recommendations seek to increase efficiency and transparency in the publication workflow through improved invoicing and reporting processes and metadata delivery, including article-based machine readable information about the OA sponsor.
+Through this study we demonstrated the utility and benefits of such  publisher-provided metadata about hybrid OA funding and highlighted the need for extending these to include comprehensive information about the licensing and the particular OA type.
+
+Usually identifying delayed OA articles and distinguishing them from other types of OA is challenging since there is no publicly available list of such journals.
+For instance, Taylor & Francis’ (2021) website indicates that some of their journals participate in a delayed or promotional OA program but no further information is provided as to which journals these are.
+However, Elsevier provides a list of journals enrolled in the "Open Archive" programme and further tags delayed OA articles in the machine-readable XML version of the article.
+This enables easy identification and more accurate assessments of hybrid OA, which allows to take into account the substantial amount of delayed OA content in future hybrid OA studies using Elsevier data.
+
+While this study advances our knowledge about hybrid OA uptake and invoicing, the limitations leave room for future research.
+This study focused on only one publisher, which limits the generalizability of our findings since Elsevier’s attributes such as composition of the journal portfolio, mix of business models, pricing, and promotion of various options cannot be assumed to be representative of scholarly journal publishing in general.
+Though the invoicing data extracted from the XML files enables unprecedented accuracy and comprehensiveness in the study of hybrid OA, it likely does not disclose the actual OA sponsor in all cases.
+Most articles were invoiced to authors, but it remains unclear how often authors themselves pay for APCs or invoices are covered through institutional OA funds or research grants.
+Further research into this topic would increase our understanding of OA funding streams outside of publishing agreements.
+Further, our study demonstrates that to achieve comprehensive mapping and understanding of the financial flows involved in OA funding, complex, in-depth country-specific meso-level analyses is needed.
+Such studies would ideally draw on various data sources, such as publicly available bibliometrics and metadata, information about significant national research funders and their OA policies, information about the national library and possible consortia landscape including their publishing agreements, which might require Freedom of Information Requests.
+
+This study provides a snapshot of hybrid OA for the largest journal publisher prior to the impact of the implementation of Plan S requirements.
+While many Plan S signatories have already had strong OA policies, this harmonized approach is likely to affect publishing decisions of funded authors,the licencing of their articles, and the offerings and pricing of publishers.
+Since the new requirements are applicable to research funded from 2021 onwards, a comparative study focused on articles  invoiced to Plan S signatories would be a fruitful endeavor.
+
+In a similar vein, future studies could  combine invoice information with research funding information since these contribute different pieces to the funding stream puzzle.
+This could also provide more insight into the actual APC funding streams behind the author-invoiced articles.
+
+# Conclusion
+
+The primary aim of this empirical study was to investigate Elsevier’s hybrid OA publishing from 2015-2019 to better understand the volume and invoicing of hybrid OA and present a novel, data-driven approach for such analyses.
+Our results indicate that although the number of hybrid OA articles has increased over time, hybrid OA uptake has remained low.
+Notably, hybrid APCs were most often invoiced directly to the authors or through institutional agreements, where only a few funding bodies were the main driver of hybrid OA.
+Finally, our findings highlight that publisher-provided metadata about the invoicing channels  of (hybrid) OA can facilitate research into and increase the understanding of the financial flows of OA publishing.
+
+Ever since its inception, hybrid OA has been a challenging subject to study due to the lack of standardised ways for publishers to flag such content and APC funding data being limited to studies based on surveys and articles’ acknowledgements sections.
+This study presented a novel approach to studying OA APC invoicing  that is based on publicly available publisher-provided metadata, which can be used on on its own or in combination with other public data sources to gain more detailed and comprehensive insights into hybrid OA uptake and invoicing than ever before.
+If more publishers reported OA invoicing data on the article level and in a machine-readable format, this would increase transparency and enable  enhanced monitoring of the scholarly journal landscape over time.
+As hybrid OA has increasingly become a central element of OA policies of research funders and libraries,  customer organisations could require that invoicing information is added to the article-level metadata.
+As long as publishers do not provide this data in a structured and comprehensive format, they prevent consumers from benchmarking prices and therefore hinder price competition as the only alternative to gain comparative insight is self-reported APC expenditures that research funders and libraries voluntarily report to the Open APC initiative.
+The Open APC initiative is a valuable service but it cannot replace publisher-provided data since global participation is limited.
+
+From recent science policy developments in Europe it appears that Big Deals have gained support and remain firmly in place in the form of transformative agreements.
+However, in addition to research funders and libraries that increasingly make publishing contracts public, it is on the publisher to provide  metadata as presented here to improve the transparency in scholarly publishing by linking invoicing data to bibliometrics.
+From this study we can affirm that hybrid OA is complex as the financial flows come from various stakeholders and invoicing forms, it is not just libraries or consortia, but authors and funders are also directly involved.
+As of right now, providing  machine-readable article-level invoicing data is not common amongst publishers, which limits the potential for broader studies.
+
+# Data Availability
+
+All data and code used for this study is openly available on GitHub, including version history:
+
+<https://github.com/njahn82/elsevier_hybrid_invoicing>
+
+Formal analyis is written in R Markdown and thus replicable.
+
+# References

### 2021-01-29 analysis/manuscript.tex
version pushed to overleaf for final polishing

@@
 
-\title{Your APA6-Style Manuscript}
-\shorttitle{}
-\author{You}
-\affiliation{Somewhere}
-
-\abstract{Your abstract here.}
+\title{Transparency to hybrid open access invoicing through publisher-provided metadata: A longitudinal article-level study of Elsevier}
+\shorttitle{Hybrid OA Elsevier}
+% \author{Najko Jahn}
+% \affiliation{State and University Library Göttingen, University of Göttingen, Germany}
+% \author{Lisa Matthias}
+% \affiliation{tba}
+% \author{Mikael Laakso}
+% \affiliation{tba}
+%
+
+\abstract{The financial flows in scholarly journal publishing have become increasingly complex with the growth of open access (OA), and the lack of data and transparency into these flows has been persistent.
+Opaqueness has concerned especially hybrid OA, where subscription-based journals also publish OA content if an optional fee is paid.
+By leveraging Elsevier’s recently introduced article metadata concerning invoiced recipients, this study provides the first longitudinal publisher-level study of uptake and invoicing of hybrid OA articles.
+The results show that the share of OA articles in Elseviers hybrid journals has grown slowly, from 2.6\% of all articles on average in 2015 to 3.7\% in 2019.
+About a third of hybrid OA articles are invoiced directly through an agreement with research funders or other type of organisational sponsor (e.g.
+national library consortia), with the majority invoiced to authors.
+83\% of agreement-invoiced hybrid OA articles are to European organisations, suggesting a strong international skewness.
+The study discovered a large discrepancy in choice of publication license between author-invoiced articles (dominantly CC BY-NC-ND) and agreement-invoiced articles (dominantly CC BY).
+The results demonstrate the value of publisher-provided metadata covering invoiced recipients of articles, which would be even more useful through wider adoption among publishers.}
 
 \begin{document}
+\maketitle
 
@@
 formal agreements between publishers and institutions that allow
-affiliated authors to publish OA free-of-charge. However, there have
-been concerns that the hybrid model allows for
+affiliated authors to publish OA free-of-charge (Laakso \& Björk, 2016).
+However, there have been concerns that the hybrid model allows for
 double-dipping---receiving two payments for one article, the APC and
@@
 With the recent introduction of transformative agreements, an evolving
-term for contracts to shift library spendings for subscriptions to OA
+term for contracts that shift library spendings for subscriptions to OA
 (Borrego et al., 2020; Hinchliffe, 2019), the demand for
@@
 academic libraries could improve their assessments and adjustments of
-contracts with publishers (Antelman, 2017). Likewise, research
-organizations and funders could improve compliance monitoring with OA
-mandates, while ensuring that the different actors were not charged
-multiple times (Larivière \& Sugimoto, 2018). However, as formal
-bibliometrics and financial studies have noted, the lack of publicly
-available and standardized data about hybrid OA publishing is
+contracts with publishers (Antelman, 2017; Schimmer et al., 2015).
+Likewise, research organizations and funders could improve compliance
+monitoring with OA mandates, while ensuring that the different actors
+were not charged multiple times (Larivière \& Sugimoto, 2018). However,
+as formal bibliometrics and financial studies have noted, the lack of
+publicly available and standardized data about hybrid OA publishing is
 considerable (Laakso \& Björk, 2016; Lawson, 2015; Pinfield et al.,
@@
 This current study investigates the extent to which publisher-provided
-data can be utilized to assess hybrid OA. The study takes Elsevier, the
-largest scholarly journal publisher in terms of article volume
+data can be utilized to research hybrid OA. The study takes Elsevier,
+the largest scholarly journal publisher in terms of article volume
 (Larivière et al., 2015), as the starting point for an empirical
 analysis of hybrid OA uptake and invoicing. Our approach is based on
-openly available data sources and brings together scholarly data from
-Crossref and machine-readable APC invoicing data from Elsevier OA
-full-texts, which enables us to analyze the development of Elsevier's
+openly available data sources and brings together scholarly metadata
+from Crossref, a DOI registration agency for scholarly works, and
+machine-readable APC invoicing data from Elsevier OA full-texts. The
+so-compiled data enable us to analyze the development of Elsevier's
 hybrid journal portfolio from 2015-19, and to compare it with that of
 Elsevier's full and delayed OA program. Further, we investigate whether
-hybrid publishing fees were waived or charged to the authors or academic
-institutions or research funders with Elsevier OA agreements. In
-particular, we identify academic consortia and research funders that
-were directly invoiced by Elsevier, and compare their hybrid OA rates by
-license and subject. As such, our findings have implications for
-research and OA policy and highlight the potential of publisher-provided
-data for large-scale, instantaneous analysis of hybrid OA uptake and
-invoicing.
+hybrid publishing fees were waived or charged. In particular, we
+identify academic consortia and research funders that were directly
+invoiced by Elsevier, and compare their hybrid OA rates by license and
+subject. As such, our findings have implications for research and OA
+policy and highlight the potential of publisher-provided data for
+large-scale, instantaneous analysis of hybrid OA uptake and invoicing.
 
@@
 while pointing towards a growing trend in recent years that recorded a
-hybrid OA share of 9.4\% of all articles published in 2015. While
-Piwowar et al.'s (2018) and Martín-Martín et al.'s (2018) methods are
-well-suited for large-scale and repeated analyses of the prevalence of
-hybrid OA, their main limitation is that they lack information about
-publishing costs.
-
-Moreover, Unpaywall's hybrid OA detection method has been refined
-several times since its launch in 2017, reflecting growing albeit
-incomprehensive license information in Crossref metadata. To overcome
-these shortcomings, the developers extended previous web-scraping
-approaches to better distinguish between immediate and delayed OA based
-on licenses at the date of observation Piwowar et al. (2019). Relatedly,
-in October 2020, Elsevier's license for delayed OA content was added to
-Unpaywall's list of license patterns (Unpaywall, n.d.-c). Finally, at
-the journal-level, a more fine-grained identification of full OA
-journals using OA article information was developed to better
-distinguish them from subscription journals including hybrid journals
-(Unpaywall, n.d.-b). Because of these ongoing changes due to
-non-standardized data, uptake studies reporting hybrid OA uptake based
-on Unpaywall data cannot easily be compared with each other.
+hybrid OA share of 9.4\% of all articles published in 2015.
+
+Since its launch, Unpaywall has become widely used as an OA evidence
+source in bibliometric research and rankings(Huang et al., 2020;
+Robinson-Garcia et al., 2020). However, results cannot easily be
+compared with each other over time because of Unpaywall's ongoing
+changes due to non-standardized data to discover hybrid OA. These
+changes mainly reflect the growing albeit incomprehensive license
+information in Crossref metadata. To overcome these shortcomings, the
+developers extended previous web-scraping approaches to better
+distinguish between immediate and delayed OA based on licenses at the
+date of observation Piwowar et al. (2019). Relatedly, in October 2020,
+Elsevier's license for delayed OA content was added to Unpaywall's list
+of license patterns (Unpaywall, n.d.-c). At the journal-level, a more
+fine-grained identification of full OA journals using OA article
+information was developed to better distinguish them from subscription
+journals including hybrid journals (Unpaywall, n.d.-b). With its servie
+unsub, Unpaywall recently launched a paid service, where libraries can
+compare hybrid OA uptake rates with internal financial and usage data to
+assess the value of journal subscriptions (Chawla, 2020).
 
@@
 participating institutions that are almost exclusively based in Europe
-and the dependency on participating organizations for data integrity.
-Moreover, since Open APC data does not include waivers or free APCs, it
-cannot present a complete picture of all publishing activities and
-expenses. This, however, would pose less of an issue when analyzing just
-hybrid OA.
-
-\hypertarget{data-driven-workflows-for-hybrid-oa}{%
-\subsection{Data-driven workflows for hybrid
-OA}\label{data-driven-workflows-for-hybrid-oa}}
-
-In September 2018, coalition S presented a set of principles known as
-Plan S and announced. that from 2021 onwards, signatories would no
-longer fund publications in hybrid OA journals. This principle includes
-so-called ``mirror'' and sister journals---hybrid journals with fully OA
+and the dependency on participating organizations for data integrity
+(Pieper \& Broschinski, 2018). Moreover, since Open APC data does not
+include waivers or free APCs, it cannot present a complete picture of
+all publishing activities and expenses. This, however, would pose less
+of an issue when analyzing just hybrid OA.
+
+\hypertarget{the-role-of-publisher-provided-data}{%
+\subsection{The role of publisher-provided
+data}\label{the-role-of-publisher-provided-data}}
+
+Uptake and financial studies have OA policy implications for hybrid OA.
+Most prominently, the cOAlition S, an international network of research
+funders, released its widely discussed Plan S. In September 2018, the
+cOAlition S announced that from 2021 onwards, signatories would no
+longer fund articles in hybrid OA journals. This principle includes
+so-called ``mirror'' and sister journals--hybrid journals with fully OA
 counterparts or split into their fully OA and subscription parts---that
@@
 hybrid journals with transformative agreements (Frantsvåg \& Strømme,
-2019). While the scale and means of these transformative agreements may
-differ, many address the lack of transparency surrounding hybrid OA by
-making publisher-provided metadata a defining element of publishing
-workflows between institutions and publishers (Geschuhn \& Stone, 2017;
-Marques et al., 2019)
-
-Elsevier, one of the world's largest scholarly publishers, provides
-article-level metadata to Crossref, including license information, which
-can be used to identify OA articles, and links to HTML and XML full-text
-versions for text and data mining purposes. Surprisingly, besides the
-full-text article, the XML files also contain invoicing information and
-the article's OA status. Table \ref{tab:els_xml} summarizes the key
-metadata elements, which comprise the type of entity to which the APC
-was invoiced (\texttt{openaccessSponsorType}) and the name of that
-entity if available (\texttt{openaccessSponsorName}). Moreover, the
-article-level metadata indicates if OA was provided through Elsevier's
-Open Archive program (\texttt{openArchiveArticle}), which enables free
-access to participating subscription journals after an embargo period
-(i.e., delayed OA; Elsevier (n.d.-b)). As we have been unable to
-identify similar efforts by other large publishers (e.g., Springer
-Nature), this increased transparency through openly available
-bibliometric and financial data at the article level seems unique to
-Elsevier.
+2019). The Plan S refers to the Efficiency and Standards for Article
+Charges (ESAC) guidelines (Geschuhn \& Stone, 2017) as critical
+component of transformative agreements, emphasizing the need for
+publisher-provided metadata to address the lack of transparency
+surrounding hybrid OA. However, studies confirm that publisher with such
+agreements lack essential data about hybrid OA uptake and invoicing, or
+are not able or willing to provide such information publicly and
+instantaneously (Marques et al., 2019; Monaghan et al., 2020), which, in
+turn, not just limit the monitoring of hybrid OA, but also research
+studies on its size and financial implications.
+
+Here, we focus on Elsevier. The publisher provides article-level
+metadata to Crossref, including license information, which can be used
+to identify OA articles, and links to HTML and XML full-text versions
+for text and data mining purposes. Surprisingly, besides the full-text
+article, the XML files also contain invoicing information and the
+article's OA status. Table \ref{tab:els_xml} summarizes the key metadata
+elements, which comprise the type of entity to which the APC was
+invoiced (\texttt{openaccessSponsorType}) and the name of that entity if
+available (\texttt{openaccessSponsorName}). Moreover, the article-level
+metadata indicates if OA was provided through Elsevier's Open Archive
+program (\texttt{openArchiveArticle}), which enables free access to
+participating subscription journals after an embargo period (i.e.,
+delayed OA; Elsevier (n.d.-b)). As we have been unable to identify
+similar efforts by other large publishers (e.g., Springer Nature), this
+increased transparency through openly available bibliometric and
+financial data at the article level seems unique to Elsevier.
 
 \begin{table}[H]
-\caption{\label{tab:els_xml}Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.}
+\caption{\label{tab:els_xml} Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019}
 \centering
@@
 \end{table}
-
-Table 1: Overview of metadata provided in XML-based Elsevier full-texts
-representing information about open access status and invoicing. A
-sample XML is available here:
-\url{https://api.elsevier.com/content/article/PII:S0169409X18301479?httpAccept=text/xml}
 
@@
 
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+Over the last five years, Elsevier recorded growth in the uptake of
+hybrid OA for each of our three indicators: The number of hybrid OA
+articles published per year doubled, the number of hybrid journals with
+at least one OA article grew by 21\%, and the share of hybrid OA
+articles relative to closed-accessed articles in these journals
+increased from 2.6\% to 3.7\%. As Laakso \& Björk (2016), we found only
+a weak relationship between citation metrics like the SNIP and hybrid OA
+uptake and observed differences in the uptake of hybrid OA among journal
+subject areas. In particular, we found the highest count of hybrid OA
+articles in physical sciences journals (see Table
+\ref{tab:subject_area_table}). This was followed by the life sciences
+and health sciences, whereas the social sciences had the lowest count of
+hybrid articles. This order mostly reflects the disciplines' overall
+publication output. According to the Open Science Monitor by the
+European Commission, the physical sciences and mathematics publish the
+most articles, followed by the health sciences, life sciences, and
+social sciences and humanities.
+
+Such disciplinary differences in hybrid OA prevalence become more
+meaningful when considering the relative share of OA articles to
+closed-access articles. In line with previous research, we found that
+Elsevier journals from the life sciences and social sciences (Jubb et
+al., 2017; Kramer \& Bosman, 2018; Laakso \& Björk, 2016) recorded
+greater than typical hybrid OA uptake, whereas physical sciences
+journals generally had a lower than typical uptake (Kramer \& Bosman,
+2018; Laakso \& Björk, 2016; Martín-Martín et al., 2018). In a
+systematic review of OA publishing patterns, Severin et al. (2020)
+linked such disciplinary differences to socio-cultural and technological
+factors that shape publishing cultures and practices. Among other
+aspects, the authors pointed to the availability of prestigious
+journals, the possibility to provide OA through alternative routes, and
+ease of access to funding sources to cover publishing fees. As such,
+looking at hybrid OA uptake in isolation can only provide limited
+insight, so that placing it within the broader context of disciplinary
+publishing and OA practices is necessary. For instance, since many
+branches of the physical sciences have established self-archiving
+practices and the technical infrastructure to support these (e.g., the
+arXiv; Severin et al. (2020), Björk et al. (2010), Laakso \& Björk
+(2016)), researchers can provide OA through repositories (i.e., green
+OA), and indeed, according to the Open Science Monitor the physical
+sciences provide OA primarily through the green route (31.2\% total OA
+share, 24\% OA if only considering green OA). Hence, the low share of
+hybrid OA in our sample might be because physical scientists perceive
+much less of a need for hybrid OA. Further, Severin et al. (2020) note
+project-based funding structures enable OA publishing in fully OA and
+hybrid journals as they allow for easy integration of publishing costs.
+Since these funding structures are typical for the life sciences
+(ibid.), we would expect higher rates of OA. Data from the Open Science
+Monitor seems to support this notion as the life sciences record
+particularly high shares of full OA articles and comparatively high
+shares of hybrid OA in some fields, such as biology and agricultural
+science. While we expected the observed hybrid OA shares for the life
+sciences and physical sciences, we were surprised by the high share of
+hybrid OA belonging to the AJSC subject area social sciences, which also
+includes the humanities, since articles ``social scientists have
+reported to face significant difficulties in access to grant funding for
+{[}publication charges{]}, as most research in the social sciences is
+not done by means of project-specific funding'' Severin et al. (2020, p.
+15). However, previous research has found comparatively low awareness
+and uptake of green OA among researchers in the field (Björk et al.,
+2010; Creaser et al., 2010), which might encourage hybrid OA uptake.
+Another possible explanation,is that Elsevier has a number of high
+prestigious social science and humanities journals (Larivière et al.,
+2015), which means that our findings are not representative of the
+general population of SSH journals. Likewise, this limitation could be
+extended to our findings in general that whilst not representative of
+the population of hybrid OA publishing, they provide valuable insight
+into a key player of scholarly publishing.
+
+Finally, OA policies in favor of hybrid OA in combination with invoicing
+agreements with Elsevier might be a driver. For instance, we found that
+nearly half of articles in arts and humanities were facilitated through
+agreements, which was above-average. This is in line with previous
+research, suggesting that OA policies and funder mandates have a
+stronger effect on OA uptake than disciplinary publication cultures
+(Huang et al., 2020; Larivière \& Sugimoto, 2018).
+
+Our overall investigation of invoice channels and recipients of hybrid
+OA APCs found that the majority of Elsevier hybrid articles were
+invoiced to the author (n=41,725; 58.2\%). However, it is important to
+note that although Elsevier's metadata classifies the sponsor type as
+``author'', this does not necessarily mean that authors paid for APCs
+out of their own pocket but rather just that the APC was not invoiced
+through an agreement. Hence, it is possible that APCs invoiced to the
+author were covered by institutional funds or as part of research
+grants. For instance, a recent Springer Nature survey found that hybrid
+OA was predominantly supported through institutional and funder sources
+(71\%), followed by OA agreements (34\%), while only 6\% were paid from
+personal funds or savings (Monaghan et al., 2020). We also observed
+notable differences in licensing among invoice types. The majority of
+hybrid OA articles, regardless of the invoice type, were licensed under
+the more restrictive CC BY-NC-ND license. There is a lack of dedicated
+studies on licence selection, but indications so far suggest that when
+given a choice authors tend to select more restrictive variants for
+their published works (Fraser et al., 2020; Noorden, 2013; Rowley et
+al., 2017). On the other hand, when hybrid articles were invoiced
+through agreements, most articles were licensed under the more liberal
+CC BY license. As several funding bodies mandate CC BY licenses,
+including the UK research funders that account for 62\% of our
+``agreement'' subsample, this result is perhaps not surprising but
+suggests the effectiveness of such agreements.
+
+Based on our findings, around a third of the articles were invoiced
+directly through OA agreements (e.g., research funders, national library
+consortia). Strikingly, sponsors based in the UK accounted for more than
+half of these , and a quarter was charged to sponsors from other
+European countries. This finding reiterates reports from previous
+studies that the UK's OA profile differs markedly from other countries
+(cf.~ref, ref) and points to the effects of science policy. To increase
+OA to research publications, the UK implemented centralized APC funding
+and embraced hybrid OA as a transition model. Within five years of the
+publication of the Finch Report in 2012, the UK recorded an 18\%
+increase in immediate OA, coupled with a rise in hybrid OA from 2.7\% in
+2012 to 15.4\% of all articles in 2016 (Jubb et al., 2017), which can be
+attributed to the availability of APC funding from Research Councils UK
+(RCUK). Indeed, RCUK block grants have been the largest single source of
+OA funds in the UK---the Wellcome Trust, and more recently
+transformative agreements (Jubb et al., 2017; Tickel, 2018). However,
+since then publishing expenditures of UK universities have been rising
+rapidly (ref), so that several universities stopped supporting hybrid OA
+in late 2018-2019 (Castro, 2019; University of Birmingham, n.d.;
+University of Oxford, 2019; Walker, 2019). This development might
+explain the slight decrease in hybrid articles invoiced to the UK our
+data showed around that time (see Figure
+\ref{fig:invoice_sponsor_country}).
+
+An essential part in the absence of transparency in scholarly publishing
+and especially in hybrid OA has been the lack of transparency in
+publishers' article management. This has prohibited consumers to get a
+comprehensive and comparative overview of their publishing activities
+and the associated costs and, thus, limited their negotiating power.
+However, certain initiatives in Europe like the ESAC Initiative have
+begun to press for improved and more transparent workflows for
+transformative agreements (Geschuhn \& Stone, 2017). In particular, the
+ESAC Workflow Recommendations seek to increase efficiency and
+transparency in the publication workflow through improved invoicing and
+reporting processes and metadata delivery, including article-based
+machine readable information about the OA sponsor. Through this study we
+demonstrated the utility and benefits of such publisher-provided
+metadata about hybrid OA funding and highlighted the need for extending
+these to include comprehensive information about the licensing and the
+particular OA type.
+
+Usually identifying delayed OA articles and distinguishing them from
+other types of OA is challenging since there is no publicly available
+list of such journals. For instance, Taylor \& Francis' (2021) website
+indicates that some of their journals participate in a delayed or
+promotional OA program but no further information is provided as to
+which journals these are. However, Elsevier provides a list of journals
+enrolled in the ``Open Archive'' programme and further tags delayed OA
+articles in the machine-readable XML version of the article. This
+enables easy identification and more accurate assessments of hybrid OA,
+which allows to take into account the substantial amount of delayed OA
+content in future hybrid OA studies using Elsevier data.
+
+While this study advances our knowledge about hybrid OA uptake and
+invoicing, the limitations leave room for future research. This study
+focused on only one publisher, which limits the generalizability of our
+findings since Elsevier's attributes such as composition of the journal
+portfolio, mix of business models, pricing, and promotion of various
+options cannot be assumed to be representative of scholarly journal
+publishing in general. Though the invoicing data extracted from the XML
+files enables unprecedented accuracy and comprehensiveness in the study
+of hybrid OA, it likely does not disclose the actual OA sponsor in all
+cases. Most articles were invoiced to authors, but it remains unclear
+how often authors themselves pay for APCs or invoices are covered
+through institutional OA funds or research grants. Further research into
+this topic would increase our understanding of OA funding streams
+outside of publishing agreements. Further, our study demonstrates that
+to achieve comprehensive mapping and understanding of the financial
+flows involved in OA funding, complex, in-depth country-specific
+meso-level analyses is needed. Such studies would ideally draw on
+various data sources, such as publicly available bibliometrics and
+metadata, information about significant national research funders and
+their OA policies, information about the national library and possible
+consortia landscape including their publishing agreements, which might
+require Freedom of Information Requests.
+
+This study provides a snapshot of hybrid OA for the largest journal
+publisher prior to the impact of the implementation of Plan S
+requirements. While many Plan S signatories have already had strong OA
+policies, this harmonized approach is likely to affect publishing
+decisions of funded authors,the licencing of their articles, and the
+offerings and pricing of publishers. Since the new requirements are
+applicable to research funded from 2021 onwards, a comparative study
+focused on articles invoiced to Plan S signatories would be a fruitful
+endeavor.
+
+In a similar vein, future studies could combine invoice information with
+research funding information since these contribute different pieces to
+the funding stream puzzle. This could also provide more insight into the
+actual APC funding streams behind the author-invoiced articles.
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+The primary aim of this empirical study was to investigate Elsevier's
+hybrid OA publishing from 2015-2019 to better understand the volume and
+invoicing of hybrid OA and present a novel, data-driven approach for
+such analyses. Our results indicate that although the number of hybrid
+OA articles has increased over time, hybrid OA uptake has remained low.
+Notably, hybrid APCs were most often invoiced directly to the authors or
+through institutional agreements, where only a few funding bodies were
+the main driver of hybrid OA. Finally, our findings highlight that
+publisher-provided metadata about the invoicing channels of (hybrid) OA
+can facilitate research into and increase the understanding of the
+financial flows of OA publishing.
+
+Ever since its inception, hybrid OA has been a challenging subject to
+study due to the lack of standardised ways for publishers to flag such
+content and APC funding data being limited to studies based on surveys
+and articles' acknowledgements sections. This study presented a novel
+approach to studying OA APC invoicing that is based on publicly
+available publisher-provided metadata, which can be used on on its own
+or in combination with other public data sources to gain more detailed
+and comprehensive insights into hybrid OA uptake and invoicing than ever
+before. If more publishers reported OA invoicing data on the article
+level and in a machine-readable format, this would increase transparency
+and enable enhanced monitoring of the scholarly journal landscape over
+time. As hybrid OA has increasingly become a central element of OA
+policies of research funders and libraries, customer organisations could
+require that invoicing information is added to the article-level
+metadata. As long as publishers do not provide this data in a structured
+and comprehensive format, they prevent consumers from benchmarking
+prices and therefore hinder price competition as the only alternative to
+gain comparative insight is self-reported APC expenditures that research
+funders and libraries voluntarily report to the Open APC initiative. The
+Open APC initiative is a valuable service but it cannot replace
+publisher-provided data since global participation is limited.
+
+From recent science policy developments in Europe it appears that Big
+Deals have gained support and remain firmly in place in the form of
+transformative agreements. However, in addition to research funders and
+libraries that increasingly make publishing contracts public, it is on
+the publisher to provide metadata as presented here to improve the
+transparency in scholarly publishing by linking invoicing data to
+bibliometrics. From this study we can affirm that hybrid OA is complex
+as the financial flows come from various stakeholders and invoicing
+forms, it is not just libraries or consortia, but authors and funders
+are also directly involved. As of right now, providing machine-readable
+article-level invoicing data is not common amongst publishers, which
+limits the potential for broader studies.
+
+\hypertarget{data-availability}{%
+\section{Data Availability}\label{data-availability}}
+
+All data and code used for this study is openly available on GitHub,
+including version history:
+
+\url{https://github.com/njahn82/elsevier_hybrid_invoicing}
+
+Formal analyis is written in R Markdown and thus replicable.
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
 \hypertarget{refs}{}
@@
 
+\leavevmode\hypertarget{ref-Bj_rk_2010}{}%
+Björk, B.-C., Welling, P., Laakso, M., Majlender, P., Hedlund, T., \&
+Guðnason, G. (2010). Open access to the scientific journal literature:
+Situation 2009. \emph{PLoS ONE}, \emph{5}(6), e11273.
+\url{https://doi.org/10.1371/journal.pone.0011273}
+
 \leavevmode\hypertarget{ref-Borrego_2020}{}%
@@
 
+\leavevmode\hypertarget{ref-Castro_2019}{}%
+Castro, P. de. (2019). \emph{Running a no-hybrid open access funding
+policy: Some results}. Open Access Strathclyde.
+\url{https://strathoa.tumblr.com/post/185703709380/running-a-no-hybrid-open-access-funding-policy}
+
 \leavevmode\hypertarget{ref-crminer}{}%
@@
 
+\leavevmode\hypertarget{ref-Chawla_2020}{}%
+Chawla, D. (2020). This tool is saving universities millions of dollars
+in journal subscriptions. \emph{Science}.
+\url{https://doi.org/10.1126/science.abd7483}
+
+\leavevmode\hypertarget{ref-Creaser_2010}{}%
+Creaser, C., Fry, J., Greenwood, H., Oppenheim, C., Probets, S., Spezi,
+V., \& White, S. (2010). Authors' awareness and attitudes toward open
+access repositories. \emph{New Review of Academic Librarianship},
+\emph{16}(sup1), 145--161.
+\url{https://doi.org/10.1080/13614533.2010.518851}
+
 \leavevmode\hypertarget{ref-Crossref_2020}{}%
@@
 \url{https://doi.org/10.3390/publications7020026}
+
+\leavevmode\hypertarget{ref-Fraser_2020}{}%
+Fraser, N., Brierley, L., Dey, G., Polka, J. K., Pálfy, M., Nanni, F.,
+\& Coates, J. A. (2020). \emph{Preprinting the COVID-19 pandemic}.
+bioRxiv. \url{https://doi.org/10.1101/2020.05.22.111294}
 
@@
 
+\leavevmode\hypertarget{ref-Huang_2020}{}%
+Huang, C.-K. (Karl), Neylon, C., Hosking, R., Montgomery, L., Wilson, K.
+S., Ozaygen, A., \& Brookes-Kenworthy, C. (2020). Evaluating the impact
+of open access policies on research institutions. \emph{eLife},
+\emph{9}. \url{https://doi.org/10.7554/elife.57067}
+
 \leavevmode\hypertarget{ref-Jahn_2016}{}%
@@
 
+\leavevmode\hypertarget{ref-Jubb_2017}{}%
+Jubb, M., Plume, A., Oeben, S., Brammer, L., Johnson, R., Bütün, C., \&
+Pinfield, S. (2017). \emph{Monitoring the transition to open access:
+December 2017}.
+\url{https://www.universitiesuk.ac.uk/policy-and-analysis/reports/Documents/2017/monitoring-transition-open-access-2017.pdf}
+
 \leavevmode\hypertarget{ref-Kirkman_2018}{}%
@@
 \url{http://hdl.handle.net/20.500.11937/77026}
+
+\leavevmode\hypertarget{ref-Kramer_2018}{}%
+Kramer, B., \& Bosman, J. (2018). \emph{Towards a plan S gap analysis:
+open access potential across disciplines using Web of Science and DOAJ}
+{[}Data set{]}. Zenodo. \url{https://doi.org/10.5281/zenodo.1979937}
 
@@
 
+\leavevmode\hypertarget{ref-Van_Noorden_2013}{}%
+Noorden, R. V. (2013). Researchers opt to limit uses of open-access
+publications. \emph{Nature}.
+\url{https://doi.org/10.1038/nature.2013.12384}
+
 \leavevmode\hypertarget{ref-Frascati}{}%
@@
 \url{http://urn.kb.se/resolve?urn=urn\%3Anbn\%3Ase\%3Ahb\%3Adiva-15501}
+
+\leavevmode\hypertarget{ref-Pieper_2018}{}%
+Pieper, D., \& Broschinski, C. (2018). OpenAPC: A contribution to a
+transparent and reproducible monitoring of fee-based open access
+publishing across institutions and nations. \emph{Insights the UKSG
+Journal}, \emph{31}. \url{https://doi.org/10.1629/uksg.439}
 
@@
 
+\leavevmode\hypertarget{ref-Robinson_Garcia_2020}{}%
+Robinson-Garcia, N., Costas, R., \& Leeuwen, T. N. van. (2020). Open
+access uptake by universities worldwide. \emph{PeerJ}, \emph{8}, e9410.
+\url{https://doi.org/10.7717/peerj.9410}
+
+\leavevmode\hypertarget{ref-Rowley_2017}{}%
+Rowley, J., Johnson, F., Sbaffi, L., Frass, W., \& Devine, E. (2017).
+Academics\\
+textquotesingle behaviors and attitudes towards open access publishing
+in scholarly journals. \emph{Journal of the Association for Information
+Science and Technology}, \emph{68}(5), 1201--1211.
+\url{https://doi.org/10.1002/asi.23710}
+
+\leavevmode\hypertarget{ref-Schimmer_2015}{}%
+Schimmer, R., Geschuhn, K., \& Vogler, A. (2015). \emph{Disrupting the
+subscription journals'business model for the necessary large-scale
+transformation to open access}. Max Planck Digital Library.
+\url{https://doi.org/10.17617/1.3}
+
+\leavevmode\hypertarget{ref-Severin_2020}{}%
+Severin, A., Egger, M., Eve, M. P., \& Hürlimann, D. (2020).
+Discipline-specific open access publishing practices and barriers to
+change: An evidence-based review. \emph{F1000Research}, \emph{7}.
+\url{https://doi.org/10.12688/f1000research.17328.2}
+
 \leavevmode\hypertarget{ref-Shieber_2009}{}%
@@
 
+\leavevmode\hypertarget{ref-Tickel_2018}{}%
+Tickel, A. (2018). \emph{Open access to research publications -- 2018}.
+\url{https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/774956/Open-access-to-research-publications-2018.pdf}
+
+\leavevmode\hypertarget{ref-birmingham}{}%
+University of Birmingham. (n.d.). \emph{UKRI open access block grant}.
+\url{https://intranet.birmingham.ac.uk/as/libraryservices/library/research/open-access/funding/ukri-open-access-block-grant.aspx}
+
+\leavevmode\hypertarget{ref-oxford_2019}{}%
+University of Oxford. (2019). \emph{Change to oxford's policy for rcuk
+oa block grant 2nd december 2019}.
+\url{http://openaccess.ox.ac.uk/2019/12/02/change-to-oxfords-policy-for-rcuk-oa-block-grant-2nd-december-2019/}
+
 \leavevmode\hypertarget{ref-Unpaywall_para}{}%
@@
 \url{http://web.archive.org/web/20210127192625/https://support.unpaywall.org/support/solutions/articles/44002063718-what-is-an-oa-license-}.
+
+\leavevmode\hypertarget{ref-walker_2019}{}%
+Walker, D. (2019). \emph{Research councils uk open access funding
+2019-2020}. Library \& Archives Service at The London School of Hygiene
+\& Tropical Medicine.
+\url{https://blogs.lshtm.ac.uk/library/2019/03/05/research-councils-uk-open-access-funding-2019-2020/}
 

### 2021-01-31 analysis/011_scopus_subject.Rmd
scopus whole counting

---
title: "011_scpopus_subject_all.Rmd"
output: 
  github_document: 
    df_print: kable
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r}
library(tidyverse)
library(janitor)
```

read in scopus journal indicators

```{r}
scopus_19 <- readxl::read_excel(here::here("data", "scopus_jn.xlsx"), sheet = 2,
                                .name_repair = make_clean_names)
ajsc_mapped <- read_csv(here::here("data", "asjc_mapped.csv")) %>%
  mutate(top_level_code = as.character(top_level_code))
```

normalize scopus jn indicators

```{r}
scopus_norm <- scopus_19 %>%
  mutate(top_level_code = substr(as.character(scopus_asjc_code_sub_subject_area), 1, 2)) %>%
  inner_join(ajsc_mapped, by = "top_level_code")
```

## All

```{r}  
scopus_norm %>%
  distinct(subject_area, scopus_source_id, scholarly_output) %>%
  group_by(subject_area) %>% 
  summarise(n = sum(scholarly_output)) %>%
  mutate(prop = n / sum(n)) %>%
  arrange(desc(n))
```

## Elsevier

```{r}  
scopus_norm %>%
  filter(publisher == "Elsevier") %>%
  distinct(subject_area, scopus_source_id, scholarly_output) %>%
  group_by(subject_area) %>% 
  summarise(n = sum(scholarly_output)) %>%
  mutate(prop = n / sum(n))
```

By oa

```{r}  
scopus_norm %>%
  filter(publisher == "Elsevier") %>%
  distinct(subject_area, scopus_source_id, scholarly_output, open_access) %>%
  group_by(subject_area, open_access) %>% 
  summarise(n = sum(scholarly_output))
```

### 2021-02-03 analysis/manuscript.Rmd
save tmp

@@
 In September 2018, the cOAlition S announced that from 2021 onwards, signatories would no longer fund articles in hybrid OA journals.
-This principle includes so-called "mirror" and sister journals--hybrid journals with fully OA counterparts or split into their fully OA and subscription parts—that have launched since Plan S was announced but makes an exception for hybrid journals with transformative agreements [@Frantsv_g_2019].
+This principle includes so-called "mirror" and sister journals--hybrid journals with fully OA counterparts or split into their fully OA and subscription parts—that have launched since Plan S was announced but makes an exception for hybrid journals with transformative agreements [@Plan_s; @Frantsv_g_2019; @McNutt_2019; @Harrison_2019].
 The Plan S refers to the Efficiency and Standards for Article Charges (ESAC) guidelines [@Geschuhn_2017] as critical component of transformative agreements, emphasizing the need for publisher-provided metadata to address the lack of transparency surrounding hybrid OA.
@@
 
-First, we obtained hybrid journals through an Elsevier APC price list from May 2020 provided as captured by Matthias (2020).
+First, we obtained hybrid journals through an Elsevier APC price list from May 2020 provided as captured by @Matthias_2019.
 We only included hybrid journals that were published under the hybrid model throughout the entire period of our analysis.
-Hence, we excluded journals that flipped from OA to hybrid and from hybrid to OA.
+Hence, we excluded journals that flipped from OA to hybrid and from hybrid to OA [@Matthias_2020].
 This approach allowed us to ensure that the article-level differences we observed were not due changes at the journal-level (i.e., increases or decreases in number).
@@
 This finding reiterates reports from previous studies that the UK's OA profile differs markedly from other countries (cf.
-ref, ref) and points to the effects of science policy.
+@Jubb_2015; @Jubb_2017) and points to the effects of science policy.
 To increase OA to research publications, the UK implemented centralized APC funding and embraced hybrid OA as a transition model.

### 2021-02-03 analysis/manuscript.tex
save tmp

@@
 \shorttitle{Hybrid OA Elsevier}
-% \author{Najko Jahn}
-% \affiliation{State and University Library Göttingen, University of Göttingen, Germany}
-% \author{Lisa Matthias}
-% \affiliation{tba}
-% \author{Mikael Laakso}
-% \affiliation{tba}
-%
+\author{Najko Jahn}
+\affiliation{State and University Library Göttingen, University of Göttingen, Germany}
+\author{Lisa Matthias}
+\affiliation{}
+\author{Mikael Laakso}
 
@@
 The study discovered a large discrepancy in choice of publication license between author-invoiced articles (dominantly CC BY-NC-ND) and agreement-invoiced articles (dominantly CC BY).
-The results demonstrate the value of publisher-provided metadata covering invoiced recipients of articles, which would be even more useful through wider adoption among publishers.}
+The results demonstrate the value of publisher-provided metadata covering invoiced recipients of articles, which would be even more useful through wider adoption among publishers.
+}
 
 \begin{document}
+
 \maketitle
@@
 hybrid journals with transformative agreements (Frantsvåg \& Strømme,
-2019). The Plan S refers to the Efficiency and Standards for Article
-Charges (ESAC) guidelines (Geschuhn \& Stone, 2017) as critical
-component of transformative agreements, emphasizing the need for
-publisher-provided metadata to address the lack of transparency
-surrounding hybrid OA. However, studies confirm that publisher with such
-agreements lack essential data about hybrid OA uptake and invoicing, or
-are not able or willing to provide such information publicly and
-instantaneously (Marques et al., 2019; Monaghan et al., 2020), which, in
-turn, not just limit the monitoring of hybrid OA, but also research
-studies on its size and financial implications.
+2019; Harrison, 2019; McNutt, 2019; S, n.d.). The Plan S refers to the
+Efficiency and Standards for Article Charges (ESAC) guidelines (Geschuhn
+\& Stone, 2017) as critical component of transformative agreements,
+emphasizing the need for publisher-provided metadata to address the lack
+of transparency surrounding hybrid OA. However, studies confirm that
+publisher with such agreements lack essential data about hybrid OA
+uptake and invoicing, or are not able or willing to provide such
+information publicly and instantaneously (Marques et al., 2019; Monaghan
+et al., 2020), which, in turn, not just limit the monitoring of hybrid
+OA, but also research studies on its size and financial implications.
 
@@
 First, we obtained hybrid journals through an Elsevier APC price list
-from May 2020 provided as captured by Matthias (2020). We only included
-hybrid journals that were published under the hybrid model throughout
-the entire period of our analysis. Hence, we excluded journals that
-flipped from OA to hybrid and from hybrid to OA. This approach allowed
-us to ensure that the article-level differences we observed were not due
-changes at the journal-level (i.e., increases or decreases in number).
-Overall, we identified 1,970 unique hybrid journals.
+from May 2020 provided as captured by Matthias et al. (2019). We only
+included hybrid journals that were published under the hybrid model
+throughout the entire period of our analysis. Hence, we excluded
+journals that flipped from OA to hybrid and from hybrid to OA (Matthias,
+2020). This approach allowed us to ensure that the article-level
+differences we observed were not due changes at the journal-level (i.e.,
+increases or decreases in number). Overall, we identified 1,970 unique
+hybrid journals.
 
@@
 studies that the UK's OA profile differs markedly from other countries
-(cf.~ref, ref) and points to the effects of science policy. To increase
-OA to research publications, the UK implemented centralized APC funding
-and embraced hybrid OA as a transition model. Within five years of the
-publication of the Finch Report in 2012, the UK recorded an 18\%
-increase in immediate OA, coupled with a rise in hybrid OA from 2.7\% in
-2012 to 15.4\% of all articles in 2016 (Jubb et al., 2017), which can be
-attributed to the availability of APC funding from Research Councils UK
-(RCUK). Indeed, RCUK block grants have been the largest single source of
-OA funds in the UK---the Wellcome Trust, and more recently
-transformative agreements (Jubb et al., 2017; Tickel, 2018). However,
-since then publishing expenditures of UK universities have been rising
-rapidly (ref), so that several universities stopped supporting hybrid OA
-in late 2018-2019 (Castro, 2019; University of Birmingham, n.d.;
-University of Oxford, 2019; Walker, 2019). This development might
-explain the slight decrease in hybrid articles invoiced to the UK our
-data showed around that time (see Figure
-\ref{fig:invoice_sponsor_country}).
+(cf.~Jubb et al. (2015); Jubb et al. (2017)) and points to the effects
+of science policy. To increase OA to research publications, the UK
+implemented centralized APC funding and embraced hybrid OA as a
+transition model. Within five years of the publication of the Finch
+Report in 2012, the UK recorded an 18\% increase in immediate OA,
+coupled with a rise in hybrid OA from 2.7\% in 2012 to 15.4\% of all
+articles in 2016 (Jubb et al., 2017), which can be attributed to the
+availability of APC funding from Research Councils UK (RCUK). Indeed,
+RCUK block grants have been the largest single source of OA funds in the
+UK---the Wellcome Trust, and more recently transformative agreements
+(Jubb et al., 2017; Tickel, 2018). However, since then publishing
+expenditures of UK universities have been rising rapidly (ref), so that
+several universities stopped supporting hybrid OA in late 2018-2019
+(Castro, 2019; University of Birmingham, n.d.; University of Oxford,
+2019; Walker, 2019). This development might explain the slight decrease
+in hybrid articles invoiced to the UK our data showed around that time
+(see Figure \ref{fig:invoice_sponsor_country}).
 
@@
 
+\leavevmode\hypertarget{ref-Harrison_2019}{}%
+Harrison, P. (2019). \emph{What are mirror journals, and can they offer
+a new world of open access?} Elsevier B.V.
+\url{https://web.archive.org/web/20210109033031/https://www.elsevier.com/connect/what-are-mirror-journals-and-can-they-offer-a-new-world-of-open-access}
+
 \leavevmode\hypertarget{ref-Hendricks_2020}{}%
@@
 \url{https://doi.org/10.7717/peerj.2323}
+
+\leavevmode\hypertarget{ref-Jubb_2015}{}%
+Jubb, M., Goldstein, S., Amin, M., Plume, A., Oeben, S., Aisati, M.,
+Pinfield, S., Bath, P., Salter, J., Johnson, R., \& Fosci, M. (2015).
+\emph{Monitoring the transition to open access: A report for the
+universities uk open access co-ordination group}.
+\url{https://web.archive.org/web/20190523005240/https://www.acu.ac.uk/research-information-network/monitoring-transition-to-open-access}
 
@@
 \emph{12}(3), 819--841. \url{https://doi.org/10.1016/j.joi.2018.06.012}
+
+\leavevmode\hypertarget{ref-Matthias_2020}{}%
+Matthias, L. (2020). \emph{Publisher oa portfolios 2.0} (Version 2.0)
+{[}Data set{]}. Zenodo. \url{https://doi.org/10.5281/zenodo.3841568}
+
+\leavevmode\hypertarget{ref-Matthias_2019}{}%
+Matthias, L., Jahn, N., \& Laakso, M. (2019). The two-way street of open
+access journal publishing: Flip it and reverse it. \emph{Publications},
+\emph{7}(2), 23. \url{https://doi.org/10.3390/publications7020023}
+
+\leavevmode\hypertarget{ref-McNutt_2019}{}%
+McNutt, M. (2019). Opinion: \\
+textquotedblleftPlan s\\
+textquotedblright falls short for society publishers\\
+textemdashand for the researchers they serve. \emph{Proceedings of the
+National Academy of Sciences}, \emph{116}(7), 2400--2403.
+\url{https://doi.org/10.1073/pnas.1900359116}
 
@@
 
+\leavevmode\hypertarget{ref-Plan_s}{}%
+S. (n.d.). \emph{Plan s: Principles and implementation}.
+\url{https://web.archive.org/web/20210129165306/https://www.coalition-s.org/addendum-to-the-coalition-s-guidance-on-the-implementation-of-plan-s/principles-and-implementation/}.
+
 \leavevmode\hypertarget{ref-Severin_2020}{}%

### 2021-02-03 install.R
testing binder setup

@@ -1,0 +1,6 @@
+# only packages used in manuscript.Rmd
+my_packages <- c("tidyverse", "here", "cowplot", "gt", 
+                 "janitor", "kableExtra", "stringr")
+install.packages(my_packages,
+                 repos='https://mran.microsoft.com/snapshot/2021-01-31/', 
+                 dependencies=TRUE)

### 2021-02-03 install.R
fix binder setup

@@ -1,6 +1,4 @@
 # only packages used in manuscript.Rmd
 my_packages <- c("tidyverse", "here", "cowplot", "gt", 
                  "janitor", "kableExtra", "stringr")
-install.packages(my_packages,
-                 repos='https://mran.microsoft.com/snapshot/2021-01-31/', 
-                 dependencies=TRUE)
+install.packages(my_packages, dependencies=TRUE)

### 2021-02-04 install.R
add latex supprt

@@ -1,4 +1,5 @@
 # only packages used in manuscript.Rmd
 my_packages <- c("tidyverse", "here", "cowplot", "gt", 
-                 "janitor", "kableExtra", "stringr")
+                 "janitor", "kableExtra", "stringr", "tinytex")
 install.packages(my_packages, dependencies=TRUE)
+tinytex::install_tinytex()

### 2021-02-04 analysis/manuscript.Rmd
minor tex fixes for binder

@@
     template: apa6.sty
-  html_document:
-    fig_caption: yes
-    force_captions: yes
-    number_sections: no
-    theme: readable
 csl: apa.csl
@@
   fig.pos = "H", 
-  out.extra = ""
+  out.extra = "",
+  dev= c("png", "cairo_ps")
 )
@@
 This order mostly reflects the disciplines' overall publication output.
-According to the Open Science Monitor by the European Commission, the physical sciences and mathematics publish the most articles, followed by the health sciences, life sciences, and social sciences and humanities.
+According to the Open Science Monitor by the European Commission [@OS_Monitor], the physical sciences and mathematics publish the most articles, followed by the health sciences, life sciences, and social sciences and humanities.
 

### 2021-02-04 analysis/manuscript.tex
minor tex fixes for binder

@@
 \usepackage{xcolor}
+% make captions italic
 
@@
 
-\title{Transparency to hybrid open access invoicing through publisher-provided metadata: A longitudinal article-level study of Elsevier}
+\title{\textbf{Transparency to hybrid open access  through publisher-provided metadata: An article-level study of Elsevier}}
 \shorttitle{Hybrid OA Elsevier}
-\author{Najko Jahn}
-\affiliation{State and University Library Göttingen, University of Göttingen, Germany}
-\author{Lisa Matthias}
-\affiliation{}
-\author{Mikael Laakso}
-
-\abstract{The financial flows in scholarly journal publishing have become increasingly complex with the growth of open access (OA), and the lack of data and transparency into these flows has been persistent.
-Opaqueness has concerned especially hybrid OA, where subscription-based journals also publish OA content if an optional fee is paid.
-By leveraging Elsevier’s recently introduced article metadata concerning invoiced recipients, this study provides the first longitudinal publisher-level study of uptake and invoicing of hybrid OA articles.
-The results show that the share of OA articles in Elseviers hybrid journals has grown slowly, from 2.6\% of all articles on average in 2015 to 3.7\% in 2019.
-About a third of hybrid OA articles are invoiced directly through an agreement with research funders or other type of organisational sponsor (e.g.
-national library consortia), with the majority invoiced to authors.
-83\% of agreement-invoiced hybrid OA articles are to European organisations, suggesting a strong international skewness.
-The study discovered a large discrepancy in choice of publication license between author-invoiced articles (dominantly CC BY-NC-ND) and agreement-invoiced articles (dominantly CC BY).
-The results demonstrate the value of publisher-provided metadata covering invoiced recipients of articles, which would be even more useful through wider adoption among publishers.
+\threeauthors{Najko Jahn}{Lisa Matthias}{Mikael Laakso}
+\threeaffiliations{Göttingen State and University Library, University of Göttingen\\
+Platz der Göttinger Sieben 1, 37073 Göttingen, Germany\\
+najko.jahn@sub.uni-goettingen.de
+}{Department of Political Science, John F. Kennedy Institute, Freie Universität Berlin \\
+Lansstraße 7-9, 14195 Berlin, Germany \\
+l.a.matthia@gmail.com
+}{Information Systems Science, Hanken School of Economics \\
+Arkadiankatu 22, 00100 Helsinki, Finland \\
+mikael.laakso@hanken.fi
 }
 
+\authornote{Correspondence concerning this article should be addressed to Najko Jahn}
+% \affiliation{State and University Library Göttingen, University of Göttingen, Germany}
+% \author{Lisa Matthias}
+% \affiliation{tba}
+% \author{Mikael Laakso}
+% \affiliation{tba}
+%
+
+\abstract{With the growth of open access (OA), the financial flows in scholarly journal
+publishing have become increasingly complex, but comprehensive data and
+transparency into these flows are still lacking. The opaqueness is especially
+concerning for hybrid OA, where subscription-based journals publish individual
+articles as OA if an optional fee is paid. This study addresses the lack of
+transparency by leveraging Elsevier article metadata and provides the first
+longitudinal publisher-level study of hybrid OA uptake and invoicing. Our
+results show that Elsevier's hybrid OA uptake has grown steadily but slowly from
+2015-2019, doubling the number of hybrid OA articles published per year and
+increasing the share of OA articles in Elsevier's hybrid journals from 2.6\% to
+3.7\% of all articles. Further, we find that most hybrid OA articles were
+invoiced directly to authors, followed by articles invoiced through agreements
+with research funders, institutions, or consortia, with only a few funding
+bodies driving hybrid OA uptake. As such, our findings point to the role of
+publishing agreements and OA policies in hybrid OA publishing. Our results
+further demonstrate the value of publisher-provided metadata improve the
+transparency in scholarly publishing by linking invoicing data to bibliometrics.}
+
 \begin{document}
-
 \maketitle
@@
 publication output. According to the Open Science Monitor by the
-European Commission, the physical sciences and mathematics publish the
-most articles, followed by the health sciences, life sciences, and
-social sciences and humanities.
+European Commission (European Commission, 2019), the physical sciences
+and mathematics publish the most articles, followed by the health
+sciences, life sciences, and social sciences and humanities.
 
@@
 
+\leavevmode\hypertarget{ref-OS_Monitor}{}%
+European Commission. (2019). \emph{Open science monitor: Trends for open
+access to publications}.
+\url{https://ec.europa.eu/info/research-and-innovation/strategy/goals-research-and-innovation-policy/open-science/open-science-monitor/trends-open-access-publications_en}
+
 \leavevmode\hypertarget{ref-Frantsv_g_2019}{}%
@@
 \leavevmode\hypertarget{ref-McNutt_2019}{}%
-McNutt, M. (2019). Opinion: \\
-textquotedblleftPlan s\\
-textquotedblright falls short for society publishers\\
-textemdashand for the researchers they serve. \emph{Proceedings of the
+McNutt, M. (2019). Opinion: Plan S falls short for society
+publishers--and for the researchers they serve. \emph{Proceedings of the
 National Academy of Sciences}, \emph{116}(7), 2400--2403.

### 2021-05-02 analysis/manuscript.Rmd
starting revision

@@
 ```
+
 ```{r hybrid_oa_per_year}

### 2021-05-02 analysis/manuscript.tex
starting revision

@@
 that allows journals to gradually convert to full OA and reduce
-subscription costs as the OA uptake increases (Prosser, 2003). Many
-large subscription publishers have since incorporated this OA model, and
-while the uptake among authors increased only slowly during the first
-decade (Björk, 2012), it has gradually gained popularity in recent
-years. This development was driven by the implementation of OA policies
-by research institutions and funders, the allocation of OA funds, and
-formal agreements between publishers and institutions that allow
-affiliated authors to publish OA free-of-charge (Laakso \& Björk, 2016).
-However, there have been concerns that the hybrid model allows for
-double-dipping---receiving two payments for one article, the APC and
-subscription fees (Prosser, 2015; Shieber, 2009). While publishers
+subscription costs as the OA uptake increases (D. C. Prosser, 2003).
+Many large subscription publishers have since incorporated this OA
+model, and while the uptake among authors increased only slowly during
+the first decade (Björk, 2012), it has gradually gained popularity in
+recent years. This development was driven by the implementation of OA
+policies by research institutions and funders, the allocation of OA
+funds, and formal agreements between publishers and institutions that
+allow affiliated authors to publish OA free-of-charge (Laakso \& Björk,
+2016). However, there have been concerns that the hybrid model allows
+for double-dipping---receiving two payments for one article, the APC and
+subscription fees (D. Prosser, 2015; Shieber, 2009). While publishers
 assure they combine subscription fees and APCs to adjust pricing
@@
 fundamental role in monitoring hybrid OA, yet time and again, previous
-studies have underscored the difficulties involved in this
-(cf.~(Akbaritabar \& Stahlschmidt, 2019; Björk, 2012; Emery, 2013;
-Laakso \& Björk, 2016; Leeuwen et al., 2018). Insights into the volume
-of hybrid OA and the various levels and sources of APC funding could
-create a better understanding of the costs involved in scholarly
-publishing and, thus, strengthen institutions' bargaining power in
-negotiations with publishers, and promote greater equity in scholarly
-publishing. In the following section, we examine how previous research
-has tackled the challenge of identifying hybrid OA content and what
-these studies taught us about the uptake and funding sources of hybrid
-OA.
+studies have underscored the difficulties involved in this (cf.
+(Akbaritabar \& Stahlschmidt, 2019; Björk, 2012; Emery, 2013; Laakso \&
+Björk, 2016; Leeuwen et al., 2018). Insights into the volume of hybrid
+OA and the various levels and sources of APC funding could create a
+better understanding of the costs involved in scholarly publishing and,
+thus, strengthen institutions' bargaining power in negotiations with
+publishers, and promote greater equity in scholarly publishing. In the
+following section, we examine how previous research has tackled the
+challenge of identifying hybrid OA content and what these studies taught
+us about the uptake and funding sources of hybrid OA.
 
@@
 information was developed to better distinguish them from subscription
-journals including hybrid journals (Unpaywall, n.d.-b). With its servie
+journals including hybrid journals (Unpaywall, n.d.-a). With its servie
 unsub, Unpaywall recently launched a paid service, where libraries can
@@
 at a discounted fee or for free. These agreements can serve as the basis
-for studies on the uptake of hybrid OA or cost-benefit analyses
-(cf.~Olsson (2018); Lawson (2017)) but are often limited in scope as
-they tend to focus on a single country and rely on data that seldom is
+for studies on the uptake of hybrid OA or cost-benefit analyses (cf.
+Olsson (2018); Lawson (2017)) but are often limited in scope as they
+tend to focus on a single country and rely on data that seldom is
 publicly available. Hence, while being valuable for specific use-cases,
@@
 of reviewers, following Unpaywall's paratext recognition approach
-(Unpaywall, n.d.-a), which we expanded to include patterns indicating
+(Unpaywall, n.d.-b), which we expanded to include patterns indicating
 corrections. Furthermore, we categorized the articles by subject
@@
 invoices billed to the corresponding author, as part of publishing
-agreements with funding bodies (hereinafter ``agreements''; cf.~Elsevier
+agreements with funding bodies (hereinafter ``agreements''; cf. Elsevier
 (n.d.-a)), fee waivers (e.g., in ``cases of genuine need'' or due to
@@
 we combined invoice recipients Elsevier listed as ``authors'' and
-``third-party sponsor'', and those we had coded as business enterprises
-(OECD, 2015, p. 97), into ``Others''. Moreover, we categorized invoice
+``third-party sponsor,'' and those we had coded as business enterprises
+(OECD, 2015, p. 97), into ``Others.'' Moreover, we categorized invoice
 recipients according to the country or countries representing their
@@
 note that although Elsevier's metadata classifies the sponsor type as
-``author'', this does not necessarily mean that authors paid for APCs
+``author,'' this does not necessarily mean that authors paid for APCs
 out of their own pocket but rather just that the APC was not invoiced
@@
 studies that the UK's OA profile differs markedly from other countries
-(cf.~Jubb et al. (2015); Jubb et al. (2017)) and points to the effects
+(cf. Jubb et al. (2015); Jubb et al. (2017)) and points to the effects
 of science policy. To increase OA to research publications, the UK
@@
 \hypertarget{refs}{}
-\begin{cslreferences}
+\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-Akbaritabar_2019}{}%
@@
 unpaywall information to identify gold, hidden gold, hybrid and delayed
-open access publications in the KB publication corpus}. SocArXiv.
+open access publications in the {KB} publication corpus}. SocArXiv.
 \url{https://doi.org/10.31235/osf.io/sdzft}
@@
 Antelman, K. (2017). Leveraging the growth of open access in library
-collection decision making. \emph{At the helm: leading transformation},
-411--422.
+collection decision making. \emph{{At the helm: leading
+transformation}}, 411--422.
 
@@
 Guðnason, G. (2010). Open access to the scientific journal literature:
-Situation 2009. \emph{PLoS ONE}, \emph{5}(6), e11273.
+Situation 2009. \emph{{PLoS} {ONE}}, \emph{5}(6), e11273.
 \url{https://doi.org/10.1371/journal.pone.0011273}
@@
 \leavevmode\hypertarget{ref-crminer}{}%
-Chamberlain, S. (2020). \emph{crminer: Fetch scholary full text from
-CrossRef}. \url{https://CRAN.R-project.org/package=crminer}
+Chamberlain, S. (2020). \emph{{crminer}: Fetch scholary full text from
+{CrossRef}}. \url{https://CRAN.R-project.org/package=crminer}
 
@@
 Chamberlain, S., Zhu, H., Jahn, N., Boettiger, C., \& Ram, K. (2020).
-\emph{rcrossref: Client for various CrossRef APIs}.
+\emph{{rcrossref}: Client for various {CrossRef} {APIs}}.
 \url{https://CRAN.R-project.org/package=rcrossref}
@@
 \leavevmode\hypertarget{ref-Els_Agreements}{}%
-Elsevier. (n.d.-a). \emph{Agreements}.
+Elsevier. (n.d.-a). \emph{{Agreements}}.
 \url{https://web.archive.org/web/20210127152729/https://www.elsevier.com/open-access/agreements}.
@@
 \leavevmode\hypertarget{ref-Els_Archive}{}%
-Elsevier. (n.d.-b). \emph{Open archive}.
+Elsevier. (n.d.-b). \emph{{Open archive}}.
 \url{http://web.archive.org/web/20210127201740/https://www.elsevier.com/open-access/open-archive}.
@@
 \leavevmode\hypertarget{ref-Els_videos}{}%
-Elsevier. (n.d.-c). \emph{Publishing Journey videos: How do I complete
-the Rights and Access form?}
+Elsevier. (n.d.-c). \emph{{Publishing Journey videos: How do I complete
+the Rights and Access form?}}
 \url{https://web.archive.org/web/20210127154144/https://service.elsevier.com/app/answers/detail/a_id/29789/supporthub/publishing/track/APN2ZgoIDv8a~RNiGvwa~yKgpv0qOS75Mv9e~zj~PP_X/}.
@@
 \leavevmode\hypertarget{ref-lund}{}%
-Elsevier. (n.d.-d). \emph{Publishing options: BIBSAM institute
-associated authors}.
+Elsevier. (n.d.-d). \emph{{Publishing options: BIBSAM institute
+associated authors}}.
 \url{https://web.archive.org/web/20210127153520/https://www.ub.lu.se/en/sites/ub.lu.se.en/files/publicering_under_elsevier_avtalet.pdf}.
@@
 \leavevmode\hypertarget{ref-vsnu}{}%
-Elsevier. (n.d.-e). \emph{Publishing options for: Dutch universities \&
-institutes associated authors}.
+Elsevier. (n.d.-e). \emph{{Publishing options for: Dutch universities \&
+institutes associated authors}}.
 \url{https://web.archive.org/web/20190530214833/https://www.openaccess.nl/sites/www.openaccess.nl/files/documenten/elsevier_-_vsnu_workflow_updated_18july_v2.pdf}.
@@
 Emery, J. (2013). Mining for gold: Identifying the librarians' toolkit
-for managing hybrid open access. \emph{Insights: The UKSG Journal},
+for managing hybrid open access. \emph{Insights: The {UKSG} Journal},
 \emph{26}(2), 115--119. \url{https://doi.org/10.1629/2048-7754.65}
@@
 Fraser, N., Brierley, L., Dey, G., Polka, J. K., Pálfy, M., Nanni, F.,
-\& Coates, J. A. (2020). \emph{Preprinting the COVID-19 pandemic}.
-bioRxiv. \url{https://doi.org/10.1101/2020.05.22.111294}
+\& Coates, J. A. (2020). \emph{Preprinting the {COVID}-19 pandemic}.
+{bioRxiv}. \url{https://doi.org/10.1101/2020.05.22.111294}
 
@@
 Geschuhn, K., \& Stone, G. (2017). It's the workflows, stupid! What is
-required to make ``offsetting'' work for the open access transition.
-\emph{Insights the UKSG Journal}, \emph{30}(3), 103--114.
+required to make {`offsetting'} work for the open access transition.
+\emph{Insights the {UKSG} Journal}, \emph{30}(3), 103--114.
 \url{https://doi.org/10.1629/uksg.391}
@@
 S., Ozaygen, A., \& Brookes-Kenworthy, C. (2020). Evaluating the impact
-of open access policies on research institutions. \emph{eLife},
+of open access policies on research institutions. \emph{{eLife}},
 \emph{9}. \url{https://doi.org/10.7554/elife.57067}
@@
 Jahn, N., \& Tullney, M. (2016). A study of institutional spending on
-open access publication fees in germany. \emph{PeerJ}, \emph{4}, e2323.
-\url{https://doi.org/10.7717/peerj.2323}
+open access publication fees in germany. \emph{{PeerJ}}, \emph{4},
+e2323. \url{https://doi.org/10.7717/peerj.2323}
 
@@
 \emph{Monitoring the transition to open access: A report for the
-universities uk open access co-ordination group}.
+universities UK open access co-ordination group}.
 \url{https://web.archive.org/web/20190523005240/https://www.acu.ac.uk/research-information-network/monitoring-transition-to-open-access}
@@
 \leavevmode\hypertarget{ref-Kirkman_2018}{}%
-Kirkman, N. S. (2018). \emph{A study of open access publishing by nhmrc
-grant recipients} {[}Curtin University{]}.
+Kirkman, N. S. (2018). \emph{A study of open access publishing by NHMRC
+grant recipients} {[}{Curtin University}{]}.
 \url{http://hdl.handle.net/20.500.11937/77026}
@@
 \leavevmode\hypertarget{ref-Kramer_2018}{}%
-Kramer, B., \& Bosman, J. (2018). \emph{Towards a plan S gap analysis:
-open access potential across disciplines using Web of Science and DOAJ}
+Kramer, B., \& Bosman, J. (2018). \emph{{Towards a plan S gap analysis:
+open access potential across disciplines using Web of Science and DOAJ}}
 {[}Data set{]}. Zenodo. \url{https://doi.org/10.5281/zenodo.1979937}
@@
 Larivière, V., Haustein, S., \& Mongeon, P. (2015). The oligopoly of
-academic publishers in the digital era. \emph{PLOS ONE}, \emph{10}(6),
-e0127502. \url{https://doi.org/10.1371/journal.pone.0127502}
+academic publishers in the digital era. \emph{{PLOS} {ONE}},
+\emph{10}(6), e0127502.
+\url{https://doi.org/10.1371/journal.pone.0127502}
 
@@
 
+\leavevmode\hypertarget{ref-Lawson_2015}{}%
+Lawson, S. (2015). "Total cost of ownership" of scholarly communication:
+Managing subscription and {APC} payments together. \emph{Learned
+Publishing}, \emph{28}(1), 9--13. \url{https://doi.org/10.1087/20150103}
+
 \leavevmode\hypertarget{ref-Lawson_2018}{}%
@@
 \url{https://doi.org/10.6084/M9.FIGSHARE.5383861.V1}
-
-\leavevmode\hypertarget{ref-Lawson_2015}{}%
-Lawson, S. (2015). "Total cost of ownership" of scholarly communication:
-Managing subscription and APC payments together. \emph{Learned
-Publishing}, \emph{28}(1), 9--13. \url{https://doi.org/10.1087/20150103}
 
@@
 Monitoring agreements with open access elements: Why article-level
-metadata are important. \emph{Insights the UKSG Journal}, \emph{32}.
+metadata are important. \emph{Insights the {UKSG} Journal}, \emph{32}.
 \url{https://doi.org/10.1629/uksg.489}
@@
 \leavevmode\hypertarget{ref-Matthias_2020}{}%
-Matthias, L. (2020). \emph{Publisher oa portfolios 2.0} (Version 2.0)
+Matthias, L. (2020). \emph{Publisher OA portfolios 2.0} (Version 2.0)
 {[}Data set{]}. Zenodo. \url{https://doi.org/10.5281/zenodo.3841568}
@@
 \leavevmode\hypertarget{ref-McNutt_2019}{}%
-McNutt, M. (2019). Opinion: Plan S falls short for society
+McNutt, M. (2019). Opinion: {Plan S} falls short for society
 publishers--and for the researchers they serve. \emph{Proceedings of the
@@
 Mittermaier, B. (2015). Double dipping in hybrid open access -- chimera
-or reality? \emph{ScienceOpen Research}.
+or reality? \emph{{ScienceOpen} Research}.
 \url{https://doi.org/10.14293/s2199-1006.1.sor-socsci.aowntu.v1}
@@
 OECD. (2015). \emph{Frascati manual 2015: Guidelines for collecting and
-reporting data on research and experimental development}. OECD
-Publishing. \url{https://doi.org/10.1787/24132764}
+reporting data on research and experimental development}. {OECD
+Publishing}. \url{https://doi.org/10.1787/24132764}
 
@@
 \leavevmode\hypertarget{ref-Pieper_2018}{}%
-Pieper, D., \& Broschinski, C. (2018). OpenAPC: A contribution to a
+Pieper, D., \& Broschinski, C. (2018). {OpenAPC}: A contribution to a
 transparent and reproducible monitoring of fee-based open access
-publishing across institutions and nations. \emph{Insights the UKSG
+publishing across institutions and nations. \emph{Insights the {UKSG}
 Journal}, \emph{31}. \url{https://doi.org/10.1629/uksg.439}
@@
 Norlander, B., Farley, A., West, J., \& Haustein, S. (2018). The state
-of OA: A large-scale analysis of the prevalence and impact of open
-access articles. \emph{PeerJ}, \emph{6}, e4375.
+of {OA}: A large-scale analysis of the prevalence and impact of open
+access articles. \emph{{PeerJ}}, \emph{6}, e4375.
 \url{https://doi.org/10.7717/peerj.4375}
@@
 \leavevmode\hypertarget{ref-Piwowar_2019}{}%
-Piwowar, H., Priem, J., \& Orr, R. (2019). \emph{The future of OA: A
+Piwowar, H., Priem, J., \& Orr, R. (2019). \emph{The future of {OA}: A
 large-scale analysis projecting open access publication and readership}.
-bioRxiv. \url{https://doi.org/10.1101/795310}
+{bioRxiv}. \url{https://doi.org/10.1101/795310}
 
@@
 \leavevmode\hypertarget{ref-Prosser_2015}{}%
-Prosser, D. (2015). \emph{The costs of double dipping}.
+Prosser, D. (2015). \emph{{The costs of double dipping}}.
 \url{https://web.archive.org/web/20210128164930/https://www.rluk.ac.uk/the-costs-of-double-dipping/}.
@@
 Robinson-Garcia, N., Costas, R., \& Leeuwen, T. N. van. (2020). Open
-access uptake by universities worldwide. \emph{PeerJ}, \emph{8}, e9410.
-\url{https://doi.org/10.7717/peerj.9410}
+access uptake by universities worldwide. \emph{{PeerJ}}, \emph{8},
+e9410. \url{https://doi.org/10.7717/peerj.9410}
 
@@
 Rowley, J., Johnson, F., Sbaffi, L., Frass, W., \& Devine, E. (2017).
-Academics\\
-textquotesingle behaviors and attitudes towards open access publishing
+Academics{\hfill\break
+textquotesingle} behaviors and attitudes towards open access publishing
 in scholarly journals. \emph{Journal of the Association for Information
@@
 
+\leavevmode\hypertarget{ref-Plan_s}{}%
+S, cOAlition. (n.d.). \emph{Plan s: Principles and implementation}.
+\url{https://web.archive.org/web/20210129165306/https://www.coalition-s.org/addendum-to-the-coalition-s-guidance-on-the-implementation-of-plan-s/principles-and-implementation/}.
+
 \leavevmode\hypertarget{ref-Schimmer_2015}{}%
-Schimmer, R., Geschuhn, K., \& Vogler, A. (2015). \emph{Disrupting the
+Schimmer, R., Geschuhn, K., \& Vogler, A. (2015). \emph{{Disrupting the
 subscription journals'business model for the necessary large-scale
-transformation to open access}. Max Planck Digital Library.
+transformation to open access}}. Max Planck Digital Library.
 \url{https://doi.org/10.17617/1.3}
-
-\leavevmode\hypertarget{ref-Plan_s}{}%
-S. (n.d.). \emph{Plan s: Principles and implementation}.
-\url{https://web.archive.org/web/20210129165306/https://www.coalition-s.org/addendum-to-the-coalition-s-guidance-on-the-implementation-of-plan-s/principles-and-implementation/}.
 
@@
 Shieber, S. M. (2009). Equity for open-access journal publishing.
-\emph{PLoS Biology}, \emph{7}(8), e1000165.
+\emph{{PLoS} Biology}, \emph{7}(8), e1000165.
 \url{https://doi.org/10.1371/journal.pbio.1000165}
@@
 \leavevmode\hypertarget{ref-oxford_2019}{}%
-University of Oxford. (2019). \emph{Change to oxford's policy for rcuk
-oa block grant 2nd december 2019}.
+University of Oxford. (2019). \emph{Change to oxford's policy for RCUK
+OA block grant 2nd december 2019}.
 \url{http://openaccess.ox.ac.uk/2019/12/02/change-to-oxfords-policy-for-rcuk-oa-block-grant-2nd-december-2019/}
 
+\leavevmode\hypertarget{ref-Unpaywall_types}{}%
+Unpaywall. (n.d.-a). \emph{{What do the types of oa\_status (green,
+gold, hybrid, and bronze) mean?}}
+\url{http://web.archive.org/web/20210127192910/https://support.unpaywall.org/support/solutions/articles/44001777288}.
+
 \leavevmode\hypertarget{ref-Unpaywall_para}{}%
-Unpaywall. (n.d.-a). \emph{What does is\_paratext mean in the API?}
+Unpaywall. (n.d.-b). \emph{{What does is\_paratext mean in the API?}}
 \url{https://web.archive.org/web/20201126131640/https://support.unpaywall.org/support/solutions/articles/44001894783}.
 
-\leavevmode\hypertarget{ref-Unpaywall_types}{}%
-Unpaywall. (n.d.-b). \emph{What do the types of oa\_status (green, gold,
-hybrid, and bronze) mean?}
-\url{http://web.archive.org/web/20210127192910/https://support.unpaywall.org/support/solutions/articles/44001777288}.
-
 \leavevmode\hypertarget{ref-Unpaywall_oa_license}{}%
-Unpaywall. (n.d.-c). \emph{What is an OA license?}
+Unpaywall. (n.d.-c). \emph{{What is an OA license?}}
 \url{http://web.archive.org/web/20210127192625/https://support.unpaywall.org/support/solutions/articles/44002063718-what-is-an-oa-license-}.
@@
 \leavevmode\hypertarget{ref-walker_2019}{}%
-Walker, D. (2019). \emph{Research councils uk open access funding
+Walker, D. (2019). \emph{Research councils UK open access funding
 2019-2020}. Library \& Archives Service at The London School of Hygiene
@@
 Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
-Welcome to the tidyverse. \emph{Journal of Open Source Software},
+Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
 \emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
@@
 \leavevmode\hypertarget{ref-bigrquery}{}%
-Wickham, H., \& Bryan, J. (2020). \emph{bigrquery: An interface to
-Google's BigQuery API}.
+Wickham, H., \& Bryan, J. (2020). \emph{{bigrquery}: An interface to
+{Google's BigQuery API}}.
 \url{https://CRAN.R-project.org/package=bigrquery}
@@
 \leavevmode\hypertarget{ref-xml2}{}%
-Wickham, H., Hester, J., \& Ooms, J. (2020). \emph{xml2: Parse XML}.
+Wickham, H., Hester, J., \& Ooms, J. (2020). \emph{{xml2}: Parse {XML}}.
 \url{https://CRAN.R-project.org/package=xml2}
-\end{cslreferences}
+
+\end{CSLReferences}
 

### 2021-05-03 analysis/manuscript.Rmd
kableExtra error

@@
   kableExtra::kbl(booktabs = TRUE,
-                  caption = " Relative subject profile of centrally invoiced OA funding bodies.", align = c("l|rrrrr|r")) %>%
+                  caption = " Relative subject profile of centrally invoiced OA funding bodies.", align = c("lrrrrrr")) %>%
     row_spec(6:7, hline_after = TRUE) %>%

### 2021-05-03 analysis/manuscript.Rmd
Making this figure larger (with all panels below each other) to make easier to read

@@
 Visualized as a box-and-whiskers plot limited to an OA share of 20\\%; the red line represents the overall median (2.5), jittered dots represent individual journals.
-Subject fields are ordered based on the median OA uptake.', out.width="90%", fig.width=9}
+Subject fields are ordered based on the median OA uptake.', fig.asp=1.3, out.width="90%"}
 p_subject <- ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
@@
   geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
-  facet_wrap(~subject_area, scales = "free_y") +
+  facet_wrap(~subject_area, scales = "free_y", nrow = 4) +
   geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
@@
   coord_flip(ylim=c(0,0.2)) +
-  labs(x = NULL, y = "Five-Year Hybrid OA Uptake Rate per Journal") +
-  theme_minimal_vgrid() 
+  labs(x = NULL, y = "Five-Year Hybrid OA Uptake Rate\nper Journal") +
+  theme_minimal_vgrid(font_size = 10) 
 p_subject

### 2021-05-09 analysis/manuscript.Rmd
add journal counts to each subject in subject field  vs. oa uptake box plot

--- 2021-05-03 analysis/manuscript.Rmd
@@
 Visualized as a box-and-whiskers plot limited to an OA share of 20\\%; the red line represents the overall median (2.5), jittered dots represent individual journals.
-Subject fields are ordered based on the median OA uptake.', out.width="90%", fig.width=9}
+Subject fields are ordered based on the median OA uptake.', fig.asp=1.3, out.width="90%"}
+# with labels
+subject_all_years <- 
+  inner_join(subject_all_years, subject_area_all_jn, by = "subject_area") %>%
+  mutate(subject_area = glue::glue("{subject_area} ({journals_n} Journals)"))
+
 p_subject <- ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
@@
   coord_flip(ylim=c(0,0.2)) +
-  labs(x = NULL, y = "Five-Year Hybrid OA Uptake Rate per Journal") +
-  theme_minimal_vgrid() 
+  labs(x = NULL, y = "Five-Year Hybrid OA Uptake Rate\nper Journal") +
+  theme_minimal_vgrid(font_size = 10) +
+  theme(strip.text = element_text(face="bold"))
 p_subject

### 2021-05-09 analysis/manuscript.Rmd
fix grow numbers by invoicing channel

@@
   filter(oa_sponsor_type == "ElsevierWaived") 
-```
-
-Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has increasingly invoiced hybrid OA APCs directly to authors compared to research funders or academic consortia ("Agreement").
-The share of fee-waived articles remained relatively stable, but we found different types of waivers.
+sponsored_articles_prop <- sponsored_articles %>%
+  group_by(issued_year, fct_source) %>%
+  summarise(n = sum(n)) %>%
+  mutate(perc = n / sum(n) * 100)
+```
+
+Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has constantly invoiced hybrid OA APCs directly to authors compared to research funders or academic consortia ("Agreement").
+Between 2015 and 2019, the proportion of author-invoicing increased from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Author") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Author") %>% pull(perc)`%, while the share of invoicing through agreements slightly changed from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Agreement") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Agreement") %>% pull(perc)`%.
+Also, the share of fee-waived articles remained relatively stable, but we found different types of waivers.
 Around `r filter(els_waived_df, !is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)* 100`% of fee waivers were linked to an OA sponsor, which indicates that a third party paid the APC.

### 2021-05-09 analysis/manuscript.tex
fix latex compile issue https://github.com/mpark/wg21/issues/54
Making this figure larger (with all panels below each other) to make easier to read
fix grow numbers by invoicing channel

@@
 \setlength{\cslhangindent}{1.5em}
-\newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+\newenvironment{CSLReferences}%
+  {}%
   {\par}
-
-\renewcommand{\baselinestretch}{1.5}
-\setlength\parindent{24pt}
 
@@
 Figure \ref{fig:invoiceoverview} illustrates that over the years,
-Elsevier has increasingly invoiced hybrid OA APCs directly to authors
-compared to research funders or academic consortia (``Agreement''). The
-share of fee-waived articles remained relatively stable, but we found
-different types of waivers. Around 51.7\% of fee waivers were linked to
-an OA sponsor, which indicates that a third party paid the APC. For
-instance, 853 waivers were linked to the French Académie des Sciences,
-presumably covering OA publication in its society journals for
-affiliated authors. The remaining 48.3\% of waived articles did not
-disclose any OA sponsor. Moreover, Figure \ref{fig:invoiceoverview}
-compares the invoicing channels based on open content license variants.
-When Elsevier invoiced authors directly, most OA articles in hybrid
-journals were published under a non-commercial license (n=32,086;
-76.9\%), whereas most articles billed as part of agreements were
-licensed under the more permissive CC BY license (n=18,331; 75.6\%).
+Elsevier has constantly invoiced hybrid OA APCs directly to authors
+compared to research funders or academic consortia (``Agreement'').
+Between 2015 and 2019, the proportion of author-invoicing increased from
+57\% to 61.2\%, while the share of invoicing through agreements slightly
+changed from 31.6\% to 32.4\%. Also, the share of fee-waived articles
+remained relatively stable, but we found different types of waivers.
+Around 51.7\% of fee waivers were linked to an OA sponsor, which
+indicates that a third party paid the APC. For instance, 853 waivers
+were linked to the French Académie des Sciences, presumably covering OA
+publication in its society journals for affiliated authors. The
+remaining 48.3\% of waived articles did not disclose any OA sponsor.
+Moreover, Figure \ref{fig:invoiceoverview} compares the invoicing
+channels based on open content license variants. When Elsevier invoiced
+authors directly, most OA articles in hybrid journals were published
+under a non-commercial license (n=32,086; 76.9\%), whereas most articles
+billed as part of agreements were licensed under the more permissive CC
+BY license (n=18,331; 75.6\%).
 
@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{l|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|r}
+\begin{tabular}[t]{lrrrrrr}
 \toprule

### 2021-05-09 analysis/manuscript.Rmd
group field per subject area, consider also to provide percentages per subject area

@@
 
+# define heat-colors
 make_pal <- scales::col_numeric(
@@
 )
-
-funder_overview_subject %>%
-  mutate_at(3:6, function(x){
+# prepare row-wise grouping per subject area
+subject_table <- jn_subjects %>% 
+  distinct(subject_area, top_level) %>%
+  inner_join(funder_overview_subject, by = "top_level") %>%
+  mutate_at(4:7, function(x){
     round(x*100,0)
   }) %>%
-  mutate_at(3:6, function(x){
+  mutate_at(4:7, function(x){
     cell_spec(x, background = make_pal(x))}) %>%
+  mutate_at(3, function(x){
+    cell_spec(x, background = make_pal(x))})
+
+subject_table %>%
+  group_by(subject_area) %>%
+  arrange(top_level, .by_group = TRUE) %>% 
+  ungroup() %>%
+  select(-subject_area) %>%
             kbl(booktabs = TRUE, "latex", format.args = list(big.mark = ","), digits = 1, escape = FALSE,
@@
   add_header_above(c(" " = 2, "Invoice channel (in %)" = 4)) %>%
+  # manual, puh, not good 
+  pack_rows("Health Sciences", 1, 5) %>%
+  pack_rows("Life Sciences", 6, 10) %>%
+  pack_rows("Health Sciences", 11, 20) %>%
+  pack_rows("Health Sciences", 21, 26) %>%
   kable_styling(latex_options = c("scale_down"))

### 2021-05-09 analysis/manuscript.tex
group field per subject area, consider also to provide percentages per subject area

@@
 \midrule
-Nursing & 1,504 & \cellcolor[HTML]{DC3977}{82} & \cellcolor[HTML]{FFFFFF}{16} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FCDE9C}{1}\\
-Decision Sciences & 785 & \cellcolor[HTML]{E6576F}{72} & \cellcolor[HTML]{FCC088}{27} & \cellcolor[HTML]{FFFFFF}{0} & \cellcolor[HTML]{FCDE9C}{1}\\
-Pharmacology, Toxicology and Pharmaceutics & 4,075 & \cellcolor[HTML]{EA646F}{69} & \cellcolor[HTML]{FCC088}{27} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FAA476}{2}\\
-Business, Management and Accounting & 1,751 & \cellcolor[HTML]{EC686E}{68} & \cellcolor[HTML]{FBAF7D}{29} & \cellcolor[HTML]{FFF0D2}{2} & \cellcolor[HTML]{FCDE9C}{1}\\
-Dentistry & 288 & \cellcolor[HTML]{EC686E}{68} & \cellcolor[HTML]{FFFAF1}{17} & \cellcolor[HTML]{F89774}{10} & \cellcolor[HTML]{DC3977}{5}\\
-Health Professions & 674 & \cellcolor[HTML]{ED6D6E}{67} & \cellcolor[HTML]{FCD898}{24} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{DC3977}{5}\\
-Medicine & 23,623 & \cellcolor[HTML]{F0756E}{65} & \cellcolor[HTML]{FCD092}{25} & \cellcolor[HTML]{FCC98E}{6} & \cellcolor[HTML]{E34F6F}{4}\\
-Agricultural and Biological Sciences & 7,956 & \cellcolor[HTML]{F38170}{63} & \cellcolor[HTML]{F9A075}{31} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FCDE9C}{1}\\
-Computer Science & 3,185 & \cellcolor[HTML]{F38170}{63} & \cellcolor[HTML]{F27F70}{36} & \cellcolor[HTML]{FFFFFF}{0} & \cellcolor[HTML]{FCDE9C}{1}\\
-Economics, Econometrics and Finance & 2,071 & \cellcolor[HTML]{F48771}{62} & \cellcolor[HTML]{F79373}{33} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FFFFFF}{0}\\
-Veterinary & 2,091 & \cellcolor[HTML]{F58D72}{61} & \cellcolor[HTML]{FCD092}{25} & \cellcolor[HTML]{F1766E}{13} & \cellcolor[HTML]{FCDE9C}{1}\\
-Social Sciences & 6,306 & \cellcolor[HTML]{F79273}{60} & \cellcolor[HTML]{F27F70}{36} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FFFFFF}{0}\\
-Earth and Planetary Sciences & 4,532 & \cellcolor[HTML]{F89874}{59} & \cellcolor[HTML]{F48671}{35} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FFFFFF}{0}\\
-Biochemistry, Genetics and Molecular Biology & 15,903 & \cellcolor[HTML]{F89874}{59} & \cellcolor[HTML]{F48671}{35} & \cellcolor[HTML]{FDE1A5}{4} & \cellcolor[HTML]{FAA476}{2}\\
-Immunology and Microbiology & 5,542 & \cellcolor[HTML]{F89874}{59} & \cellcolor[HTML]{F58D72}{34} & \cellcolor[HTML]{FCC98E}{6} & \cellcolor[HTML]{FAA476}{2}\\
-Environmental Science & 9,000 & \cellcolor[HTML]{F99D75}{58} & \cellcolor[HTML]{EF726E}{38} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FCDE9C}{1}\\
-Psychology & 3,087 & \cellcolor[HTML]{FAA376}{57} & \cellcolor[HTML]{EC686E}{40} & \cellcolor[HTML]{FFF0D2}{2} & \cellcolor[HTML]{FCDE9C}{1}\\
-Neuroscience & 5,906 & \cellcolor[HTML]{FBAA7A}{56} & \cellcolor[HTML]{EC686E}{40} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FCDE9C}{1}\\
-Arts and Humanities & 1,565 & \cellcolor[HTML]{FCB882}{54} & \cellcolor[HTML]{E4536F}{44} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FFFFFF}{0}\\
-Mathematics & 2,105 & \cellcolor[HTML]{FCBF87}{53} & \cellcolor[HTML]{EF726E}{38} & \cellcolor[HTML]{FBAF7D}{8} & \cellcolor[HTML]{FFFFFF}{0}\\
-Engineering & 8,405 & \cellcolor[HTML]{FCD395}{50} & \cellcolor[HTML]{E24C70}{46} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FCDE9C}{1}\\
-Materials Science & 4,838 & \cellcolor[HTML]{FCDA99}{49} & \cellcolor[HTML]{E14971}{47} & \cellcolor[HTML]{FDE1A5}{4} & \cellcolor[HTML]{FFFFFF}{0}\\
-Chemistry & 4,033 & \cellcolor[HTML]{FCE0A1}{48} & \cellcolor[HTML]{E24C70}{46} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FCDE9C}{1}\\
-Energy & 4,286 & \cellcolor[HTML]{FEE7B8}{46} & \cellcolor[HTML]{DC3977}{52} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FCDE9C}{1}\\
-Chemical Engineering & 2,977 & \cellcolor[HTML]{FFEBC4}{45} & \cellcolor[HTML]{E14971}{47} & \cellcolor[HTML]{FCBC85}{7} & \cellcolor[HTML]{FCDE9C}{1}\\
-Physics and Astronomy & 5,859 & \cellcolor[HTML]{FFFFFF}{40} & \cellcolor[HTML]{EF726E}{38} & \cellcolor[HTML]{DC3977}{22} & \cellcolor[HTML]{FFFFFF}{0}\\
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Health Sciences}}\\
+\hspace{1em}Dentistry & \cellcolor[HTML]{FFFFFF}{288} & \cellcolor[HTML]{EC686E}{68} & \cellcolor[HTML]{FFFAF1}{17} & \cellcolor[HTML]{F89774}{10} & \cellcolor[HTML]{DC3977}{5}\\
+\hspace{1em}Health Professions & \cellcolor[HTML]{FFFCF7}{674} & \cellcolor[HTML]{ED6D6E}{67} & \cellcolor[HTML]{FCD898}{24} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{DC3977}{5}\\
+\hspace{1em}Medicine & \cellcolor[HTML]{DC3977}{23623} & \cellcolor[HTML]{F0756E}{65} & \cellcolor[HTML]{FCD092}{25} & \cellcolor[HTML]{FCC98E}{6} & \cellcolor[HTML]{E34F6F}{4}\\
+\hspace{1em}Nursing & \cellcolor[HTML]{FFF6E5}{1504} & \cellcolor[HTML]{DC3977}{82} & \cellcolor[HTML]{FFFFFF}{16} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Veterinary & \cellcolor[HTML]{FFF2D9}{2091} & \cellcolor[HTML]{F58D72}{61} & \cellcolor[HTML]{FCD092}{25} & \cellcolor[HTML]{F1766E}{13} & \cellcolor[HTML]{FCDE9C}{1}\\
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Life Sciences}}\\
+\hspace{1em}Agricultural and Biological Sciences & \cellcolor[HTML]{FCB983}{7956} & \cellcolor[HTML]{F38170}{63} & \cellcolor[HTML]{F9A075}{31} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Biochemistry, Genetics and Molecular Biology & \cellcolor[HTML]{EC686E}{15903} & \cellcolor[HTML]{F89874}{59} & \cellcolor[HTML]{F48671}{35} & \cellcolor[HTML]{FDE1A5}{4} & \cellcolor[HTML]{FAA476}{2}\\
+\hspace{1em}Immunology and Microbiology & \cellcolor[HTML]{FCD797}{5542} & \cellcolor[HTML]{F89874}{59} & \cellcolor[HTML]{F58D72}{34} & \cellcolor[HTML]{FCC98E}{6} & \cellcolor[HTML]{FAA476}{2}\\
+\hspace{1em}Neuroscience & \cellcolor[HTML]{FCD294}{5906} & \cellcolor[HTML]{FBAA7A}{56} & \cellcolor[HTML]{EC686E}{40} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Pharmacology, Toxicology and Pharmaceutics & \cellcolor[HTML]{FEE4AF}{4075} & \cellcolor[HTML]{EA646F}{69} & \cellcolor[HTML]{FCC088}{27} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FAA476}{2}\\
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Health Sciences}}\\
+\hspace{1em}Chemical Engineering & \cellcolor[HTML]{FFECC6}{2977} & \cellcolor[HTML]{FFEBC4}{45} & \cellcolor[HTML]{E14971}{47} & \cellcolor[HTML]{FCBC85}{7} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Chemistry & \cellcolor[HTML]{FEE4AF}{4033} & \cellcolor[HTML]{FCE0A1}{48} & \cellcolor[HTML]{E24C70}{46} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Computer Science & \cellcolor[HTML]{FFEAC1}{3185} & \cellcolor[HTML]{F38170}{63} & \cellcolor[HTML]{F27F70}{36} & \cellcolor[HTML]{FFFFFF}{0} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Earth and Planetary Sciences & \cellcolor[HTML]{FDE1A5}{4532} & \cellcolor[HTML]{F89874}{59} & \cellcolor[HTML]{F48671}{35} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FFFFFF}{0}\\
+\hspace{1em}Energy & \cellcolor[HTML]{FDE3AA}{4286} & \cellcolor[HTML]{FEE7B8}{46} & \cellcolor[HTML]{DC3977}{52} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Engineering & \cellcolor[HTML]{FBB380}{8405} & \cellcolor[HTML]{FCD395}{50} & \cellcolor[HTML]{E24C70}{46} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Environmental Science & \cellcolor[HTML]{FBAC7B}{9000} & \cellcolor[HTML]{F99D75}{58} & \cellcolor[HTML]{EF726E}{38} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Materials Science & \cellcolor[HTML]{FCDF9E}{4838} & \cellcolor[HTML]{FCDA99}{49} & \cellcolor[HTML]{E14971}{47} & \cellcolor[HTML]{FDE1A5}{4} & \cellcolor[HTML]{FFFFFF}{0}\\
+\hspace{1em}Mathematics & \cellcolor[HTML]{FFF2D8}{2105} & \cellcolor[HTML]{FCBF87}{53} & \cellcolor[HTML]{EF726E}{38} & \cellcolor[HTML]{FBAF7D}{8} & \cellcolor[HTML]{FFFFFF}{0}\\
+\hspace{1em}Physics and Astronomy & \cellcolor[HTML]{FCD394}{5859} & \cellcolor[HTML]{FFFFFF}{40} & \cellcolor[HTML]{EF726E}{38} & \cellcolor[HTML]{DC3977}{22} & \cellcolor[HTML]{FFFFFF}{0}\\
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Health Sciences}}\\
+\hspace{1em}Arts and Humanities & \cellcolor[HTML]{FFF6E4}{1565} & \cellcolor[HTML]{FCB882}{54} & \cellcolor[HTML]{E4536F}{44} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FFFFFF}{0}\\
+\hspace{1em}Business, Management and Accounting & \cellcolor[HTML]{FFF4E0}{1751} & \cellcolor[HTML]{EC686E}{68} & \cellcolor[HTML]{FBAF7D}{29} & \cellcolor[HTML]{FFF0D2}{2} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Decision Sciences & \cellcolor[HTML]{FFFBF4}{785} & \cellcolor[HTML]{E6576F}{72} & \cellcolor[HTML]{FCC088}{27} & \cellcolor[HTML]{FFFFFF}{0} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Economics, Econometrics and Finance & \cellcolor[HTML]{FFF2D9}{2071} & \cellcolor[HTML]{F48771}{62} & \cellcolor[HTML]{F79373}{33} & \cellcolor[HTML]{FCD697}{5} & \cellcolor[HTML]{FFFFFF}{0}\\
+\hspace{1em}Psychology & \cellcolor[HTML]{FFEBC3}{3087} & \cellcolor[HTML]{FAA376}{57} & \cellcolor[HTML]{EC686E}{40} & \cellcolor[HTML]{FFF0D2}{2} & \cellcolor[HTML]{FCDE9C}{1}\\
+\hspace{1em}Social Sciences & \cellcolor[HTML]{FCCD91}{6306} & \cellcolor[HTML]{F79273}{60} & \cellcolor[HTML]{F27F70}{36} & \cellcolor[HTML]{FFE8BB}{3} & \cellcolor[HTML]{FFFFFF}{0}\\
 \bottomrule

### 2021-05-09 analysis/manuscript.Rmd
typo

@@
   pack_rows("Life Sciences", 6, 10) %>%
-  pack_rows("Health Sciences", 11, 20) %>%
-  pack_rows("Health Sciences", 21, 26) %>%
+  pack_rows("Physical Sciences", 11, 20) %>%
+  pack_rows("Social Sciences", 21, 26) %>%
   kable_styling(latex_options = c("scale_down"))

### 2021-05-09 analysis/manuscript.Rmd
group programmatically

--- 2021-05-09 analysis/manuscript.Rmd
@@
 
-subject_table %>%
+subject_table_ <- subject_table %>%
   group_by(subject_area) %>%
   arrange(top_level, .by_group = TRUE) %>% 
-  ungroup() %>%
+  ungroup() 
+subject_table_ %>%
   select(-subject_area) %>%
@@
   add_header_above(c(" " = 2, "Invoice channel (in %)" = 4)) %>%
-  # manual, puh, not good 
-  pack_rows("Health Sciences", 1, 5) %>%
-  pack_rows("Life Sciences", 6, 10) %>%
-  pack_rows("Health Sciences", 11, 20) %>%
-  pack_rows("Health Sciences", 21, 26) %>%
+  pack_rows(index = table(subject_table_$subject_area)) %>%
   kable_styling(latex_options = c("scale_down"))

### 2021-05-10 analysis/manuscript.Rmd
subject field naming

@@
   filter(subject_area != "Multidisciplinary")
+# distinct journals per subject
+journals_per_subject <- subject_all_years %>%
+  summarise(n_journals = n_distinct(issn)) %>%
+  mutate(field_label = glue::glue("{description} (n={n_journals})"))
 
@@
 # with labels
-subject_all_years <- 
-  inner_join(subject_all_years, subject_area_all_jn, by = "subject_area") %>%
-  mutate(subject_area = glue::glue("{subject_area} ({journals_n} Journals)"))
-
-p_subject <- ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
+subject_all_years_ <- 
+  inner_join(subject_all_years, journals_per_subject, by = "description")
+
+p_subject <- ggplot(subject_all_years_, aes(fct_reorder(field_label, n_oa / n, .fun = median), n_oa / n)) + 
   geom_boxplot(aes(color = subject_area)) +
@@
    kable_styling(latex_options = c("scale_down"))
+```
+
+```{r}
+# sector/discipline
+hybrid_invoice %>%
+  filter(!is.na(`Primary research focus`)) %>%
+  count(Sector, `Primary research focus`) %>%
+  mutate(prop = n / sum(n) * 100) %>% 
+  pivot_wider(id_cols = c("Primary research focus"), names_from = Sector, values_from = n) %>%
+  kableExtra::kbl(booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down"))
+
 ```
@@
 On the other hand, APCs that were invoiced as part of transformative agreements were not publicly shared.
+
+```{r ccby_vs_openapc}
+ggplot(ind_sponsor_df, aes(in_open_apc, cc_by, size = articles)) +
+  geom_point(alpha = 0.5) +
+  theme_minimal_grid() +
+  scale_x_continuous(limits = c(0,1), labels = scales::percent_format()) +
+  scale_y_continuous(limits = c(0,1), labels = scales::percent_format()) +
+  labs(y = "CC BY Share", x = "Spending info in Open APC", size = "Invoiced\nOA Articles") +
+  theme(legend.position="right", legend.justification = "right")
+```
 

### 2021-05-10 analysis/manuscript.tex
typo
subject field naming

@@
 \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Health Sciences}}\\
+\multicolumn{6}{l}{\textbf{Physical Sciences}}\\
 \hspace{1em}Chemical Engineering & \cellcolor[HTML]{FFECC6}{2977} & \cellcolor[HTML]{FFEBC4}{45} & \cellcolor[HTML]{E14971}{47} & \cellcolor[HTML]{FCBC85}{7} & \cellcolor[HTML]{FCDE9C}{1}\\
@@
 \addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Health Sciences}}\\
+\multicolumn{6}{l}{\textbf{Social Sciences}}\\
 \hspace{1em}Arts and Humanities & \cellcolor[HTML]{FFF6E4}{1565} & \cellcolor[HTML]{FCB882}{54} & \cellcolor[HTML]{E4536F}{44} & \cellcolor[HTML]{FFF7E8}{1} & \cellcolor[HTML]{FFFFFF}{0}\\
@@
 \midrule
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrr}
+\toprule
+Primary research focus & government & higher education & other & private non-profit\\
+\midrule
+Broad & 2849 & 4003 & NA & 6\\
+Health Science & 2706 & NA & 3 & 4678\\
+Life Sciences & 2591 & NA & NA & 3\\
+Physical Sciences & 5169 & 13 & NA & 14\\
+SSH & 1209 & 1 & NA & 1\\
 \bottomrule
@@
 shared.
+
+\begin{center}\includegraphics[width=0.7\linewidth,]{manuscript_files/figure-latex/ccby_vs_openapc-1} \end{center}
 

### 2021-05-17 analysis/manuscript.Rmd
revise geographical distribution on central invoicing receipients

@@
 hybrid_volume <- readr::read_csv(here::here("data", "hybrid_articles.csv")) 
-
+# add open apc info
+hybrid_volume <- hybrid_volume %>%
+      mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
 # aggregate data
@@
   # manual changes
-  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "European Union", Country)) %>%
+  mutate(Country = ifelse(grepl("EU", Country), "European Union", Country)) %>%
   mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%

### 2021-05-17 analysis/manuscript.tex
revise geographical distribution on central invoicing receipients

@@
 UK-based research funders and institutions (n=14,344; 59.2\%), followed
-by the Netherlands (n=2,835; 11.7\%) and the European Union (n=2,164;
-8.9\%; EU). Figure \ref{fig:invoice_sponsor_country} provides a closer
+by the Netherlands (n=2,835; 11.7\%) and the European Union (n=1,990;
+8.2\%; EU). Figure \ref{fig:invoice_sponsor_country} provides a closer
 look at the geographical distribution by license prevalence,

### 2021-05-19 analysis/manuscript.Rmd
sv tmp

@@
 
-The rise of open access (OA) to journal articles has added complexity to scholarly publishing, particularly concerning transparency of economic dimensions.
-Financial transparency in journal publishing has long been lacking since Big Deal subscription contracts between academic institutions and large academic publishers usually include confidentiality clauses that prohibit publicly sharing information about the scope and pricing of such agreements [@Frazier_2001; @Bergstrom_2014; @Larivi_re_2015].
-Further, the distributed process of paying for accessing and publishing scholarly articles, which can involve several actors at the same time, including authors, research organizations, libraries, and research funders, adds to the opacity of the financial flows of scholarly publishing [@Lawson_2016].
-
-The problem with this lack of transparency becomes more apparent in the case of hybrid OA.
-In this OA business model, individual articles can be made openly available for a fee, also known as article-processing charge (APC), while the journal as a whole remains subscription access.
-Hence, hybrid OA was originally introduced as a low-risk transitional model that allows journals to gradually convert to full OA and reduce subscription costs as the OA uptake increases [@Prosser_2003].
-Many large subscription publishers have since incorporated this OA model, and while the uptake among authors increased only slowly during the first decade [@Bj_rk_2012], it has gradually gained popularity in recent years.
-This development was driven by the implementation of OA policies by research institutions and funders, the allocation of OA funds, and formal agreements between publishers and institutions that allow affiliated authors to publish OA free-of-charge [@Laakso_2016].
-However, there have been concerns that the hybrid model allows for double-dipping—receiving two payments for one article, the APC and subscription fees [@Shieber_2009; @Prosser_2015].
-While publishers assure they combine subscription fees and APCs to adjust pricing [@Mittermaier_2015], without transparency around the uptake of hybrid OA and revenue streams, such claims are impossible to evaluate.
-
-With the recent introduction of transformative agreements, an evolving term for contracts that shift library spendings for subscriptions to OA [@Borrego_2020; @Hinchliffe_2019], the demand for publisher-provided data has even increased.
-If publishers would enable transparency about the OA uptake in hybrid journals and APC invoicing, academic libraries could improve their assessments and adjustments of contracts with publishers [@antelmann_2017; @Schimmer_2015].
-Likewise, research organizations and funders could improve compliance monitoring with OA mandates, while ensuring that the different actors were not charged multiple times [@Larivi_re_2018].
-However, as formal bibliometrics and financial studies have noted, the lack of publicly available and standardized data about hybrid OA publishing is considerable [@Laakso_2016; @Lawson_2015; @Pinfield_2016]
-
-This current study investigates the extent to which publisher-provided data can be utilized to research hybrid OA.
-The study takes Elsevier, the largest scholarly journal publisher in terms of article volume [@Larivi_re_2015], as the starting point for an empirical analysis of hybrid OA uptake and invoicing.
-Our approach is based on openly available data sources and brings together scholarly metadata from Crossref, a DOI registration agency for scholarly works, and machine-readable APC invoicing data from Elsevier OA full-texts.
-The so-compiled data enable us to analyze the development of Elsevier’s hybrid journal portfolio from 2015-19, and to compare it with that of Elsevier’s full and delayed OA program.
-Further, we investigate whether hybrid publishing fees were waived or charged.
-In particular, we identify academic consortia and research funders that were directly invoiced by Elsevier, and compare their hybrid OA rates by license and subject.
-As such, our findings have implications for research and OA policy and highlight the potential of publisher-provided data for large-scale, instantaneous analysis of hybrid OA uptake and invoicing.
+The rise of open access (OA) has added complexity to
+scholarly publishing, particularly concerning transparency of economic
+dimensions. Financial transparency in journal publishing has long been
+lacking because Big Deal subscription contracts between academic
+institutions and large academic publishers usually include
+confidentiality clauses that prohibit publicly sharing
+the scope or pricing of such agreements [@Frazier_2001; @Bergstrom_2014; @Larivi_re_2015]. Further, the distributed payment processes
+to access and publish scholarly articles add to the opacity of
+the financial flows of scholarly publishing [@Lawson_2016] as they can
+involve several actors simultaneously, including authors, research
+institutions, libraries, and funders.
+
+The problem with this lack of transparency becomes more apparent in the
+case of hybrid OA. In this OA business model, individual articles can be
+made openly available by paying an article processing charge
+(APC), while the journal as a whole remains subscription-access. Hence,
+hybrid OA was originally introduced as a low-risk transitional model
+that allows journals to gradually convert to full OA and reduce
+subscription costs as the OA uptake increases [@Prosser_2003]. Many
+large subscription publishers have since incorporated hybrid OA, but there have been concerns that the hybrid model allows for double dipping---receiving two payments for one article, the APC and
+subscription fees [@Shieber_2009; @Prosser_2015]. While publishers
+assure they adjust their pricing
+[@Mittermaier_2015], without transparency around the uptake of hybrid OA
+and both revenue streams, such claims are impossible to evaluate.
+This is especially concerning because hybrid OA has recently gained popularity after a decade of only slow uptake [@Bj_rk_2012].
+The main drivers of this development have been research institutions and funders that implemented  OA policies, allocated OA funds, and established
+formal agreements with publishers that allow
+affiliated authors to publish OA free-of-charge [@Laakso_2016].
+
+The recent introduction of transformative agreements, an evolving
+concept describing contracts that shift library spending from subscriptions to OA
+[@Borrego_2020; @Hinchliffe_2019], might perpetuate the lack of transparency because pricing is often based on traditional subscription costs. Hence, the demand for
+publisher-provided data has increased. More
+transparency about hybrid OA uptake and funding could  facilitate the assessment and adjustment of publisher contracts [@Schimmer_2015], enhance OA mandate compliance monitoring, and avoid double-dipping [@Larivi_re_2018].
+However, previous studies have noted an absence of transparency in hybrid OA publishing and a considerable lack of
+publicly available and standardized data [@Laakso_2016; @Lawson_2015; @Pinfield_2016].
+
+This paper investigates to what extent publisher-provided
+data can increase the transparency around hybrid OA by taking Elsevier,
+the largest scholarly journal publisher [@Larivi_re_2015], as the starting point for an empirical
+analysis of hybrid OA uptake and invoicing. Our approach is based on
+openly available data sources and combines scholarly metadata
+from Crossref, a DOI registration agency, with
+ APC invoicing data from Elsevier. The
+compiled data enable us to analyze the 
+hybrid OA uptake in Elsevier journals between  2015 and 2019 by license and subject, and to compare it with
+Elsevier's full and delayed OA program. Further, we determine whether
+hybrid APCs were waived or charged to authors or as part of publishing agreements. In the latter case, we 
+examine the invoiced academic consortia and research funders. As such, our findings have implications for research and OA
+policy and highlight the potential of publisher-provided data for
+large-scale, instantaneous analysis of hybrid OA uptake and invoicing.
 
@@
 
-The ability to identify and discover hybrid OA articles plays a fundamental role in monitoring hybrid OA, yet time and again, previous studies have underscored the difficulties involved in this (cf.
-[@Akbaritabar_2019; @van_Leeuwen_2018; @Laakso_2016; @Emery_2013; @Bj_rk_2012].
-Insights into the volume of hybrid OA and the various levels and sources of APC funding could create a better understanding of the costs involved in scholarly publishing and, thus, strengthen institutions' bargaining power in negotiations with publishers, and promote greater equity in scholarly publishing.
-In the following section, we examine how previous research has tackled the challenge of identifying hybrid OA content and what these studies taught us about the uptake and funding sources of hybrid OA.
-
 ## Prevalence / uptake studies
@@
 
-In the first article-level study on hybrid OA, @Laakso_2016 reported that between 2007 and 2013, the five largest publishers--Elsevier, Sage, Springer, Taylor & Francis, and Wiley--recorded growing numbers of hybrid journals that coincided with a twenty-fold increase in hybrid OA articles (from 666 to 13,994).
-As a result, hybrid journals with at least one OA article more than doubled in number (1,082 in 2009, 2,714 in 2013) but fell by 31% relative to the number of hybrid journals in total.
-Although Laakso and Björk's (2016) presented a comprehensive overview tracking the growth of hybrid OA over time and across multiple publishers, their exploratory approach is unsuitable for repeated use as it required a considerable amount of manual data collection and cleaning.
-Minimizing such manual tasks, @Nelson_2017 narrowed their focus to one publisher--the American Chemical Society (ACS)--and requested a list of all hybrid OA articles published from 2006-2011 directly from the ACS.
-To assess the hybrid OA uptake, the authors aggregated and compared the number of OA (n=814) and subscription articles (n=27,621), finding that over the five years, 2.9% of ACS articles were published as hybrid OA.
-
-Contrary to the previous two studies, @Kirkman_2018 took a research funder as a starting point, reporting a 26% share of hybrid OA for articles funded by the Australian National Health and Medical Research Council from 2012-2014 (816 of 3,190).
-Kirkman collected research articles using the Funding Agency and Funding Text search fields on Web of Science, and classified articles as OA when they were freely available on the publisher website, and further distinguished hybrid OA articles based on journal-level information from the publisher website, such as hybrid OA journal and APC lists.
-Another approach is presented by @P_l_nen_2020, who used current research information system (CRIS) data—institutional publication data—to study the extent of OA among Finnish research publications from 2016-2017 (n=34,507 research articles).
-As part of the analysis, the authors identified hybrid OA articles through a dedicated OA status metadata field that has been mandatory to report in Finland since 2016 and found a 7% share of hybrid OA.
-However, the drawbacks of Kirkman's (2018) and Pölönen et al.'s (2020) approaches are that such data sources might not be openly available or, regarding CRIS data, might not be available at all since not all countries maintain comprehensive institutional publication data.
-
-Other studies assessed the share of (hybrid) OA for the entire corpus of scholarly research articles.
-For instance, using research articles with an assigned DOI and indexed in the Web of Science Core Collection (n=2,610,305) as a benchmark, @Mart_n_Mart_n_2018 assessed how many research articles from 2009 and 2014 are freely available through Google Scholar.
-After identifying OA articles through licensing information from Crossref, the authors estimated the share of hybrid OA at around 0.5% in 2009 and 1.5% in 2014.
-In a similar vein, @Piwowar_2018 drew on Crossref to identify content licenses.
-The authors followed a data-driven approach based on Unpaywall data, which is openly available and contained 66 million Crossref DOIs at the time, to determine how much of the scholarly literature is OA.
-Based on a random sample of 100,000 Crossref DOIs, the authors found 3.6% of hybrid OA, while pointing towards a growing trend in recent years that recorded a hybrid OA share of 9.4% of all articles published in 2015.
-
-Since its launch, Unpaywall has become widely used as an OA evidence source in bibliometric research and rankings[@Robinson_Garcia_2020; @Huang_2020].
-However, results cannot easily be compared with each other over time because of Unpaywall's ongoing changes due to non-standardized data to discover hybrid OA.
-These changes mainly reflect the growing albeit incomprehensive license information in Crossref metadata.
-To overcome these shortcomings, the developers extended previous web-scraping approaches to better distinguish between immediate and delayed OA based on licenses at the date of observation @Piwowar_2019.
-Relatedly, in October 2020, Elsevier's license for delayed OA content was added to Unpaywall's list of license patterns [@Unpaywall_oa_license].
-At the journal-level, a more fine-grained identification of full OA journals using OA article information was developed to better distinguish them from subscription journals including hybrid journals [@Unpaywall_types].
-With its servie unsub, Unpaywall recently launched a paid service, where libraries can compare hybrid OA uptake rates with internal financial and usage data to assess the value of journal subscriptions [@Chawla_2020].
+In the first article-level study on hybrid OA, @Laakso_2016
+reported that between 2007 and 2013, the five largest
+publishers--Elsevier, Sage, Springer, Taylor \& Francis, and
+Wiley--recorded growing numbers of hybrid journals that coincided with a
+twenty-fold increase in hybrid OA (from 666 to 13,994). As a
+result, hybrid journals with at least one OA article more than doubled
+in number (1,082 in 2009, 2,714 in 2013) but fell by 31\% relative to
+the total number of hybrid journals. Laakso and Björk's
+(2016) exploratory approach yielded a comprehensive overview tracking the growth of hybrid
+OA over time and across multiple publishers, but due to the amount of
+manual data collection and cleaning it
+is unsuitable for repeated use. Minimizing such manual tasks,
+Nelson \& Eggett (2017) narrowed their focus to one publisher--the
+American Chemical Society (ACS)--and requested a list of all hybrid OA
+articles published from 2006-2011 from the ACS. To assess the
+hybrid OA uptake, the authors aggregated and compared the number of OA
+(n=814) and subscription articles (n=27,621), finding that over the five
+years, 2.9\% of ACS articles were published as hybrid OA.
+
+Contrary to the previous two studies, @Kirkman_2018 took a research
+funder as a starting point, reporting a 26\% share of hybrid OA for
+articles funded by the Australian National Health and Medical Research
+Council from 2012-2014 (816 of 3,190). Kirkman collected research
+articles using the Funding Agency and Funding Text search fields on Web
+of Science, and classified articles as OA when they were freely
+available on the publisher website, and further distinguished hybrid OA
+articles based on journal-level information from the publisher website,
+such as hybrid OA journal and APC lists. Another approach is presented
+by @P_l_nen_2020, who used current research information system
+(CRIS) data---institutional publication data---to study the extent of OA
+among Finnish research publications from 2016-2017 (n=34,507 research
+articles). As part of the analysis, the authors identified hybrid OA
+articles through a dedicated OA status metadata field that has been
+mandatory to report in Finland since 2016 and found a 7\% share of
+hybrid OA. However, the drawbacks of Kirkman's (2018) and Pölönen et
+al.'s (2020) approaches are that such data sources might not be openly
+available or, regarding CRIS data, might not be available at all because
+not all countries maintain comprehensive institutional publication data.
+
+Other studies assessed the share of (hybrid) OA for the entire corpus of research articles. Using DOI-assigned research articles indexed in the Web of Science Core Collection
+ as a benchmark, @Mart_n_Mart_n_2018 investigated how
+many research articles from 2009 and 2014 (n=2,610,305 in total) are freely available through
+Google Scholar. After identifying OA articles through licensing
+information from Crossref, the authors estimated the share of hybrid OA
+at around 0.5\% in 2009 and 1.5\% in 2014. Similarly, in a study assessing the OA level among all scholarly articles and that experienced by Unpaywall users, Piwowar et
+al. (2018) obtained license information through Crossref and web-scraping. The study found 3.6\% of 100,000 random Crossref DOIs were hybrid OA
+and pointed towards a growing trend in recent years that recorded a
+hybrid OA share of 9.4\% of all articles published in 2015.
+
+Since its launch, Unpaywall has become widely used in bibliometric research and rankings [@Robinson_Garcia_2020; @Huang_2020].
+However, the lack of standardized and comprehensive publisher-provided data has required several updates to Unpaywall to improve hybrid OA identification and differentiation from  delayed OA [@Piwowar_2019; @Unpaywall_oa_license; @Unpaywall_types], illustrating ongoing challenges in tracking and comparing hybrid OA prevalence over time.
+After changes in Unpaywall's approach to detect hybrid OA, Piwowar et al.
+(2019) estimated a hybrid OA share of 4\% amongst all articles published by July 2019.
 
@@
 
-Some countries like Sweden and the United Kingdom (UK) have negotiated national publishing agreements that allow authors to publish hybrid OA at a discounted fee or for free.
-These agreements can serve as the basis for studies on the uptake of hybrid OA or cost-benefit analyses (cf.
-@Olsson1271866; @Lawson_2018) but are often limited in scope as they tend to focus on a single country and rely on data that seldom is publicly available.
-Hence, while being valuable for specific use-cases, such studies can neither provide a comprehensive understanding of how the scholarly publishing system functions beyond national boundaries nor facilitate collective sharing of financial information about hybrid OA.
-
-Pinfield et al.'s (2016) analysis of APC payment records provided by 23 UK higher education institutions revealed a sharp increase in central payments from 2007-2013, which was largely attributed to the introduction of block grants by Research Councils UK (RCUK) and non-compliance sanctions by the Wellcome Trust.
-Moreover, the study showed that OA fees were paid almost exclusively through RCUK and Wellcome block grants (92%), and only a small number of APCs were paid through internal funding (7%).
-In contrast, a recent Springer Nature survey found that authors draw on a range of funding sources to cover OA fees, such as dedicated institutional OA funds, block grants, OA agreements, or research grants [@Monaghan_2020].
-Most hybrid OA authors were supported through dedicated institutional OA funds (43%, excluding block grants) and OA agreements with Springer Nature (34%).
-The differences between these two studies may be due to changes in policy and funding arrangements—in particular, the introduction of OA agreements—since @Pinfield_2016, or due to Monaghan et al.'s (2020) more regionally diverse sample, indicating regional differences in APC funding.
-
-Regional and policy differences in APC payments also came to light in a study by @Jahn_2016 that analyzed APC records from 30 German higher education and research institutions crowd-sourced by the Open APC initiative, the Austrian Science Fund (FWF), Jisc, and the Wellcome Trust.
-In particular, the study revealed large differences in the amount of hybrid OA funded from 2014-2015.
-Whilst hybrid OA accounted for less than 1% of APCs paid by German institutions (23 of 3,846), the three non-German research funders recorded a hybrid share of75% (11,533 of 15,779).
-According to @Jahn_2016, this could point towards differences in science policy, such as hybrid OA being supported by the three non-German research funders but not by Germany's largest national funder (DFG).
-Another possibility is that German hybrid OA fees were paid from budgets not reported to the Open APC initiative, from where the authors acquired the German APC data.
-Among these unreported funds are research grants and research unit budgets, which author surveys identified  as funding sources for hybrid OA fees [@Monaghan_2020; @Graaf_2017].
-As such, Jahn and Tullney's (2016) findings could reflect the complexities and potential limitations of institutional OA spending data that Pinfield et al.
-(2016) and Monaghan et al.
-(2020) attributed to incomplete or missing records.
-
-As this brief review has shown, the most comprehensive studies approach hybrid OA funding from an institutional perspective that focuses on research institutions or funders and draws on internal records or crowd-sourced data like from the Open APC initiative.
-The advantages of Open APC data are that they reflect the actual APCs paid rather than list prices; they are not limited to certain publishers or journals; they are provided in a standardized format regardless of the publisher.
-On the other hand, the drawbacks include the limited number of participating institutions that are almost exclusively based in Europe and the dependency on participating organizations for data integrity [@Pieper_2018].
-Moreover, since Open APC data does not include waivers or free APCs, it cannot present a complete picture of all publishing activities and expenses.
-This, however, would pose less of an issue when analyzing just hybrid OA.
-
-## The role of publisher-provided data
-
-Uptake and financial studies have OA policy implications for hybrid OA.
-Most prominently, the cOAlition S, an international network of research funders, released its widely discussed Plan S.
-In September 2018, the cOAlition S announced that from 2021 onwards, signatories would no longer fund articles in hybrid OA journals.
-This principle includes so-called "mirror" and sister journals--hybrid journals with fully OA counterparts or split into their fully OA and subscription parts—that have launched since Plan S was announced but makes an exception for hybrid journals with transformative agreements [@Plan_s; @Frantsv_g_2019; @McNutt_2019; @Harrison_2019].
-The Plan S refers to the Efficiency and Standards for Article Charges (ESAC) guidelines [@Geschuhn_2017] as critical component of transformative agreements, emphasizing the need for publisher-provided metadata to address the lack of transparency surrounding hybrid OA.
-However, studies confirm that publisher with such agreements lack essential data about hybrid OA uptake and invoicing, or are not able or willing to provide such information publicly and instantaneously [@Monaghan_2020; @Marques_2019], which, in turn, not just limit the monitoring of hybrid OA, but also research studies on its size and financial implications.
-
-Here, we focus on Elsevier.
-The publisher provides article-level metadata to Crossref, including license information, which can be used to identify OA articles, and links to HTML and XML full-text versions for text and data mining purposes.
-Surprisingly, besides the full-text article, the XML files also contain invoicing information and the article's OA status.
-Table \ref{tab:els_xml} summarizes the key metadata elements, which comprise the type of entity to which the APC was invoiced (`openaccessSponsorType`) and the name of that entity if available (`openaccessSponsorName`).
-Moreover, the article-level metadata indicates if OA was provided through Elsevier's Open Archive program (`openArchiveArticle`), which enables free access to participating subscription journals after an embargo period (i.e., delayed OA; @Els_Archive).
-As we have been unable to identify similar efforts by other large publishers (e.g., Springer Nature), this increased transparency through openly available bibliometric and financial data at the article level seems unique to Elsevier.
+Pinfield et al.'s (2016) analysis of APC payment records provided by 23
+United Kingdom (UK) higher education institutions revealed a sharp increase in central
+payments from 2007-2013, which was largely attributed to the
+introduction of block grants by Research Councils UK (RCUK) and
+non-compliance sanctions by the Wellcome Trust. Moreover, the study
+showed that OA fees were paid almost exclusively through block grants (92\%), and only a small number of APCs were paid
+through internal funding (7\%). In contrast, a recent Springer Nature
+survey found that authors draw on a range of funding sources to cover OA
+fees, such as dedicated institutional OA funds, block grants, OA
+agreements, or research grants (Monaghan et al., 2020). Most hybrid OA
+authors were supported through dedicated institutional OA funds (43\%,
+excluding block grants) and OA agreements with Springer Nature (34\%).
+The differences between these two studies might be due to Monaghan et al.'s
+(2020) focus on Springer Nature articles or due to} different policy and funding arrangements---considering the introduction of OA
+agreements since  @Pinfield_2016 and Monaghan et al.'s
+(2020) more regionally diverse sample.
+
+Regional and policy differences in APC payments also came to light in a
+study by @Jahn_2016 that analyzed APC records from 30 German
+higher education and research institutions, the Austrian Science Fund (FWF), Jisc, and the Wellcome
+Trust. In particular, the study revealed large differences in the amount
+of hybrid OA funded from 2014-2015. Whilst hybrid OA accounted for less
+than 1\% of APCs paid by German institutions (23 of 3,846), the three
+non-German research funders recorded a hybrid share of 75\% (11,533 of
+15,779). According to @Jahn_2016, this could point towards
+differences in science policy, such as hybrid OA being supported by the
+three non-German research funders but not by Germany's largest national
+funder, the German Research Foundation (DFG). Another possibility is that German hybrid OA fees were
+paid from budgets not reported to the Open APC initiative, a crowd-sourcing effort [@Pieper_2018] from where
+the authors acquired data. Among these unreported funds
+are research grants and research unit budgets, which author surveys
+identified as APC funding sources [@Monaghan_2020; @Graaf_2017]. As such, Jahn and Tullney's (2016) findings could reflect
+the complexities and potential limitations of institutional OA spending
+data that @Pinfield_2016 and @Monaghan_2020 attributed
+to incomplete or missing records.
+
+In recent years, national consortia in Europe have negotiated publishing agreements covering hybrid OA fees for affiliated authors (Borrego et al., 2020).
+While these have improved invoicing workflows, internal assessments of hybrid OA uptake and invoicing remain challenging because transparent and comparative data have largely remained absent [@Marques_2020].
+Such publicly available information might be scarce because publishers and institutions lack or withhold such data, or due to confidentiality clauses [@Marques_2020; @Monaghan_2020].
+Several European consortia have collaborated through the Efficiency and Standards for Article Charges (ESAC) Initiative and Knowledge Exchange to standardize metadata requirements across publishers as great variation has been found between different countries and publishers [@Marques_2019].
+
+## Research questions and aims
+
+In this paper, we focus on Elsevier, a prominent example in recent hybrid OA uptake and financial studies [@Pinfield_2016; @@Laakso_2016].
+Elsevier's OA portfolio presents the challenges in examining hybrid OA described above.
+For instance, distinguishing between different OA types.
+Elsevier supports delayed [@Els_Archive], hybrid, and full OA, including so-called mirror journals---full OA counterparts of hybrid journals addressing OA policies opposed to hybrid OA [@Harrison_2019].
+Further, Elsevier processes APC invoices through various channels, such as agreements with research funders and library consortia or, in the absence thereof, the authors (Elsevier,
+n.d.-a, n.d.-c). Surprisingly, we found Elsevier article-level metadata embedded in XML full-texts indicating the articles' OA status and invoicing.  Leveraging this publicly available data, we 
+address the lack of transparency around hybrid OA noted in previous studies.
+In particular, we use this novel approach to answer the following
+questions:
+
+* What was the uptake of Elsevier's hybrid OA publishing option between 2015 and 2019?
+* Through which channels were hybrid APCs invoiced, and who were the recipients?
+
+# Methodology
+
+For this study, we collected data relating to Elsevier's hybrid OA
+option by drawing on multiple freely available data sources. We
+identified Elsevier hybrid journals through Elsevier's APC list and
+supplemented our sample with Crossref metadata and text-mined invoicing
+data to investigate the invoicing of immediate OA articles provided
+under a Creative Commons (CC) license in subscription-based journals. Figure
+\ref{fig:workflow} visualizes the automated workflow we used to collect
+data from Elsevier and Crossref.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.95\textwidth]{../figure/els_flow.pdf}
+
+}
+
+\caption{Data Collection Workflow to Obtain Article-Level OA and Invoicing Data.
+}\label{fig:workflow}
+\end{figure}
+
+First, we identified hybrid journals through an Elsevier APC price list
+from May 2020 provided by @Matthias_2020. We excluded journals that transitioned from hybrid to full OA and
+reverse-flip journals that flipped 
+from full to hybrid OA [@Matthias_2019]. Overall, we identified 1,970 unique hybrid journals that published between 2015 and 2019.
+
+Next, we used an openly available Crossref database snapshot [@Crossref_2020], which contains all Crossref records registered until March 2020,
+to calculate the journals' combined article volume for the five-year
+period from 2015-2019. We only included articles published in regular
+issues aside from supplements containing conference contributions like
+meeting abstracts, indicated by non-numeric pagination. We excluded
+non-scholarly journal content, such as the table of contents, following Unpaywall's paratext recognition approach
+ [@Unpaywall_para], which we expanded to include patterns indicating
+corrections. Furthermore, we categorized the articles by subject
+according to the All Science Journal Classification code (ASJC) and
+added citation indicators from the Scopus Source Title list (version
+June 2020) based on the journal they were published in.
+
+We identified OA articles
+through CC licenses in Crossref metadata records (Hendricks et al., 2020) and then downloaded the XML
+version of all CC-licensed articles published in a hybrid journal. From
+the XML files, we obtained the articles' OA status and invoicing information (see Table \ref{tab:els_xml}). We determined whether articles were immediate or delayed
+OA using the XML node \texttt{openArchiveArticle}, which indicates if an article was made freely available under Elsevier's Open Archive program, and measured the
+uptake of hybrid OA.
+Moreover, we obtained the invoice channels and recipients of hybrid OA APCs.
+Using the \texttt{openaccessSponsorType} node, we distinguished between four invoice channels, including
+invoices billed to authors, as part of publishing
+agreements with funding bodies (cf. @Els_Agreements), exempted through fee waivers (e.g., in ``cases of genuine need'' or due to
+society or university sponsorships, cf.~Elsevier,
+n.d.-c), and other types not
+specified by Elsevier. Finally, when hybrid APCs were invoiced as part
+of agreements, we identified invoice recipients through the
+\texttt{openaccessSponsorName} node.
+It is worth noting that while this article focuses on OA articles published in hybrid journals, the same data is also available for articles published in full OA journals.
+Elsevier does not provide the invoiced APC in either case.
 
 \begin{table}[H]
-\caption{\label{tab:els_xml} Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019}
+\caption{\label{tab:els_xml}Metadata in Elsevier XML Full-Texts.}
 \centering
@@
 \toprule
-XML Node & Description\\
+XML node & Description\\
 \midrule
-\texttt{openaccessArticle} & Indicates if the article is open
-access. \\
-\texttt{openaccessSponsorType} & Invoice recipient type.\\
-\texttt{openaccessSponsorName} & Entity that received Elsevier's APC
-invoice.\\
-\texttt{openArchiveArticle} & Indicates if open access was provided through
-Elsevier's Open Archive programme.\tabularnewline \\
+\texttt{openaccessArticle} & Was the article open
+access? \\
+\texttt{openaccessSponsorType} & Invoice channel.\\
+\texttt{openaccessSponsorName} & Invoice recipient.\\
+\texttt{openArchiveArticle} & Was open access provided through
+Elsevier's Open Archive program?\tabularnewline \\
 \bottomrule
@@
 
-In this study, we utilize Elsevier's publicly available metadata to analyze the uptake and invoicing of their hybrid OA publishing option.
-In particular, we use this novel methodology to answer the following questions:
-
-- What was the uptake of Elsevier's hybrid OA publishing option between 2015 and 2019?
-
-- Through which channels were hybrid APCs invoiced, and who were the recipients?
-
-# Methodology
-
-For this study, we collected data relating to Elsevier's hybrid OA option by drawing on multiple freely available data sources.
-We identified Elsevier hybrid journals through Elsevier’s APC list and supplemented our sample with Crossref metadata and text-mined invoicing data to investigate the financing of immediate OA articles provided under a Creative Commons license in subscription-based journals.
-Figure \ref{fig:workflow} visualizes the automated workflow we used to collect data from Elsevier and Crossref.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.95\textwidth]{../figure/els_flow.pdf}
-
-}
-
-\caption{Data Analytics Workflow to obtain article-level OA and invoicing data using freely available data sources provided by Elsevier and Crossref.
-}\label{fig:workflow}
-\end{figure}
-
-First, we obtained hybrid journals through an Elsevier APC price list from May 2020 provided as captured by @Matthias_2019.
-We only included hybrid journals that were published under the hybrid model throughout the entire period of our analysis.
-Hence, we excluded journals that flipped from OA to hybrid and from hybrid to OA [@Matthias_2020].
-This approach allowed us to ensure that the article-level differences we observed were not due changes at the journal-level (i.e., increases or decreases in number).
-Overall, we identified 1,970 unique hybrid journals.
-
-Next, we used an openly available Crossref database snapshot [@Crossref_2020], which contains all Crossref records registered until March 2020, to calculate the journals' combined article volume for the five-year period from 2015-2019.
-We only included articles published in regular issues aside from supplements containing conference contributions like meeting abstracts, indicated by non-numeric pagination.
-We excluded non-scholarly journal content, such as the table of contents and lists of reviewers, following Unpaywall's paratext recognition approach [@Unpaywall_para], which we expanded to include patterns indicating corrections.
-Furthermore, we categorized the articles by subject according to the All Science Journal Classification code (ASJC) and added citation indicators from the Scopus Source Title list (version June 2020) based on the journal they were published in.
-
-Crossref metadata also contains license information that indicates if and under which terms the publisher has made articles openly available on their website [@Hendricks_2020].
-We identified OA articles through Creative Commons (CC) licenses and then downloaded the XML version of all CC-licensed articles published in a hybrid journal.
-From the XML files, we determined whether or not the articles were immediate OA using the XML node `openArchiveArticle` and measured the uptake of hybrid OA through three bibliometric indicators [@Nelson_2017; @Laakso_2016]-—the number of hybrid journals with at least one OA article, the number of hybrid OA articles, and the number of hybrid OA articles relative to the number of closed-accessed articles in hybrid journals with at least one OA article.
-
-Moreover, we obtained APC invoicing metadata from the XML (see Table 1), which enabled us to identify the invoice channels and recipients of hybrid OA APCs.
-Based on the `openaccessSponsorType` node in the XML files, we distinguish between four invoice channels, including invoices billed to the corresponding author, as part of publishing agreements with funding bodies (hereinafter "agreements"; cf.
-@Els_Agreements), fee waivers (e.g., in "cases of genuine need" or due to society or university sponsorships, cf.
-ref), and other types not specified by Elsevier.
-Finally, when hybrid APCs were invoiced as part of OA agreements, we identified invoice recipients through the `openaccessSponsorName` node.
-
-We manually classified invoice recipients based on their institutional sectors, countries, and primary research areas.
-Following the OECD's Frascati Manual [@Frascati, p.
-91], we coded for four sectors, business enterprise, government, higher education, and private non-profit.
-The government sector consists of federal, state, and municipal governments, plus non-profit institutions controlled by government units [@Frascati, p.
-100].
-The higher education sector comprises institutions that perform research and development and provide formal tertiary education services, such as universities or research institutes [@Frascati, p.
-101].
-Legal or social entities that produce goods and services without generating profit, such as learned societies or charities, constitute the private non-profit sector [@Frascati, p.
-288].
-Due to the low article volume, we combined invoice recipients Elsevier listed as "authors" and "third-party sponsor", and those we had coded as business enterprises [@Frascati, p.
-97], into "Others".
-Moreover, we categorized invoice recipients according to the country or countries representing their scope of funding and based on the following primary research areas health sciences, life sciences, physical sciences and mathematics, social sciences and humanities, unknown, and broad (i.e., multiple research areas).
-
-It is important to note that we could not find any official comprehensive documentation of Elsevier's XML schema , at the time of writing.
-However, our conceptualization of the variables and interpretation of the data was informed by a comprehensive review of Elsevier's websites about OA, including a list of funding agreements and details about Elsevier's Open Archive program,and Elsevier’s article submission process in the context of dedicated OA agreements [@lund; @vsnu; @Els_videos].
-For validation, we matched the invoicing data provided by Elsevier to data from  the Open APC initiative.
-
-Throughout the mostly automated data gathering and analysis process, we used tools from the Tidyverse [@tidyverse] for the R programming language [@r].
-To allow for efficient data manipulation and retrieval, we imported the Crossref dump to Google BigQuery, applying the rcrossref [@rcrossref] parsers to extract relevant metadata fields.
-We interfaced Google BigQuery with SQL with the bigrquery client [@bigrquery].
-We used crminer [@crminer] to obtain the XML files from Elsevier and the xml2 package [@xml2] to parse relevant metadata.
-To ensure our results are reproducible, we provided all code for obtaining the data and performing the analyses with detailed documentation as RMarkdown notebooks on GitHub (<https://github.com/njahn82/elsevier_hybrid_invoicing>).
+We manually classified invoice recipients based on their institutional
+sectors, countries, and primary research areas. Following the OECD's
+Frascati Manual [@Frascati, p. 91], we coded for four sectors: business
+enterprise, government, higher education, and private non-profit. 
+Due to the low article volume,
+we combined the business enterprise sector with invoice recipients Elsevier listed as "authors" and
+"third-party sponsor"
+ into "Others". Moreover, we categorized invoice
+recipients according to the countries representing their
+scope of funding and based on the following primary research areas
+health sciences, life sciences, physical sciences and mathematics,
+social sciences and humanities, broad (i.e., multiple
+research areas), and unknown. Further, we compared Elsevier's invoicing data
+with institutional spending data from the Open APC initiative.
+
+Throughout this mostly automated data gathering and analysis process, we
+used tools from the Tidyverse [@tidyverse] for the R
+programming language [@r]. To allow for efficient data
+manipulation and retrieval, we imported the Crossref dump to Google
+BigQuery, applying the rcrossref [@rcrossref] parsers to
+extract relevant metadata fields. We used crminer
+[@crminer] to obtain the XML-full texts from Elsevier.
 
@@
 
-This section first presents the results of our analysis of hybrid OA uptake in Elsevier's journal portfolio with a view to licensing, disciplinary differences, and citation impact.
-Then, we present a descriptive analysis of Elsevier's invoicing data, highlighting licensing and disciplinary differences for different invoicing channels, institutional sectors, and individual invoice recipients.
+This section first presents the results of our analysis of hybrid OA
+uptake in Elsevier's journal portfolio with a view to licensing,
+disciplinary differences, and citation impact. Then, we present a
+descriptive analysis of Elsevier's invoicing data, highlighting
+licensing and disciplinary differences for invoicing channels and invoice recipients.
 
@@
 Together these journals published `r five_years_ind %>% pull(oa_n)` OA articles, which represents `r five_years_ind %>% pull(oa_uptake)*100`% of their total article volume (n=`r five_years_ind %>% pull(articles_n)`).
-Table \ref{tab:overview} presents these findings for each year and the aggregated five-year period, illustrating a moderate growth in hybrid OA over time.
-With each year, the number of hybrid journals with at least one OA article increased, as did the number of immediate OA articles in these journals, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019.
-However, since the total number of articles published in these journals also grew over time, the relative share of OA articles in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.
+Table \ref{tab:overview} presents  findings for each year and the aggregated five-year period, illustrating moderate growth in hybrid OA over time.
+Each year, the number of hybrid journals
+with at least one OA article increased, as did the number of immediate
+OA articles in these journals, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019.
+However, since the journals' total article output also grew over time, the relative share of OA articles
+in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.
 
@@
   ~ind_group, ~ind, ~` `,
-    "Elsevier Hybrid Journals with ≥ 1 OA article", "n_journals", "Total",
-  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_n", "Total ",
-  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_mean", "Avg.",
-  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_sd", "SD",
-  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_n", "Total",
-  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_mean", "Avg.",
-  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_sd", "SD",
-  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake", "Total (%)",
-  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_mean", "Avg.",
-  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_sd", "SD"
+    "Elsevier Hybrid Journals with ≥ 1 OA article (n)", "n_journals", "Total",
+  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "articles_n", "Total ",
+  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "articles_mean", "Avg.",
+  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "articles_sd", "SD",
+  "OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "oa_n", "Total",
+  "OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "oa_mean", "Mean",
+  "OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", "oa_sd", "SD",
+  "Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", "oa_uptake", "Total",
+  "Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", "oa_uptake_mean", "Mean",
+  "Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", "oa_uptake_sd", "SD"
 )
@@
   kableExtra::kbl(booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
-                  caption = "Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.",
+                  caption = "Elsevier Hybrid OA Uptake  2015-2019.",
                   position = "H"
@@
   kable_styling() %>%
-  pack_rows("Elsevier Hybrid Journals with ≥ 1 OA article", 1, 1) %>%
-  pack_rows("Articles in Elsevier Hybrid Journals with ≥ 1 OA Article", 2, 4) %>%
-  pack_rows("OA Articles in Elsevier Hybrid Journals with ≥ 1 OA Article", 5, 7) %>%
-  pack_rows("OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA Article", 8, 10)
+  pack_rows("Elsevier Hybrid Journals with ≥ 1 OA article (n)", 1, 1) %>%
+  pack_rows("Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", 2, 4) %>%
+  pack_rows("OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)", 5, 7) %>%
+  pack_rows("Share of Elsevier Hybrid Journals with ≥ 1 OA article (%)", 8, 10)
 ```
@@
 This is in line with previous research, suggesting that OA policies and funder mandates have a stronger effect on OA uptake than disciplinary publication cultures [@Huang_2020; @Larivi_re_2018].
+While these aspects are important drivers of both full and hybrid OA, faculty members generally prioritize journal prestige and the journal impact factor [@Niles_2020] in publishing decisions, which, in combination with the availability of OA funding, might encourage publishing hybrid OA [@van_der_graaf_maurits_2017].
 

### 2021-05-19 analysis/manuscript.Rmd
align text

@@
 agreements with funding bodies (cf. @Els_Agreements), exempted through fee waivers (e.g., in ``cases of genuine need'' or due to
-society or university sponsorships, cf.~Elsevier,
-n.d.-c), and other types not
+society or university sponsorships, cf. @Els_Pricing), and other types not
 specified by Elsevier. Finally, when hybrid APCs were invoiced as part
@@
 
-Figure \ref{fig:boxuptake} presents the hybrid OA uptake by year, showing considerable variation among Elsevier's hybrid journals.
-The box-and-whiskers plot visualizes the median hybrid OA uptake (horizontal line inside the box), the range of the majority of journals (the box), and individual outliers (dots).
-Notably, the upper quartiles and whiskers stretch farther over the years, which indicates that the range of uptake among journals with an above-average proportion of OA articles increased over the years.
-Nevertheless, the prevalence of hybrid OA remained relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.
-
-```{r boxuptake, fig.cap = "Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 20\\%."}
+Table \ref{tab:overview} also shows that the average OA share per journal was higher than the overall percentage.
+To consider these variations across journals, Figure \ref{fig:boxuptake} illustrates the range of OA uptake rates as a box-and-whiskers plot by year.
+The bold line shows the median OA prevalence across journals, which slightly rose from `r filter(perc_df, issued_year == 2015) %>% pull(median) * 100`% in 2015 to `r filter(perc_df, issued_year == 2019) %>% pull(median) * 100`% in 2019}.
+Notably, the upper quartiles and
+whiskers stretch farther over the years, which indicates that the range
+of uptake among journals with an above-average proportion of OA articles
+increased over the years. Nevertheless, the prevalence of hybrid OA
+remained relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.
+
+```{r boxuptake, fig.cap = "Open Access Uptake per Elsevier Hybrid Journal (Per Year)."}
 hybrid_oa_volume %>%
@@
 ```
-
-### Share of the Total Volume published in OA{.unnumbered}
 
@@
 
-During the same period, Elsevier published `r all_oa` OA articles in total (i.e., in hybrid and full OA journals), representing `r all_oa / all_els * 100`% of the total article volume (n=`r all_els`).
-Looking at Figure \ref{fig:oa_volume_fig}, it is apparent that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind that of Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and full OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`; `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.
-
-```{r oa_volume_fig, fig.cap="Open Access (OA) in Elsevier journals in terms of article volume between 2015-2019.
-Figure A presents the aggregated OA article volume of Elsevier journals from 2015-2019 by OA type.
-Figure B presents the share of OA articles relative to Elsevier's overall article output by OA type and open content license from 2015-2019."}
+During the same period, Elsevier published OA to `r all_oa` articles in total (i.e., full, hybrid, and delayed OA combined), representing `r all_oa / all_els * 100`% of the total article volume (n=`r all_els`).
+Looking at Figure \ref{fig:oa_volume_fig}, it is apparent that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged delayed OA (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`; `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and full OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`; `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.
+
+```{r oa_volume_fig, fig.cap="Elsevier OA Article Volume 2015-2019 by (A) OA Type and (B) Open Content License."}
 # see analysis/006_license.Rmd how the figure was created
@@
 Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`; `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to full OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`; `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`; `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) .
-Most  delayed OA articles were provided under an Elsevier user license (Els-User), which prohibits reuse for commercial purposes.
+Most  delayed OA articles were provided under an Elsevier user license (Els-User), which prohibits reuse for commercial purposes and redistribution.
 
@@
 
-Next, we present disciplinary variations in hybrid OA.
-Table \ref{tab:subject_area_table} presents the high-level findings by AJSC subject area.
-In the five-year period we studied, the physical sciences (`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and health sciences (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) were the most prevalent in terms of hybrid journals with at least one OA article.
-In terms of article volume, most OA articles were published in hybrid journals belonging to the life sciences (`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles), while social sciences  (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles) journals published the lowest number of OA articles.
-However, it is important to note that there is a large overlap between the journals’ subject areas.
-Therefore, Table  \ref{tab:subject_area_table} also assigns hybrid journals and OA articles fractionally to a subject area and compares it to the full counting, where journals assigned more than one subject in the AJSC classification were counted once for each field.
+Next, we present the hybrid OA uptake of different disciplines. Table
+\ref{tab:subject_area_table} presents the high-level findings by ASJC
+subject area. Between 2015 and 2019, the physical sciences
+(`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and health sciences  (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) recorded the most
+hybrid journals with at least one OA article. However,
+the life sciences ((`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles) published the most hybrid OA articles, while the
+social sciences published the least (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles).
+It is important to
+note, though, that there is a large overlap between the journals' subject areas.
+Therefore, Table \ref{tab:subject_area_table} also presents fractional counts for hybrid journals and OA articles, which equally weights counting with respect to the total number of ASJC subject areas a journal belongs to.
+Journals with multiple ASJC codes that are members of the same subject area were counted once.
 For the remaining results, we only present the full counting.
@@
   kbl(booktabs = TRUE, col.names = c("Subject area", "Journals", "OA articles", "Journals", "OA articles"),
-      caption = "Subject area overview using full and fractional counting.
-Table reports the number of Elsevier hybrid journals with at least one open access (OA) article, and OA articles published from 2015-2019.
-One multidisciplinary with 61 OA articles and 11 journals with 37 OA articles that  could not be matched to the Scopus Source Title list used to obtain ASJC codes are not included.",
+      caption = "Full and Fractional Counting of Subject Areas (2015-2019).",
       format.args = list(big.mark = ",")) %>%
@@
 Notably, with the exception of environmental science and earth and planetary sciences, journals from the physical sciences show a lower median proportion of hybrid OA articles, whereas most life sciences and social sciences journals ranked above average.
-Figure X also shows large variations among the 26 subject fields, with the highest OA rates in immunology and microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), environmental science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), whilst  chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), materials science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`) recorded the lowest OA rate.
-
-```{r oa_sub_uptake, fig.cap='Open access (OA) uptake in Elsevier hybrid journals in percent grouped by subject area and field between 2015-19.
-Visualized as a box-and-whiskers plot limited to an OA share of 20\\%; the red line represents the overall median (2.5), jittered dots represent individual journals.
-Subject fields are ordered based on the median OA uptake.', fig.asp=1.3, out.width="90%"}
+Figure \ref{fig:oa_sub_uptake}  also shows large variations among the 26 subject fields, with the highest OA rates in immunology and microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), environmental science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), whilst  chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), materials science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`) recorded the lowest OA rate.
+
+```{r oa_sub_uptake, fig.cap='OA Uptake in Elsevier Hybrid Journals by Subject Area and Field (2015-2019).', fig.asp=1.3, out.width="90%"}
 # with labels
@@
 p_subject
-```
-
-### Citation Impact{.unnumbered}
-
-```{r}
-hybrid_subject_19 <- hybrid_subject %>% 
-  mutate(prop = oa /articles) %>%
-  filter(issued_year == "2019", !is.na(percentile)
-         , prop != 0) %>%
-  distinct(issn, prop, percentile, snip, sjr)
-
-cor_test_res <- cor.test(hybrid_subject_19$prop, hybrid_subject_19$percentile, method = "spearman") 
-```
-
-Furthermore, we were interested in the relationship between the journals' field-specific impact and their OA uptake.
-Spearman's rho correlation coefficient was used to assess the relationship between the 2019 Source Normalized Impact per Paper (SNIP) value calculated by Scopus and the journals' OA uptake that year.
-Considering only journals with at least one OA article, we found a weak but positive correlation between journal impact and OA uptake ($r_s$=`r as.character(round(cor_test_res$estimate,4))`, $p<0.001$), suggesting that the relationship is not very strong, based on our data (see Figure \ref{fig:sniptest}).
-
-```{r sniptest, fig.cap="Source Normalized Impact per Paper (SNIP) versus open access (OA) uptake 2019.
-Each dot represents a hybrid journal with at least one published OA article in 2019.
-There is only a weak, positive correlation between the two variables ($r_s$=0.1679, $p<0.001$)"}
-ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = 0.2) + 
-  geom_smooth(method = "loess", se = TRUE) +
-  coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
-  labs(x = "Hybrid OA Uptake", y = "SNIP") +
-  theme_cowplot()
 ```
@@
 
-As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APC were most often invoiced to authors (n=`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) and to a lesser extent as part of agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%).
+As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APCs were most often invoiced to authors (n=`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) and to a lesser extent as part of agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%).
 Interestingly, we also found a small number of cases where hybrid APCs were waived (n=`r filter(sponsored_art, fct_source == "Fee waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee waived") %>% pull(prop)`%).
+A brief note of caution regarding APCs invoiced to the authors is due here: Although author invoicing was indicated in the \texttt{openaccessSponsorType} article metadata field used for this study, this does not necessarily mean that authors paid the APC out of their own pocket.
+Rather, this implies that the authors' institution or funder did not have a central invoicing agreement in place at the time of publication.
 
@@
   kbl(booktabs = TRUE, col.names = c("Invoicing channel", "Hybrid OA articles (n)", "Percentage"),
-      caption = "Invoice channels for hybrid OA in Elsevier journals between 2015-19.", digits = 1,
+      caption = "Invoice Channels for Elsevier Hybrid OA Articles (2015-2019).", digits = 1,
       format.args = list(big.mark = ",")) %>%
@@
 
-Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has constantly invoiced hybrid OA APCs directly to authors compared to research funders or academic consortia ("Agreement").
-Between 2015 and 2019, the proportion of author-invoicing increased from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Author") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Author") %>% pull(perc)`%, while the share of invoicing through agreements slightly changed from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Agreement") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Agreement") %>% pull(perc)`%.
-Also, the share of fee-waived articles remained relatively stable, but we found different types of waivers.
+Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has consistently invoiced more authors than funders or academic consortia ("Agreement"), although both shares increased.
+In particular, the proportion of author-invoicing increased from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Author") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Author") %>% pull(perc)`%, whereas the share of invoicing through agreements slightly
+increased from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Agreement") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Agreement") %>% pull(perc)`%, respectively.
+The share of fee-waived articles also remained relatively stable, but we found different types of waivers.
 Around `r filter(els_waived_df, !is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)* 100`% of fee waivers were linked to an OA sponsor, which indicates that a third party paid the APC.
@@
 
-```{r invoiceoverview, fig.cap="Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoice channel Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant.
-Grey bars show the total number of hybrid OA articles published in Elsevier journals from 2015-2019."}
+```{r invoiceoverview, fig.cap="Elsevier Hybrid OA Articles by Invoice Channel and CC License (Per Year)."}
 # plot
@@
 We also observed large differences in OA invoicing among subject fields.
-Table \ref{tab:invocing_by_subject} shows the number of OA articles by subject field and invoice channel.
-For articles in nursing, decision dciences, and pharmacology, toxicology and pharmaceutics, Elsevier predominantly invoiced authors, whereas most energy and chemical engineering articles were invoiced through agreements.
-Likewise, the majority of articles in materials science, chemistry and physics and astronomy were not invoiced to authors but facilitated through agreements or waived.
-The large share of waived APCs in physics and astronomy can be attributed to a single 2015 issue of Nuclear and Particle Physics Proceedings, which offered sponsored OA.
+Table \ref{tab:invocing_by_subject} shows the number of OA articles by
+subject field and invoice channel. For articles in nursing, decision
+sciences, and pharmacology, toxicology and pharmaceutics, Elsevier
+predominantly invoiced authors, whereas most energy and chemical
+engineering articles were invoiced through agreements. Likewise, the
+majority of articles in materials science, chemistry and physics and
+astronomy were not invoiced to authors but facilitated through
+agreements or waived. The large share of waived APCs in physics and
+astronomy can be attributed to a single 2015 issue of Nuclear and
+Particle Physics Proceedings.
 
@@
 Elsevier’s data offer more insight into invoicing.
-Overall, we identified 62 sponsoring institutions and organizations that received invoices for hybrid publishing fees as part of an agreement with Elsevier.
-We found that, by a large margin, most invoices were issued to UK-based research funders and institutions (n=`r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%), followed by the Netherlands (n=`r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%) and the European Union (n=`r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%; EU).
-Figure \ref{fig:invoice_sponsor_country} provides a closer look at the geographical distribution by license prevalence, highlighting the dominance of CC -BY licensed articles invoiced to sponsoring institutions in the UK, the United States, and, to a much lesser extent, the Netherlands.
-In contrast, the majority of articles invoiced to sponsoring institutions from Norway or representing the EU were published under a non-commercial license.
-
-```{r invoice_sponsor_country, fig.cap="Number of centrally invoiced OA articles in Elsevier hybrid journals by country of recipient.
-Colored bars represent the Creative Commons (CC) license variant under which the article was published between 2015 and 2019."}
+Overall, we identified `r filter(sponsor_df, !is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` academic institutions and funders that received
+invoices as part of publishing agreements. By a large margin, most invoices were issued to
+UK-based research funders and institutions (n=`r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "United Kingdom") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%), followed by the Netherlands (n=`r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "The Netherlands") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%) and the European Union (n=`r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow()`; `r filter(sponsor_df, country_factor == "European Union") %>% distinct(doi) %>% nrow() / sponsor_df %>% distinct(doi) %>% nrow()* 100`%; EU), which in the context of this paper refers to institutions established by the European Commission, and not the  individual member states.
+Invoicing data comprises two EU institutions, the research funder European Research Council and the health agency European Center for Disease Prevention and Control (ECDS).
+Figure \ref{fig:invoice_sponsor_country} presents the geographical distribution by license prevalence,
+highlighting the dominance of CC BY licensed articles invoiced to institutions in the UK, the United States (US), and, to a much
+lesser extent, the Netherlands. In contrast, most articles
+invoiced to institutions from Norway or representing the EU
+were published under a non-commercial license.
+
+```{r invoice_sponsor_country, fig.cap="Centrally Invoiced Elsevier Hybrid OA Articles by Country and CC License (2015-2019)."}
 sponsor_df %>%
@@
 
-Figure \ref{fig:sponsor_sector} shows the yearly distribution of central invoiced articles by year and institutional sector.
-While UK-based OA invoices were mainly addressed to discipline-specific governmental and non-profit research funders, invoices to the Netherlands and Norway were issued to national academic consortia representing the higher education sector.
-In 2019, Elsevier also launched similar agreements in countries with lower publication output including Hungary and Poland.
-Besides, we found that invoiced funding bodies from the UK and the US mainly represented discipline-specific funders, while the invoiced funding bodies from other countries were focused on a broad variety of disciplines (see Table \ref{tab:subject_profile}).
-
-```{r sponsor_sector, fig.cap=' Development of centrally invoiced OA articles in hybrid journals by country. Colored bars represent the institutional sector OA funding bodies belong to.'}
+Figure \ref{fig:sponsor_sector} shows the yearly distribution of centrally
+invoiced articles by year and institutional sector. While UK-based OA
+invoices were mainly addressed to discipline-specific governmental and
+non-profit research funders, invoices to the Netherlands and Norway were
+issued to national academic consortia representing the higher education
+sector. In 2019, Elsevier also launched similar agreements in countries
+with lower publication output including Hungary and Poland. We
+found that invoice recipients from the UK and the US mainly
+represented discipline-specific funders, while invoice recipients from other countries focused on a broad variety of
+disciplines.
+
+```{r sponsor_sector, fig.cap='Centrally Invoiced Elsevier Hybrid OA Articles by Country and Sector (Per Year).'}
 sponsor_df %>%
@@
 
-```{r subject_profile}
-sponsor_country %>%
-  ungroup() %>%
-  pivot_wider(id_cols = c("country_factor"), names_from = `Primary research focus`, values_from = prop) %>%
-  mutate_all(~replace(., is.na(.), 0)) %>%
-  janitor::adorn_totals(c("col", "row")) %>%
-  mutate_if(is.numeric, ~paste0(., "%")) %>%
-  rename(`Country/Subject area` = country_factor) %>%
-  kableExtra::kbl(booktabs = TRUE,
-                  caption = " Relative subject profile of centrally invoiced OA funding bodies.", align = c("lrrrrrr")) %>%
-    row_spec(6:7, hline_after = TRUE) %>%
-   kable_styling(latex_options = c("scale_down"))
-```
-
-```{r}
-# sector/discipline
-hybrid_invoice %>%
-  filter(!is.na(`Primary research focus`)) %>%
-  count(Sector, `Primary research focus`) %>%
-  mutate(prop = n / sum(n) * 100) %>% 
-  pivot_wider(id_cols = c("Primary research focus"), names_from = Sector, values_from = n) %>%
-  kableExtra::kbl(booktabs = TRUE) %>%
-  kable_styling(latex_options = c("scale_down"))
-
-```
-
-### Institutional ranking{.unnumbered}
-
 ```{r open_apc_check}
@@
 Table \ref{tab:sponsor_top_10} presents the top ten of `r hybrid_invoice %>% filter(!is.na(oa_sponsor_cleaned)) %>% distinct(oa_sponsor_cleaned) %>% nrow()` sponsoring bodies in our sample.
-Together, they accounted for around 80% of centrally invoiced APCs.
-The table also highlights the number of distinct journals and the share of CC BY licensed articles.
-Notably, UK-based research funders and the US-based Bill and Melinda Gates Foundation mainly sponsored articles published under a CC BY license.
-In contrast, articles funded by VSNU, the European Research Council, and Norway Institutes had a much lower proportion of CC BY.
+Together, they accounted for around 80\% of
+centrally invoiced APCs. The table also highlights the number of
+distinct journals and the share of CC BY licensed articles. Notably,
+UK-based research funders and the US-based Bill and Melinda Gates
+Foundation mainly received invoices for CC BY licensed articles. In
+contrast, articles invoiced to VSNU, the European Research Council, and
+Norway Institutes had a much lower proportion of CC BY.
 
@@
       linesep = "",
-      caption = " Top 10 Elsevier hybrid OA sponsors by centrally invoiced articles 2015-2019.
-The table furthermore presents the number of Elsevier hybrid journals with at least one invoiced article and two indicators of compliance: proportion of  CC BY licensed articles  and disclosed spending through the Open APC Initiative (OAPC.") %>%
+      caption = "Elsevier Hybrid OA Invoice Recipients  (2015-2019).") %>%
   row_spec(11:12, hline_after = TRUE) %>%
@@
 
-```{r ccby_vs_openapc}
-ggplot(ind_sponsor_df, aes(in_open_apc, cc_by, size = articles)) +
-  geom_point(alpha = 0.5) +
-  theme_minimal_grid() +
-  scale_x_continuous(limits = c(0,1), labels = scales::percent_format()) +
-  scale_y_continuous(limits = c(0,1), labels = scales::percent_format()) +
-  labs(y = "CC BY Share", x = "Spending info in Open APC", size = "Invoiced\nOA Articles") +
-  theme(legend.position="right", legend.justification = "right")
-```
-
-Figure \ref{fig:field_funder_matrix} shows the subject profile of OA sponsors based on to the journals’ subject fields of centrally invoiced articles.
-As expected, the figure shows that subject-specific funders, such as the British Engineering and Physical Sciences Research Council (EPSRC), tended to focus on a limited set of subject fields.
-In contrast, national consortia, such as VSNU and Norwegian Institutes (NO), financed OA articles  for a broader range of fields.
-However, within these consortia, the subject distribution also varied.
-VSNU OA spending, for example, had a strong focus on articles relating to medicine, environmental science, and energy, presumably reflecting disciplines with a generally strong publication output at Dutch universities.
-Norwegian universities mostly published hybrid OA in engineering, environmental science and agricultural and biological sciences journals.
-
-```{r}
-#short funder labels
-funder_codes <- tribble(
-  ~oa_sponsor, ~oa_sponsor_short,
-  "Engineering and Physical Sciences Research Council", "EPSRC",
-  "VSNU", "VSNU",
-  "Wellcome Trust", "Wellcome",
-  "European Research Council", "ERC",
-  "Medical Research Council", "MRC",
-  "Natural Environment Research Council", "NERC",
-  "Biotechnology and Biological Sciences Research Council", "BBSRC",
-  "Bill & Melinda Gates Foundation", "Gates",
-  "Economic and Social Research Council", "ESRC",
-  "Norway Institutes", "NO",
-  "Other", "Other")
-```
-
-```{r field_funder_matrix, out.width="90%", fig.width=8, fig.height=12}
-field_funder <- hybrid_df %>%
-  filter(oa_sponsor_type == "FundingBody") %>%
-  left_join(funder_codes, by = c("oa_sponsor_name" = "oa_sponsor")) %>%
-  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_short), "Other", oa_sponsor_short)) %>%
-  mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
-  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
-  left_join(jn_subjects, by = "issn") %>%
-  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
-  group_by(oa_sponsor, top_level) %>%
-  summarise(n_oa_type = n_distinct(doi)) %>%
-  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
-  mutate(oa_sponsor = fct_relevel(oa_sponsor, "Other", after = Inf))
-# plot
-pp <- ggplot(field_funder,
-       aes(
-         x = factor(oa_sponsor),
-         y = reorder(top_level, desc(top_level)),
-         colour = prop,
-         size = n_oa_type
-       )) +
-  geom_point() +
-  scale_color_viridis_c(
-    labels = scales::percent_format(accuracy = 5L),
-    option = "magma",
-    name = "Percentage per Funder",
-    breaks =  scales::extended_breaks(),
-    begin = 0.15,
-    end = 0.85
-  ) +
-  scale_size("OA Articles") +
-  scale_x_discrete(position = "top") +
-  labs(x = NULL, y = NULL) +
-  theme_minimal_grid(font_size = 12) +
-  theme(
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    panel.background = element_blank()
-  ) +
-  theme(legend.position = "top", legend.justification = "right") +
-  theme(legend.direction = "horizontal", 
-        legend.box = "horizontal")
-ggsave(here::here("figure", "funder_subject.png"), pp, dpi = 300, width = 9.5, height = 9)
-```
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.95\textwidth]{../figure/funder_subject.png}
-
-}
-
-\caption{Top 10 Elsevier hybrid OA institutional invoice recipients in terms of articles published by journal subject.
-EPSRC, Engineering and Physical Sciences Research Council; VSNU, VSNU; Wellcome, Wellcome Trust; ERC, European Research Council; ERC, European Research Council; MRC, Medical Research Council; NERC, Natural Environment Research Council; BBSRC, Biotechnology and Biological Sciences Research Council; Gates, Bill and Melinda Gates Foundation; ESRC, Economic and Social Research Council; NO, Norway Institutes.
-}\label{fig:field_funder_matrix}
-\end{figure}
-
 # Discussion
 
-Over the last five years, Elsevier recorded growth in the uptake of hybrid OA for each of our three indicators: The number of hybrid OA articles published per year doubled, the number of hybrid journals with at least one OA article grew by 21%, and the share of hybrid OA articles relative to closed-accessed articles in these journals increased from 2.6% to 3.7%.
-As @Laakso_2016, we found only a weak relationship between citation metrics like the SNIP and hybrid OA uptake and observed differences in the uptake of hybrid OA among journal subject areas.
-In particular, we found the highest count of hybrid OA articles in physical sciences journals (see Table \ref{tab:subject_area_table}).
-This was followed by the life sciences and health sciences, whereas the social sciences had the lowest count of hybrid articles.
-This order mostly reflects the disciplines' overall publication output.
-According to the Open Science Monitor by the European Commission [@OS_Monitor], the physical sciences and mathematics publish the most articles, followed by the health sciences, life sciences, and social sciences and humanities.
-
-Such disciplinary differences in hybrid OA prevalence become more meaningful when considering the relative share of OA articles to closed-access articles.
-In line with previous research, we found that Elsevier journals from the life sciences and social sciences [@Jubb_2017; @Kramer_2018; @Laakso_2016] recorded greater than typical hybrid OA uptake, whereas physical sciences journals generally had a lower than typical uptake [@Kramer_2018; @Mart_n_Mart_n_2018; @Laakso_2016].
-In a systematic review of OA publishing patterns, @Severin_2020 linked such disciplinary differences to socio-cultural and technological factors that shape publishing cultures and practices.
-Among other aspects, the authors pointed to the availability of prestigious journals, the possibility to provide OA through alternative routes, and ease of access to funding sources to cover publishing fees.
-As such, looking at hybrid OA uptake in isolation can only provide limited insight, so that placing it within the broader context of disciplinary publishing and OA practices is necessary.
-For instance, since many branches of the physical sciences have established self-archiving practices and the technical infrastructure to support these (e.g., the arXiv; @Severin_2020, @Bj_rk_2010, @Laakso_2016), researchers can provide OA through repositories (i.e., green OA), and indeed, according to the Open Science Monitor the physical sciences provide OA primarily through the green route (31.2% total OA share, 24% OA if only considering green OA).
-Hence, the low share of hybrid OA in our sample might be because physical scientists perceive much less of a need for hybrid OA.
-Further, @Severin_2020 note project-based funding structures enable OA publishing in fully OA and hybrid journals as they allow for easy integration of publishing costs.
-Since these funding structures are typical for the life sciences (ibid.), we would expect higher rates of OA.
-Data from the Open Science Monitor seems to support this notion as the life sciences record particularly high shares of full OA articles and comparatively high  shares of hybrid OA in some fields, such as biology and agricultural science.
-While we expected the observed hybrid OA shares for the life sciences and physical sciences, we were surprised by the high share of hybrid OA belonging to the AJSC subject area social sciences, which also includes the humanities, since articles "social scientists have reported to face significant difficulties in access to grant funding for [publication charges], as most research in the social sciences is not done by means of project-specific funding" @Severin_2020 [p.
-15].
-However, previous research has found comparatively low awareness and uptake of green OA among researchers in the  field [@Bj_rk_2010; @Creaser_2010], which might encourage hybrid OA uptake.
-Another possible explanation,is that Elsevier has a number of high prestigious social science and humanities journals [@Larivi_re_2015], which means that our findings are not representative of the general population of SSH journals.
-Likewise, this limitation could be extended to our findings in general that whilst not representative of the population of hybrid OA publishing, they provide valuable insight into a key player of scholarly publishing.
-
-Finally, OA policies in favor of hybrid OA in combination with invoicing agreements with Elsevier might be a driver.
-For instance, we found that nearly half of articles in arts and humanities were facilitated through agreements, which was above-average.
-This is in line with previous research, suggesting that OA policies and funder mandates have a stronger effect on OA uptake than disciplinary publication cultures [@Huang_2020; @Larivi_re_2018].
-While these aspects are important drivers of both full and hybrid OA, faculty members generally prioritize journal prestige and the journal impact factor [@Niles_2020] in publishing decisions, which, in combination with the availability of OA funding, might encourage publishing hybrid OA [@van_der_graaf_maurits_2017].
-
-Our overall investigation of invoice channels and recipients of hybrid OA APCs found that the majority of Elsevier hybrid articles were invoiced to the author (n=41,725; 58.2%).
-However, it is important to note that although Elsevier's metadata classifies the sponsor type as "author", this does not necessarily mean that authors paid for APCs out of their own pocket but rather just that the APC was not invoiced through an agreement.
-Hence, it is possible that APCs invoiced to the author were covered by institutional funds or as part of research grants.
-For instance, a recent Springer Nature survey found that hybrid OA was predominantly supported through institutional and funder sources (71%), followed by OA agreements (34%), while only 6% were paid from personal funds or savings [@Monaghan_2020].
-We also observed notable differences in licensing among invoice types.
-The majority of hybrid OA articles, regardless of the invoice type, were licensed under the more restrictive CC BY-NC-ND license.
-There is a lack of dedicated studies on licence selection, but indications so far suggest that when given a choice authors tend to select more restrictive variants for their published works [@Fraser_2020; @Rowley_2017; @Van_Noorden_2013].
-On the other hand, when hybrid articles were invoiced through agreements, most articles were licensed under the more liberal CC BY license.
-As several funding bodies mandate CC BY licenses, including the UK research funders that account for 62% of our "agreement" subsample, this result is perhaps not surprising but suggests the effectiveness of such agreements.
-
-Based on our findings, around a third of the articles were invoiced directly through OA agreements (e.g., research funders, national library consortia).
-Strikingly, sponsors based in the UK accounted for more than half of these , and a quarter was charged to sponsors from other European countries.
-This finding reiterates reports from previous studies that the UK's OA profile differs markedly from other countries (cf.
-@Jubb_2015; @Jubb_2017) and points to the effects of science policy.
-To increase OA to research publications, the UK implemented centralized APC funding and embraced hybrid OA as a transition model.
-Within five years of the publication of the Finch Report in 2012, the UK recorded an 18% increase in immediate OA, coupled with a rise in hybrid OA from 2.7% in 2012 to 15.4% of all articles in 2016 [@Jubb_2017], which can be attributed to the availability of APC funding from Research Councils UK (RCUK).
-Indeed, RCUK block grants have been the largest single source of OA funds in the UK—the Wellcome Trust, and more recently transformative agreements [@Tickel_2018; @Jubb_2017].
-However, since then publishing expenditures of UK universities have been rising rapidly (ref), so that several universities stopped supporting hybrid OA in late 2018-2019 [@oxford_2019; @walker_2019; @Castro_2019; @birmingham].
-This development might explain the slight decrease in hybrid articles invoiced to the UK our data showed around that time (see Figure \ref{fig:invoice_sponsor_country}).
-
-An essential part in the absence of transparency in scholarly publishing and especially in hybrid OA has been the lack of transparency in publishers’ article management.
-This has prohibited consumers to get a comprehensive and comparative overview of their publishing activities and the associated costs and, thus, limited their negotiating power.
-However, certain initiatives in Europe like the ESAC Initiative have begun to press for improved and more transparent workflows for transformative agreements [@Geschuhn_2017].
-In particular, the ESAC Workflow Recommendations seek to increase efficiency and transparency in the publication workflow through improved invoicing and reporting processes and metadata delivery, including article-based machine readable information about the OA sponsor.
-Through this study we demonstrated the utility and benefits of such  publisher-provided metadata about hybrid OA funding and highlighted the need for extending these to include comprehensive information about the licensing and the particular OA type.
-
-Usually identifying delayed OA articles and distinguishing them from other types of OA is challenging since there is no publicly available list of such journals.
-For instance, Taylor & Francis’ (2021) website indicates that some of their journals participate in a delayed or promotional OA program but no further information is provided as to which journals these are.
-However, Elsevier provides a list of journals enrolled in the "Open Archive" programme and further tags delayed OA articles in the machine-readable XML version of the article.
-This enables easy identification and more accurate assessments of hybrid OA, which allows to take into account the substantial amount of delayed OA content in future hybrid OA studies using Elsevier data.
-
-While this study advances our knowledge about hybrid OA uptake and invoicing, the limitations leave room for future research.
-This study focused on only one publisher, which limits the generalizability of our findings since Elsevier’s attributes such as composition of the journal portfolio, mix of business models, pricing, and promotion of various options cannot be assumed to be representative of scholarly journal publishing in general.
-Though the invoicing data extracted from the XML files enables unprecedented accuracy and comprehensiveness in the study of hybrid OA, it likely does not disclose the actual OA sponsor in all cases.
-Most articles were invoiced to authors, but it remains unclear how often authors themselves pay for APCs or invoices are covered through institutional OA funds or research grants.
-Further research into this topic would increase our understanding of OA funding streams outside of publishing agreements.
-Further, our study demonstrates that to achieve comprehensive mapping and understanding of the financial flows involved in OA funding, complex, in-depth country-specific meso-level analyses is needed.
-Such studies would ideally draw on various data sources, such as publicly available bibliometrics and metadata, information about significant national research funders and their OA policies, information about the national library and possible consortia landscape including their publishing agreements, which might require Freedom of Information Requests.
-
-This study provides a snapshot of hybrid OA for the largest journal publisher prior to the impact of the implementation of Plan S requirements.
-While many Plan S signatories have already had strong OA policies, this harmonized approach is likely to affect publishing decisions of funded authors,the licencing of their articles, and the offerings and pricing of publishers.
-Since the new requirements are applicable to research funded from 2021 onwards, a comparative study focused on articles  invoiced to Plan S signatories would be a fruitful endeavor.
-
-In a similar vein, future studies could  combine invoice information with research funding information since these contribute different pieces to the funding stream puzzle.
-This could also provide more insight into the actual APC funding streams behind the author-invoiced articles.
+From 2015-2019, Elsevier recorded growth in the uptake of
+hybrid OA: The number of hybrid OA
+articles published per year doubled, the number of hybrid journals with
+at least one OA article grew by 21\%, and the share of hybrid OA
+articles relative to closed-accessed articles in these journals
+increased from 2.6\% to 3.7\%. As  @Laakso_2016, we observed disciplinary differences. In particular, we found the highest count of hybrid OA
+articles in physical sciences journals (see Table
+\ref{tab:subject_area_table}). This was followed by the life sciences
+and health sciences, whereas the social sciences had the lowest count. This order mostly reflects the disciplines' overall
+publication output. According to the Open Science Monitor [@OS_Monitor], the physical sciences publish the
+most articles, followed by the health sciences, life sciences, and
+social sciences.
+
+Disciplinary differences in hybrid OA prevalence become more
+meaningful when considering the relative share of OA articles to
+closed-access articles. In line with previous research, we found that
+Elsevier journals from the life sciences and social sciences  [@Jubb_2017; @Kramer_2018; @Laakso_2016] recorded
+greater than typical hybrid OA uptake, whereas physical sciences
+journals generally had a lower than typical uptake [@Kramer_2018; @Mart_n_Mart_n_2018; @Laakso_2016]. In a
+systematic review of disciplinary OA publishing patterns, @Severin_2020
+highlighted the importance of socio-cultural and technological
+factors in shaping publishing cultures and practices.  For instance, since many
+branches of the physical sciences have established self-archiving
+practices [@Severin_2020, @Bj_rk_2010, @Laakso_2016], researchers can provide OA through repositories and might perceive
+much less of a need for hybrid OA.
+In contrast, hybrid OA in the life sciences could be enabled through project-based funding structures as they allow for easy integration of publishing costs @Severin_2020.
+On the other hand, the high hybrid OA uptake that we observed among the social sciences could point to the influential role of OA policies and invoicing agreements [@Huang_2020; @Larivi_re_2018].
+While these aspects
+are important drivers of both full and hybrid OA, faculty members
+generally prioritize journal prestige and the journal impact factor
+ [@Niles_2020] in publishing decisions, which, in combination with
+the availability of OA funding, might encourage publishing hybrid OA
+[@van_der_graaf_maurits_2017].
+
+Our invoicing data analysis found that most APCs were
+invoiced to the author (n=41,725; 58.2\%). However, it is important to
+emphasize although Elsevier's metadata classifies the sponsor type as
+``author'', this does not necessarily mean that authors paid for APCs themselves but rather that APCs were not invoiced
+through publishing agreements. It is possible that these APCs were covered through institutional funds or  research
+grants. This would align with a recent Springer Nature survey that found hybrid
+OA was predominantly supported through institutional and funder sources
+(71\%), followed by OA agreements (34\%), while only 6\% were paid from
+personal funds or savings [@Monaghan_2020].
+
+Further, we observed
+notable differences in licensing. Most
+hybrid OA articles invoiced to author were licensed under
+the more restrictive CC BY-NC-ND license. Previous research, while lacking dedicated
+studies on license selection,  suggests that authors tend to select more restrictive license variants when
+given a choice [@Fraser_2020; @Rowley_2017; @Van_Noorden_2013]. On the other hand, we found when hybrid OA was invoiced
+through agreements, most articles were licensed under the more liberal
+CC BY license. As several funding bodies mandate CC BY licenses,
+including the UK research funders that account for 62\% of our
+"agreement" subsample, this result is perhaps not surprising but
+suggests the effectiveness of such agreements.
+
+Based on our findings, around a third of the articles were invoiced
+ through OA agreements (e.g., research funders, national library
+consortia). The predominance of UK funding bodies in this subsample reiterates reports from previous
+studies that the UK's OA profile differs markedly from other countries
+(cf. @Jubb_2015; @Jubb_2017), pointing to the impact of science policy. To promote 
+OA, the UK implemented centralized APC funding
+and embraced hybrid OA as a transition model. Within five years of the
+publication of the Finch Report in 2012, the UK recorded an 18\%
+increase in immediate OA, coupled with a rise in hybrid OA from 2.7\% in
+2012 to 15.4\% of all articles in 2016 [@Jubb_2017], which can be
+attributed to the availability of funding from 
+RCUK. Indeed, RCUK block grants have been the largest single source of
+APC funds in the UK---the Wellcome Trust, and more recently
+transformative agreements [@Tickel_2018; @Jubb_2017]. However,
+since then publishing expenditures of UK universities have been rising
+rapidly (Jubb et al., 2017), so that several universities stopped supporting hybrid OA
+in late 2018-2019 [@walker_2019; @birmingham]. This development might
+explain the slight decrease in hybrid articles invoiced to the UK our
+data showed around that time (see Figure
+\ref{fig:sponsor_sector}).
+
+Through this study we
+demonstrated the utility and benefits of publisher-provided
+metadata about hybrid OA invoicing and highlighted the need for extending
+these to include comprehensive information about licensing and APC waivers.
+Metadata guidance should also consider the substantial amount of delayed OA content, which needs to be distinguished from hybrid OA.
+Hence, our study substantiates the recommendations from the ESAC Initiative that seek to increase efficiency and
+transparency through improved invoicing and
+reporting processes and metadata about OA funding [@Geschuhn_2017; @Marques_2019].
+
+While this study advances our knowledge about hybrid OA uptake and
+invoicing, the limitations leave room for future research. We
+focused on only one publisher, which limits the generalizability of our
+findings because Elsevier's journal
+portfolio, mix of business models, pricing, and promotion of various
+options cannot be assumed to be representative of scholarly journal
+publishing in general.
+Relatedly, our study is limited to articles published in hybrid OA journals and, hence, is not representative of invoices for articles in full OA journals.
+To develop a full picture of OA invoicing, the methodology presented in this paper can be used for future studies in this direction.
+Further, although Elsevier's invoicing data improves transparency, it seems likely that not all actual OA funding bodies are disclosed.
+For instance, most articles were invoiced to authors, but it remains unclear
+if the APCs were paid by the authors themselves or
+through institutional OA funds or research grants. Research into
+this topic would improve our understanding of OA funding 
+outside of publishing agreements. Moreover, our study demonstrates that
+comprehensive mapping of the financial
+flows of OA publishing requires complex, in-depth country-specific
+analyses. Such studies would ideally draw on
+various data sources to consider  research funders, the 
+consortia landscape, OA policies, and  publishing agreements Future research could analyze funder policies in combination with institutional agreements (including which journals are covered by the agreements) and hybrid OA invoicing metadata to provide more insight into publishing patterns and CC licence choice.
+A further avenue of research would be dedicated studies on the strengths and weaknesses of data provided by publishers and the Open APC initiative, and consider an integrative approach that combines both datastreams to reap the benefits of each while mitigating their individual drawbacks.
+
+The policy implications and the underlying data of this study present opportunities for actors in the scholarly publishing landscape to rely less on assumptions and secondary data, and instead establish a direct connection between the publication output and the related financial information.
+This could increase accountability in spending public and charitable funds, and improve OA cost monitoring through enhanced transparency, precision, and timeliness.
+Analyses like we have presented here can inform library consortia in negotiations through better understanding their publisher-specific OA output, determining the amount of delayed OA content, and facilitating comparative analyses across different consortia.
+The study could also inform research funders and research institutions about license selection implications if there is not any strict policy in place.
+This study provides a snapshot of hybrid OA for Elsevier, the largest journal
+publisher, prior to the impact of the implementation of Plan S, an initiative to accelerate the transition to OA that will no longer support hybrid OA [@Plan_s].
+While many Plan S signatories have already had strong OA
+policies, this harmonized approach is likely to affect publishing
+decisions of funded authors, the licencing of their articles, and the
+offerings and pricing of publishers at a larger scale than before. Because the new requirements apply to research funded from 2021 onward, a comparative study
+on articles invoiced to Plan S signatories would be a fruitful
+endeavor.
+Plan S does not fund hybrid OA publication unless the host journal is explicitly committed to convert to full OA, however, with the growth of national and institutional agreements that enable affiliated authors to publish hybrid OA, such restrictions will influence researchers differently.
+Invoicing data as used for this study can be useful in exploring later on, by looking at research funding statements and invoicing channels in relation to each other.
 
@@
 
-The primary aim of this empirical study was to investigate Elsevier’s hybrid OA publishing from 2015-2019 to better understand the volume and invoicing of hybrid OA and present a novel, data-driven approach for such analyses.
-Our results indicate that although the number of hybrid OA articles has increased over time, hybrid OA uptake has remained low.
-Notably, hybrid APCs were most often invoiced directly to the authors or through institutional agreements, where only a few funding bodies were the main driver of hybrid OA.
-Finally, our findings highlight that publisher-provided metadata about the invoicing channels  of (hybrid) OA can facilitate research into and increase the understanding of the financial flows of OA publishing.
-
-Ever since its inception, hybrid OA has been a challenging subject to study due to the lack of standardised ways for publishers to flag such content and APC funding data being limited to studies based on surveys and articles’ acknowledgements sections.
-This study presented a novel approach to studying OA APC invoicing  that is based on publicly available publisher-provided metadata, which can be used on on its own or in combination with other public data sources to gain more detailed and comprehensive insights into hybrid OA uptake and invoicing than ever before.
-If more publishers reported OA invoicing data on the article level and in a machine-readable format, this would increase transparency and enable  enhanced monitoring of the scholarly journal landscape over time.
-As hybrid OA has increasingly become a central element of OA policies of research funders and libraries,  customer organisations could require that invoicing information is added to the article-level metadata.
-As long as publishers do not provide this data in a structured and comprehensive format, they prevent consumers from benchmarking prices and therefore hinder price competition as the only alternative to gain comparative insight is self-reported APC expenditures that research funders and libraries voluntarily report to the Open APC initiative.
-The Open APC initiative is a valuable service but it cannot replace publisher-provided data since global participation is limited.
-
-From recent science policy developments in Europe it appears that Big Deals have gained support and remain firmly in place in the form of transformative agreements.
-However, in addition to research funders and libraries that increasingly make publishing contracts public, it is on the publisher to provide  metadata as presented here to improve the transparency in scholarly publishing by linking invoicing data to bibliometrics.
-From this study we can affirm that hybrid OA is complex as the financial flows come from various stakeholders and invoicing forms, it is not just libraries or consortia, but authors and funders are also directly involved.
-As of right now, providing  machine-readable article-level invoicing data is not common amongst publishers, which limits the potential for broader studies.
+The primary aim of this empirical study was to investigate Elsevier's
+hybrid OA publishing from 2015-2019 to better understand the volume and
+invoicing of hybrid OA and to present a novel, data-driven approach for
+such analyses. Our results indicate that although the number of hybrid
+OA articles has increased over time, its uptake has remained low.
+Notably, hybrid APCs were most often invoiced directly to the authors, followed by agreements, where only a few funding bodies were
+the primary drivers of hybrid OA. Finally, our findings highlight that
+publisher-provided metadata about the invoicing channels of (hybrid) OA
+can facilitate research into and increase the understanding of the
+financial flows of OA publishing.
+
+Since the beginning, hybrid OA has been a challenging subject to
+study due to the lack of standardized ways publishers flag such
+content and APC funding data being limited self-reported data, surveys,
+and other secondary sources. This study presented a novel
+approach to studying APC invoicing that is based on publicly
+available publisher-provided metadata, which can be used on on its own
+or in combination with other public data sources to gain more detailed
+and comprehensive insights into hybrid OA uptake and invoicing. If more publishers reported OA invoicing on the article
+level and in a machine-readable format, this would increase transparency
+and improve monitoring of the scholarly journal landscape over
+time. As hybrid OA has  become a central element of OA
+policies of research funders and libraries, consumer organizations could
+require that invoicing information is added to the article-level
+metadata. As long as publishers do not provide this data in a structured
+and comprehensive format, they prevent benchmarking prices and therefore hinder competition.
+
+From recent science policy developments in Europe it appears that Big
+Deals have gained support and remain firmly in place in the form of
+transformative agreements.
+Moreover, such agreements seem to become an increasingly dominant feature of scholarly publishing as indicated by the number of contracts that Elsevier and other publishers have successfully negotiated and secured since we conducted the analyses presented here.[^1] Through this study we can affirm that hybrid OA is complex
+as the financial flows involve research funders, libraries, consortia, and authors. However, it is on
+ publishers to increase the transparency of OA publishing, including hybrid OA and transformative agreements, by providing instantaneous open data about OA uptake and invoicing.
 
@@
 
-All data and code used for this study is openly available on GitHub, including version history:
+The source code data analysis is available on GitHub:
 
@@
 
-Formal analyis is written in R Markdown and thus replicable.
-
 # References
+
+[^1]: See, for instance, the announcements from the University of California: <https://www.universityofcalifornia.edu/news/uc-s-deal-elsevier-what-it-took-what-it-means-why-it-matters> and the Danish National Licensing Consortium: <https://www.kb.dk/en/about-us/licensing/national-licensing-consortium/elsevier-agreement-2021-24>

### 2021-05-19 analysis/manuscript.tex
align text

--- 2021-05-09 analysis/manuscript.tex
@@
 
-\title{\textbf{Transparency to hybrid open access  through publisher-provided metadata: An article-level study of Elsevier}}
+% tightlist
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\title{\textbf{Towards transparency of hybrid open access through publisher-provided metadata: An article-level study of Elsevier}}
 \shorttitle{Hybrid OA Elsevier}
@@
 \abstract{With the growth of open access (OA), the financial flows in scholarly journal
-publishing have become increasingly complex, but comprehensive data and
-transparency into these flows are still lacking. The opaqueness is especially
+publishing have become increasingly complex, but comprehensive data on and
+transparency of these flows are still lacking. The opacity is especially
 concerning for hybrid OA, where subscription-based journals publish individual
 articles as OA if an optional fee is paid. This study addresses the lack of
-transparency by leveraging Elsevier article metadata and provides the first
-longitudinal publisher-level study of hybrid OA uptake and invoicing. Our
+transparency by leveraging Elsevier article metadata and provides the first publisher-level study of hybrid OA uptake and invoicing. Our
 results show that Elsevier's hybrid OA uptake has grown steadily but slowly from
@@
 publishing agreements and OA policies in hybrid OA publishing. Our results
-further demonstrate the value of publisher-provided metadata improve the
-transparency in scholarly publishing by linking invoicing data to bibliometrics.}
+further demonstrate the value of publisher-provided metadata to improve the
+transparency in scholarly publishing.}
 
@@
 
-The rise of open access (OA) to journal articles has added complexity to
-scholarly publishing, particularly concerning transparency of economic
-dimensions. Financial transparency in journal publishing has long been
-lacking since Big Deal subscription contracts between academic
-institutions and large academic publishers usually include
-confidentiality clauses that prohibit publicly sharing information about
-the scope and pricing of such agreements (Bergstrom et al., 2014;
-Frazier, 2001; Larivière et al., 2015). Further, the distributed process
-of paying for accessing and publishing scholarly articles, which can
-involve several actors at the same time, including authors, research
-organizations, libraries, and research funders, adds to the opacity of
-the financial flows of scholarly publishing (Lawson et al., 2016).
+The rise of open access (OA) has added complexity to scholarly
+publishing, particularly concerning transparency of economic dimensions.
+Financial transparency in journal publishing has long been lacking
+because Big Deal subscription contracts between academic institutions
+and large academic publishers usually include confidentiality clauses
+that prohibit publicly sharing the scope or pricing of such agreements
+(Bergstrom et al., 2014; Frazier, 2001; Larivière et al., 2015).
+Further, the distributed payment processes to access and publish
+scholarly articles add to the opacity of the financial flows of
+scholarly publishing (Lawson et al., 2016) as they can involve several
+actors simultaneously, including authors, research institutions,
+libraries, and funders.
 
@@
 case of hybrid OA. In this OA business model, individual articles can be
-made openly available for a fee, also known as article-processing charge
-(APC), while the journal as a whole remains subscription access. Hence,
-hybrid OA was originally introduced as a low-risk transitional model
-that allows journals to gradually convert to full OA and reduce
-subscription costs as the OA uptake increases (D. C. Prosser, 2003).
-Many large subscription publishers have since incorporated this OA
-model, and while the uptake among authors increased only slowly during
-the first decade (Björk, 2012), it has gradually gained popularity in
-recent years. This development was driven by the implementation of OA
-policies by research institutions and funders, the allocation of OA
-funds, and formal agreements between publishers and institutions that
-allow affiliated authors to publish OA free-of-charge (Laakso \& Björk,
-2016). However, there have been concerns that the hybrid model allows
-for double-dipping---receiving two payments for one article, the APC and
+made openly available by paying an article processing charge (APC),
+while the journal as a whole remains subscription-access. Hence, hybrid
+OA was originally introduced as a low-risk transitional model that
+allows journals to gradually convert to full OA and reduce subscription
+costs as the OA uptake increases (D. C. Prosser, 2003). Many large
+subscription publishers have since incorporated hybrid OA, but there
+have been concerns that the hybrid model allows for double
+dipping---receiving two payments for one article, the APC and
 subscription fees (D. Prosser, 2015; Shieber, 2009). While publishers
-assure they combine subscription fees and APCs to adjust pricing
-(Mittermaier, 2015), without transparency around the uptake of hybrid OA
-and revenue streams, such claims are impossible to evaluate.
-
-With the recent introduction of transformative agreements, an evolving
-term for contracts that shift library spendings for subscriptions to OA
-(Borrego et al., 2020; Hinchliffe, 2019), the demand for
-publisher-provided data has even increased. If publishers would enable
-transparency about the OA uptake in hybrid journals and APC invoicing,
-academic libraries could improve their assessments and adjustments of
-contracts with publishers (Antelman, 2017; Schimmer et al., 2015).
-Likewise, research organizations and funders could improve compliance
-monitoring with OA mandates, while ensuring that the different actors
-were not charged multiple times (Larivière \& Sugimoto, 2018). However,
-as formal bibliometrics and financial studies have noted, the lack of
-publicly available and standardized data about hybrid OA publishing is
-considerable (Laakso \& Björk, 2016; Lawson, 2015; Pinfield et al.,
-2016)
-
-This current study investigates the extent to which publisher-provided
-data can be utilized to research hybrid OA. The study takes Elsevier,
-the largest scholarly journal publisher in terms of article volume
-(Larivière et al., 2015), as the starting point for an empirical
-analysis of hybrid OA uptake and invoicing. Our approach is based on
-openly available data sources and brings together scholarly metadata
-from Crossref, a DOI registration agency for scholarly works, and
-machine-readable APC invoicing data from Elsevier OA full-texts. The
-so-compiled data enable us to analyze the development of Elsevier's
-hybrid journal portfolio from 2015-19, and to compare it with that of
-Elsevier's full and delayed OA program. Further, we investigate whether
-hybrid publishing fees were waived or charged. In particular, we
-identify academic consortia and research funders that were directly
-invoiced by Elsevier, and compare their hybrid OA rates by license and
-subject. As such, our findings have implications for research and OA
+assure they adjust their pricing (Mittermaier, 2015), without
+transparency around the uptake of hybrid OA and both revenue streams,
+such claims are impossible to evaluate. This is especially concerning
+because hybrid OA has recently gained popularity after a decade of only
+slow uptake (Björk, 2012). The main drivers of this development have
+been research institutions and funders that implemented OA policies,
+allocated OA funds, and established formal agreements with publishers
+that allow affiliated authors to publish OA free-of-charge (Laakso \&
+Björk, 2016).
+
+The recent introduction of transformative agreements, an evolving
+concept describing contracts that shift library spending from
+subscriptions to OA (Borrego et al., 2020; Hinchliffe, 2019), might
+perpetuate the lack of transparency because pricing is often based on
+traditional subscription costs. Hence, the demand for publisher-provided
+data has increased. More transparency about hybrid OA uptake and funding
+could facilitate the assessment and adjustment of publisher contracts
+(Schimmer et al., 2015), enhance OA mandate compliance monitoring, and
+avoid double-dipping (Larivière \& Sugimoto, 2018). However, previous
+studies have noted an absence of transparency in hybrid OA publishing
+and a considerable lack of publicly available and standardized data
+(Laakso \& Björk, 2016; Lawson, 2015; Pinfield et al., 2016).
+
+This paper investigates to what extent publisher-provided data can
+increase the transparency around hybrid OA by taking Elsevier, the
+largest scholarly journal publisher (Larivière et al., 2015), as the
+starting point for an empirical analysis of hybrid OA uptake and
+invoicing. Our approach is based on openly available data sources and
+combines scholarly metadata from Crossref, a DOI registration agency,
+with APC invoicing data from Elsevier. The compiled data enable us to
+analyze the hybrid OA uptake in Elsevier journals between 2015 and 2019
+by license and subject, and to compare it with Elsevier's full and
+delayed OA program. Further, we determine whether hybrid APCs were
+waived or charged to authors or as part of publishing agreements. In the
+latter case, we examine the invoiced academic consortia and research
+funders. As such, our findings have implications for research and OA
 policy and highlight the potential of publisher-provided data for
@@
 \section{Background}\label{background}}
-
-The ability to identify and discover hybrid OA articles plays a
-fundamental role in monitoring hybrid OA, yet time and again, previous
-studies have underscored the difficulties involved in this (cf.
-(Akbaritabar \& Stahlschmidt, 2019; Björk, 2012; Emery, 2013; Laakso \&
-Björk, 2016; Leeuwen et al., 2018). Insights into the volume of hybrid
-OA and the various levels and sources of APC funding could create a
-better understanding of the costs involved in scholarly publishing and,
-thus, strengthen institutions' bargaining power in negotiations with
-publishers, and promote greater equity in scholarly publishing. In the
-following section, we examine how previous research has tackled the
-challenge of identifying hybrid OA content and what these studies taught
-us about the uptake and funding sources of hybrid OA.
 
@@
 Wiley--recorded growing numbers of hybrid journals that coincided with a
-twenty-fold increase in hybrid OA articles (from 666 to 13,994). As a
-result, hybrid journals with at least one OA article more than doubled
-in number (1,082 in 2009, 2,714 in 2013) but fell by 31\% relative to
-the number of hybrid journals in total. Although Laakso and Björk's
-(2016) presented a comprehensive overview tracking the growth of hybrid
-OA over time and across multiple publishers, their exploratory approach
-is unsuitable for repeated use as it required a considerable amount of
-manual data collection and cleaning. Minimizing such manual tasks,
-Nelson \& Eggett (2017) narrowed their focus to one publisher--the
-American Chemical Society (ACS)--and requested a list of all hybrid OA
-articles published from 2006-2011 directly from the ACS. To assess the
-hybrid OA uptake, the authors aggregated and compared the number of OA
-(n=814) and subscription articles (n=27,621), finding that over the five
-years, 2.9\% of ACS articles were published as hybrid OA.
+twenty-fold increase in hybrid OA (from 666 to 13,994). As a result,
+hybrid journals with at least one OA article more than doubled in number
+(1,082 in 2009, 2,714 in 2013) but fell by 31\% relative to the total
+number of hybrid journals. Laakso and Björk's (2016) exploratory
+approach yielded a comprehensive overview tracking the growth of hybrid
+OA over time and across multiple publishers, but due to the amount of
+manual data collection and cleaning it is unsuitable for repeated use.
+Minimizing such manual tasks, Nelson \& Eggett (2017) narrowed their
+focus to one publisher--the American Chemical Society (ACS)--and
+requested a list of all hybrid OA articles published from 2006-2011 from
+the ACS. To assess the hybrid OA uptake, the authors aggregated and
+compared the number of OA (n=814) and subscription articles (n=27,621),
+finding that over the five years, 2.9\% of ACS articles were published
+as hybrid OA.
 
@@
 al.'s (2020) approaches are that such data sources might not be openly
-available or, regarding CRIS data, might not be available at all since
+available or, regarding CRIS data, might not be available at all because
 not all countries maintain comprehensive institutional publication data.
@@
 Other studies assessed the share of (hybrid) OA for the entire corpus of
-scholarly research articles. For instance, using research articles with
-an assigned DOI and indexed in the Web of Science Core Collection
-(n=2,610,305) as a benchmark, Martín-Martín et al. (2018) assessed how
-many research articles from 2009 and 2014 are freely available through
-Google Scholar. After identifying OA articles through licensing
-information from Crossref, the authors estimated the share of hybrid OA
-at around 0.5\% in 2009 and 1.5\% in 2014. In a similar vein, Piwowar et
-al. (2018) drew on Crossref to identify content licenses. The authors
-followed a data-driven approach based on Unpaywall data, which is openly
-available and contained 66 million Crossref DOIs at the time, to
-determine how much of the scholarly literature is OA. Based on a random
-sample of 100,000 Crossref DOIs, the authors found 3.6\% of hybrid OA,
-while pointing towards a growing trend in recent years that recorded a
-hybrid OA share of 9.4\% of all articles published in 2015.
-
-Since its launch, Unpaywall has become widely used as an OA evidence
-source in bibliometric research and rankings(Huang et al., 2020;
-Robinson-Garcia et al., 2020). However, results cannot easily be
-compared with each other over time because of Unpaywall's ongoing
-changes due to non-standardized data to discover hybrid OA. These
-changes mainly reflect the growing albeit incomprehensive license
-information in Crossref metadata. To overcome these shortcomings, the
-developers extended previous web-scraping approaches to better
-distinguish between immediate and delayed OA based on licenses at the
-date of observation Piwowar et al. (2019). Relatedly, in October 2020,
-Elsevier's license for delayed OA content was added to Unpaywall's list
-of license patterns (Unpaywall, n.d.-c). At the journal-level, a more
-fine-grained identification of full OA journals using OA article
-information was developed to better distinguish them from subscription
-journals including hybrid journals (Unpaywall, n.d.-a). With its servie
-unsub, Unpaywall recently launched a paid service, where libraries can
-compare hybrid OA uptake rates with internal financial and usage data to
-assess the value of journal subscriptions (Chawla, 2020).
+research articles. Using DOI-assigned research articles indexed in the
+Web of Science Core Collection as a benchmark, Martín-Martín et al.
+(2018) investigated how many research articles from 2009 and 2014
+(n=2,610,305 in total) are freely available through Google Scholar.
+After identifying OA articles through licensing information from
+Crossref, the authors estimated the share of hybrid OA at around 0.5\%
+in 2009 and 1.5\% in 2014. Similarly, in a study assessing the OA level
+among all scholarly articles and that experienced by Unpaywall users,
+Piwowar et al.~(2018) obtained license information through Crossref and
+web-scraping. The study found 3.6\% of 100,000 random Crossref DOIs were
+hybrid OA and pointed towards a growing trend in recent years that
+recorded a hybrid OA share of 9.4\% of all articles published in 2015.
+
+Since its launch, Unpaywall has become widely used in bibliometric
+research and rankings (Huang et al., 2020; Robinson-Garcia et al.,
+2020). However, the lack of standardized and comprehensive
+publisher-provided data has required several updates to Unpaywall to
+improve hybrid OA identification and differentiation from delayed OA
+(Piwowar et al., 2019; Unpaywall, n.d.-c, n.d.-a), illustrating ongoing
+challenges in tracking and comparing hybrid OA prevalence over time.
+After changes in Unpaywall's approach to detect hybrid OA, Piwowar et
+al.~(2019) estimated a hybrid OA share of 4\% amongst all articles
+published by July 2019.
 
@@
 
-Some countries like Sweden and the United Kingdom (UK) have negotiated
-national publishing agreements that allow authors to publish hybrid OA
-at a discounted fee or for free. These agreements can serve as the basis
-for studies on the uptake of hybrid OA or cost-benefit analyses (cf.
-Olsson (2018); Lawson (2017)) but are often limited in scope as they
-tend to focus on a single country and rely on data that seldom is
-publicly available. Hence, while being valuable for specific use-cases,
-such studies can neither provide a comprehensive understanding of how
-the scholarly publishing system functions beyond national boundaries nor
-facilitate collective sharing of financial information about hybrid OA.
-
 Pinfield et al.'s (2016) analysis of APC payment records provided by 23
-UK higher education institutions revealed a sharp increase in central
-payments from 2007-2013, which was largely attributed to the
-introduction of block grants by Research Councils UK (RCUK) and
-non-compliance sanctions by the Wellcome Trust. Moreover, the study
-showed that OA fees were paid almost exclusively through RCUK and
-Wellcome block grants (92\%), and only a small number of APCs were paid
-through internal funding (7\%). In contrast, a recent Springer Nature
-survey found that authors draw on a range of funding sources to cover OA
-fees, such as dedicated institutional OA funds, block grants, OA
-agreements, or research grants (Monaghan et al., 2020). Most hybrid OA
-authors were supported through dedicated institutional OA funds (43\%,
-excluding block grants) and OA agreements with Springer Nature (34\%).
-The differences between these two studies may be due to changes in
-policy and funding arrangements---in particular, the introduction of OA
-agreements---since Pinfield et al. (2016), or due to Monaghan et al.'s
-(2020) more regionally diverse sample, indicating regional differences
-in APC funding.
+United Kingdom (UK) higher education institutions revealed a sharp
+increase in central payments from 2007-2013, which was largely
+attributed to the introduction of block grants by Research Councils UK
+(RCUK) and non-compliance sanctions by the Wellcome Trust. Moreover, the
+study showed that OA fees were paid almost exclusively through block
+grants (92\%), and only a small number of APCs were paid through
+internal funding (7\%). In contrast, a recent Springer Nature survey
+found that authors draw on a range of funding sources to cover OA fees,
+such as dedicated institutional OA funds, block grants, OA agreements,
+or research grants (Monaghan et al., 2020). Most hybrid OA authors were
+supported through dedicated institutional OA funds (43\%, excluding
+block grants) and OA agreements with Springer Nature (34\%). The
+differences between these two studies might be due to Monaghan et al.'s
+(2020) focus on Springer Nature articles or due to\} different policy
+and funding arrangements---considering the introduction of OA agreements
+since Pinfield et al. (2016) and Monaghan et al.'s (2020) more
+regionally diverse sample.
 
@@
 study by Jahn \& Tullney (2016) that analyzed APC records from 30 German
-higher education and research institutions crowd-sourced by the Open APC
-initiative, the Austrian Science Fund (FWF), Jisc, and the Wellcome
-Trust. In particular, the study revealed large differences in the amount
-of hybrid OA funded from 2014-2015. Whilst hybrid OA accounted for less
-than 1\% of APCs paid by German institutions (23 of 3,846), the three
-non-German research funders recorded a hybrid share of75\% (11,533 of
-15,779). According to Jahn \& Tullney (2016), this could point towards
-differences in science policy, such as hybrid OA being supported by the
-three non-German research funders but not by Germany's largest national
-funder (DFG). Another possibility is that German hybrid OA fees were
-paid from budgets not reported to the Open APC initiative, from where
-the authors acquired the German APC data. Among these unreported funds
-are research grants and research unit budgets, which author surveys
-identified as funding sources for hybrid OA fees (Graaf, 2017; Monaghan
-et al., 2020). As such, Jahn and Tullney's (2016) findings could reflect
-the complexities and potential limitations of institutional OA spending
-data that Pinfield et al.~(2016) and Monaghan et al.~(2020) attributed
-to incomplete or missing records.
-
-As this brief review has shown, the most comprehensive studies approach
-hybrid OA funding from an institutional perspective that focuses on
-research institutions or funders and draws on internal records or
-crowd-sourced data like from the Open APC initiative. The advantages of
-Open APC data are that they reflect the actual APCs paid rather than
-list prices; they are not limited to certain publishers or journals;
-they are provided in a standardized format regardless of the publisher.
-On the other hand, the drawbacks include the limited number of
-participating institutions that are almost exclusively based in Europe
-and the dependency on participating organizations for data integrity
-(Pieper \& Broschinski, 2018). Moreover, since Open APC data does not
-include waivers or free APCs, it cannot present a complete picture of
-all publishing activities and expenses. This, however, would pose less
-of an issue when analyzing just hybrid OA.
-
-\hypertarget{the-role-of-publisher-provided-data}{%
-\subsection{The role of publisher-provided
-data}\label{the-role-of-publisher-provided-data}}
-
-Uptake and financial studies have OA policy implications for hybrid OA.
-Most prominently, the cOAlition S, an international network of research
-funders, released its widely discussed Plan S. In September 2018, the
-cOAlition S announced that from 2021 onwards, signatories would no
-longer fund articles in hybrid OA journals. This principle includes
-so-called ``mirror'' and sister journals--hybrid journals with fully OA
-counterparts or split into their fully OA and subscription parts---that
-have launched since Plan S was announced but makes an exception for
-hybrid journals with transformative agreements (Frantsvåg \& Strømme,
-2019; Harrison, 2019; McNutt, 2019; S, n.d.). The Plan S refers to the
-Efficiency and Standards for Article Charges (ESAC) guidelines (Geschuhn
-\& Stone, 2017) as critical component of transformative agreements,
-emphasizing the need for publisher-provided metadata to address the lack
-of transparency surrounding hybrid OA. However, studies confirm that
-publisher with such agreements lack essential data about hybrid OA
-uptake and invoicing, or are not able or willing to provide such
-information publicly and instantaneously (Marques et al., 2019; Monaghan
-et al., 2020), which, in turn, not just limit the monitoring of hybrid
-OA, but also research studies on its size and financial implications.
-
-Here, we focus on Elsevier. The publisher provides article-level
-metadata to Crossref, including license information, which can be used
-to identify OA articles, and links to HTML and XML full-text versions
-for text and data mining purposes. Surprisingly, besides the full-text
-article, the XML files also contain invoicing information and the
-article's OA status. Table \ref{tab:els_xml} summarizes the key metadata
-elements, which comprise the type of entity to which the APC was
-invoiced (\texttt{openaccessSponsorType}) and the name of that entity if
-available (\texttt{openaccessSponsorName}). Moreover, the article-level
-metadata indicates if OA was provided through Elsevier's Open Archive
-program (\texttt{openArchiveArticle}), which enables free access to
-participating subscription journals after an embargo period (i.e.,
-delayed OA; Elsevier (n.d.-b)). As we have been unable to identify
-similar efforts by other large publishers (e.g., Springer Nature), this
-increased transparency through openly available bibliometric and
-financial data at the article level seems unique to Elsevier.
-
-\begin{table}[H]
-\caption{\label{tab:els_xml} Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019}
-\centering
-\begin{tabular}[t]{lp{10cm}}
-\toprule
-XML Node & Description\\
-\midrule
-\texttt{openaccessArticle} & Indicates if the article is open
-access. \\
-\texttt{openaccessSponsorType} & Invoice recipient type.\\
-\texttt{openaccessSponsorName} & Entity that received Elsevier's APC
-invoice.\\
-\texttt{openArchiveArticle} & Indicates if open access was provided through
-Elsevier's Open Archive programme.\tabularnewline \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-In this study, we utilize Elsevier's publicly available metadata to
-analyze the uptake and invoicing of their hybrid OA publishing option.
-In particular, we use this novel methodology to answer the following
+higher education and research institutions, the Austrian Science Fund
+(FWF), Jisc, and the Wellcome Trust. In particular, the study revealed
+large differences in the amount of hybrid OA funded from 2014-2015.
+Whilst hybrid OA accounted for less than 1\% of APCs paid by German
+institutions (23 of 3,846), the three non-German research funders
+recorded a hybrid share of 75\% (11,533 of 15,779). According to Jahn \&
+Tullney (2016), this could point towards differences in science policy,
+such as hybrid OA being supported by the three non-German research
+funders but not by Germany's largest national funder, the German
+Research Foundation (DFG). Another possibility is that German hybrid OA
+fees were paid from budgets not reported to the Open APC initiative, a
+crowd-sourcing effort (Pieper \& Broschinski, 2018) from where the
+authors acquired data. Among these unreported funds are research grants
+and research unit budgets, which author surveys identified as APC
+funding sources (Graaf, 2017; Monaghan et al., 2020). As such, Jahn and
+Tullney's (2016) findings could reflect the complexities and potential
+limitations of institutional OA spending data that Pinfield et al.
+(2016) and Monaghan et al. (2020) attributed to incomplete or missing
+records.
+
+In recent years, national consortia in Europe have negotiated publishing
+agreements covering hybrid OA fees for affiliated authors (Borrego et
+al., 2020). While these have improved invoicing workflows, internal
+assessments of hybrid OA uptake and invoicing remain challenging because
+transparent and comparative data have largely remained absent (Marques
+\& Stone, 2020). Such publicly available information might be scarce
+because publishers and institutions lack or withhold such data, or due
+to confidentiality clauses (Marques \& Stone, 2020; Monaghan et al.,
+2020). Several European consortia have collaborated through the
+Efficiency and Standards for Article Charges (ESAC) Initiative and
+Knowledge Exchange to standardize metadata requirements across
+publishers as great variation has been found between different countries
+and publishers (Marques et al., 2019).
+
+\hypertarget{research-questions-and-aims}{%
+\subsection{Research questions and
+aims}\label{research-questions-and-aims}}
+
+In this paper, we focus on Elsevier, a prominent example in recent
+hybrid OA uptake and financial studies (Pinfield et al., 2016; @ Laakso
+\& Björk, 2016). Elsevier's OA portfolio presents the challenges in
+examining hybrid OA described above. For instance, distinguishing
+between different OA types. Elsevier supports delayed (Elsevier,
+n.d.-b), hybrid, and full OA, including so-called mirror journals---full
+OA counterparts of hybrid journals addressing OA policies opposed to
+hybrid OA (Harrison, 2019). Further, Elsevier processes APC invoices
+through various channels, such as agreements with research funders and
+library consortia or, in the absence thereof, the authors (Elsevier,
+n.d.-a, n.d.-c). Surprisingly, we found Elsevier article-level metadata
+embedded in XML full-texts indicating the articles' OA status and
+invoicing. Leveraging this publicly available data, we address the lack
+of transparency around hybrid OA noted in previous studies. In
+particular, we use this novel approach to answer the following
 questions:
@@
 \begin{itemize}
+\tightlist
 \item
@@
 supplemented our sample with Crossref metadata and text-mined invoicing
-data to investigate the financing of immediate OA articles provided
-under a Creative Commons license in subscription-based journals. Figure
-\ref{fig:workflow} visualizes the automated workflow we used to collect
-data from Elsevier and Crossref.
+data to investigate the invoicing of immediate OA articles provided
+under a Creative Commons (CC) license in subscription-based journals.
+Figure \ref{fig:workflow} visualizes the automated workflow we used to
+collect data from Elsevier and Crossref.
 
@@
 
-\caption{Data Analytics Workflow to obtain article-level OA and invoicing data using freely available data sources provided by Elsevier and Crossref.
+\caption{Data Collection Workflow to Obtain Article-Level OA and Invoicing Data.
 }\label{fig:workflow}
@@
 
-First, we obtained hybrid journals through an Elsevier APC price list
-from May 2020 provided as captured by Matthias et al. (2019). We only
-included hybrid journals that were published under the hybrid model
-throughout the entire period of our analysis. Hence, we excluded
-journals that flipped from OA to hybrid and from hybrid to OA (Matthias,
-2020). This approach allowed us to ensure that the article-level
-differences we observed were not due changes at the journal-level (i.e.,
-increases or decreases in number). Overall, we identified 1,970 unique
-hybrid journals.
+First, we identified hybrid journals through an Elsevier APC price list
+from May 2020 provided by Matthias (2020). We excluded journals that
+transitioned from hybrid to full OA and reverse-flip journals that
+flipped from full to hybrid OA (Matthias et al., 2019). Overall, we
+identified 1,970 unique hybrid journals that published between 2015 and
+2019.
 
@@
 meeting abstracts, indicated by non-numeric pagination. We excluded
-non-scholarly journal content, such as the table of contents and lists
-of reviewers, following Unpaywall's paratext recognition approach
-(Unpaywall, n.d.-b), which we expanded to include patterns indicating
-corrections. Furthermore, we categorized the articles by subject
-according to the All Science Journal Classification code (ASJC) and
-added citation indicators from the Scopus Source Title list (version
-June 2020) based on the journal they were published in.
-
-Crossref metadata also contains license information that indicates if
-and under which terms the publisher has made articles openly available
-on their website (Hendricks et al., 2020). We identified OA articles
-through Creative Commons (CC) licenses and then downloaded the XML
-version of all CC-licensed articles published in a hybrid journal. From
-the XML files, we determined whether or not the articles were immediate
-OA using the XML node \texttt{openArchiveArticle} and measured the
-uptake of hybrid OA through three bibliometric indicators (Laakso \&
-Björk, 2016; Nelson \& Eggett, 2017)-\/---the number of hybrid journals
-with at least one OA article, the number of hybrid OA articles, and the
-number of hybrid OA articles relative to the number of closed-accessed
-articles in hybrid journals with at least one OA article.
-
-Moreover, we obtained APC invoicing metadata from the XML (see Table 1),
-which enabled us to identify the invoice channels and recipients of
-hybrid OA APCs. Based on the \texttt{openaccessSponsorType} node in the
-XML files, we distinguish between four invoice channels, including
-invoices billed to the corresponding author, as part of publishing
-agreements with funding bodies (hereinafter ``agreements''; cf. Elsevier
-(n.d.-a)), fee waivers (e.g., in ``cases of genuine need'' or due to
-society or university sponsorships, cf.~ref), and other types not
-specified by Elsevier. Finally, when hybrid APCs were invoiced as part
-of OA agreements, we identified invoice recipients through the
-\texttt{openaccessSponsorName} node.
+non-scholarly journal content, such as the table of contents, following
+Unpaywall's paratext recognition approach (Unpaywall, n.d.-b), which we
+expanded to include patterns indicating corrections. Furthermore, we
+categorized the articles by subject according to the All Science Journal
+Classification code (ASJC) and added citation indicators from the Scopus
+Source Title list (version June 2020) based on the journal they were
+published in.
+
+We identified OA articles through CC licenses in Crossref metadata
+records (Hendricks et al., 2020) and then downloaded the XML version of
+all CC-licensed articles published in a hybrid journal. From the XML
+files, we obtained the articles' OA status and invoicing information
+(see Table \ref{tab:els_xml}). We determined whether articles were
+immediate or delayed OA using the XML node \texttt{openArchiveArticle},
+which indicates if an article was made freely available under Elsevier's
+Open Archive program, and measured the uptake of hybrid OA. Moreover, we
+obtained the invoice channels and recipients of hybrid OA APCs. Using
+the \texttt{openaccessSponsorType} node, we distinguished between four
+invoice channels, including invoices billed to authors, as part of
+publishing agreements with funding bodies (cf. Elsevier (n.d.-a)),
+exempted through fee waivers (e.g., in ``cases of genuine need'\,' or
+due to society or university sponsorships, cf. Elsevier (n.d.-c)), and
+other types not specified by Elsevier. Finally, when hybrid APCs were
+invoiced as part of agreements, we identified invoice recipients through
+the \texttt{openaccessSponsorName} node. It is worth noting that while
+this article focuses on OA articles published in hybrid journals, the
+same data is also available for articles published in full OA journals.
+Elsevier does not provide the invoiced APC in either case.
+
+\begin{table}[H]
+\caption{\label{tab:els_xml}Metadata in Elsevier XML Full-Texts.}
+\centering
+\begin{tabular}[t]{lp{10cm}}
+\toprule
+XML node & Description\\
+\midrule
+\texttt{openaccessArticle} & Was the article open
+access? \\
+\texttt{openaccessSponsorType} & Invoice channel.\\
+\texttt{openaccessSponsorName} & Invoice recipient.\\
+\texttt{openArchiveArticle} & Was open access provided through
+Elsevier's Open Archive program?\tabularnewline \\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 sectors, countries, and primary research areas. Following the OECD's
-Frascati Manual (OECD, 2015, p. 91), we coded for four sectors, business
-enterprise, government, higher education, and private non-profit. The
-government sector consists of federal, state, and municipal governments,
-plus non-profit institutions controlled by government units (OECD, 2015,
-p. 100). The higher education sector comprises institutions that perform
-research and development and provide formal tertiary education services,
-such as universities or research institutes (OECD, 2015, p. 101). Legal
-or social entities that produce goods and services without generating
-profit, such as learned societies or charities, constitute the private
-non-profit sector (OECD, 2015, p. 288). Due to the low article volume,
-we combined invoice recipients Elsevier listed as ``authors'' and
-``third-party sponsor,'' and those we had coded as business enterprises
-(OECD, 2015, p. 97), into ``Others.'' Moreover, we categorized invoice
-recipients according to the country or countries representing their
-scope of funding and based on the following primary research areas
-health sciences, life sciences, physical sciences and mathematics,
-social sciences and humanities, unknown, and broad (i.e., multiple
-research areas).
-
-It is important to note that we could not find any official
-comprehensive documentation of Elsevier's XML schema , at the time of
-writing. However, our conceptualization of the variables and
-interpretation of the data was informed by a comprehensive review of
-Elsevier's websites about OA, including a list of funding agreements and
-details about Elsevier's Open Archive program,and Elsevier's article
-submission process in the context of dedicated OA agreements (Elsevier,
-n.d.-d, n.d.-e, n.d.-c). For validation, we matched the invoicing data
-provided by Elsevier to data from the Open APC initiative.
-
-Throughout the mostly automated data gathering and analysis process, we
+Frascati Manual (OECD, 2015, p. 91), we coded for four sectors: business
+enterprise, government, higher education, and private non-profit. Due to
+the low article volume, we combined the business enterprise sector with
+invoice recipients Elsevier listed as ``authors'' and ``third-party
+sponsor'' into ``Others.'' Moreover, we categorized invoice recipients
+according to the countries representing their scope of funding and based
+on the following primary research areas health sciences, life sciences,
+physical sciences and mathematics, social sciences and humanities, broad
+(i.e., multiple research areas), and unknown. Further, we compared
+Elsevier's invoicing data with institutional spending data from the Open
+APC initiative.
+
+Throughout this mostly automated data gathering and analysis process, we
 used tools from the Tidyverse (Wickham et al., 2019) for the R
@@
 BigQuery, applying the rcrossref (Chamberlain et al., 2020) parsers to
-extract relevant metadata fields. We interfaced Google BigQuery with SQL
-with the bigrquery client (Wickham \& Bryan, 2020). We used crminer
-(Chamberlain, 2020) to obtain the XML files from Elsevier and the xml2
-package (Wickham et al., 2020) to parse relevant metadata. To ensure our
-results are reproducible, we provided all code for obtaining the data
-and performing the analyses with detailed documentation as RMarkdown
-notebooks on GitHub
-(\url{https://github.com/njahn82/elsevier_hybrid_invoicing}).
+extract relevant metadata fields. We used crminer (Chamberlain, 2020) to
+obtain the XML-full texts from Elsevier.
 
@@
 descriptive analysis of Elsevier's invoicing data, highlighting
-licensing and disciplinary differences for different invoicing channels,
-institutional sectors, and individual invoice recipients.
+licensing and disciplinary differences for invoicing channels and
+invoice recipients.
 
@@
 articles, which represents 3\% of their total article volume
-(n=2,422,087). Table \ref{tab:overview} presents these findings for each
-year and the aggregated five-year period, illustrating a moderate growth
-in hybrid OA over time. With each year, the number of hybrid journals
-with at least one OA article increased, as did the number of immediate
-OA articles in these journals, which nearly doubled from 10,672 in 2015
-to 19,311 in 2019. However, since the total number of articles published
-in these journals also grew over time, the relative share of OA articles
-in hybrid journals only increased slightly from 2.6\% to 3.7\%.
+(n=2,422,087). Table \ref{tab:overview} presents findings for each year
+and the aggregated five-year period, illustrating moderate growth in
+hybrid OA over time. Each year, the number of hybrid journals with at
+least one OA article increased, as did the number of immediate OA
+articles in these journals, which nearly doubled from 10,672 in 2015 to
+19,311 in 2019. However, since the journals' total article output also
+grew over time, the relative share of OA articles in hybrid journals
+only increased slightly from 2.6\% to 3.7\%.
 
@@
 
-\caption{\label{tab:overview}Overview of hybrid open access (OA) uptake across Elsevier hybrid journals with at least one OA article 2015-2019.}
+\caption{\label{tab:overview}Elsevier Hybrid OA Uptake  2015-2019.}
 \centering
@@
 \addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{Elsevier Hybrid Journals with ≥ 1 OA article}}\\
+\multicolumn{7}{l}{\textbf{Elsevier Hybrid Journals with ≥ 1 OA article (n)}}\\
 \hspace{1em}Total & 1,317.0 & 1,364.0 & 1,401.0 & 1,501.0 & 1,600.0 & 1,755.0\\
 \addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{Articles in Elsevier Hybrid Journals with ≥ 1 OA Article}}\\
+\multicolumn{7}{l}{\textbf{Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)}}\\
 \hspace{1em}Total & 406,701.0 & 422,423.0 & 436,418.0 & 468,952.0 & 518,062.0 & 2,422,087.0\\
@@
 \addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{OA Articles in Elsevier Hybrid Journals with ≥ 1 OA Article}}\\
+\multicolumn{7}{l}{\textbf{OA Articles in Elsevier Hybrid Journals with ≥ 1 OA article (n)}}\\
 \hspace{1em}Total & 10,672.0 & 12,729.0 & 13,361.0 & 15,570.0 & 19,311.0 & 71,643.0\\
-\hspace{1em}Avg. & 8.1 & 9.3 & 9.5 & 10.4 & 12.1 & 40.8\\
+\hspace{1em}Mean & 8.1 & 9.3 & 9.5 & 10.4 & 12.1 & 40.8\\
 \hspace{1em}SD & 12.9 & 19.7 & 15.6 & 16.2 & 20.1 & 68.9\\
 \addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA Article}}\\
-\hspace{1em}Total (\%) & 2.6 & 3.0 & 3.1 & 3.3 & 3.7 & 3.0\\
-\hspace{1em}Avg. & 3.7 & 4.3 & 4.4 & 4.7 & 5.4 & 3.8\\
+\multicolumn{7}{l}{\textbf{Share of Elsevier Hybrid Journals with ≥ 1 OA article (\%)}}\\
+\hspace{1em}Total & 2.6 & 3.0 & 3.1 & 3.3 & 3.7 & 3.0\\
+\hspace{1em}Mean & 3.7 & 4.3 & 4.4 & 4.7 & 5.4 & 3.8\\
 \hspace{1em}SD & 4.5 & 5.9 & 5.1 & 4.9 & 5.7 & 4.1\\
@@
 
-Figure \ref{fig:boxuptake} presents the hybrid OA uptake by year,
-showing considerable variation among Elsevier's hybrid journals. The
-box-and-whiskers plot visualizes the median hybrid OA uptake (horizontal
-line inside the box), the range of the majority of journals (the box),
-and individual outliers (dots). Notably, the upper quartiles and
+Table \ref{tab:overview} also shows that the average OA share per
+journal was higher than the overall percentage. To consider these
+variations across journals, Figure \ref{fig:boxuptake} illustrates the
+range of OA uptake rates as a box-and-whiskers plot by year. The bold
+line shows the median OA prevalence across journals, which slightly rose
+from 2.4\% in 2015 to 3.7\% in 2019\}. Notably, the upper quartiles and
 whiskers stretch farther over the years, which indicates that the range
@@
 
-\caption{Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 20\%.}\label{fig:boxuptake}
+\caption{Open Access Uptake per Elsevier Hybrid Journal (Per Year).}\label{fig:boxuptake}
 \end{figure}
 
-\hypertarget{share-of-the-total-volume-published-in-oa}{%
-\subsubsection*{Share of the Total Volume published in
-OA}\label{share-of-the-total-volume-published-in-oa}}
-\addcontentsline{toc}{subsubsection}{Share of the Total Volume published
-in OA}
-
-During the same period, Elsevier published 328,601 OA articles in total
-(i.e., in hybrid and full OA journals), representing 12.4\% of the total
-article volume (n=2,643,474). Looking at Figure \ref{fig:oa_volume_fig},
-it is apparent that the OA article volume of hybrid journals (n=71,643;
-21.8\%) lagged behind that of Elsevier's Open Archive program
-(n=163,643; 49.8\%) and full OA journals (n=93,315; 28.4\%), which
-included 817 articles published in 38 mirror journals.
+During the same period, Elsevier published OA to 328,601 articles in
+total (i.e., full, hybrid, and delayed OA combined), representing 12.4\%
+of the total article volume (n=2,643,474). Looking at Figure
+\ref{fig:oa_volume_fig}, it is apparent that the OA article volume of
+hybrid journals (n=71,643; 21.8\%) lagged delayed OA (n=163,643; 49.8\%)
+and full OA journals (n=93,315; 28.4\%), which included 817 articles
+published in 38 mirror journals.
 
@@
 
-\caption{Open Access (OA) in Elsevier journals in terms of article volume between 2015-2019.
-Figure A presents the aggregated OA article volume of Elsevier journals from 2015-2019 by OA type.
-Figure B presents the share of OA articles relative to Elsevier's overall article output by OA type and open content license from 2015-2019.}\label{fig:oa_volume_fig}
+\caption{Elsevier OA Article Volume 2015-2019 by (A) OA Type and (B) Open Content License.}\label{fig:oa_volume_fig}
 \end{figure}
@@
 1\%) . Most delayed OA articles were provided under an Elsevier user
-license (Els-User), which prohibits reuse for commercial purposes.
+license (Els-User), which prohibits reuse for commercial purposes and
+redistribution.
 
@@
 
-Next, we present disciplinary variations in hybrid OA. Table
-\ref{tab:subject_area_table} presents the high-level findings by AJSC
-subject area. In the five-year period we studied, the physical sciences
-(712 hybrid journals, 29,584 OA articles) and health sciences (634
-hybrid journals, 25,119 OA articles) were the most prevalent in terms of
-hybrid journals with at least one OA article. In terms of article
-volume, most OA articles were published in hybrid journals belonging to
-the life sciences (538 hybrid journals, 31,383 OA articles), while
-social sciences (372 hybrid journals, 11,204 articles) journals
-published the lowest number of OA articles. However, it is important to
-note that there is a large overlap between the journals' subject areas.
-Therefore, Table \ref{tab:subject_area_table} also assigns hybrid
-journals and OA articles fractionally to a subject area and compares it
-to the full counting, where journals assigned more than one subject in
-the AJSC classification were counted once for each field. For the
+Next, we present the hybrid OA uptake of different disciplines. Table
+\ref{tab:subject_area_table} presents the high-level findings by ASJC
+subject area. Between 2015 and 2019, the physical sciences (712 hybrid
+journals, 29,584 OA articles) and health sciences (634 hybrid journals,
+25,119 OA articles) recorded the most hybrid journals with at least one
+OA article. However, the life sciences ((538 hybrid journals, 31,383 OA
+articles) published the most hybrid OA articles, while the social
+sciences published the least (372 hybrid journals, 11,204 articles). It
+is important to note, though, that there is a large overlap between the
+journals' subject areas. Therefore, Table \ref{tab:subject_area_table}
+also presents fractional counts for hybrid journals and OA articles,
+which equally weights counting with respect to the total number of ASJC
+subject areas a journal belongs to. Journals with multiple ASJC codes
+that are members of the same subject area were counted once. For the
 remaining results, we only present the full counting.
@@
 
-\caption{\label{tab:subject_area_table}Subject area overview using full and fractional counting.
-Table reports the number of Elsevier hybrid journals with at least one open access (OA) article, and OA articles published from 2015-2019.
-One multidisciplinary with 61 OA articles and 11 journals with 37 OA articles that  could not be matched to the Scopus Source Title list used to obtain ASJC codes are not included.}
+\caption{\label{tab:subject_area_table}Full and Fractional Counting of Subject Areas (2015-2019).}
 \centering
@@
 lower median proportion of hybrid OA articles, whereas most life
-sciences and social sciences journals ranked above average. Figure X
-also shows large variations among the 26 subject fields, with the
-highest OA rates in immunology and microbiology (Mdn=5.1), environmental
-science (Mdn=4.8), and neuroscience (Mdn=4.3), whilst chemistry
-(Mdn=1.4), materials science (Mdn=1.4), and dentistry (Mdn=0.6) recorded
-the lowest OA rate.
+sciences and social sciences journals ranked above average. Figure
+\ref{fig:oa_sub_uptake} also shows large variations among the 26 subject
+fields, with the highest OA rates in immunology and microbiology
+(Mdn=5.1), environmental science (Mdn=4.8), and neuroscience (Mdn=4.3),
+whilst chemistry (Mdn=1.4), materials science (Mdn=1.4), and dentistry
+(Mdn=0.6) recorded the lowest OA rate.
 
@@
 
-\caption{Open access (OA) uptake in Elsevier hybrid journals in percent grouped by subject area and field between 2015-19.
-Visualized as a box-and-whiskers plot limited to an OA share of 20\%; the red line represents the overall median (2.5), jittered dots represent individual journals.
-Subject fields are ordered based on the median OA uptake.}\label{fig:oa_sub_uptake}
-\end{figure}
-
-\hypertarget{citation-impact}{%
-\subsubsection*{Citation Impact}\label{citation-impact}}
-\addcontentsline{toc}{subsubsection}{Citation Impact}
-
-Furthermore, we were interested in the relationship between the
-journals' field-specific impact and their OA uptake. Spearman's rho
-correlation coefficient was used to assess the relationship between the
-2019 Source Normalized Impact per Paper (SNIP) value calculated by
-Scopus and the journals' OA uptake that year. Considering only journals
-with at least one OA article, we found a weak but positive correlation
-between journal impact and OA uptake (\(r_s\)=0.1679, \(p<0.001\)),
-suggesting that the relationship is not very strong, based on our data
-(see Figure \ref{fig:sniptest}).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth,]{manuscript_files/figure-latex/sniptest-1}
-
-}
-
-\caption{Source Normalized Impact per Paper (SNIP) versus open access (OA) uptake 2019.
-Each dot represents a hybrid journal with at least one published OA article in 2019.
-There is only a weak, positive correlation between the two variables ($r_s$=0.1679, $p<0.001$)}\label{fig:sniptest}
+\caption{OA Uptake in Elsevier Hybrid Journals by Subject Area and Field (2015-2019).}\label{fig:oa_sub_uptake}
 \end{figure}
@@
 
-As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APC
+As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APCs
 were most often invoiced to authors (n=41,725; 58.2\%) and to a lesser
@@
 found a small number of cases where hybrid APCs were waived (n=4,345;
-6.1\%).
+6.1\%). A brief note of caution regarding APCs invoiced to the authors
+is due here: Although author invoicing was indicated in the
+\texttt{openaccessSponsorType} article metadata field used for this
+study, this does not necessarily mean that authors paid the APC out of
+their own pocket. Rather, this implies that the authors' institution or
+funder did not have a central invoicing agreement in place at the time
+of publication.
 
@@
 
-\caption{\label{tab:invoicing_overview}Invoice channels for hybrid OA in Elsevier journals between 2015-19.}
+\caption{\label{tab:invoicing_overview}Invoice Channels for Elsevier Hybrid OA Articles (2015-2019).}
 \centering
@@
 Figure \ref{fig:invoiceoverview} illustrates that over the years,
-Elsevier has constantly invoiced hybrid OA APCs directly to authors
-compared to research funders or academic consortia (``Agreement'').
-Between 2015 and 2019, the proportion of author-invoicing increased from
-57\% to 61.2\%, while the share of invoicing through agreements slightly
-changed from 31.6\% to 32.4\%. Also, the share of fee-waived articles
-remained relatively stable, but we found different types of waivers.
-Around 51.7\% of fee waivers were linked to an OA sponsor, which
-indicates that a third party paid the APC. For instance, 853 waivers
-were linked to the French Académie des Sciences, presumably covering OA
-publication in its society journals for affiliated authors. The
-remaining 48.3\% of waived articles did not disclose any OA sponsor.
+Elsevier has consistently invoiced more authors than funders or academic
+consortia (``Agreement''), although both shares increased. In
+particular, the proportion of author-invoicing increased from 57\% to
+61.2\%, whereas the share of invoicing through agreements slightly
+increased from 31.6\% to 32.4\%, respectively. The share of fee-waived
+articles also remained relatively stable, but we found different types
+of waivers. Around 51.7\% of fee waivers were linked to an OA sponsor,
+which indicates that a third party paid the APC. For instance, 853
+waivers were linked to the French Académie des Sciences, presumably
+covering OA publication in its society journals for affiliated authors.
+The remaining 48.3\% of waived articles did not disclose any OA sponsor.
 Moreover, Figure \ref{fig:invoiceoverview} compares the invoicing
@@
 
-\caption{Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoice channel Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant.
-Grey bars show the total number of hybrid OA articles published in Elsevier journals from 2015-2019.}\label{fig:invoiceoverview}
+\caption{Elsevier Hybrid OA Articles by Invoice Channel and CC License (Per Year).}\label{fig:invoiceoverview}
 \end{figure}
@@
 subject field and invoice channel. For articles in nursing, decision
-dciences, and pharmacology, toxicology and pharmaceutics, Elsevier
+sciences, and pharmacology, toxicology and pharmaceutics, Elsevier
 predominantly invoiced authors, whereas most energy and chemical
@@
 astronomy can be attributed to a single 2015 issue of Nuclear and
-Particle Physics Proceedings, which offered sponsored OA.
+Particle Physics Proceedings.
 
@@
 Elsevier's data offer more insight into invoicing. Overall, we
-identified 62 sponsoring institutions and organizations that received
-invoices for hybrid publishing fees as part of an agreement with
-Elsevier. We found that, by a large margin, most invoices were issued to
-UK-based research funders and institutions (n=14,344; 59.2\%), followed
-by the Netherlands (n=2,835; 11.7\%) and the European Union (n=2,164;
-8.9\%; EU). Figure \ref{fig:invoice_sponsor_country} provides a closer
-look at the geographical distribution by license prevalence,
-highlighting the dominance of CC -BY licensed articles invoiced to
-sponsoring institutions in the UK, the United States, and, to a much
-lesser extent, the Netherlands. In contrast, the majority of articles
-invoiced to sponsoring institutions from Norway or representing the EU
+identified 63 academic institutions and funders that received invoices
+as part of publishing agreements. By a large margin, most invoices were
+issued to UK-based research funders and institutions (n=14,344; 59.2\%),
+followed by the Netherlands (n=2,835; 11.7\%) and the European Union
+(n=1,990; 8.2\%; EU), which in the context of this paper refers to
+institutions established by the European Commission, and not the
+individual member states. Invoicing data comprises two EU institutions,
+the research funder European Research Council and the health agency
+European Center for Disease Prevention and Control (ECDS). Figure
+\ref{fig:invoice_sponsor_country} presents the geographical distribution
+by license prevalence, highlighting the dominance of CC BY licensed
+articles invoiced to institutions in the UK, the United States (US),
+and, to a much lesser extent, the Netherlands. In contrast, most
+articles invoiced to institutions from Norway or representing the EU
 were published under a non-commercial license.
@@
 
-\caption{Number of centrally invoiced OA articles in Elsevier hybrid journals by country of recipient.
-Colored bars represent the Creative Commons (CC) license variant under which the article was published between 2015 and 2019.}\label{fig:invoice_sponsor_country}
+\caption{Centrally Invoiced Elsevier Hybrid OA Articles by Country and CC License (2015-2019).}\label{fig:invoice_sponsor_country}
 \end{figure}
 
-Figure \ref{fig:sponsor_sector} shows the yearly distribution of central
-invoiced articles by year and institutional sector. While UK-based OA
-invoices were mainly addressed to discipline-specific governmental and
-non-profit research funders, invoices to the Netherlands and Norway were
-issued to national academic consortia representing the higher education
-sector. In 2019, Elsevier also launched similar agreements in countries
-with lower publication output including Hungary and Poland. Besides, we
-found that invoiced funding bodies from the UK and the US mainly
-represented discipline-specific funders, while the invoiced funding
-bodies from other countries were focused on a broad variety of
-disciplines (see Table \ref{tab:subject_profile}).
+Figure \ref{fig:sponsor_sector} shows the yearly distribution of
+centrally invoiced articles by year and institutional sector. While
+UK-based OA invoices were mainly addressed to discipline-specific
+governmental and non-profit research funders, invoices to the
+Netherlands and Norway were issued to national academic consortia
+representing the higher education sector. In 2019, Elsevier also
+launched similar agreements in countries with lower publication output
+including Hungary and Poland. We found that invoice recipients from the
+UK and the US mainly represented discipline-specific funders, while
+invoice recipients from other countries focused on a broad variety of
+disciplines.
 
@@
 
-\caption{ Development of centrally invoiced OA articles in hybrid journals by country. Colored bars represent the institutional sector OA funding bodies belong to.}\label{fig:sponsor_sector}
+\caption{Centrally Invoiced Elsevier Hybrid OA Articles by Country and Sector (Per Year).}\label{fig:sponsor_sector}
 \end{figure}
-
-\begin{table}
-
-\caption{\label{tab:subject_profile} Relative subject profile of centrally invoiced OA funding bodies.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lrrrrrr}
-\toprule
-Country/Subject area & Health science & Life sciences & Physical sciences & SSH & Broad & Total\\
-\midrule
-United Kingdom & 25\% & 11\% & 21\% & 5\% & 0\% & 62\%\\
-The Netherlands & 0\% & 0\% & 0\% & 0\% & 12\% & 12\%\\
-European Union & 0\% & 0\% & 0\% & 0\% & 9\% & 9\%\\
-United States & 6\% & 0\% & 0\% & 0\% & 1\% & 7\%\\
-Norway & 0\% & 0\% & 0\% & 0\% & 3\% & 3\%\\
-\addlinespace
-Other & 1\% & 0\% & 1\% & 0\% & 5\% & 7\%\\
-\midrule
-Total & 32\% & 11\% & 22\% & 5\% & 30\% & 100\%\\
-\midrule
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\hypertarget{institutional-ranking}{%
-\subsubsection*{Institutional ranking}\label{institutional-ranking}}
-\addcontentsline{toc}{subsubsection}{Institutional ranking}
 
@@
 UK-based research funders and the US-based Bill and Melinda Gates
-Foundation mainly sponsored articles published under a CC BY license. In
-contrast, articles funded by VSNU, the European Research Council, and
+Foundation mainly received invoices for CC BY licensed articles. In
+contrast, articles invoiced to VSNU, the European Research Council, and
 Norway Institutes had a much lower proportion of CC BY.
@@
 
-\caption{\label{tab:sponsor_top_10} Top 10 Elsevier hybrid OA sponsors by centrally invoiced articles 2015-2019.
-The table furthermore presents the number of Elsevier hybrid journals with at least one invoiced article and two indicators of compliance: proportion of  CC BY licensed articles  and disclosed spending through the Open APC Initiative (OAPC.}
+\caption{\label{tab:sponsor_top_10}Elsevier Hybrid OA Invoice Recipients  (2015-2019).}
 \centering
@@
 
-Figure \ref{fig:field_funder_matrix} shows the subject profile of OA
-sponsors based on to the journals' subject fields of centrally invoiced
-articles. As expected, the figure shows that subject-specific funders,
-such as the British Engineering and Physical Sciences Research Council
-(EPSRC), tended to focus on a limited set of subject fields. In
-contrast, national consortia, such as VSNU and Norwegian Institutes
-(NO), financed OA articles for a broader range of fields. However,
-within these consortia, the subject distribution also varied. VSNU OA
-spending, for example, had a strong focus on articles relating to
-medicine, environmental science, and energy, presumably reflecting
-disciplines with a generally strong publication output at Dutch
-universities. Norwegian universities mostly published hybrid OA in
-engineering, environmental science and agricultural and biological
-sciences journals.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.95\textwidth]{../figure/funder_subject.png}
-
-}
-
-\caption{Top 10 Elsevier hybrid OA institutional invoice recipients in terms of articles published by journal subject.
-EPSRC, Engineering and Physical Sciences Research Council; VSNU, VSNU; Wellcome, Wellcome Trust; ERC, European Research Council; ERC, European Research Council; MRC, Medical Research Council; NERC, Natural Environment Research Council; BBSRC, Biotechnology and Biological Sciences Research Council; Gates, Bill and Melinda Gates Foundation; ESRC, Economic and Social Research Council; NO, Norway Institutes.
-}\label{fig:field_funder_matrix}
-\end{figure}
-
 \hypertarget{discussion}{%
@@
 
-Over the last five years, Elsevier recorded growth in the uptake of
-hybrid OA for each of our three indicators: The number of hybrid OA
-articles published per year doubled, the number of hybrid journals with
-at least one OA article grew by 21\%, and the share of hybrid OA
-articles relative to closed-accessed articles in these journals
-increased from 2.6\% to 3.7\%. As Laakso \& Björk (2016), we found only
-a weak relationship between citation metrics like the SNIP and hybrid OA
-uptake and observed differences in the uptake of hybrid OA among journal
-subject areas. In particular, we found the highest count of hybrid OA
-articles in physical sciences journals (see Table
+From 2015-2019, Elsevier recorded growth in the uptake of hybrid OA: The
+number of hybrid OA articles published per year doubled, the number of
+hybrid journals with at least one OA article grew by 21\%, and the share
+of hybrid OA articles relative to closed-accessed articles in these
+journals increased from 2.6\% to 3.7\%. As Laakso \& Björk (2016), we
+observed disciplinary differences. In particular, we found the highest
+count of hybrid OA articles in physical sciences journals (see Table
 \ref{tab:subject_area_table}). This was followed by the life sciences
-and health sciences, whereas the social sciences had the lowest count of
-hybrid articles. This order mostly reflects the disciplines' overall
-publication output. According to the Open Science Monitor by the
-European Commission (European Commission, 2019), the physical sciences
-and mathematics publish the most articles, followed by the health
-sciences, life sciences, and social sciences and humanities.
-
-Such disciplinary differences in hybrid OA prevalence become more
-meaningful when considering the relative share of OA articles to
-closed-access articles. In line with previous research, we found that
-Elsevier journals from the life sciences and social sciences (Jubb et
-al., 2017; Kramer \& Bosman, 2018; Laakso \& Björk, 2016) recorded
-greater than typical hybrid OA uptake, whereas physical sciences
-journals generally had a lower than typical uptake (Kramer \& Bosman,
-2018; Laakso \& Björk, 2016; Martín-Martín et al., 2018). In a
-systematic review of OA publishing patterns, Severin et al. (2020)
-linked such disciplinary differences to socio-cultural and technological
-factors that shape publishing cultures and practices. Among other
-aspects, the authors pointed to the availability of prestigious
-journals, the possibility to provide OA through alternative routes, and
-ease of access to funding sources to cover publishing fees. As such,
-looking at hybrid OA uptake in isolation can only provide limited
-insight, so that placing it within the broader context of disciplinary
-publishing and OA practices is necessary. For instance, since many
-branches of the physical sciences have established self-archiving
-practices and the technical infrastructure to support these (e.g., the
-arXiv; Severin et al. (2020), Björk et al. (2010), Laakso \& Björk
-(2016)), researchers can provide OA through repositories (i.e., green
-OA), and indeed, according to the Open Science Monitor the physical
-sciences provide OA primarily through the green route (31.2\% total OA
-share, 24\% OA if only considering green OA). Hence, the low share of
-hybrid OA in our sample might be because physical scientists perceive
-much less of a need for hybrid OA. Further, Severin et al. (2020) note
-project-based funding structures enable OA publishing in fully OA and
-hybrid journals as they allow for easy integration of publishing costs.
-Since these funding structures are typical for the life sciences
-(ibid.), we would expect higher rates of OA. Data from the Open Science
-Monitor seems to support this notion as the life sciences record
-particularly high shares of full OA articles and comparatively high
-shares of hybrid OA in some fields, such as biology and agricultural
-science. While we expected the observed hybrid OA shares for the life
-sciences and physical sciences, we were surprised by the high share of
-hybrid OA belonging to the AJSC subject area social sciences, which also
-includes the humanities, since articles ``social scientists have
-reported to face significant difficulties in access to grant funding for
-{[}publication charges{]}, as most research in the social sciences is
-not done by means of project-specific funding'' Severin et al. (2020, p.
-15). However, previous research has found comparatively low awareness
-and uptake of green OA among researchers in the field (Björk et al.,
-2010; Creaser et al., 2010), which might encourage hybrid OA uptake.
-Another possible explanation,is that Elsevier has a number of high
-prestigious social science and humanities journals (Larivière et al.,
-2015), which means that our findings are not representative of the
-general population of SSH journals. Likewise, this limitation could be
-extended to our findings in general that whilst not representative of
-the population of hybrid OA publishing, they provide valuable insight
-into a key player of scholarly publishing.
-
-Finally, OA policies in favor of hybrid OA in combination with invoicing
-agreements with Elsevier might be a driver. For instance, we found that
-nearly half of articles in arts and humanities were facilitated through
-agreements, which was above-average. This is in line with previous
-research, suggesting that OA policies and funder mandates have a
-stronger effect on OA uptake than disciplinary publication cultures
-(Huang et al., 2020; Larivière \& Sugimoto, 2018).
-
-Our overall investigation of invoice channels and recipients of hybrid
-OA APCs found that the majority of Elsevier hybrid articles were
-invoiced to the author (n=41,725; 58.2\%). However, it is important to
-note that although Elsevier's metadata classifies the sponsor type as
-``author,'' this does not necessarily mean that authors paid for APCs
-out of their own pocket but rather just that the APC was not invoiced
-through an agreement. Hence, it is possible that APCs invoiced to the
-author were covered by institutional funds or as part of research
-grants. For instance, a recent Springer Nature survey found that hybrid
-OA was predominantly supported through institutional and funder sources
-(71\%), followed by OA agreements (34\%), while only 6\% were paid from
-personal funds or savings (Monaghan et al., 2020). We also observed
-notable differences in licensing among invoice types. The majority of
-hybrid OA articles, regardless of the invoice type, were licensed under
-the more restrictive CC BY-NC-ND license. There is a lack of dedicated
-studies on licence selection, but indications so far suggest that when
-given a choice authors tend to select more restrictive variants for
-their published works (Fraser et al., 2020; Noorden, 2013; Rowley et
-al., 2017). On the other hand, when hybrid articles were invoiced
-through agreements, most articles were licensed under the more liberal
-CC BY license. As several funding bodies mandate CC BY licenses,
-including the UK research funders that account for 62\% of our
+and health sciences, whereas the social sciences had the lowest count.
+This order mostly reflects the disciplines' overall publication output.
+According to the Open Science Monitor (European Commission, 2019), the
+physical sciences publish the most articles, followed by the health
+sciences, life sciences, and social sciences.
+
+Disciplinary differences in hybrid OA prevalence become more meaningful
+when considering the relative share of OA articles to closed-access
+articles. In line with previous research, we found that Elsevier
+journals from the life sciences and social sciences (Jubb et al., 2017;
+Kramer \& Bosman, 2018; Laakso \& Björk, 2016) recorded greater than
+typical hybrid OA uptake, whereas physical sciences journals generally
+had a lower than typical uptake (Kramer \& Bosman, 2018; Laakso \&
+Björk, 2016; Martín-Martín et al., 2018). In a systematic review of
+disciplinary OA publishing patterns, Severin et al. (2020) highlighted
+the importance of socio-cultural and technological factors in shaping
+publishing cultures and practices. For instance, since many branches of
+the physical sciences have established self-archiving practices Laakso
+\& Björk (2016), researchers can provide OA through repositories and
+might perceive much less of a need for hybrid OA. In contrast, hybrid OA
+in the life sciences could be enabled through project-based funding
+structures as they allow for easy integration of publishing costs
+Severin et al. (2020). On the other hand, the high hybrid OA uptake that
+we observed among the social sciences could point to the influential
+role of OA policies and invoicing agreements (Huang et al., 2020;
+Larivière \& Sugimoto, 2018). While these aspects are important drivers
+of both full and hybrid OA, faculty members generally prioritize journal
+prestige and the journal impact factor (Niles et al., 2020) in
+publishing decisions, which, in combination with the availability of OA
+funding, might encourage publishing hybrid OA (Graaf et al., 2017).
+
+Our invoicing data analysis found that most APCs were invoiced to the
+author (n=41,725; 58.2\%). However, it is important to emphasize
+although Elsevier's metadata classifies the sponsor type as
+``author'\,', this does not necessarily mean that authors paid for APCs
+themselves but rather that APCs were not invoiced through publishing
+agreements. It is possible that these APCs were covered through
+institutional funds or research grants. This would align with a recent
+Springer Nature survey that found hybrid OA was predominantly supported
+through institutional and funder sources (71\%), followed by OA
+agreements (34\%), while only 6\% were paid from personal funds or
+savings (Monaghan et al., 2020).
+
+Further, we observed notable differences in licensing. Most hybrid OA
+articles invoiced to author were licensed under the more restrictive CC
+BY-NC-ND license. Previous research, while lacking dedicated studies on
+license selection, suggests that authors tend to select more restrictive
+license variants when given a choice (Fraser et al., 2021; Noorden,
+2013; Rowley et al., 2017). On the other hand, we found when hybrid OA
+was invoiced through agreements, most articles were licensed under the
+more liberal CC BY license. As several funding bodies mandate CC BY
+licenses, including the UK research funders that account for 62\% of our
 ``agreement'' subsample, this result is perhaps not surprising but
@@
 Based on our findings, around a third of the articles were invoiced
-directly through OA agreements (e.g., research funders, national library
-consortia). Strikingly, sponsors based in the UK accounted for more than
-half of these , and a quarter was charged to sponsors from other
-European countries. This finding reiterates reports from previous
-studies that the UK's OA profile differs markedly from other countries
-(cf. Jubb et al. (2015); Jubb et al. (2017)) and points to the effects
-of science policy. To increase OA to research publications, the UK
-implemented centralized APC funding and embraced hybrid OA as a
+through OA agreements (e.g., research funders, national library
+consortia). The predominance of UK funding bodies in this subsample
+reiterates reports from previous studies that the UK's OA profile
+differs markedly from other countries (cf. Jubb et al. (2015); Jubb et
+al. (2017)), pointing to the impact of science policy. To promote OA,
+the UK implemented centralized APC funding and embraced hybrid OA as a
 transition model. Within five years of the publication of the Finch
@@
 articles in 2016 (Jubb et al., 2017), which can be attributed to the
-availability of APC funding from Research Councils UK (RCUK). Indeed,
-RCUK block grants have been the largest single source of OA funds in the
-UK---the Wellcome Trust, and more recently transformative agreements
-(Jubb et al., 2017; Tickel, 2018). However, since then publishing
-expenditures of UK universities have been rising rapidly (ref), so that
-several universities stopped supporting hybrid OA in late 2018-2019
-(Castro, 2019; University of Birmingham, n.d.; University of Oxford,
-2019; Walker, 2019). This development might explain the slight decrease
-in hybrid articles invoiced to the UK our data showed around that time
-(see Figure \ref{fig:invoice_sponsor_country}).
-
-An essential part in the absence of transparency in scholarly publishing
-and especially in hybrid OA has been the lack of transparency in
-publishers' article management. This has prohibited consumers to get a
-comprehensive and comparative overview of their publishing activities
-and the associated costs and, thus, limited their negotiating power.
-However, certain initiatives in Europe like the ESAC Initiative have
-begun to press for improved and more transparent workflows for
-transformative agreements (Geschuhn \& Stone, 2017). In particular, the
-ESAC Workflow Recommendations seek to increase efficiency and
-transparency in the publication workflow through improved invoicing and
-reporting processes and metadata delivery, including article-based
-machine readable information about the OA sponsor. Through this study we
-demonstrated the utility and benefits of such publisher-provided
-metadata about hybrid OA funding and highlighted the need for extending
-these to include comprehensive information about the licensing and the
-particular OA type.
-
-Usually identifying delayed OA articles and distinguishing them from
-other types of OA is challenging since there is no publicly available
-list of such journals. For instance, Taylor \& Francis' (2021) website
-indicates that some of their journals participate in a delayed or
-promotional OA program but no further information is provided as to
-which journals these are. However, Elsevier provides a list of journals
-enrolled in the ``Open Archive'' programme and further tags delayed OA
-articles in the machine-readable XML version of the article. This
-enables easy identification and more accurate assessments of hybrid OA,
-which allows to take into account the substantial amount of delayed OA
-content in future hybrid OA studies using Elsevier data.
+availability of funding from RCUK. Indeed, RCUK block grants have been
+the largest single source of APC funds in the UK---the Wellcome Trust,
+and more recently transformative agreements (Jubb et al., 2017; Tickel,
+2018). However, since then publishing expenditures of UK universities
+have been rising rapidly (Jubb et al., 2017), so that several
+universities stopped supporting hybrid OA in late 2018-2019 (University
+of Birmingham, n.d.; Walker, 2019). This development might explain the
+slight decrease in hybrid articles invoiced to the UK our data showed
+around that time (see Figure \ref{fig:sponsor_sector}).
+
+Through this study we demonstrated the utility and benefits of
+publisher-provided metadata about hybrid OA invoicing and highlighted
+the need for extending these to include comprehensive information about
+licensing and APC waivers. Metadata guidance should also consider the
+substantial amount of delayed OA content, which needs to be
+distinguished from hybrid OA. Hence, our study substantiates the
+recommendations from the ESAC Initiative that seek to increase
+efficiency and transparency through improved invoicing and reporting
+processes and metadata about OA funding (Geschuhn \& Stone, 2017;
+Marques et al., 2019).
 
 While this study advances our knowledge about hybrid OA uptake and
-invoicing, the limitations leave room for future research. This study
-focused on only one publisher, which limits the generalizability of our
-findings since Elsevier's attributes such as composition of the journal
-portfolio, mix of business models, pricing, and promotion of various
-options cannot be assumed to be representative of scholarly journal
-publishing in general. Though the invoicing data extracted from the XML
-files enables unprecedented accuracy and comprehensiveness in the study
-of hybrid OA, it likely does not disclose the actual OA sponsor in all
-cases. Most articles were invoiced to authors, but it remains unclear
-how often authors themselves pay for APCs or invoices are covered
-through institutional OA funds or research grants. Further research into
-this topic would increase our understanding of OA funding streams
-outside of publishing agreements. Further, our study demonstrates that
-to achieve comprehensive mapping and understanding of the financial
-flows involved in OA funding, complex, in-depth country-specific
-meso-level analyses is needed. Such studies would ideally draw on
-various data sources, such as publicly available bibliometrics and
-metadata, information about significant national research funders and
-their OA policies, information about the national library and possible
-consortia landscape including their publishing agreements, which might
-require Freedom of Information Requests.
-
-This study provides a snapshot of hybrid OA for the largest journal
-publisher prior to the impact of the implementation of Plan S
-requirements. While many Plan S signatories have already had strong OA
-policies, this harmonized approach is likely to affect publishing
-decisions of funded authors,the licencing of their articles, and the
-offerings and pricing of publishers. Since the new requirements are
-applicable to research funded from 2021 onwards, a comparative study
-focused on articles invoiced to Plan S signatories would be a fruitful
-endeavor.
-
-In a similar vein, future studies could combine invoice information with
-research funding information since these contribute different pieces to
-the funding stream puzzle. This could also provide more insight into the
-actual APC funding streams behind the author-invoiced articles.
+invoicing, the limitations leave room for future research. We focused on
+only one publisher, which limits the generalizability of our findings
+because Elsevier's journal portfolio, mix of business models, pricing,
+and promotion of various options cannot be assumed to be representative
+of scholarly journal publishing in general. Relatedly, our study is
+limited to articles published in hybrid OA journals and, hence, is not
+representative of invoices for articles in full OA journals. To develop
+a full picture of OA invoicing, the methodology presented in this paper
+can be used for future studies in this direction. Further, although
+Elsevier's invoicing data improves transparency, it seems likely that
+not all actual OA funding bodies are disclosed. For instance, most
+articles were invoiced to authors, but it remains unclear if the APCs
+were paid by the authors themselves or through institutional OA funds or
+research grants. Research into this topic would improve our
+understanding of OA funding outside of publishing agreements. Moreover,
+our study demonstrates that comprehensive mapping of the financial flows
+of OA publishing requires complex, in-depth country-specific analyses.
+Such studies would ideally draw on various data sources to consider
+research funders, the consortia landscape, OA policies, and publishing
+agreements Future research could analyze funder policies in combination
+with institutional agreements (including which journals are covered by
+the agreements) and hybrid OA invoicing metadata to provide more insight
+into publishing patterns and CC licence choice. A further avenue of
+research would be dedicated studies on the strengths and weaknesses of
+data provided by publishers and the Open APC initiative, and consider an
+integrative approach that combines both datastreams to reap the benefits
+of each while mitigating their individual drawbacks.
+
+The policy implications and the underlying data of this study present
+opportunities for actors in the scholarly publishing landscape to rely
+less on assumptions and secondary data, and instead establish a direct
+connection between the publication output and the related financial
+information. This could increase accountability in spending public and
+charitable funds, and improve OA cost monitoring through enhanced
+transparency, precision, and timeliness. Analyses like we have presented
+here can inform library consortia in negotiations through better
+understanding their publisher-specific OA output, determining the amount
+of delayed OA content, and facilitating comparative analyses across
+different consortia. The study could also inform research funders and
+research institutions about license selection implications if there is
+not any strict policy in place. This study provides a snapshot of hybrid
+OA for Elsevier, the largest journal publisher, prior to the impact of
+the implementation of Plan S, an initiative to accelerate the transition
+to OA that will no longer support hybrid OA (cOAlition S, n.d.). While
+many Plan S signatories have already had strong OA policies, this
+harmonized approach is likely to affect publishing decisions of funded
+authors, the licencing of their articles, and the offerings and pricing
+of publishers at a larger scale than before. Because the new
+requirements apply to research funded from 2021 onward, a comparative
+study on articles invoiced to Plan S signatories would be a fruitful
+endeavor. Plan S does not fund hybrid OA publication unless the host
+journal is explicitly committed to convert to full OA, however, with the
+growth of national and institutional agreements that enable affiliated
+authors to publish hybrid OA, such restrictions will influence
+researchers differently. Invoicing data as used for this study can be
+useful in exploring later on, by looking at research funding statements
+and invoicing channels in relation to each other.
 
@@
 hybrid OA publishing from 2015-2019 to better understand the volume and
-invoicing of hybrid OA and present a novel, data-driven approach for
+invoicing of hybrid OA and to present a novel, data-driven approach for
 such analyses. Our results indicate that although the number of hybrid
-OA articles has increased over time, hybrid OA uptake has remained low.
-Notably, hybrid APCs were most often invoiced directly to the authors or
-through institutional agreements, where only a few funding bodies were
-the main driver of hybrid OA. Finally, our findings highlight that
+OA articles has increased over time, its uptake has remained low.
+Notably, hybrid APCs were most often invoiced directly to the authors,
+followed by agreements, where only a few funding bodies were the primary
+drivers of hybrid OA. Finally, our findings highlight that
 publisher-provided metadata about the invoicing channels of (hybrid) OA
@@
 
-Ever since its inception, hybrid OA has been a challenging subject to
-study due to the lack of standardised ways for publishers to flag such
-content and APC funding data being limited to studies based on surveys
-and articles' acknowledgements sections. This study presented a novel
-approach to studying OA APC invoicing that is based on publicly
-available publisher-provided metadata, which can be used on on its own
-or in combination with other public data sources to gain more detailed
-and comprehensive insights into hybrid OA uptake and invoicing than ever
-before. If more publishers reported OA invoicing data on the article
-level and in a machine-readable format, this would increase transparency
-and enable enhanced monitoring of the scholarly journal landscape over
-time. As hybrid OA has increasingly become a central element of OA
-policies of research funders and libraries, customer organisations could
-require that invoicing information is added to the article-level
+Since the beginning, hybrid OA has been a challenging subject to study
+due to the lack of standardized ways publishers flag such content and
+APC funding data being limited self-reported data, surveys, and other
+secondary sources. This study presented a novel approach to studying APC
+invoicing that is based on publicly available publisher-provided
+metadata, which can be used on on its own or in combination with other
+public data sources to gain more detailed and comprehensive insights
+into hybrid OA uptake and invoicing. If more publishers reported OA
+invoicing on the article level and in a machine-readable format, this
+would increase transparency and improve monitoring of the scholarly
+journal landscape over time. As hybrid OA has become a central element
+of OA policies of research funders and libraries, consumer organizations
+could require that invoicing information is added to the article-level
 metadata. As long as publishers do not provide this data in a structured
-and comprehensive format, they prevent consumers from benchmarking
-prices and therefore hinder price competition as the only alternative to
-gain comparative insight is self-reported APC expenditures that research
-funders and libraries voluntarily report to the Open APC initiative. The
-Open APC initiative is a valuable service but it cannot replace
-publisher-provided data since global participation is limited.
+and comprehensive format, they prevent benchmarking prices and therefore
+hinder competition.
 
@@
 Deals have gained support and remain firmly in place in the form of
-transformative agreements. However, in addition to research funders and
-libraries that increasingly make publishing contracts public, it is on
-the publisher to provide metadata as presented here to improve the
-transparency in scholarly publishing by linking invoicing data to
-bibliometrics. From this study we can affirm that hybrid OA is complex
-as the financial flows come from various stakeholders and invoicing
-forms, it is not just libraries or consortia, but authors and funders
-are also directly involved. As of right now, providing machine-readable
-article-level invoicing data is not common amongst publishers, which
-limits the potential for broader studies.
+transformative agreements. Moreover, such agreements seem to become an
+increasingly dominant feature of scholarly publishing as indicated by
+the number of contracts that Elsevier and other publishers have
+successfully negotiated and secured since we conducted the analyses
+presented here.\footnote{See, for instance, the announcements from the
+  University of California:
+  \url{https://www.universityofcalifornia.edu/news/uc-s-deal-elsevier-what-it-took-what-it-means-why-it-matters}
+  and the Danish National Licensing Consortium:
+  \url{https://www.kb.dk/en/about-us/licensing/national-licensing-consortium/elsevier-agreement-2021-24}}
+Through this study we can affirm that hybrid OA is complex as the
+financial flows involve research funders, libraries, consortia, and
+authors. However, it is on publishers to increase the transparency of OA
+publishing, including hybrid OA and transformative agreements, by
+providing instantaneous open data about OA uptake and invoicing.
 
@@
 
-All data and code used for this study is openly available on GitHub,
-including version history:
+The source code data analysis is available on GitHub:
 
 \url{https://github.com/njahn82/elsevier_hybrid_invoicing}
-
-Formal analyis is written in R Markdown and thus replicable.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-Akbaritabar_2019}{}%
-Akbaritabar, A., \& Stahlschmidt, S. (2019). \emph{Applying crossref and
-unpaywall information to identify gold, hidden gold, hybrid and delayed
-open access publications in the {KB} publication corpus}. SocArXiv.
-\url{https://doi.org/10.31235/osf.io/sdzft}
-
-\leavevmode\hypertarget{ref-antelmann_2017}{}%
-Antelman, K. (2017). Leveraging the growth of open access in library
-collection decision making. \emph{{At the helm: leading
-transformation}}, 411--422.
-
 \leavevmode\hypertarget{ref-Bergstrom_2014}{}%
@@
 
-\leavevmode\hypertarget{ref-Castro_2019}{}%
-Castro, P. de. (2019). \emph{Running a no-hybrid open access funding
-policy: Some results}. Open Access Strathclyde.
-\url{https://strathoa.tumblr.com/post/185703709380/running-a-no-hybrid-open-access-funding-policy}
-
 \leavevmode\hypertarget{ref-crminer}{}%
@@
 
-\leavevmode\hypertarget{ref-Chawla_2020}{}%
-Chawla, D. (2020). This tool is saving universities millions of dollars
-in journal subscriptions. \emph{Science}.
-\url{https://doi.org/10.1126/science.abd7483}
-
-\leavevmode\hypertarget{ref-Creaser_2010}{}%
-Creaser, C., Fry, J., Greenwood, H., Oppenheim, C., Probets, S., Spezi,
-V., \& White, S. (2010). Authors' awareness and attitudes toward open
-access repositories. \emph{New Review of Academic Librarianship},
-\emph{16}(sup1), 145--161.
-\url{https://doi.org/10.1080/13614533.2010.518851}
+\leavevmode\hypertarget{ref-Plan_s}{}%
+cOAlition S. (n.d.). \emph{Plan s: Principles and implementation}.
+\url{https://web.archive.org/web/20210129165306/https://www.coalition-s.org/addendum-to-the-coalition-s-guidance-on-the-implementation-of-plan-s/principles-and-implementation/}.
 
@@
 
-\leavevmode\hypertarget{ref-Els_videos}{}%
-Elsevier. (n.d.-c). \emph{{Publishing Journey videos: How do I complete
-the Rights and Access form?}}
-\url{https://web.archive.org/web/20210127154144/https://service.elsevier.com/app/answers/detail/a_id/29789/supporthub/publishing/track/APN2ZgoIDv8a~RNiGvwa~yKgpv0qOS75Mv9e~zj~PP_X/}.
-
-\leavevmode\hypertarget{ref-lund}{}%
-Elsevier. (n.d.-d). \emph{{Publishing options: BIBSAM institute
-associated authors}}.
-\url{https://web.archive.org/web/20210127153520/https://www.ub.lu.se/en/sites/ub.lu.se.en/files/publicering_under_elsevier_avtalet.pdf}.
-
-\leavevmode\hypertarget{ref-vsnu}{}%
-Elsevier. (n.d.-e). \emph{{Publishing options for: Dutch universities \&
-institutes associated authors}}.
-\url{https://web.archive.org/web/20190530214833/https://www.openaccess.nl/sites/www.openaccess.nl/files/documenten/elsevier_-_vsnu_workflow_updated_18july_v2.pdf}.
-
-\leavevmode\hypertarget{ref-Emery_2013}{}%
-Emery, J. (2013). Mining for gold: Identifying the librarians' toolkit
-for managing hybrid open access. \emph{Insights: The {UKSG} Journal},
-\emph{26}(2), 115--119. \url{https://doi.org/10.1629/2048-7754.65}
+\leavevmode\hypertarget{ref-Els_Pricing}{}%
+Elsevier. (n.d.-c). \emph{{Pricing}}.
+\url{https://web.archive.org/web/20210127152857/https://www.elsevier.com/about/policies/pricing}.
 
@@
 
-\leavevmode\hypertarget{ref-Frantsv_g_2019}{}%
-Frantsvåg, J. E., \& Strømme, T. E. (2019). Few open access journals are
-compliant with plan s. \emph{Publications}, \emph{7}(2), 26.
-\url{https://doi.org/10.3390/publications7020026}
-
 \leavevmode\hypertarget{ref-Fraser_2020}{}%
 Fraser, N., Brierley, L., Dey, G., Polka, J. K., Pálfy, M., Nanni, F.,
-\& Coates, J. A. (2020). \emph{Preprinting the {COVID}-19 pandemic}.
-{bioRxiv}. \url{https://doi.org/10.1101/2020.05.22.111294}
+\& Coates, J. A. (2021). The evolving role of preprints in the
+dissemination of {COVID}-19 research and their impact on the science
+communication landscape. In \emph{{PLOS} Biology} (No. 4; Vol. 19, p.
+e3000959). Public Library of Science ({PLoS}).
+\url{https://doi.org/10.1371/journal.pbio.3000959}
 
@@
 \leavevmode\hypertarget{ref-Graaf_2017}{}%
-Graaf, M. V. D. (2017). \emph{Paying for open access: The author's
+Graaf, M. van der. (2017). \emph{Paying for open access: The author's
 perspective}. Zenodo. \url{https://doi.org/10.5281/ZENODO.438037}
+
+\leavevmode\hypertarget{ref-van_der_graaf_maurits_2017}{}%
+Graaf, M. van der, Johnson, R., \& Chiarelli, A. (2017). \emph{The role
+of hybrid open access in extending author choice}. Zenodo.
+\url{https://doi.org/10.5281/zenodo.3958621}
 
@@
 \url{https://web.archive.org/web/20210109033031/https://www.elsevier.com/connect/what-are-mirror-journals-and-can-they-offer-a-new-world-of-open-access}
-
-\leavevmode\hypertarget{ref-Hendricks_2020}{}%
-Hendricks, G., Tkaczyk, D., Lin, J., \& Feeney, P. (2020). Crossref: The
-sustainable source of community-owned scholarly metadata.
-\emph{Quantitative Science Studies}, \emph{1}(1), 414--427.
-\url{https://doi.org/10.1162/qss_a_00022}
 
@@
 
-\leavevmode\hypertarget{ref-Lawson_2018}{}%
-Lawson, S. (2017). \emph{Report on offset agreements: Evaluating current
-jisc collections deals. Year 2 -- evaluating 2016 deals}. figshare.
-\url{https://doi.org/10.6084/M9.FIGSHARE.5383861.V1}
-
 \leavevmode\hypertarget{ref-Lawson_2016}{}%
@@
 
-\leavevmode\hypertarget{ref-van_Leeuwen_2018}{}%
-Leeuwen, T. N. van, Tatum, C., \& Wouters, P. F. (2018). Exploring
-possibilities to use bibliometric data to monitor gold open access
-publishing at the national level. \emph{Journal of the Association for
-Information Science and Technology}, \emph{69}(9), 1161--1173.
-\url{https://doi.org/10.1002/asi.24029}
+\leavevmode\hypertarget{ref-Marques_2020}{}%
+Marques, M., \& Stone, G. (2020). Transitioning to open access: An
+evaluation of the {UK} springer compact agreement pilot 2016--2018.
+\emph{College {\&} Research Libraries}, \emph{81}(6), 913--927.
+\url{https://doi.org/10.5860/crl.81.6.913}
 
@@
 
-\leavevmode\hypertarget{ref-McNutt_2019}{}%
-McNutt, M. (2019). Opinion: {Plan S} falls short for society
-publishers--and for the researchers they serve. \emph{Proceedings of the
-National Academy of Sciences}, \emph{116}(7), 2400--2403.
-\url{https://doi.org/10.1073/pnas.1900359116}
-
 \leavevmode\hypertarget{ref-Mittermaier_2015}{}%
@@
 
-\leavevmode\hypertarget{ref-Nelson_2017}{}%
-Nelson, G. M., \& Eggett, D. L. (2017). Citations, mandates, and money:
-Author motivations to publish in chemistry hybrid open access journals.
-\emph{Journal of the Association for Information Science and
-Technology}, \emph{68}(10), 2501--2510.
-\url{https://doi.org/10.1002/asi.23897}
+\leavevmode\hypertarget{ref-Niles_2020}{}%
+Niles, M. T., Schimanski, L. A., McKiernan, E. C., \& Alperin, J. P.
+(2020). Why we publish where we do: Faculty publishing values and their
+relationship to review, promotion and tenure expectations. \emph{{PLOS}
+{ONE}}, \emph{15}(3), e0228914.
+\url{https://doi.org/10.1371/journal.pone.0228914}
 
@@
 Publishing}. \url{https://doi.org/10.1787/24132764}
-
-\leavevmode\hypertarget{ref-Olsson1271866}{}%
-Olsson, L. (2018). \emph{Evaluation of offset agreements -- report 4 :
-Springer compact} (p. 20). Kungliga biblioteket.
-\url{http://urn.kb.se/resolve?urn=urn\%3Anbn\%3Ase\%3Ahb\%3Adiva-15501}
 
@@
 
-\leavevmode\hypertarget{ref-Piwowar_2018}{}%
-Piwowar, H., Priem, J., Larivière, V., Alperin, J. P., Matthias, L.,
-Norlander, B., Farley, A., West, J., \& Haustein, S. (2018). The state
-of {OA}: A large-scale analysis of the prevalence and impact of open
-access articles. \emph{{PeerJ}}, \emph{6}, e4375.
-\url{https://doi.org/10.7717/peerj.4375}
-
 \leavevmode\hypertarget{ref-Piwowar_2019}{}%
@@
 
-\leavevmode\hypertarget{ref-Plan_s}{}%
-S, cOAlition. (n.d.). \emph{Plan s: Principles and implementation}.
-\url{https://web.archive.org/web/20210129165306/https://www.coalition-s.org/addendum-to-the-coalition-s-guidance-on-the-implementation-of-plan-s/principles-and-implementation/}.
-
 \leavevmode\hypertarget{ref-Schimmer_2015}{}%
@@
 \url{https://intranet.birmingham.ac.uk/as/libraryservices/library/research/open-access/funding/ukri-open-access-block-grant.aspx}
-
-\leavevmode\hypertarget{ref-oxford_2019}{}%
-University of Oxford. (2019). \emph{Change to oxford's policy for RCUK
-OA block grant 2nd december 2019}.
-\url{http://openaccess.ox.ac.uk/2019/12/02/change-to-oxfords-policy-for-rcuk-oa-block-grant-2nd-december-2019/}
 
@@
 
-\leavevmode\hypertarget{ref-bigrquery}{}%
-Wickham, H., \& Bryan, J. (2020). \emph{{bigrquery}: An interface to
-{Google's BigQuery API}}.
-\url{https://CRAN.R-project.org/package=bigrquery}
-
-\leavevmode\hypertarget{ref-xml2}{}%
-Wickham, H., Hester, J., \& Ooms, J. (2020). \emph{{xml2}: Parse {XML}}.
-\url{https://CRAN.R-project.org/package=xml2}
-
 \end{CSLReferences}

### 2021-05-19 analysis/006_license.Rmd

@@
   expand = expansion(mult = c(0, 0.0005)),
-    labels = scales::percent_format(accuracy = 1L),
-    breaks = c(0, 0.02, 0.04, 0.06, 0.08)
-  ) +
+   labels = scales::percent_format(accuracy = 1),
+   limits=c(0,.1),
+   breaks = c(0, 0.03, 0.06, 0.09)) +
   scale_fill_manual(values = 
@@
 )
-ggsave(here::here("figure", "license_portfolio.png"), p, dpi = 300, width = 6.5, height = 6)
+ggsave(here::here("figure", "license_portfolio.png"), p, dpi = 600, width = 6.5, height = 6)
 ```
